---
title: "Biomass removal"
output: html_document
date: "2023-10-20"
---


```{r setup, include=FALSE}
library(terra)
library(tidyverse)
library(here)
```

# extraction of biomass through fisheries
Prepare the catch data from Watson et al. 2018 (version 5.0). We will calculate total catch and FOFM catch for industrial and non-industrial catch (landings + IUU + discards). 

We use the data from:

1. Watson, R. A. and Tidd, A. 2018. Mapping nearly a century and a half of global marine fishing: 1869–2015. Marine Policy, 93, pp. 171-177. [(Paper URL)](https://doi.org/10.1016/j.marpol.2018.04.023)

```{r}
r <- rast(ncol=720, nrow=360)
values(r) <- 1:259200

catch <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/watson_fisheries/v5.0/Catch2015_2019.csv")


test <- catch %>%
  mutate(biomass = ReportedIND + 
           ReportedNIND) %>%
  select(IYear, Cell, biomass) %>%
  group_by(IYear, Cell) %>%
  summarise(biomass = sum(biomass)) %>%
  filter(IYear == 2017)   
sum(test$biomass)
# NOTE: in 2017 total was 126,256,514 tonnes of catch. Higher than what is reported by SAU,
# but on the high end of what is considered reasonable based on uncertainty around unreported and discards.
# Reported is 82,612,263, and SAUP reports 80,000,000
all_catch_average <- catch %>%
  mutate(biomass = ReportedIND + IUUIND + DiscardsIND + 
+            ReportedNIND + IUUNIND + DiscardsNIND) %>%
  select(IYear, Cell, biomass) %>%
  group_by(IYear, Cell) %>%
  summarise(biomass = sum(biomass)) %>%
  group_by(Cell) %>%
  summarise(biomass = mean(biomass))


# Create an empty raster based on the template
all_catch_raster <- r
values(all_catch_raster) <- NA

# Fill the raster with biomass values
values(all_catch_raster)[all_catch_average$Cell] <- all_catch_average$biomass

plot(all_catch_raster)
global(all_catch_raster, "sum", na.rm=TRUE)

```


getting the data behind figure 2.
```{r testing something}

m_map_path <- "/home/shares/ohi/stressors_2021/_raw_data/Bianchi_fish_biomass/MAKE_FIGURES/datasets"

fig2 <- readMat(file.path(m_map_path, 'fig2_workspace.mat'))
str(fig2)


# Install the 'rhdf5' package
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("rhdf5")

# Load the 'rhdf5' package
library(rhdf5)

# Read the .mat file
file_contents <- h5ls(file.path(m_map_path, 'LME_mask.mat'))
file_contents
lmemask <- h5read(file.path(m_map_path, 'LME_mask.mat'), 'LME_mask')
lmemask[lmemask == 0] <- NA
lmemask[!is.na(lmemask)] <- 1


lmemask_rep <- array(lmemask, dim = c(dim(lmemask), length(fig2$saup10.year)))
## Saup time series data
saup <- readMat(file.path(m_map_path, 'LME_mask.mat'))
str(biomass_rast)

lme_rast <- readMat(file.path(m_map_path, 'saup10.mat'))
str(biomass_rast)

## saup time data
iio <- which.max(fig2$saup10.time)
mean_maxh_year <- mean(fig2$composite["other", , ][[1]][[1]][[1]])
saup10_year1 = fig2$saup10.year - fig2$saup10.year[iio] + mean_maxh_year
saup10_year2 = fig2$saup10.year - fig2$saup10.year[iio]  
npoints <- ncol((fig2$composite["year", , ])$year)
composite.elist.nrun = fig2$composite["elist", , ][[1]][[2]]
afish <- matrix(0, nrow=121, ncol=composite.elist.nrun) ##matrix(0, nrow = npoints*2+1, ncol = composite.elist.nrun)
acntr = npoints+1
stime <- seq(from = -npoints, to = npoints, by = 1)

#1: ??
for (indr in 1:31) { #indr=1
    ypeak = mean(fig2$composite["other", , ][[1]][[1]][[1]])[indr]
   start_idx <- ypeak -100  #max(acntr - ypeak + 1, 1)
    end_idx <- ypeak + 20 # min(acntr, length(fish[, indr]))
    num_elements <- end_idx - start_idx 
    # Assign values to afish
   afish[, indr] <- fish[start_idx:end_idx, indr]
}

fish_mean <- rowMeans(afish)

# fish and harvest data
dim(fig2$composite)
time_element <- fig2$composite["time", , ]

fish <- time_element$time[[1]][[1]]
fish <- apply(fish, c(1, 3), sum)/1e15

harvest <- time_element$time[[1]][[2]]
s2y = 365*86400
hfact = s2y/1e12
harvest <- apply(harvest, c(1,3), sum)*hfact

fish <- fish[1:420, ]
harvest <- harvest[1:420, ]
rowMeans(fish)

other_element <- fig2$composite["other", , ]
year <- other_element$other[[1]][[1]]
mean_maxh_year = mean(composite.other.glo_lme.year)

# end of testing something

```



# Fish biomass preindustrial and max catch
 
peak catch is 1996
```{r}
# Install and load the R.matlab package
#install.packages("R.matlab")
library(R.matlab)

m_map_path <- "/home/shares/ohi/stressors_2021/_raw_data/Bianchi_fish_biomass/MAKE_FIGURES/datasets"
biomass_rast <- readMat(file.path(m_map_path, 'comp3_p10_r10_fvar6_ks24_p50_c7.mat'))
str(biomass_rast)


  # function to format the raster data

format_mat <- function(unfished = FALSE){

if(unfished==TRUE){  
  mean_biomass <- biomass_rast$comp3["y200.fish.mean", , ]$y200.fish.mean}else{         
  mean_biomass <- biomass_rast$comp3["yglo.fish.mean", , ]$yglo.fish.mean
}
  lat <- biomass_rast$comp3["lat", , ]$lat
  lon <- biomass_rast$comp3["lon", , ]$lon
  lon <- ifelse(lon > 180, lon - 360, lon) # adjust longitude display
  
# Define extent and empty SpatRaster
ext <- ext(min(lon), max(lon), min(lat), max(lat))
biomass_raster <- rast(ncol=length(lon), nrow=length(lat), ext=ext)
biomass_matrix <- as.matrix(mean_biomass)
biomass_matrix <- biomass_matrix[nrow(biomass_matrix):1, ] # turn right side up!
# Now adjust longitude!
left_half <- biomass_matrix[, which(lon <= 0)]
right_half <- biomass_matrix[, which(lon > 0)]

# Combine the two halves in the right order
biomass_matrix_adjusted <- cbind(left_half, right_half)
values(biomass_raster) <- biomass_matrix_adjusted
biomass_raster
}

fished_biomass <- format_mat(unfished=FALSE)
unfished_biomass <- format_mat(unfished=TRUE)


# stop fishing when biomass < 2.5 tonnes/km2

# Assume raster_stack is your RasterStack with 50 layers
# Calculate the percent change from year to year
pct_change <- terra::app(c(fished_biomass, unfished_biomass), fun = function(x) {
  # Calculate average percent change
  ((x[1]/x[2])^(1/50)) - 1
  
})

plot(pct_change)
# pct_change now contains the average annual percent change for each cell

change_2050 <- app(c(fished_biomass, pct_change), function(x){
  x[1] * (1 + x[2]) ** 54
})

pressure_2050 <- 1- change_2050/unfished_biomass
plot(pressure_2050)
# Plot the SpatRaster
plot(biomass_raster, main="Mean Biomass", xlab="Longitude", ylab="Latitude")

```


Combine the rasters to determine proportion removed on yearly basis:

```{r}
biomass_raster <- terra::resample(biomass_raster, all_catch_raster, method="bilinear")
# convert from tonnes per km2 to tonnes per cell

# get area of cell (terra returns raster side)
cell_area <- cellSize(biomass_raster, unit="km")
biomass_tonnes <- biomass_raster * cell_area
biomass_tonnes <- biomass_tonnes * 2 #estimate of all biomass, not just targeted
prop <- all_catch_raster/biomass_tonnes
prop[prop>1] <- 1 

writeRaster(all_catch_raster, here("prep/fisheries/data/tonnes_catch.tif"))
writeRaster(prop, here("prep/fisheries/data/prop_catch_biomass.tif"))
writeRaster(biomass_tonnes, here("prep/fisheries/data/tonnes_biomass.tif"))

```

