---
title: "High Seas fishery pressures"
output: html_document
date: "2024-03-20"
---

step1: extract biomass data for regions.
step2: Modeling fishing pressure on the high seas. 

```{r setup, include=FALSE}

library(countrycode)
library(tidyverse)
library(here)


## plotting high seas against EEZs to compare patterns
pressure_avg <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv")) %>%
  mutate(ant_typ = ifelse(!is.na(uninhabited), "eez_uninhabited", ant_typ))

biomass_rgn <- read_csv(here("prep/fisheries/data/country_biomass_2000_2010.csv"))

## some general patterns:
ggplot(pressure_avg, aes(x=avg_percent_change, group=ant_typ, fill=ant_typ)) +
  geom_histogram()
ggplot(pressure_avg, aes(x=slope, group=ant_typ, fill=ant_typ)) +
  geom_histogram()
ggplot(filter(pressure_avg, prop_biomass<2), aes(x=prop_biomass, group=ant_typ, fill=ant_typ)) +
  geom_histogram()

pressure_avg %>% 
  group_by(ant_typ) %>%
  summarize(historic_biomass = sum(biomass_2010, na.rm=TRUE))

## image of overall high seas biomass
hs_biomass_mean <- biomass_rgn %>%
  filter(ant_typ %in% c("fao", "eez-ccamlr")) %>%
  select(year, biomass) %>%
  group_by(year) %>%
  summarize(biomass=sum(biomass, na.rm=TRUE))

ggplot(hs_biomass_mean, aes(x=year, y=biomass)) + 
  geom_line() 

check <- hs_biomass_mean %>%
  arrange(year) %>%
  mutate(yearly_percent_change = (biomass - lag(biomass))/lag(biomass)) 

ggplot(check, aes(x=year, y=yearly_percent_change)) + 
  geom_line() 


```


Parameters for future predictions.
The mean and sd of the annual rate of change in biomass across fao regions is used for random sampling future change.
```{r}

hs_rgn_avg <- pressure_avg %>%
  filter(ant_typ %in% c("fao", "eez-ccamlr"))

```


Creating the high seas, region specific data:
```{r}

nsims <- 500

hs_biomass_rgn <- biomass_rgn %>%
  filter(ant_typ %in% c("fao", "eez-ccamlr"))
  
scenario_year_template <- expand.grid(scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"),
                                      year = 2000:2100,
                                      rgn_id = unique(hs_biomass_rgn$rgn_id),
                                      simulation = paste0("V", 1:nsims))

hs_average <- pressure_avg %>%
  select(rgn_id, iso3c, rgn_nam, biomass_2005, biomass_2010, avg_percent_change, historic_1900)

hs_scenario <- left_join(scenario_year_template, hs_average, by=c("rgn_id")) %>%
  mutate(biomass = NA, prop=NA)

## add in random annual change
hs_scenario$rate_change <- rnorm(dim(hs_scenario)[1], mean = mean(hs_rgn_avg$avg_percent_change), sd = sd(hs_rgn_avg$avg_percent_change))
hs_scenario <- hs_scenario %>%
  mutate(rate_change = ifelse(scenario == "SSP1", 0.001, rate_change)) 
```

Function to update yearly biomass based on randomized sample:
```{r}

update_biomass <- function(df) {
  # df <- new_data_sim
  #print(df)
  
  df$prop[1] <- df$biomass_2010[1] / df$historic_1900[1]
  df$biomass[1] <- df$biomass_2005[1]  # starting value is 2005

  scenario <- df$scenario2[1]
  
  # print(scenario)

    if (scenario == "SSP1") {
        upper_threshold <- 0.99
      } else if (scenario == "SSP2") {
      lower_threshold <- 0.75
    } else if (scenario == "SSP3" || scenario == "SSP4" || scenario == "SSP5") {
      lower_threshold <- 0.60
    } else {
    stop("Error: Unrecognized scenario value.")
    }
   
  for (i in 2:nrow(df)) { # i=2
    
    potential_biomass <- df$biomass[i - 1] * (1 + df$rate_change[i-1])
    potential_prop <- potential_biomass / df$historic_1900[i]


     adjusted_rate_change <- df$rate_change[i-1]

     
    # Constraints
     if (scenario == "SSP1" && potential_prop > upper_threshold) {
      adjusted_rate_change <- 0 
         }
      if (scenario != "SSP1" && potential_prop < lower_threshold) {
      adjusted_rate_change <- 0         
        }
   
     ## Make the biomass approach thresholds more slowly  
  #      if (scenario == "SSP1") {
  #    adjust_factor <- (potential_prop/upper_threshold - 1) * (-1)
  #   }
  #    if (scenario != "SSP1") {
  #    adjust_factor <- (potential_prop/lower_threshold - 1) * (-1)
  #  }
    adjust_factor <- 1
    # Apply the adjusted percent change and update yearly values
    df$biomass[i] <- df$biomass[i - 1] * (1 + adjusted_rate_change * adjust_factor)
    df$prop[i] <- df$biomass[i] / df$historic_1900[i]
  }

  return(df)
}

```

Run function
```{r}

# Apply the function to each group
new_data_sim <- hs_scenario %>%
  filter(year>=2005) %>%
  mutate(scenario2 = scenario) %>%
  arrange(rgn_id, rgn_nam, simulation, scenario, year) %>%
  group_by(iso3c, rgn_id, rgn_nam, simulation, scenario) %>%
  group_modify(~ update_biomass(.x)) %>%
  ungroup()  %>%
  select(iso3c, rgn_id, rgn_nam, scenario, simulation, year, biomass)

write_csv(new_data_sim, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/random_hs_%s.csv", Sys.Date()))
new_data_sim <- read_csv(sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/random_hs_%s.csv", "2024-06-18"))
```


Plot the data
```{r}

rgn <- 273

rgn_biomass_stats <- new_data_sim %>% 
  filter(rgn_id==rgn) %>%
  group_by(year, scenario) %>%
  summarize(
    median_biomass = median(biomass),
    p10 = quantile(biomass, 0.1),
    p90 = quantile(biomass, 0.9),
    .groups = 'drop'
  )

thresholds <- filter(pressure_avg, rgn_id==rgn)
name <- pressure_avg$rgn_nam[pressure_avg$rgn_id==rgn]
  
ggplot(rgn_biomass_stats, aes(x = year, y = median_biomass, color = scenario)) +
  geom_line() +
  geom_point() +
# geom_ribbon(aes(ymin = p10, ymax = p90, fill = scenario), alpha = 0.2) +
  labs(title = paste(name, "Median Biomass by scenario"),
       x = "Year",
       y = "Biomass") +
  geom_hline(yintercept = thresholds$historic_1900, col="blue") + 
  geom_hline(yintercept = thresholds$historic_1900 * 0.75, col="green") + 
  geom_hline(yintercept = thresholds$historic_1900 * 0.60, col="purple") + 
  theme_minimal()

#global
global_biomass_stats <- new_data_sim %>%
  group_by(year, scenario, simulation) %>%
  summarize(total_biomass = sum(biomass)) %>%
  ungroup() %>%
  group_by(year, scenario) %>%
  summarize(
    median_biomass = median(total_biomass),
    p10 = quantile(total_biomass, 0.1),
    p90 = quantile(total_biomass, 0.9),
    .groups = 'drop'
  )

ggplot(filter(global_biomass_stats, year <= 2100), aes(x = year, y = median_biomass, color = scenario)) +
  geom_line() +
  geom_point() +
# geom_ribbon(aes(ymin = p10, ymax = p90, fill = scenario), alpha = 0.2) +
  labs(title = "Global: Median Biomass for Global by Scenario",
       x = "Year",
       y = "Biomass") +
  theme_minimal()


```


compare to observed
```{r}
simulation <- new_data_sim %>%
  group_by(year, scenario, simulation) %>%
  summarize(total_biomass = sum(biomass)) %>%
  ungroup() %>%
  filter(year %in% 2005:2010)


# observed data: this file created in step1
country_biomass_change <- read_csv("prep/fisheries/data/country_biomass_2000_2010.csv") %>%
  filter(rgn_id %in% unique(new_data_sim$rgn_id))

observed <- country_biomass_change %>%
  group_by(year) %>%
  summarize(observed_biomass = sum(biomass, na.rm=TRUE))

ggplot() + 
  geom_line(data=filter(simulation, scenario == "SSP4"), aes(x=year, y=total_biomass, group=simulation), alpha=0.2) +
  geom_point(data=filter(observed, year>=2005), aes(x=year, y=observed_biomass), col="red")




```