---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

step1: extract biomass data for eezs.
step2: data and models to see if we can predict changes to fisheries pressures
This is an information gathering step that can be skipped. 

Outcome: governance was the only variable that really seemed to predict fishing pressures. 
```{r setup, include=FALSE}

library(countrycode)
library(tidyverse)
library(here)

pressure_eez <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv"))

# cut out uninhabited/sparsely habited islands and FAO major fishing areas beyond the EEZs,
# these are estimated elsewhere.
pressure_eez <- pressure_eez %>%
  filter(ant_typ == "eez"|iso3c=="DIS") %>%
  filter(is.na(uninhabited)) %>%
  filter(!is.na(slope)) %>%
  select(-uninhabited)

# n = 184
```


Gathering variables with future scenario data that might be useful.
```{r}

gdp <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "GDP|PPP") %>%
  filter(MODEL %in% c("OECD Env-Growth", "IIASA GDP")) %>%
  separate(SCENARIO, into=c("scenario", "model_info"), sep="_", extra="merge") %>%
  select(MODEL, scenario, iso3c=REGION, unit=UNIT, `1950`:`2100`) %>%
  pivot_longer(
     cols = `1950`:`2100`,  # Adjust this to cover all the year columns
    names_to = "year",
    values_to = "gdp_ppp",
    names_transform = list(Year = as.integer) 
  ) %>%
  group_by(scenario, iso3c, unit) %>%
  summarize(gdp_ppp = mean(gdp_ppp, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(!is.na(gdp_ppp)) %>%
  filter(scenario == "SSP1") %>%
  select(-scenario)
  
pop <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "Population") %>%
  filter(MODEL %in% c("IIASA GDP", "IIASA-WiC POP", "NCAR", "OECD Env-Growth")) %>%
  separate(SCENARIO, into=c("scenario", "model_info"), sep="_", extra="merge") %>%
  select(MODEL, scenario, iso3c=REGION, unit=UNIT, `1950`:`2100`) %>%
  pivot_longer(
     cols = `1950`:`2100`,  # Adjust this to cover all the year columns
    names_to = "year",
    values_to = "pop",
    names_transform = list(Year = as.integer) 
  ) %>%
  group_by(scenario, iso3c, unit) %>%
  summarize(pop = mean(pop, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(!is.na(pop)) %>%
  filter(scenario == "SSP1") %>%
  select(-scenario)

pressure_eez2 <- left_join(pressure_eez, gdp, by="iso3c") %>%
         left_join(pop, by="iso3c") 

## some extra data
gf_regions <- read_csv(here("prep/fisheries/data/country_pop_gf.csv")) %>%
  select(iso3c, pop_gf = population, gdp_gf = gdp_current_USD) %>%
  mutate(pop_gf = pop_gf/1000000,
         gdp_gf = gdp_gf/1000000000)

pressure_eez3 <- left_join(pressure_eez2, gf_regions, by="iso3c") %>%
  mutate(pop = ifelse(is.na(pop), pop_gf, pop),
         gdp_ppp = ifelse(is.na(gdp_ppp), gdp_gf, gdp_ppp)) %>%
         mutate(pc_gdp = gdp_ppp/pop) %>%
  select(-pop_gf, -gdp_gf)


gini <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/NRao_et_al_GiniProjections_2018/Gini_projections_SSPs.csv") %>%
  filter(year == c(2011)) %>%
  filter(scenario == "SSP1") %>%
  pivot_longer( cols = -c(scenario, year),  
  names_to = "iso3c",
  values_to = "gini") %>%
  select(iso3c, gini)

pressure_eez3 <- left_join(pressure_eez3, gini, by="iso3c") 

gov <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(year%in% c(2000:2010) & scenario=="Observed") %>%
  group_by(countrycode, country, scenario) %>%
  summarize(governance = mean(governance, na.rm=TRUE)) %>%
  rename("iso3c" = "countrycode") %>%
  ungroup() %>%
  select(iso3c, governance)


pressure_eez4 <- left_join(pressure_eez3, gov, by="iso3c")

eez_area <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/rgn_area.csv")

pressure_eez5 <- left_join(pressure_eez4, eez_area, by="rgn_id")

georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv")

pressure_eez6 <- left_join(pressure_eez5, georgns, by="rgn_id")


pressure_eez6 <- pressure_eez6  %>% 
  mutate(ln_gdp_ppp = log(gdp_ppp+1),
         ln_pop = log(pop + 1),
         ln_pop_ln_area = log(pop)/log(area_km2),
         ln_pc_gdp = log(pc_gdp),
         ln_area = log(area_km2)) 


```



## some models
I tried a range of models, and governance was the only one that seemed predictive.
```{r}

library(ranger)

library(caret)
# Set up training control
train_control <- trainControl(method = "LOOCV", savePredictions=TRUE)

# Train the model with leave-one-out cross-validation
data8 <- na.omit(pressure_eez6) %>%
  mutate(pressure_easing = ifelse(slope>0, "easing", "worsening"))

model <- train(slope ~  governance, 
               data = na.omit(data8), 
               method = "lm", 
               trControl = train_control)
model
summary(model)
plot(model$pred$pred ~ model$pred$obs)


model <- train(as.factor(pressure_easing) ~  governance, 
               data = na.omit(data8), 
               method = "glm", 
               trControl = train_control)
model
plot(model$pred$pred ~ model$pred$obs)

```



