---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

step1: extract biomass data for regions.
step2: data and models to see if we can predict changes to fisheries pressures
This is an information gathering step that can be skipped. 

```{r setup, include=FALSE}

library(countrycode)
library(tidyverse)
library(here)

pressure_avg <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv")) %>%
  mutate(ant_typ = ifelse(!is.na(uninhabited), "eez_uninhabited", ant_typ))

biomass_rgn <- read_csv(here("prep/fisheries/data/country_biomass_2000_2010.csv"))

## some general patterns:
ggplot(pressure_avg, aes(x=avg_percent_change, group=ant_typ, fill=ant_typ)) +
  geom_histogram()
ggplot(pressure_avg, aes(x=slope, group=ant_typ, fill=ant_typ)) +
  geom_histogram()
ggplot(filter(pressure_avg, prop_biomass<2), aes(x=prop_biomass, group=ant_typ, fill=ant_typ)) +
  geom_histogram()

pressure_avg %>% 
  group_by(ant_typ) %>%
  summarize(historic_biomass = sum(historic_1900, na.rm=TRUE))

pressure_avg %>% 
  group_by(ant_typ) %>%
  summarize(historic_biomass = sum(biomass_2010, na.rm=TRUE))

```


Combining the high seas and ccamlr regions to get future predictions.
```{r}

overall <- biomass_rgn %>%
  filter(uninhabited %in% c("x")) %>%
  select(year, biomass) %>%
  group_by(year) %>%
  summarize(biomass=sum(biomass, na.rm=TRUE))

ggplot(overall, aes(x=year, y=biomass)) + 
  geom_line() 

overall <- overall %>%
  arrange(year) %>%
  mutate(yearly_percent_change = (biomass - lag(biomass))/lag(biomass)) 

ggplot(overall, aes(x=year, y=yearly_percent_change)) + 
  geom_line() 

mean(overall$yearly_percent_change, na.rm=TRUE)
pressure_avg %>% 
  summarize(historic_biomass = sum(biomass_2010, na.rm=TRUE))

  group_by()

biomass_rgn %>%
  filter(year==2010) %>%
  group_by(ant_typ) %>%
  summarize(total_biomass = sum(biomass, na.rm=TRUE))

biomass_rgn %>%
  filter(year==2010) %>%
  group_by(ant_typ) %>%
  summarize(total_historic_biomass = sum(historic_1900, na.rm=TRUE))

# cut out eezs,
biomass_hs <- pressure_rgns %>%
  filter(ant_typ == "fao") %>%
  filter(is.na(uninhabited)) %>%
  filter(!is.na(slope)) %>%
  select(-uninhabited)

# n = 184
```


Gathering variables with future scenario data that might be useful.
```{r}

gdp <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "GDP|PPP") %>%
  filter(MODEL %in% c("OECD Env-Growth", "IIASA GDP")) %>%
  separate(SCENARIO, into=c("scenario", "model_info"), sep="_", extra="merge") %>%
  select(MODEL, scenario, iso3c=REGION, unit=UNIT, `1950`:`2100`) %>%
  pivot_longer(
     cols = `1950`:`2100`,  # Adjust this to cover all the year columns
    names_to = "year",
    values_to = "gdp_ppp",
    names_transform = list(Year = as.integer) 
  ) %>%
  group_by(scenario, iso3c, unit) %>%
  summarize(gdp_ppp = mean(gdp_ppp, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(!is.na(gdp_ppp)) %>%
  filter(scenario == "SSP1") %>%
  select(-scenario)
  
pop <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "Population") %>%
  filter(MODEL %in% c("IIASA GDP", "IIASA-WiC POP", "NCAR", "OECD Env-Growth")) %>%
  separate(SCENARIO, into=c("scenario", "model_info"), sep="_", extra="merge") %>%
  select(MODEL, scenario, iso3c=REGION, unit=UNIT, `1950`:`2100`) %>%
  pivot_longer(
     cols = `1950`:`2100`,  # Adjust this to cover all the year columns
    names_to = "year",
    values_to = "pop",
    names_transform = list(Year = as.integer) 
  ) %>%
  group_by(scenario, iso3c, unit) %>%
  summarize(pop = mean(pop, na.rm=TRUE)) %>%
  ungroup() %>%
  filter(!is.na(pop)) %>%
  filter(scenario == "SSP1") %>%
  select(-scenario)

pressure_eez2 <- left_join(pressure_eez, gdp, by="iso3c") %>%
         left_join(pop, by="iso3c") 

## some extra data
gf_regions <- read_csv(here("prep/fisheries/data/country_pop_gf.csv")) %>%
  select(iso3c, pop_gf = population, gdp_gf = gdp_current_USD) %>%
  mutate(pop_gf = pop_gf/1000000,
         gdp_gf = gdp_gf/1000000000)

pressure_eez3 <- left_join(pressure_eez2, gf_regions, by="iso3c") %>%
  mutate(pop = ifelse(is.na(pop), pop_gf, pop),
         gdp_ppp = ifelse(is.na(gdp_ppp), gdp_gf, gdp_ppp)) %>%
         mutate(pc_gdp = gdp_ppp/pop) %>%
  select(-pop_gf, -gdp_gf)


gini <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/NRao_et_al_GiniProjections_2018/Gini_projections_SSPs.csv") %>%
  filter(year == c(2011)) %>%
  filter(scenario == "SSP1") %>%
  pivot_longer( cols = -c(scenario, year),  
  names_to = "iso3c",
  values_to = "gini") %>%
  select(iso3c, gini)

pressure_eez3 <- left_join(pressure_eez3, gini, by="iso3c") 

gov <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(year%in% c(2000:2010) & scenario=="Observed") %>%
  group_by(countrycode, country, scenario) %>%
  summarize(governance = mean(governance, na.rm=TRUE)) %>%
  rename("iso3c" = "countrycode") %>%
  ungroup() %>%
  select(iso3c, governance)


pressure_eez4 <- left_join(pressure_eez3, gov, by="iso3c")

eez_area <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/rgn_area.csv")

pressure_eez5 <- left_join(pressure_eez4, eez_area, by="rgn_id")

georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv")

pressure_eez6 <- left_join(pressure_eez5, georgns, by="rgn_id")


pressure_eez6 <- pressure_eez6  %>% 
  mutate(ln_gdp_ppp = log(gdp_ppp+1),
         ln_pop = log(pop + 1),
         ln_pop_ln_area = log(pop)/log(area_km2),
         ln_pc_gdp = log(pc_gdp),
         ln_area = log(area_km2)) 


```

# some models to gapfill missing data

Currently I do not use gapfilled data, and this code would need to be updated.
```{r}
###
pressure_eez7 <- pressure_eez6
mod_gov1 <- lm(governance ~ pc_gdp + r2_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gov1)

pressure_eez7$gov_gf1 <- predict(mod_gov1, newdata = pressure_eez7 %>%
                                               mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

mod_gov2 <- lm(governance ~ pc_gdp + r1_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gov2)

pressure_eez7$gov_gf2 <- predict(mod_gov2, newdata = pressure_eez7)

pressure_eez7 <- pressure_eez7 %>%
  mutate(governance = ifelse(is.na(governance), gov_gf1, governance),
         governance = ifelse(is.na(governance), gov_gf2, governance),
         governance = ifelse(governance > 0.94154, 0.94154, governance)) # set to max observed value

mod_gini1 <- lm(gini ~ pc_gdp + governance + r2_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gini1)

pressure_eez7$gini_gf1 <- predict(mod_gini1, newdata = pressure_eez7 %>%
                                               mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

mod_gini2 <- lm(gini ~ pc_gdp + governance + r1_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gini2)

pressure_eez7$gini_gf2 <- predict(mod_gini2, newdata = pressure_eez7)

pressure_eez7 <- pressure_eez7 %>%
  mutate(gini = ifelse(is.na(gini), gini_gf1, gini),
           gini = ifelse(is.na(gini), gini_gf2, gini)) %>%
  select(-gini_gf1, -gini_gf2, -gov_gf1, -gov_gf2)


```



## some models
I tried a range of models, and governance was the only one that seemed predictive.
```{r}

library(ranger)

library(caret)
# Set up training control
train_control <- trainControl(method = "LOOCV", savePredictions=TRUE)

# Train the model with leave-one-out cross-validation
data8 <- na.omit(pressure_eez6) %>%
  mutate(pressure_easing = ifelse(slope>0, "easing", "worsening"))

model <- train(slope ~  governance, 
               data = na.omit(data8), 
               method = "lm", 
               trControl = train_control)
model
summary(model)
plot(model$pred$pred ~ model$pred$obs)


model <- train(as.factor(pressure_easing) ~  governance, 
               data = na.omit(data8), 
               method = "glm", 
               trControl = train_control)
model
plot(model$pred$pred ~ model$pred$obs)

```



## end of this script

## Prepping gov data

```{r}
# what governance looks like:
gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  group_by(scenario, year) %>%
  summarize(governance_mean = mean(governance, na.rm=TRUE),
            sd = sd(governance, na.rm=TRUE))
ggplot(gov_raw, aes(x=year, y=governance_mean, group=scenario, color=scenario)) +
  geom_point() +
  geom_line()

ggplot(data8, aes(x=governance, y=slope)) + 
  geom_point() + 
  geom_smooth()


# going to estimate yearly change for gapfilling regions with current estimates, but no future estimates
# just using the average patterns
gov_ssp_change_data <- gov_raw %>%
  group_by(scenario) %>%
  select(scenario, year, governance_mean) %>%
  filter(year >= 2015) %>%
  arrange(year) 

start_gov <- filter(gov_ssp_change_data, year==2015 & scenario != "Observed") %>%
  select(scenario, start_governance_mean = governance_mean)

gov_ssp_change <- gov_ssp_change_data %>%
  left_join(start_gov, by= "scenario") %>%
  mutate(governance_yearly_change = governance_mean - start_governance_mean) %>%
  filter(scenario != "Observed") %>%
  select(scenario, year, governance_yearly_change)

ggplot(gov_ssp_change, aes(x=year, y=governance_yearly_change, group=scenario)) + 
  geom_line()


## preparing gov_data
gov_data <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  select(iso3c=countrycode, scenario, year, governance)

correction <- gov_data %>%
  filter(year == 2015) %>%
  group_by(iso3c) %>%
  mutate(observed = ifelse(scenario == "Observed", governance, NA_real_)) %>%
  fill(observed, .direction = "downup") %>%  # Filling NA values within each group
  mutate(correction_value = governance-observed) %>%
  filter(scenario != "Observed") %>%
  select(iso3c, scenario, correction_value)
  

obs_gov <- gov_data %>%
  filter(scenario == "Observed") %>%
  filter(year>2010) %>%
  select(iso3c, year, governance)
obs_gov <- crossing(obs_gov, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))

est_gov <- gov_data %>%
  filter(scenario != "Observed") %>%
  filter(!is.na(scenario)) %>%
  filter(year > 2015) %>% 
  select(iso3c, year, governance, scenario)
  
gov_scores <- rbind(obs_gov, est_gov) %>%
  left_join(correction, by=c("iso3c", "scenario")) %>%
  mutate(correction_value = ifelse(year <= 2015, 0, correction_value)) %>%
  mutate(corrected_governance = governance - correction_value) 

check <- gov_scores %>%
  group_by(year, scenario) %>%
  summarize(gov = mean(corrected_governance, na.rm=TRUE))

ggplot(check, aes(x=year, y=gov, group=scenario)) +
  geom_line() + 
  geom_point() 

ggplot(filter(gov_scores, iso3c=="USA"), aes(x=year, y=governance, group=scenario)) +
  geom_line() + 
  geom_point() +
  geom_line(aes(x=year, y=corrected_governance, group=scenario), col="red")

gov_scores <- gov_scores %>%
  select(iso3c, year, scenario, governance=corrected_governance) %>%
  filter(!is.na(governance))


```


## sampling


```{r}

sim_data <- pressure_eez7 %>%
  filter(ant_typ == "eez") %>%
  select(ant_typ, rgn_id, rgn_nam, iso3c, avg_biomass, avg_percent_change, slope, governance) %>%
  mutate(year = 2005) 

sim_data <- na.omit(sim_data)

sample_low_gov <- sim_data %>%
  filter(governance < 0.4) %>%
  pull(slope)
low_gov_avg <- mean(sample_low_gov)
low_gov_sd <- sd(sample_low_gov)

sample_mid_gov <- sim_data %>%
  filter(governance >= 0.4 & governance < 0.75) %>%
  pull(slope)
mid_gov_avg <- mean(sample_mid_gov)
mid_gov_sd <- sd(sample_mid_gov)

sample_high_gov <- sim_data %>%
  filter(governance >= 0.75) %>%
  pull(slope)
high_gov_avg <- mean(sample_high_gov)
high_gov_sd <- sd(sample_high_gov)

low_gov_avg
mid_gov_avg
high_gov_avg

sim_data_final <- sim_data %>%
  ungroup() %>%
  select(iso3c, biomass_t0 = avg_biomass, pct_change_t0 = avg_percent_change, slope, governance, year)


## prepping and gapfilling governance scenario data

#missing_recent <- data.frame(iso3c = c("ESH", "COD", "ROU"),
#                            governance = c(mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="ESH"]),
#                            mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="COD"]),
#                            mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="ROU"])))
#missing_recent <- crossing(missing_recent, year = 2011:2015, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))
#gov_scores_gf <- rbind(gov_scores, missing_recent)

missing_gov_iso <- setdiff(sim_data_final$iso3c, gov_scores$iso3c)
missing_gov <- expand.grid(iso3c = missing_gov_iso, year = c(2011:2099), scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))

missing_gov_estimated <- filter(sim_data_final, iso3c %in% missing_gov_iso) %>%
  select(iso3c, governance)
missing_gov_estimated <- crossing(missing_gov_estimated, year= 2011:2099, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5")) %>%
  left_join(gov_ssp_change, by=c("scenario", "year")) %>%
  mutate(governance_yearly_change = ifelse(is.na(governance_yearly_change), 0, governance_yearly_change)) %>%
  mutate(governance2 = governance + governance_yearly_change) %>%
  select(iso3c, year, governance=governance2, scenario)
ggplot(filter(missing_gov_estimated, iso3c=="ABW"), aes(x=year, y=governance, group=scenario)) +
  geom_line()

gov_scores_gf <- rbind(gov_scores, missing_gov_estimated)
setdiff(gov_scores_gf$iso3c, sim_data_final$iso3c) # mostly landlocked
setdiff(sim_data_final$iso3c, gov_scores_gf$iso3c)
## now prepare sample data

########################################
## function to run randomization:
########################################
randomized_fishing <- function(scenario_selects="SSP1", sim_data_finals = sim_data_final, gov_scores_gfs=gov_scores_gf){

  for (year in 2005:2099){ #year = 2010 scenario_selects="SSP1"

scenario_select <- scenario_selects

if(year==2005){
t0 <- sim_data_finals %>%
  select(-slope) %>%
  mutate(scenario = scenario_select) %>%
  mutate(!!paste0("biomass_", year) := biomass_t0) %>%
  mutate(!!paste0("pct_change_", year) := pct_change_t0) 

} else{
  t0 <- simulation
}

  if(year %in% c(2005:2010)){
simulation <- t0 %>%
  rowwise() %>%
  mutate(
    change_pct_change_random = case_when(
      governance < 0.4 ~ rnorm(1, mean=low_gov_avg, sd=low_gov_sd),
      governance >= 0.4 & governance < 0.75 ~ rnorm(1, mean=mid_gov_avg, sd=mid_gov_sd),
      governance >= 0.75 ~ rnorm(1, mean=high_gov_avg, sd = high_gov_sd))) %>%
  mutate(pct_change_t1 = pct_change_t0 + change_pct_change_random) %>%
  mutate(!!paste0("pct_change_", year+1) := pct_change_t1) %>%
  mutate(!!paste0("change_pct_change", year+1) := change_pct_change_random) %>%
  mutate(biomass_t1 = biomass_t0 * (1 + pct_change_t1)) %>%
  mutate(!!paste0("biomass_", year+1) := biomass_t1) %>%
  mutate(year=year+1) %>%
  select(-biomass_t0, -pct_change_t0) %>%
  rename(biomass_t0 = biomass_t1, pct_change_t0=pct_change_t1) 

#cat(year, "\n", "dim = ", dim(simulation)[1], "\n") # for troubleshooting
  } else {
    
simulation <- t0 %>%
  select(-governance) %>%
  left_join(gov_scores_gfs, by=c("iso3c", "year", "scenario")) %>%
  rowwise() %>%
  mutate(
    change_pct_change_random = case_when(
      governance < 0.4 ~ rnorm(1, mean=low_gov_avg, sd=low_gov_sd),
      governance >= 0.4 & governance < 0.75 ~ rnorm(1, mean=mid_gov_avg, sd=mid_gov_sd),
      governance >= 0.75 ~ rnorm(1, mean=high_gov_avg, sd = high_gov_sd))) %>%
  mutate(pct_change_t1 = pct_change_t0 + change_pct_change_random) %>%
  mutate(!!paste0("pct_change_", year+1) := pct_change_t1) %>%
  mutate(!!paste0("change_pct_change", year+1) := change_pct_change_random) %>%
  mutate(biomass_t1 = biomass_t0 * (1 + pct_change_t1)) %>%
  mutate(!!paste0("biomass_", year+1) := biomass_t1) %>%
  mutate(year=year+1) %>%
  select(-biomass_t0, -pct_change_t0) %>%
  rename(biomass_t0 = biomass_t1, pct_change_t0=pct_change_t1) 
#cat(year, "\n", "dim = ", dim(simulation)[1], "\n") # for troubleshooting

  }
  }
  return(simulation)
}

########################################
## end of function to run randomization
########################################

summary(simulation_long)
ggplot(filter(simulation_long, year <=2060), aes(x=year, y=biomass, group=iso3c, col=governance)) +
  geom_line()

simulation_long %>%
  group_by(year) %>%
  summarize(sum(biomass, na.rm=TRUE)) %>% print(n=100)




## Run simulations

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP1", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)

scenario_select = "SSP1"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}


for(i in 1:100){
  
tmp <- randomized_fishing(scenario_selects="SSP2", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)

scenario_select = "SSP2"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP3", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select="SSP3"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP4", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select <- "SSP4"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP5", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select <- "SSP5"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

```


```{r}
tmp <- read_csv("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_SSP1_s1.csv")

files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/", full=TRUE, pattern=".csv")

rand_biomass <- lapply(files, read.csv) %>% bind_rows()

## compare to global observed: 2005:2010
## ensure estimates are not going off the rails.
compare <- rand_biomass %>%
  group_by(simulation, scenario, year) %>%
  summarize(global_biomass = sum(biomass)) %>%
  filter(year %in% 2005:2010)
compare %>%
  group_by(year) %>%
  summarize(simulated_global_biomass = sum(global_biomass))

check <- country_biomass_change
check$iso3c <- countrycode(check$rgn_nam, "country.name", "iso3c")
check <- filter(check, iso3c %in% unique(tmp$iso3c))
observed <- check %>%
  group_by(year) %>%
  summarize(observed_biomass = sum(biomass, na.rm=TRUE) - 3865822)

ggplot() + 
  geom_line(data=compare, aes(x=year, y=global_biomass, group=simulation), alpha=0.2) +
  geom_point(data=filter(observed, year>=2005), aes(x=year, y=observed_biomass), col="red")

## note that the 2005 biomass estimate is lower in simulated data data due to averaging 2000-2010 
## to arrive at this value. (consider changing)
692322460-688456638

US_means <- rand_biomass %>%
  filter(iso3c=="USA" & scenario=="SSP1" & year %in% c(2010:2050)) %>%
  group_by(year) %>%
    summarize(mean_biomass = mean(biomass))

ggplot(data=filter(rand_biomass, iso3c=="USA" & scenario=="SSP1" & year%in%c(2010:2050)), aes(x=year, y=biomass)) +
  geom_point() +
  geom_point(data=US_means, aes(x=year, y=mean_biomass), col="red") +
    geom_line(data=US_means, aes(x=year, y=mean_biomass), col="red")
  


## constrain data, to not go above or below a particular percentage.
## get reference. biomass:
ref_times <- which(as.Date(time(biomass)) >= as.Date("1900-01-01") & as.Date(time(biomass)) < as.Date("1910-01-01")) 
ref_density_rasters <- subset(biomass, ref_times)
ref_biomass_rasters <- cellSize(ref_density_rasters) * ref_density_rasters * 1e-6
ref_biomass <- mean(ref_biomass_rasters)

ref_country_biomass <- terra::zonal(ref_biomass, terra::rast(zones_raster), fun = sum, na.rm=TRUE)

ref_country_biomass <- ref_country_biomass %>% 
 rename(rgn_id = layer, ref_biomass=mean) %>%
  left_join(rgn_data, by="rgn_id") %>%
  filter(!is.na(ref_biomass)) %>%
  filter(ant_typ == "eez")

ref_country_biomass$iso3c <- countrycode(ref_country_biomass$rgn_nam, "country.name", "iso3c") %>%
  filter(!is.na(iso3c))

## Get time 0 biomass
t0_times <- which(as.Date(time(biomass)) >= as.Date("1905-01-01") & as.Date(time(biomass)) < as.Date("1906-01-01")) 
t0_density_rasters <- subset(biomass, t0_times)
t0_biomass_rasters <- cellSize(biomass_c) * biomass_c * 1e-6
t0_biomass <- mean(ref_biomass_rasters)

ref_country_biomass <- terra::zonal(ref_biomass, terra::rast(zones_raster), fun = sum, na.rm=TRUE)

ref_country_biomass <- ref_country_biomass %>% 
 rename(rgn_id = layer, ref_biomass=mean) %>%
  left_join(rgn_data, by="rgn_id") %>%
  filter(!is.na(ref_biomass)) %>%
  filter(ant_typ == "eez")

ref_country_biomass$iso3c <- countrycode(ref_country_biomass$rgn_nam, "country.name", "iso3c") %>%
  filter(!is.na(iso3c))


```