---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

```{r setup, include=FALSE}

library(sf)
library(terra)
library(countrycode)
library(broom)
library(tidyverse)
library(here)
library(rnaturalearth)

```

## Prepare region data

```{r}
## EEZ regions
eez_shp <- sf::st_read("/home/shares/ohi/git-annex/globalprep/spatial/v2017/regions_2017_update.shp") 
eez <- sf::st_transform(eez_shp, 4326)
eez_data <- read_csv(here("prep/fisheries/data/region_id.csv"))
eez_data$iso3c[duplicated(eez_data$iso3c)]

```

# BOATS: biomass model based on effort

```{r}

effort_data_files <- "/home/shares/ohi/stressors_2021/_raw_data/Bianchi_fish_biomass/BOATS_effort"

biomass <- rast(file.path(effort_data_files, "boats-v2_gfdl-mom6-cobalt2_none_obsclim_histsoc_onedeg_tcb_global_annually_1841-2010.nc"))

```

# Explore raster scale data
Isolate years between 2000 and 2010. For each cell, calculate rate of change for each year. 
Average of 2000 to 2010.

```{r}

biomass_times <- which(as.Date(time(biomass)) >= as.Date("2000-01-01") & as.Date(time(biomass)) < as.Date("2011-01-01")) 
biomass_c <- subset(biomass, biomass_times)
biomass_c <- cellSize(biomass_c) * biomass_c * 1e-6

# Initialize an empty SpatRaster to store percent change values
percent_change <- biomass_c[[1:10]]
values(percent_change) <- NA # Fill with NAs since the first layer can't calculate change

# Calculate percent change for each layer, starting from the second layer
for (i in 2:nlyr(biomass_c)) { # i = 2 
  # Calculate percent change
  pc <- (biomass_c[[i]] - biomass_c[[i-1]]) / biomass_c[[i-1]] * 100
  
  # Assign calculated values to the corresponding layer in the percent_change_raster
  percent_change[[i-1]] <- pc
}

avg_percent_change <- mean(percent_change)
plot(avg_percent_change)
plot_avg_percent_change <- avg_percent_change
plot_avg_percent_change[plot_avg_percent_change < -5] <- -5
plot_avg_percent_change[plot_avg_percent_change > 5] <- 5

colors <- c("red", "beige", "green")
color_palette <- colorRampPalette(colors)(100)
plot(plot_avg_percent_change, col=color_palette, breaks=seq(-5, 5, length.out=101))

```

## Explore data at eez/mfa scale
Determine average percent change during past 10 years. 
And the change in the average percent change during those past 10 years

```{r}
# create template raster
zones_raster <- rasterize(eez, rast(avg_percent_change), field="rgn_id")
plot(zones_raster)

setdiff(eez_data$rgn_id, unique(zones_raster)$rgn_id)
setdiff(unique(zones_raster)$rgn_id, eez_data$rgn_id)

# extract current yearly biomass for each eez
country_biomass <- terra::zonal(biomass_c, zones_raster, fun = sum, na.rm=TRUE)

names(country_biomass) <- c("rgn_id", paste0("yr", 2000:2010))

# clean data 
country_biomass_change <- country_biomass %>% 
  pivot_longer(cols = starts_with("yr"),
                names_to = "year",
                values_to = "biomass") %>%
   mutate(year = as.numeric(gsub("yr", "", year))) %>%
  left_join(eez_data, by="rgn_id")

# calculate mean percent change for every year (used to get slopes)
country_biomass_change <- country_biomass_change %>%
  group_by(ant_typ, rgn_id, rgn_nam, iso3c, uninhabited) %>%
  arrange(year) %>%
  mutate(yearly_percent_change = (biomass - lag(biomass))/lag(biomass)) 

write_csv(country_biomass_change, here("prep/fisheries/data/country_biomass_2000_2010.csv"))

country_biomass_change <- read_csv(here("prep/fisheries/data/country_biomass_2000_2010.csv"))

## explore USA
ggplot(filter(country_biomass_change, rgn_nam == "Pacific, Western Central"), aes(x=year, y=biomass)) +
  geom_point() + 
  geom_line()
ggplot(filter(country_biomass_change, rgn_nam == "Pacific, Western Central"), aes(x=year, y=yearly_percent_change)) +
  geom_point() + 
  geom_line()

summary(lm(yearly_percent_change ~ year, data=filter(country_biomass_change, rgn_nam == "Pacific, Western Central")))

# calculate slope of yearly percent change
country_slopes <- country_biomass_change %>%
  filter(!is.na(yearly_percent_change)) %>%
  group_by(rgn_id, ant_typ, rgn_nam, iso3c, uninhabited) %>%
  do(tidy(lm(yearly_percent_change ~ year, data = .))) %>%
  ungroup() %>%
  filter(term == "year") %>%
  arrange(estimate)

# calculate average annual percent change
country_avg_change <- country_biomass_change %>%
  group_by(rgn_id, ant_typ, rgn_nam, iso3c, uninhabited) %>%
  summarize(start_biomass = first(biomass[year == min(year)]),
            end_biomass = first(biomass[year == max(year)]),
            start_year = min(year),
            end_year = max(year),
            avg_biomass = mean(biomass),
            biomass_2005 = biomass[year==2005],
            biomass_2010 = biomass[year==2010]) %>%
  mutate(duration = end_year - start_year,
         avg_percent_change = ((end_biomass / start_biomass)^(1 / duration) - 1)) %>%
  select(rgn_id, ant_typ, rgn_nam, iso3c, uninhabited, avg_biomass, biomass_2005, biomass_2010, avg_percent_change) %>%
  ungroup()
  
# calculate mean biomass: another way of calculating
# should be similar to above approach, and it is (just wanted to check)
#country_avg_change2 <- country_biomass_change %>%
#  group_by(rgn_id, ant_typ, rgn_nam) %>%
#  mutate(avg_percent_change2 = mean(yearly_percent_change, na.rm=TRUE)) %>%
#  select(rgn_id, ant_typ, rgn_nam, avg_percent_change2)
#tmp <- left_join(country_avg_change, country_avg_change2)
#plot(tmp$avg_percent_change, tmp$avg_percent_change2)
#abline(0,1, col="red")

country_avg_change <- country_avg_change %>%
  left_join(country_slopes, by=c("rgn_id", "ant_typ", "rgn_nam", "iso3c", "uninhabited")) %>%
#  left_join(country_avg_change2, by=c("rgn_id", "ant_typ", "rgn_nam")) %>%
  select(rgn_id, ant_typ, rgn_nam, iso3c, uninhabited, avg_biomass, biomass_2005, biomass_2010, avg_percent_change, slope=estimate)
  
## look at all countries
ggplot(country_avg_change, aes(x=avg_percent_change, y=slope, col=uninhabited)) +
  geom_point()
ggplot(country_avg_change, aes(x=avg_percent_change, y=slope, col=ant_typ)) +
  geom_point()


## add in historic biomass
biomass_historic_times <- which(as.Date(time(biomass)) == as.Date("1900-01-01")) 
biomass_h <- subset(biomass, biomass_historic_times)
biomass_h <- cellSize(biomass_h) * biomass_h * 1e-6
country_historic_biomass <- terra::zonal(biomass_h, zones_raster, fun = sum, na.rm=TRUE)

names(country_historic_biomass) <- c("rgn_id", "historic_1900")

country_avg_change <- left_join(country_avg_change, country_historic_biomass, by="rgn_id")
country_avg_change <- country_avg_change %>%
  mutate(prop_biomass = avg_biomass/historic_1900)

ggplot(country_avg_change, aes(x=avg_percent_change, y=prop_biomass, col=uninhabited)) +
  geom_point()
filter(country_avg_change, prop_biomass>2)
ggplot(country_avg_change, aes(x=avg_percent_change, y=prop_biomass, col=ant_typ)) +
  geom_point()



## save data
write_csv(country_avg_change, here("prep/fisheries/data/region_biomass_trends.csv"))


```


Get values of biomass for eezs throughout time
(exploration)

```{r}

effort_data_files <- "/home/shares/ohi/stressors_2021/_raw_data/Bianchi_fish_biomass/BOATS_effort"

biomass <- rast(file.path(effort_data_files, "boats-v2_gfdl-mom6-cobalt2_none_obsclim_histsoc_onedeg_tcb_global_annually_1841-2010.nc"))

biomass_times<- which(as.Date(time(biomass)) >= as.Date("1900-01-01") & as.Date(time(biomass)) < as.Date("2011-01-01")) 
biomass_c <- subset(biomass, biomass_times)
biomass_c <- cellSize(biomass_c) * biomass_c * 1e-6

# Initialize an empty SpatRaster to store percent change values
percent_change <- biomass_c[[1]]

zones_raster <- rasterize(eez, rast(percent_change), field="rgn_id")
plot(zones_raster)

setdiff(eez_data$rgn_id, unique(zones_raster)$rgn_id)
setdiff(unique(zones_raster)$rgn_id, eez_data$rgn_id)

# extract current yearly biomass for each eez
country_biomass <- terra::zonal(biomass_c, zones_raster, fun = sum, na.rm=TRUE)

names(country_biomass) <- c("rgn_id", paste0("yr", 1900:2010))

# clean data 
country_biomass_change <- country_biomass %>% 
  pivot_longer(cols = starts_with("yr"),
                names_to = "year",
                values_to = "biomass") %>%
   mutate(year = as.numeric(gsub("yr", "", year))) %>%
  left_join(eez_data, by="rgn_id")

write_csv(country_biomass_change, here("prep/fisheries/data/country_1900_2010_biomass.csv"))
