---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

step1: extract biomass data for eezs.
step2: data and models to see if we can predict changes to fisheries pressures
This is mostly an information gathering step that can be skipped. 
step3: prepare governance data. I need two datasets: 
 - average governance from 2000 to 2010, with gapfilling, for estimating slope of change in fishing pressure from 2000 to 2010
 - 2010-2100 governance, with gapfilling

 
```{r setup, include=FALSE}

library(countrycode)
library(tidyverse)
library(here)
library(WDI)
```


## fisheries data
(generated in script: step 1)

```{r}

pressure_eez <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv"))

pressure_eez <- pressure_eez %>%
  filter(ant_typ == "eez"|iso3c=="DIS") %>%
  filter(is.na(uninhabited)) %>%
  filter(!is.na(slope)) %>%
  select(-uninhabited)

```


Take a look at the data.
```{r}
# what governance looks like:
gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  group_by(scenario, year) %>%
  summarize(governance_mean = mean(governance, na.rm=TRUE),
            sd = sd(governance, na.rm=TRUE))
ggplot(gov_raw, aes(x=year, y=governance_mean, group=scenario, color=scenario)) +
  geom_point() +
  geom_line()

gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") 
ggplot(filter(gov_raw, countrycode == "ARG"), aes(x=year, y=governance, group=scenario, color=scenario)) +
  geom_point() + 
  geom_line()

## note the weird jump
gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(scenario %in% c("Observed", "SSP5")) 
ggplot(gov_raw, aes(x=year, y=governance, group=countrycode)) +
  geom_point(size=0.2) +
  geom_line(size=0.2)


```

## Get the observed governance data from 2000-2010

Take the average from 2000 to 2010
This will be used to develop the predictive model.

```{r}

gov_gf <- pressure_eez %>%
  select(iso3c, rgn_nam, rgn_id)


gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(scenario == "Observed") %>%
  filter(year >= 2000 & year <= 2010) %>%
  select(iso3c = countrycode, year, governance)

gov_gf <- left_join(gov_gf, gov_raw, by="iso3c")

## first couple years of Tonga missing, fill with closest year
gov_gf <- gov_gf %>%
  mutate(governance = ifelse(iso3c=="TON" & is.na(governance), gov_gf$governance[gov_gf$iso3c=="TON" & gov_gf$year==2003], governance))
summary(gov_gf)
# N = 61 (of 184) countries with no governance, included general disputed regions

## will be working with the avg. governance from 2000 to 2010

gov_gf_avg <- gov_gf %>%
  group_by(iso3c, rgn_id, rgn_nam) %>%
  summarize(governance = mean(governance)) %>%
  ungroup()

```


Gapfill missing governance data: average from 2000 - 2010
This is used to model the relationship between governance and change in fisheries loss from 2000 to 2010. 

I will use pcGDPppp and georegion to predict *average* governance from 2000-2010.

I attempted to use the SSP gdp and pop projection data (see step2: here("prep/fisheries/data/current_future_gdp_pop.csv")), but there were quite a lot of gaps
in these data, so I am just going to use World Bank data (even though there are no future 
projections).
```{r}

### GDP

#indicator <- "NY.GDP.PCAP.PP.CD"

# Fetch data
#gdp_pc <- WDI(
#  indicator = indicator,
#  start = 2000,  # Adjust the starting year as needed
#  end = 2010,    # Adjust the ending year as needed
#  extra = TRUE   # Include additional country metadata
#)

#write_csv(gdp_pc, here("prep/fisheries/data/inputs/world_bank_pcgdp.csv"))

gdp_pc <- read_csv(here("prep/fisheries/data/inputs/world_bank_pcgdp.csv"))

gdp_pc <- gdp_pc %>%
  select(iso3c, year, pcGDP=NY.GDP.PCAP.PP.CD) %>%
  filter(!is.na(pcGDP)) %>% #most have 11 years of data, but a few don't
  mutate(round(pcGDP,0)) %>%
  group_by(iso3c) %>%
  summarize(pcGDP = mean(pcGDP))

## rough estimate from google searches (Wikipedia and such)
gf_regions <- read_csv(here("prep/fisheries/data/country_pop_gf.csv")) %>%
  filter(population > 0) %>%
  mutate(pop_gf = population/1000000,
         gdp_gf = gdp_current_USD/1000000) %>%
  mutate(pcGDP = round(gdp_gf/pop_gf, 0)) %>%
  select(iso3c, pcGDP)

new <- setdiff(gf_regions$iso3c, gdp_pc$iso3c)


gdp_pc <- rbind(gdp_pc, gf_regions[gf_regions$iso3c %in% new,])

gov_gf_avg2 <- left_join(gov_gf_avg, gdp_pc, by="iso3c")  # N=41 of 61 have gdp data
filter(gov_gf_avg2, is.na(governance)) %>% print(n=100)

## Fill in disputed regions with low governance score
gov_gf_avg2$rgn_nam[gov_gf_avg2$iso3c == "DIS"] <- "Disputed"
gov_gf_avg2$governance[gov_gf_avg2$iso3c == "DIS"] <- quantile(gov_gf_avg2$governance, c(0.1), na.rm=TRUE)


## Georegions used to gapfill missing governance
georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv") %>%
  select(rgn_id, r1_label, r2_label)


gov_gf_avg3 <- left_join(gov_gf_avg2, georgns, by="rgn_id") %>%
  mutate(log_gdp = log(pcGDP))
plot(gov_gf_avg3$governance, log(gov_gf_avg3$pcGDP))
plot(gov_gf_avg3$governance, gov_gf_avg3$pcGDP)
summary(gov_gf_avg3)

### gapfill all regions (except Micronesia) that have gdp data
mod_gov1 <- lm(governance ~ log_gdp + r2_label, na.action=na.exclude, data=gov_gf_avg3)
summary(mod_gov1) # R2= 0.72


gov_gf_avg3$gov_gf1 <- predict(mod_gov1, newdata = gov_gf_avg3 %>%
                          mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

gov_gf_avg4 <- gov_gf_avg3 %>%
  mutate(gapfilled = ifelse(is.na(governance), "gapfilled", NA)) %>%
  mutate(governance = ifelse(is.na(governance), gov_gf1, governance)) # n missing now N= 24


### gapfill the Micronesia region
mod_gov2 <- lm(governance ~ log_gdp + r1_label, data=gov_gf_avg4, na.action=na.exclude)
summary(mod_gov2) # R2 = 0.70

gov_gf_avg4$gov_gf2 <- predict(mod_gov2, newdata = gov_gf_avg4)
gov_gf_avg5 <- gov_gf_avg4 %>%
  mutate(governance = ifelse(is.na(governance), gov_gf2, governance)) # n missing n= 19


### gapfill remaining regions based on georegion, r2 label
mod_gov3 <- lm(governance ~ r2_label, data=gov_gf_avg5, na.action=na.exclude)
summary(mod_gov3) # R2 = 0.63

gov_gf_avg5$gov_gf3 <- predict(mod_gov3, newdata = gov_gf_avg5)
gov_gf_avg6 <- gov_gf_avg5 %>%
  mutate(governance = ifelse(is.na(governance), gov_gf3, governance))

summary(gov_gf_avg6) # all have governance

avg_governance_gf <- gov_gf_avg6 %>%
  select(iso3c, rgn_id, rgn_nam, governance, gapfilled)

write_csv(avg_governance_gf, here("prep/fisheries/data/avg_2000_2010_governance_gf.csv"))

```

## Gapfilling governance 2000-2015 data

In this case, we will use the same general predictive model as before to gapfill the yearly data (vs. the average data across 10 years).

Gather the observed governance data

```{r}
regions <- pressure_eez %>%
  select(iso3c, rgn_nam, rgn_id)


gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(scenario == "Observed") %>%
  filter(year >= 2000) %>%
  select(iso3c = countrycode, year, governance)

gov_gf <- left_join(regions, gov_raw, by="iso3c") %>%
  group_by(iso3c, rgn_id, rgn_nam) %>%
  complete(year = 2000:2015) %>%
  arrange(iso3c, year) %>%
  ungroup() %>%
  filter(!is.na(year))



```

Now estimate the 2000 to 2015 years based on GDP and georegion:

```{r}


# Fetch data
# indicator <- "NY.GDP.PCAP.PP.CD"
#gdp_pc <- WDI(
#  indicator = indicator,
#  start = 2000,  # Adjust the starting year as needed
#  end = 2015,    # Adjust the ending year as needed
#  extra = TRUE   # Include additional country metadata
#)

#write_csv(gdp_pc, here("prep/fisheries/data/inputs/world_bank_pcgdp_2000_2015.csv"))

gdp_pc <- read_csv(here("prep/fisheries/data/inputs/world_bank_pcgdp_2000_2015.csv"))

# a few countries don't have all years of gdp data (NA N = 405 to N = 304), gapfill here:
gdp_pc <- gdp_pc %>%
  select(iso3c, year, pcGDP=NY.GDP.PCAP.PP.CD) %>%
  group_by(iso3c) %>%
  arrange(year) %>%
  tidyr::complete(year = 2000:2015) %>%
  tidyr::fill(pcGDP, .direction = "downup") %>%
  ungroup()

gov_gf2 <- left_join(gov_gf, gdp_pc, by=c("iso3c", "year")) # n=640 pcGDp



# Add in the values I looked up from other sources # n = 352 pcGDP
## rough estimate from google searches (Wikipedia and such)
gf_regions <- read_csv(here("prep/fisheries/data/country_pop_gf.csv")) %>%
  filter(population > 0) %>%
  mutate(pop_gf = population/1000000,
         gdp_gf = gdp_current_USD/1000000) %>%
  mutate(pcGDP_est = round(gdp_gf/pop_gf, 0)) %>%
  select(iso3c, pcGDP_est)

gov_gf3 <- left_join(gov_gf2, gf_regions, by="iso3c") %>%
  mutate(pcGDP = ifelse(is.na(pcGDP), pcGDP_est, pcGDP)) %>%
  select(-pcGDP_est)

## Georegions
georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv") %>%
  select(rgn_id, r1_label, r2_label)

## NA = N 1101
gov_gf4 <- left_join(gov_gf3, georgns, by="rgn_id") %>%
  mutate(log_gdp = log(pcGDP))
plot(gov_gf4$governance, log(gov_gf4$pcGDP))
plot(gov_gf4$governance, gov_gf4$pcGDP)

### gapfill all regions (except Micronesia) that have gdp data
# tried including year, but was not significant
mod_gov1 <- lm(governance ~ log_gdp + r2_label, na.action=na.exclude, data=gov_gf4)
summary(mod_gov1) # R2= 0.73

gov_gf4$gov_gf1 <- predict(mod_gov1, 
                          newdata = gov_gf4 %>%
                          mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

# NA gov = 402
gov_gf5 <- gov_gf4 %>%  
  mutate(gapfilled = ifelse(is.na(governance), "gapfilled", NA)) %>%
  mutate(governance = ifelse(is.na(governance), gov_gf1, governance))


### gapfill the Micronesia region
mod_gov2 <- lm(governance ~ log_gdp + r1_label, data=gov_gf5, na.action=na.exclude)
summary(mod_gov2) # R2 = 0.69

gov_gf5$gov_gf2 <- predict(mod_gov2, newdata = gov_gf5)

#NA gov = 322
gov_gf6 <- gov_gf5 %>%  
  mutate(governance = ifelse(is.na(governance), gov_gf2, governance))


### gapfill remaining regions based on georegion
mod_gov3 <- lm(governance ~ r2_label, data=gov_gf6, na.action=na.exclude)
summary(mod_gov3) # R2 = 0.68

gov_gf6$gov_gf3 <- predict(mod_gov3, newdata = gov_gf6)

## NA gov N = 16 (disputed areas)
gov_gf7 <- gov_gf6 %>%  
  mutate(governance = ifelse(is.na(governance), gov_gf3, governance))

## fill in disputed areas with bottom quantile
gov_gf7 <- gov_gf7 %>%
  mutate(governance = ifelse(is.na(governance), quantile(gov_gf7$governance, c(0.1), na.rm=TRUE), governance))

governance_gf_2000_2015 <- gov_gf7 %>%
  mutate(scenario = "Observed") %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled)



```

# Gapfill missing >2015 based on modeled average yearly change

First bias correct the governance data so the observed and scenario 2015 values align. 
```{r}

gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv")

observed <- gov_raw %>%
  filter(year==2015 & scenario == "Observed") %>%
  select(iso3c=countrycode, observed_governance=governance)

scenario_correct <- gov_raw %>%
  filter(year==2015 & scenario != "Observed") %>%
  select(iso3c = countrycode, scenario, governance) %>%
  left_join(observed, by = "iso3c") %>%
  mutate(bias_correct = governance - observed_governance) %>%
  select(iso3c, scenario, bias_correct)

bias_corrected_future <- gov_raw %>%
  select(iso3c=countrycode, scenario, year, governance) %>%
  filter(year >= 2015 & scenario != "Observed") %>%
  left_join(scenario_correct, by=c("iso3c", "scenario")) %>%
  mutate(bias_correct = ifelse(is.na(bias_correct), 0, bias_correct)) %>% # COD, ESH, ROU have future data, but not past data
  mutate(governance = governance - bias_correct) %>%
  left_join(regions, by = "iso3c") %>%
  filter(!is.na(rgn_id)) %>%
  mutate(gapfilled = NA) %>%
  mutate(governance = ifelse(governance > 1, 1, governance)) %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled) 

```

Now use the average yearly change across regions for each SSP to estimate the regions that have current estimates (gapfilled) but do not have future data.
```{r}
# first figure out average yearly change for each scenario

yearly_change <- bias_corrected_future %>%
  arrange(iso3c, scenario, year) %>%
  group_by(iso3c, scenario, rgn_id, rgn_nam) %>%
  mutate(governance_change = governance - lag(governance, default = governance[1])) %>%
  ungroup()  

avg_yearly_change <- yearly_change %>%
  group_by(scenario, year) %>%
  filter(governance < 1) %>%
  summarize(avg_change = mean(governance_change)) %>%
  ungroup() %>%
  group_by(scenario) %>%
  mutate(cumulative_avg_change = cumsum(avg_change))

## figure out what is missing in the >2015 data
no_data_ids <- setdiff(governance_gf_2000_2015$iso3c, bias_corrected_future$iso3c)

scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5")
governance_2015 <- filter(governance_gf_2000_2015, iso3c %in% no_data_ids) %>%
  filter(year == 2015) %>%
  select(iso3c, rgn_id, rgn_nam, governance_2015 = governance) %>%
  crossing(scenario = scenario)

gov_future_gf <- full_join(avg_yearly_change, governance_2015, by=c("scenario")) %>%
  rowwise() %>%
  mutate(governance = governance_2015 + cumulative_avg_change) %>%
  mutate(governance = ifelse(governance > 1, 1, governance)) %>%
  mutate(gapfilled = "gapfilled") %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled)

## Now getting the observed data for 2000-2015 (duplicating for each scenario)
scenario <- c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5")
scenario_governance_gf_2000_2015 <- governance_gf_2000_2015 %>%
  select(-scenario) %>%
  expand_grid(scenario) %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled)



future_governance_gf <- rbind(bias_corrected_future, gov_future_gf) %>%
  filter(year != 2015) %>%
  rbind(scenario_governance_gf_2000_2015)

setdiff(pressure_eez$iso3c, future_governance_gf$iso3c)
## should be none! all accounted for!

write_csv(future_governance_gf, here("prep/fisheries/data/governance_2000_2100.csv"))

```
