---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

step1: extract biomass data for eezs.
step2: data and models to see if we can predict changes to fisheries pressures
This is mostly an information gathering step that can be skipped. 
step3: prepare governance data. I need two datasets: 
 - average governance from 2000 to 2010, with gapfilling, for estimating slope of change in fishing pressure from 2000 to 2010
 - 2000-2100 governance, with gapfilling
 step4: random sampling for future fishing pressure scenarios

 
```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(WDI)
```


## fisheries data
(generated in script: step 1)
(governance data: step 3)
```{r}

fish_pressure <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv"))
governance <- read_csv(here("prep/fisheries/data/avg_2000_2010_governance_gf.csv"))
pressure_data <- left_join(governance, fish_pressure, by=c("rgn_id", "rgn_nam", "iso3c")) %>%
  filter(!(is.na(slope))) # average country data

# yearly scenario governance
gov_data <- read_csv(here("prep/fisheries/data/governance_2000_2100.csv")) 
```

## Develop model to randomly sample from
Model and parameters. Predict change in fisheries pressure based on governance.
```{r}

## model to predict slope based on government
mod <- lm(slope ~ governance, data = filter(pressure_data, is.na(gapfilled)))
summary(mod) #R2 = 0.18
ggplot(mod, aes(y = slope, x=governance)) +
  geom_point() +
  geom_smooth()
  
# get the model values
residual_sd <- summary(mod)$sigma
intercept <- coef(mod)["(Intercept)"]
governance_coeff <- coef(mod)["governance"]

```

## Prepare simulation data
This collects the random error and the expected slope based on governance.

```{r}

# Number of simulations
n_sims <- 100

# Create a matrix of random errors (one row per simulation, one column per row of original data)
random_errors <- matrix(rnorm(n_sims * nrow(gov_data), mean = 0, sd = residual_sd), nrow = n_sims, ncol = nrow(gov_data))

# Calculate the expected slopes for all rows in one operation
expected_slopes <- intercept + governance_coeff * gov_data$governance

# Add the random errors to the expected slopes to get the simulated slopes
simulated_slopes <- sweep(random_errors, 2, expected_slopes, '+')

simulated_data <- as.data.frame(t(simulated_slopes))

## region, scenario, year (2005:2100), and simulated slopes
simulated_data2 <- cbind(gov_data, simulated_data) %>% 
  arrange(iso3c, scenario, year) %>%
  filter(year >= 2005)

## summarized data for each region (slope, avg_percent_change, etc.)
simulated_biomass <- fish_pressure %>%
  filter(iso3c %in% simulated_data2$iso3c) %>%
  select(rgn_id, rgn_nam, iso3c, avg_biomass, avg_percent_change, historic_1900, biomass_2005)

```

## Function for updating biomass based on governance
This is used in the next code chunk
```{r}


update_biomass <- function(df, random_column) {
  # Initialize the first year's values
  df$prop[1] <- df$biomass_2005[1] / df$historic_1900[1]
  df$avg_percent_change[1] <- 0
  df$avg_biomass[1] <- df$biomass_2005[1] # starting value is 2005
  
  # Save the starting prop for each row to ensure it does not get worse for props < 0.2
  starting_prop <- rep(0, nrow(df))
  starting_prop[1] <- df$prop[1]
  
  ## then run this for the rest of the rows:
  for (i in 2:nrow(df)) {
    # Save starting prop if it's below 0.2
    if (df$prop[i-1] < 0.2) {
      starting_prop[i] <- max(starting_prop[i-1], df$prop[i-1])
    } else {
      starting_prop[i] <- df$prop[i-1]
    }
    
    new_percent_change <- df$avg_percent_change[i - 1] + df[[random_column]][i]

    # Calculate potential next year's biomass and prop for decision making
    potential_biomass <- df$avg_biomass[i - 1] * (1 + new_percent_change)
    potential_prop <- potential_biomass / df$historic_1900[i]

    # Determine the multiplier based on the proportion constraints from the previous year
    if (df$prop[i - 1] >= 0.7) {
      multiplier <- ifelse(new_percent_change > 0, 0, 1)  # Only allow decreases
    } elseif (df$prop[i - 1] < 0.2) {
      # Prevent prop from getting worse than the start of the year for these cases
      multiplier <- ifelse(potential_prop < starting_prop[i], 0, 1)
    } else {
      # Ensure prop stays within [0.2, 0.7] range if it starts between [0.2, 0.7]
      multiplier <- 1
      if (potential_prop < 0.2) {
        multiplier <- 0  # Prevent decrease beyond 0.2
      } elseif (potential_prop > 0.7) {
        multiplier <- 0  # Prevent increase beyond 0.7
      }
    }
    
    # Apply multiplier to the change
    df$avg_percent_change[i] <- new_percent_change * multiplier

    # Update the avg_biomass and prop for the current year
    df$avg_biomass[i] <- df$avg_biomass[i - 1] * (1 + df$avg_percent_change[i])
    df$prop[i] <- df$avg_biomass[i] / df$historic_19090[i]  # Update proportion for the new year
  }
  
  return(df)
}



# Define the function to update biomass with dynamic scaling
## running this one!
update_biomass <- function(df, random_column) {
  # Initialize the first year's values
  df$prop[1] <- df$biomass_2005[1] / df$historic_1900[1]
  df$avg_percent_change[1] <- df$avg_percent_change[1]
  df$avg_biomass[1] <- df$biomass_2005[1]  # starting value is 2005
  
  ## then run this for the rest of the rows:
  for (i in 2:nrow(df)) { #i=2
    new_percent_change <- df$avg_percent_change[i - 1] + df[[random_column]][i]

    # Calculate potential next year's biomass and prop
    potential_biomass <- df$avg_biomass[i - 1] * (1 + new_percent_change)
    potential_prop <- potential_biomass / df$historic_1900[i]

    # Determine the multiplier based on the proportion constraints from the previous year
    if (df$prop[i - 1] > 0.7) {
      if (new_percent_change > 0) {
        multiplier <- 0  # Prevent increases entirely
      } else if (potential_prop < 0.7) {
        multiplier <- 0  # Prevent decrease if it would go below 0.7
      } else {
        multiplier <- 1  # Allow decrease but not below 0.7
      }
    } else if (df$prop[i - 1] < 0.2) {
      multiplier <- ifelse(new_percent_change < 0, 0, 1)  # No decrease if new change is negative
    } else {
      multiplier <- 1  # No restrictions on change
    }

    # Apply multiplier to the change
    df$avg_percent_change[i] <- new_percent_change * multiplier

    # Update the avg_biomass for the current year using the previous year's biomass and the adjusted percent change
    df$avg_biomass[i] <- df$avg_biomass[i - 1] * (1 + df$avg_percent_change[i])
    df$prop[i] <- df$avg_biomass[i] / df$historic_1900[i]  # Update proportion for the new year
  }
  
  return(df)
}

```

## Run the simulation

```{r}
# empty data frame to add to
simulated_data <- simulated_data2 %>%
    select(iso3c, scenario, year, rgn_id, rgn_nam) %>%
  filter(year >=2005)

for(i in 1:n_sims){
#i <- "1"
random_column <- paste0("V", i)

# Select and filter the necessary columns
new_data <- simulated_data2 %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, all_of(random_column)) %>%
  left_join(simulated_biomass, by = c("iso3c", "rgn_id", "rgn_nam")) %>%
  mutate(prop = NA_real_)

# Apply the function to each group
new_data_sim <- new_data %>%
  group_by(iso3c, rgn_id, rgn_nam, scenario) %>%
  arrange(iso3c, scenario, year) %>%  # Ensure the data is sorted by year
  group_modify(~ update_biomass(.x, random_column = random_column)) %>%
  ungroup() %>% # Remove grouping for final dataframe
  rename_with(~ paste0("biomass_", i), avg_biomass) %>%
  select(iso3c, rgn_id, rgn_nam, scenario, year, starts_with("biomass_"))

simulated_data <- left_join(simulated_data, new_data_sim)    
cat("simulation = ", i, "\n")

}


# Pivot the biomass columns into a long format
data_long <- simulated_data %>%
  pivot_longer(cols = starts_with("biomass_"), names_to = "biomass_type", values_to = "biomass") %>%
  filter(biomass_type != "biomass_2005")

usa_biomass_stats <- data_long %>%
  filter(iso3c=="CHN") %>%
  group_by(year, scenario) %>%
  summarize(
    median_biomass = median(biomass),
    p10 = quantile(biomass, 0.1),
    p90 = quantile(biomass, 0.9),
    .groups = 'drop'
  )

thresholds <- filter(simulated_biomass, iso3c=="CHN")

ggplot(usa_biomass_stats, aes(x = year, y = median_biomass, color = scenario)) +
  geom_line() +
  geom_point() +
# geom_ribbon(aes(ymin = p10, ymax = p90, fill = scenario), alpha = 0.2) +
  labs(title = "China Median Biomass for Global by Scenario",
       x = "Year",
       y = "Biomass") +
  geom_hline(yintercept = thresholds$historic_1900, col="blue") + 
  geom_hline(yintercept = thresholds$historic_1900 * 0.7, col="purple") + 
  geom_hline(yintercept = thresholds$historic_1900 * 0.2, col="red") + 
  theme_minimal()

#global
global_biomass_stats <- data_long %>%
  group_by(year, scenario, biomass_type) %>%
  summarize(total_biomass = sum(biomass)) %>%
  ungroup() %>%
  group_by(year, scenario) %>%
  summarize(
    median_biomass = median(total_biomass),
    p10 = quantile(total_biomass, 0.1),
    p90 = quantile(total_biomass, 0.9),
    .groups = 'drop'
  )
sum(pressure_data$historic_1900)*0.7
ggplot(filter(global_biomass_stats, year <= 2100), aes(x = year, y = median_biomass, color = scenario)) +
  geom_line() +
  geom_point() +
# geom_ribbon(aes(ymin = p10, ymax = p90, fill = scenario), alpha = 0.2) +
  labs(title = "Global: Median Biomass for Global by Scenario",
       x = "Year",
       y = "Biomass") +
  theme_minimal()

## historic is 11.3+08

## compare to global observed: 2005:2010
## ensure estimates are not going off the rails.


```


```{r}
#simulation data, created above

simulation <- data_long %>%
  group_by(year, scenario, biomass_type) %>%
  summarize(total_biomass = sum(biomass)) %>%
  ungroup() %>%
  filter(year %in% 2005:2010)


# observed data: this file created in step1
country_biomass_change <- read_csv("prep/fisheries/data/country_biomass_2000_2010.csv") %>%
  filter(iso3c %in% unique(data_long$iso3c))

observed <- country_biomass_change %>%
  group_by(year) %>%
  summarize(observed_biomass = sum(biomass, na.rm=TRUE))

ggplot() + 
  geom_line(data=simulation, aes(x=year, y=total_biomass, group=biomass_type), alpha=0.2) +
  geom_point(data=filter(observed, year>=2005), aes(x=year, y=observed_biomass), col="red")



```