---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

```{r setup, include=FALSE}

library(sf)
library(terra)
library(countrycode)
library(broom)
library(tidyverse)
library(here)
library(rnaturalearth)

```

## Prepare region data

```{r}
## EEZ regions
eez_shp <- sf::st_read("/home/shares/ohi/git-annex/globalprep/spatial/v2017/regions_2017_update.shp") 
eez_lat_lon <- sf::st_transform(zones_shp, 4326)
rgn_data <- read.csv("/home/shares/ohi/git-annex/globalprep/spatial/d2014/data/rgn_mol_raster_1km/regionData.csv") %>%
  dplyr::select(rgn_id, ant_typ, rgn_nam) %>% 
  unique()

## fao major fishing regions
fao_zones <- sf::st_read("/home/shares/ohi/stressors_2021/_raw_data/fao_boundaries/FAO_AREAS_CWP_NOCOASTLINE.shp") %>%
  filter(F_LEVEL=="MAJOR")
fao_zones_lat_lon <- sf::st_transform(fao_zones, 4326)

#intersecting the eez and fao shapfiles, takes forever, so saved outputs to use in analysis
#eez_fao <- st_intersection(eez_lat_lon, fao_zones_lat_lon)   %>%
#  mutate(ID = paste0(F_AREA, "0", rgn_id)) %>%
#  mutate(ID = as.numeric(as.character(ID)))
#st_write(eez_fao, "/home/shares/ohi/stressors_2021/_dataprep/spatial/fao_eez/fao_eez.shp", append=FALSE)


eez_fao <- sf::st_read("/home/shares/ohi/stressors_2021/_dataprep/spatial/fao_eez/fao_eez.shp")

eez_fao_data <- st_drop_geometry(eez_fao) %>%
  filter(!(rgn_type %in%  c("land", "land-noeez", "land-disputed"))) %>%
  select(ID, ant_type = type_w_ant, rgn_id, rgn_type, rgn_name, F_CODE, OCEAN, NAME_EN) %>%
  unique()

#write_csv(eez_fao_data, here("prep/fisheries/data/eez_fao_id.csv"))

eez_fao_ID <- read_csv(here("prep/fisheries/data/eez_fao_id.csv"))

```

# BOATS: biomass model based on effort

```{r}

effort_data_files <- "/home/shares/ohi/stressors_2021/_raw_data/Bianchi_fish_biomass/BOATS_effort"

biomass <- rast(file.path(effort_data_files, "boats-v2_gfdl-mom6-cobalt2_none_obsclim_histsoc_onedeg_tcb_global_annually_1841-2010.nc"))

```

# Explore raster scale data
Isolate years between 2000 and 2010. For each cell, calculate rate of change for each year. 
Average of 2000 to 2010.

```{r}

biomass_times <- which(as.Date(time(biomass)) >= as.Date("2000-01-01") & as.Date(time(biomass)) < as.Date("2011-01-01")) 
biomass_c <- subset(biomass, biomass_times)
biomass_c <- cellSize(biomass_c) * biomass_c * 1e-6

# Initialize an empty SpatRaster to store percent change values
percent_change <- biomass_c[[1:10]]
values(percent_change) <- NA # Fill with NAs since the first layer can't calculate change

# Calculate percent change for each layer, starting from the second layer
for (i in 2:nlyr(biomass_c)) { # i = 2 
  # Calculate percent change
  pc <- (biomass_c[[i]] - biomass_c[[i-1]]) / biomass_c[[i-1]] * 100
  
  # Assign calculated values to the corresponding layer in the percent_change_raster
  percent_change[[i-1]] <- pc
}

avg_percent_change <- mean(percent_change)
plot(avg_percent_change)
plot_avg_percent_change <- avg_percent_change
plot_avg_percent_change[plot_avg_percent_change < -5] <- -5
plot_avg_percent_change[plot_avg_percent_change > 5] <- 5

colors <- c("red", "beige", "green")
color_palette <- colorRampPalette(colors)(100)
plot(plot_avg_percent_change, col=color_palette, breaks=seq(-5, 5, length.out=101))

```

## Explore data at eez/mfa scale
Determine average percent change during past 10 years. 
And the change in the average percent change during those past 10 years

```{r}
# create template raster
zones_raster <- rasterize(eez_fao, rast(avg_percent_change), field="ID")
plot(zones_raster)

# extract yearly biomass for each eez
country_biomass <- terra::zonal(biomass_c, zones_raster, fun = sum, na.rm=TRUE)

names(country_biomass) <- c("ID", paste0("yr", 2000:2010))

# clean data 
country_biomass_change <- country_biomass %>% 
  pivot_longer(cols = starts_with("yr"),
                names_to = "year",
                values_to = "biomass") %>%
   mutate(year = as.numeric(gsub("yr", "", year))) %>%
  left_join(eez_fao_ID, by="ID")

# calculate mean percent change for every year (used to get slopes)
country_biomass_change <- country_biomass_change %>%
  group_by(ID, ant_type, rgn_id, rgn_name, OCEAN, NAME_EN) %>%
  arrange(year) %>%
  mutate(yearly_percent_change = (biomass - lag(biomass))/lag(biomass)) %>%
  filter(year != 2000)

## explore USA
ggplot(filter(country_biomass_change, rgn_name == "United States"), aes(x=year, y=biomass, group=ID)) +
  geom_point() + 
  geom_line()
ggplot(filter(country_biomass_change, rgn_name == "United States"), aes(x=year, y=yearly_percent_change, group=ID)) +
  geom_point() + 
  geom_line()

summary(lm(yearly_percent_change ~ year, data=filter(country_biomass_change, ID==670163)))

# calculate slope of yearly percent change
country_slopes <- country_biomass_change %>%
  filter(!is.na(yearly_percent_change)) %>%
  group_by(ID, rgn_id, ant_type, rgn_name, OCEAN, NAME_EN) %>%
  do(tidy(lm(yearly_percent_change ~ year, data = .))) %>%
  ungroup() %>%
  filter(term == "year") %>%
  arrange(estimate)

# calculate average annual percent change
country_avg_change <- country_biomass_change %>%
  group_by(ID, rgn_id, ant_type, rgn_name, OCEAN, NAME_EN) %>%
  summarize(start_biomass = first(biomass[year == min(year)]),
            end_biomass = first(biomass[year == max(year)]),
            start_year = min(year),
            end_year = max(year),
            avg_biomass = mean(biomass)) %>%
  mutate(duration = end_year - start_year,
         avg_percent_change = ((end_biomass / start_biomass)^(1 / duration) - 1)) %>%
  select(ID, rgn_id, ant_type, rgn_name, OCEAN, NAME_EN, avg_biomass, avg_percent_change)
  
# calculate mean biomass: another way of calculating
# should be similar to above approach, and it is (just wanted to check)
#country_avg_change2 <- country_biomass_change %>%
#  group_by(rgn_id, ant_typ, rgn_nam) %>%
#  mutate(avg_percent_change2 = mean(yearly_percent_change, na.rm=TRUE)) %>%
#  select(rgn_id, ant_typ, rgn_nam, avg_percent_change2)


country_avg_change <- country_avg_change %>%
  left_join(country_slopes, by=c("ID","rgn_id", "ant_type", "rgn_name", "OCEAN", "NAME_EN")) %>%
#  left_join(country_avg_change2, by=c("rgn_id", "ant_typ", "rgn_nam")) %>%
  select(ID, rgn_id, ant_type, rgn_name, OCEAN, NAME_EN, avg_biomass, avg_percent_change, slope=estimate)
  
## look at all countries
ggplot(country_avg_change, aes(x=avg_percent_change, y=slope)) +
  geom_point()

## save data
country_avg_change$iso3c <- countrycode(country_avg_change$rgn_name, "country.name", "iso3c")
write_csv(country_avg_change, here("prep/fisheries/data/region_biomass_trends.csv"))


```


Add in variables
```{r}
pressure_eez <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv")) %>%
  filter(ant_type == "eez")
# cut out territories and such with no iso3c and uninhabited (or, mostly uninhabited) islands. These have
# fisheries and should be modeled, but will need a different model.
pressure_eez <- pressure_eez %>%
  filter(!is.na(iso3c),
         iso3c!= "ATA", 
         rgn_name != "Bassas da India", 
         iso3c != "IOT",  # British Indian Ocean Territory
         iso3c != "SGS",  # South Georgia and the South Sandwich Islands 
         iso3c != "HMD",  # Heard and McDonald Islands
         iso3c != "BVT",  # Bouvet Island
         iso3c != "PCN") # Pitcairn Islands
sum(duplicated(pressure_eez$iso3c)) # regions with eezs in multiple fao major fishing regions
dups <- pressure_eez[duplicated(pressure_eez$iso3c),]


gdp <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "GDP|PPP") %>%
  filter(MODEL == "OECD Env-Growth") %>%
  filter(SCENARIO == "SSP1_v9_130325") %>%
  rowwise() %>%
  mutate(gdp_ppp = mean(c(`2000`, `2005`, `2010`), na.rm=TRUE)) %>%
  dplyr::select(iso3c = REGION, gdp_ppp)

pop <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "Population") %>%
  filter(SCENARIO == "SSP1_v9_130115") %>%
  rowwise() %>%
  mutate(avg_pop = mean(c(`2000`, `2005`, `2010`), na.rm=TRUE)) %>%
  select(MODEL, SCENARIO, REGION, avg_pop) %>%
  pivot_wider(names_from = MODEL, values_from = avg_pop) %>%
  mutate(NCAR = ifelse(is.na(NCAR), `IIASA-WiC POP`, NCAR)) %>% # estimates from two models are nearly the same, will use NCAR, but gapfill with IIASA 
  select(iso3c = REGION, population = NCAR)

pressure_eez2 <- left_join(pressure_eez, gdp, by="iso3c") %>%
         left_join(pop, by="iso3c") 

gini <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/NRao_et_al_GiniProjections_2018/Gini_projections_SSPs.csv") %>%
  filter(year == c(2011)) %>%
  filter(scenario == "SSP1") %>%
  pivot_longer( cols = -c(scenario, year),  
  names_to = "iso3c",
  values_to = "gini") %>%
  select(iso3c, gini)

pressure_eez3 <- left_join(pressure_eez2, gini, by="iso3c") 

gov <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(year%in% c(2000:2010) & scenario=="Observed") %>%
  group_by(countrycode, country, scenario) %>%
  summarize(governance = mean(governance, na.rm=TRUE)) %>%
  rename("iso3c" = "countrycode") %>%
  ungroup() %>%
  select(iso3c, governance)


pressure_eez4 <- left_join(pressure_eez3, gov, by="iso3c")

eez_area <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/rgn_area.csv")

pressure_eez5 <- left_join(pressure_eez4, eez_area, by="rgn_id")



gf_regions <- read_csv(here("prep/fisheries/data/country_pop_gf.csv")) %>%
  select(iso3c, pop_gf = population, gdp_gf = gdp_current_USD) %>%
  mutate(pop_gf = pop_gf/1000000,
         gdp_gf = gdp_gf/1000000000)

pressure_eez6 <- left_join(pressure_eez5, gf_regions, by="iso3c") %>%
  mutate(population = ifelse(is.na(population), pop_gf, population),
         gdp_ppp = ifelse(is.na(gdp_ppp), gdp_gf, gdp_ppp)) %>%
         mutate(pc_gdp = gdp_ppp/population) %>%
  select(-pop_gf, -gdp_gf)

georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv")

pressure_eez7 <- left_join(pressure_eez6, georgns, by="rgn_id")

###
mod_gov1 <- lm(governance ~ pc_gdp + r2_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gov1)

pressure_eez7$gov_gf1 <- predict(mod_gov1, newdata = pressure_eez7 %>%
                                               mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

mod_gov2 <- lm(governance ~ pc_gdp + r1_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gov2)

pressure_eez7$gov_gf2 <- predict(mod_gov2, newdata = pressure_eez7)

pressure_eez7 <- pressure_eez7 %>%
  mutate(governance = ifelse(is.na(governance), gov_gf1, governance),
         governance = ifelse(is.na(governance), gov_gf2, governance),
         governance = ifelse(governance > 0.94154, 0.94154, governance)) # set to max observed value

mod_gini1 <- lm(gini ~ pc_gdp + governance + r2_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gini1)

pressure_eez7$gini_gf1 <- predict(mod_gini1, newdata = pressure_eez7 %>%
                                               mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

mod_gini2 <- lm(gini ~ pc_gdp + governance + r1_label, data=pressure_eez7, na.action=na.exclude)
summary(mod_gini2)

pressure_eez7$gini_gf2 <- predict(mod_gini2, newdata = pressure_eez7)

pressure_eez7 <- pressure_eez7 %>%
  mutate(gini = ifelse(is.na(gini), gini_gf1, gini),
           gini = ifelse(is.na(gini), gini_gf2, gini)) %>%
  select(-gini_gf1, -gini_gf2, -gov_gf1, -gov_gf2)

pressure_eez7 <- pressure_eez7  %>% 
  mutate(ln_gdp_ppp = log(gdp_ppp+1),
         ln_population = log(population + 1),
         ln_pop_ln_area = log(population)/log(area_km2),
         ln_pc_gdp = log(pc_gdp),
         ln_area = log(area_km2)) 

```


## some predictions
```{r}

library(ranger)

library(caret)
# Set up training control
train_control <- trainControl(method = "LOOCV", savePredictions=TRUE)

# Train the model with leave-one-out cross-validation
data8 <- na.omit(pressure_eez7) %>%
  mutate(pressure_easing = ifelse(slope>0, "easing", "worsening"))

model <- train(slope ~  governance, 
               data = na.omit(data8), 
               method = "lm", 
               trControl = train_control)
model
summary(model)
plot(model$pred$pred ~ model$pred$obs)


model <- train(as.factor(pressure_easing) ~  governance, 
               data = na.omit(data8), 
               method = "glm", 
               trControl = train_control)
model
plot(model$pred$pred ~ model$pred$obs)

```

## Prepping gov data

```{r}
# what governance looks like:
gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  group_by(scenario, year) %>%
  summarize(governance_mean = mean(governance, na.rm=TRUE),
            sd = sd(governance, na.rm=TRUE))
ggplot(gov_raw, aes(x=year, y=governance_mean, group=scenario, color=scenario)) +
  geom_point() +
  geom_line()

ggplot(data8, aes(x=governance, y=slope)) + 
  geom_point() + 
  geom_smooth()


# going to estimate yearly change for gapfilling regions with current estimates, but no future estimates
# just using the average patterns
gov_ssp_change_data <- gov_raw %>%
  group_by(scenario) %>%
  select(scenario, year, governance_mean) %>%
  filter(year >= 2015) %>%
  arrange(year) 

start_gov <- filter(gov_ssp_change_data, year==2015 & scenario != "Observed") %>%
  select(scenario, start_governance_mean = governance_mean)

gov_ssp_change <- gov_ssp_change_data %>%
  left_join(start_gov, by= "scenario") %>%
  mutate(governance_yearly_change = governance_mean - start_governance_mean) %>%
  filter(scenario != "Observed") %>%
  select(scenario, year, governance_yearly_change)

ggplot(gov_ssp_change, aes(x=year, y=governance_yearly_change, group=scenario)) + 
  geom_line()


## preparing gov_data
gov_data <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  select(iso3c=countrycode, scenario, year, governance)

correction <- gov_data %>%
  filter(year == 2015) %>%
  group_by(iso3c) %>%
  mutate(observed = ifelse(scenario == "Observed", governance, NA_real_)) %>%
  fill(observed, .direction = "downup") %>%  # Filling NA values within each group
  mutate(correction_value = governance-observed) %>%
  filter(scenario != "Observed") %>%
  select(iso3c, scenario, correction_value)
  

obs_gov <- gov_data %>%
  filter(scenario == "Observed") %>%
  filter(year>2010) %>%
  select(iso3c, year, governance)
obs_gov <- crossing(obs_gov, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))

est_gov <- gov_data %>%
  filter(scenario != "Observed") %>%
  filter(!is.na(scenario)) %>%
  filter(year > 2015) %>% 
  select(iso3c, year, governance, scenario)
  
gov_scores <- rbind(obs_gov, est_gov) %>%
  left_join(correction, by=c("iso3c", "scenario")) %>%
  mutate(correction_value = ifelse(year <= 2015, 0, correction_value)) %>%
  mutate(corrected_governance = governance - correction_value) 

check <- gov_scores %>%
  group_by(year, scenario) %>%
  summarize(gov = mean(corrected_governance, na.rm=TRUE))

ggplot(check, aes(x=year, y=gov, group=scenario)) +
  geom_line() + 
  geom_point() 

ggplot(filter(gov_scores, iso3c=="USA"), aes(x=year, y=governance, group=scenario)) +
  geom_line() + 
  geom_point() +
  geom_line(aes(x=year, y=corrected_governance, group=scenario), col="red")

gov_scores <- gov_scores %>%
  select(iso3c, year, scenario, governance=corrected_governance) %>%
  filter(!is.na(governance))


```


## sampling


```{r}

sim_data <- pressure_eez7 %>%
  filter(ant_type == "eez") %>%
  select(ant_type, rgn_id, rgn_name, iso3c, avg_biomass, avg_percent_change, slope, governance) %>%
  mutate(year = 2005) 

sim_data <- na.omit(sim_data)

sample_low_gov <- sim_data %>%
  filter(governance < 0.4) %>%
  pull(slope)
low_gov_avg <- mean(sample_low_gov)
low_gov_sd <- sd(sample_low_gov)

sample_mid_gov <- sim_data %>%
  filter(governance >= 0.4 & governance < 0.75) %>%
  pull(slope)
mid_gov_avg <- mean(sample_mid_gov)
mid_gov_sd <- sd(sample_mid_gov)

sample_high_gov <- sim_data %>%
  filter(governance >= 0.75) %>%
  pull(slope)
high_gov_avg <- mean(sample_high_gov)
high_gov_sd <- sd(sample_high_gov)

low_gov_avg
mid_gov_avg
high_gov_avg

sim_data_final <- sim_data %>%
  ungroup() %>%
  select(iso3c, biomass_t0 = avg_biomass, pct_change_t0 = avg_percent_change, slope, governance, year)


## prepping and gapfilling governance scenario data

#missing_recent <- data.frame(iso3c = c("ESH", "COD", "ROU"),
#                            governance = c(mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="ESH"]),
#                            mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="COD"]),
#                            mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="ROU"])))
#missing_recent <- crossing(missing_recent, year = 2011:2015, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))
#gov_scores_gf <- rbind(gov_scores, missing_recent)

missing_gov_iso <- setdiff(sim_data_final$iso3c, gov_scores$iso3c)
missing_gov <- expand.grid(iso3c = missing_gov_iso, year = c(2011:2099), scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))

missing_gov_estimated <- filter(sim_data_final, iso3c %in% missing_gov_iso) %>%
  select(iso3c, governance)
missing_gov_estimated <- crossing(missing_gov_estimated, year= 2011:2099, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5")) %>%
  left_join(gov_ssp_change, by=c("scenario", "year")) %>%
  mutate(governance_yearly_change = ifelse(is.na(governance_yearly_change), 0, governance_yearly_change)) %>%
  mutate(governance2 = governance + governance_yearly_change) %>%
  select(iso3c, year, governance=governance2, scenario)
ggplot(filter(missing_gov_estimated, iso3c=="ABW"), aes(x=year, y=governance, group=scenario)) +
  geom_line()

gov_scores_gf <- rbind(gov_scores, missing_gov_estimated)
setdiff(gov_scores_gf$iso3c, sim_data_final$iso3c) # mostly landlocked
setdiff(sim_data_final$iso3c, gov_scores_gf$iso3c)
## now prepare sample data

########################################
## function to run randomization:
########################################
randomized_fishing <- function(scenario_selects="SSP1", sim_data_finals = sim_data_final, gov_scores_gfs=gov_scores_gf){

  for (year in 2005:2099){ #year = 2010 scenario_selects="SSP1"

scenario_select <- scenario_selects

if(year==2005){
t0 <- sim_data_finals %>%
  select(-slope) %>%
  mutate(scenario = scenario_select) %>%
  mutate(!!paste0("biomass_", year) := biomass_t0) %>%
  mutate(!!paste0("pct_change_", year) := pct_change_t0) 

} else{
  t0 <- simulation
}

  if(year %in% c(2005:2010)){
simulation <- t0 %>%
  rowwise() %>%
  mutate(
    change_pct_change_random = case_when(
      governance < 0.4 ~ rnorm(1, mean=low_gov_avg, sd=low_gov_sd),
      governance >= 0.4 & governance < 0.75 ~ rnorm(1, mean=mid_gov_avg, sd=mid_gov_sd),
      governance >= 0.75 ~ rnorm(1, mean=high_gov_avg, sd = high_gov_sd))) %>%
  mutate(pct_change_t1 = pct_change_t0 + change_pct_change_random) %>%
  mutate(!!paste0("pct_change_", year+1) := pct_change_t1) %>%
  mutate(!!paste0("change_pct_change", year+1) := change_pct_change_random) %>%
  mutate(biomass_t1 = biomass_t0 * (1 + pct_change_t1)) %>%
  mutate(!!paste0("biomass_", year+1) := biomass_t1) %>%
  mutate(year=year+1) %>%
  select(-biomass_t0, -pct_change_t0) %>%
  rename(biomass_t0 = biomass_t1, pct_change_t0=pct_change_t1) 

#cat(year, "\n", "dim = ", dim(simulation)[1], "\n") # for troubleshooting
  } else {
    
simulation <- t0 %>%
  select(-governance) %>%
  left_join(gov_scores_gfs, by=c("iso3c", "year", "scenario")) %>%
  rowwise() %>%
  mutate(
    change_pct_change_random = case_when(
      governance < 0.4 ~ rnorm(1, mean=low_gov_avg, sd=low_gov_sd),
      governance >= 0.4 & governance < 0.75 ~ rnorm(1, mean=mid_gov_avg, sd=mid_gov_sd),
      governance >= 0.75 ~ rnorm(1, mean=high_gov_avg, sd = high_gov_sd))) %>%
  mutate(pct_change_t1 = pct_change_t0 + change_pct_change_random) %>%
  mutate(!!paste0("pct_change_", year+1) := pct_change_t1) %>%
  mutate(!!paste0("change_pct_change", year+1) := change_pct_change_random) %>%
  mutate(biomass_t1 = biomass_t0 * (1 + pct_change_t1)) %>%
  mutate(!!paste0("biomass_", year+1) := biomass_t1) %>%
  mutate(year=year+1) %>%
  select(-biomass_t0, -pct_change_t0) %>%
  rename(biomass_t0 = biomass_t1, pct_change_t0=pct_change_t1) 
#cat(year, "\n", "dim = ", dim(simulation)[1], "\n") # for troubleshooting

  }
  }
  return(simulation)
}

########################################
## end of function to run randomization
########################################

summary(simulation_long)
ggplot(filter(simulation_long, year <=2060), aes(x=year, y=biomass, group=iso3c, col=governance)) +
  geom_line()

simulation_long %>%
  group_by(year) %>%
  summarize(sum(biomass, na.rm=TRUE)) %>% print(n=100)




## Run simulations

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP1", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)

scenario_select = "SSP1"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}


for(i in 1:100){
  
tmp <- randomized_fishing(scenario_selects="SSP2", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)

scenario_select = "SSP2"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP3", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select="SSP3"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP4", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select <- "SSP4"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP5", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select <- "SSP5"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

```


```{r}
tmp <- read_csv("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_SSP1_s1.csv")

files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/", full=TRUE, pattern=".csv")

rand_biomass <- lapply(files, read.csv) %>% bind_rows()

## compare to global observed: 2005:2010
## ensure estimates are not going off the rails.
compare <- rand_biomass %>%
  group_by(simulation, scenario, year) %>%
  summarize(global_biomass = sum(biomass)) %>%
  filter(year %in% 2005:2010)
compare %>%
  group_by(year) %>%
  summarize(simulated_global_biomass = sum(global_biomass))

check <- country_biomass_change
check$iso3c <- countrycode(check$rgn_nam, "country.name", "iso3c")
check <- filter(check, iso3c %in% unique(tmp$iso3c))
observed <- check %>%
  group_by(year) %>%
  summarize(observed_biomass = sum(biomass, na.rm=TRUE) - 3865822)

ggplot() + 
  geom_line(data=compare, aes(x=year, y=global_biomass, group=simulation), alpha=0.2) +
  geom_point(data=filter(observed, year>=2005), aes(x=year, y=observed_biomass), col="red")

## note that the 2005 biomass estimate is lower in simulated data data due to averaging 2000-2010 
## to arrive at this value. (consider changing)
692322460-688456638

US_means <- rand_biomass %>%
  filter(iso3c=="USA" & scenario=="SSP1" & year %in% c(2010:2050)) %>%
  group_by(year) %>%
    summarize(mean_biomass = mean(biomass))

ggplot(data=filter(rand_biomass, iso3c=="USA" & scenario=="SSP1" & year%in%c(2010:2050)), aes(x=year, y=biomass)) +
  geom_point() +
  geom_point(data=US_means, aes(x=year, y=mean_biomass), col="red") +
    geom_line(data=US_means, aes(x=year, y=mean_biomass), col="red")
  


## constrain data, to not go above or below a particular percentage.
## get reference. biomass:
ref_times <- which(as.Date(time(biomass)) >= as.Date("1900-01-01") & as.Date(time(biomass)) < as.Date("1910-01-01")) 
ref_density_rasters <- subset(biomass, ref_times)
ref_biomass_rasters <- cellSize(ref_density_rasters) * ref_density_rasters * 1e-6
ref_biomass <- mean(ref_biomass_rasters)

ref_country_biomass <- terra::zonal(ref_biomass, terra::rast(zones_raster), fun = sum, na.rm=TRUE)

ref_country_biomass <- ref_country_biomass %>% 
 rename(rgn_id = layer, ref_biomass=mean) %>%
  left_join(rgn_data, by="rgn_id") %>%
  filter(!is.na(ref_biomass)) %>%
  filter(ant_typ == "eez")

ref_country_biomass$iso3c <- countrycode(ref_country_biomass$rgn_nam, "country.name", "iso3c") %>%
  filter(!is.na(iso3c))

## Get time 0 biomass
t0_times <- which(as.Date(time(biomass)) >= as.Date("1905-01-01") & as.Date(time(biomass)) < as.Date("1906-01-01")) 
t0_density_rasters <- subset(biomass, t0_times)
t0_biomass_rasters <- cellSize(biomass_c) * biomass_c * 1e-6
t0_biomass <- mean(ref_biomass_rasters)

ref_country_biomass <- terra::zonal(ref_biomass, terra::rast(zones_raster), fun = sum, na.rm=TRUE)

ref_country_biomass <- ref_country_biomass %>% 
 rename(rgn_id = layer, ref_biomass=mean) %>%
  left_join(rgn_data, by="rgn_id") %>%
  filter(!is.na(ref_biomass)) %>%
  filter(ant_typ == "eez")

ref_country_biomass$iso3c <- countrycode(ref_country_biomass$rgn_nam, "country.name", "iso3c") %>%
  filter(!is.na(iso3c))


```