---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

step1: extract biomass data for eezs.
step2: data and models to see if we can predict changes to fisheries pressures
This is mostly an information gathering step that can be skipped. 
step3: prepare governance data. I need two datasets: 
 - average governance from 2000 to 2010, with gapfilling, for estimating slope of change in fishing pressure from 2000 to 2010
 - 2000-2100 governance, with gapfilling
step4: models for future biomass for populated EEZ, HS, and unpopulated EEZ

 
```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(WDI)
```


# Populated EEZs

### fisheries data
(generated in script: step 1)
(governance data: step 3)
```{r}

fish_pressure <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv"))
governance <- read_csv(here("prep/fisheries/data/avg_2000_2010_governance_gf.csv"))
pressure_data <- left_join(governance, fish_pressure, by=c("rgn_id", "rgn_nam", "iso3c")) %>%
  filter(!(is.na(slope))) # average country data

```


## Develop model to randomly sample from
Model and parameters. Predict change in fisheries pressure based on governance.
```{r}

## model to predict slope based on government
mod <- lm(slope ~ governance, data = filter(pressure_data, is.na(gapfilled) & avg_percent_change<0))
summary(mod) #R2 = 0.20
ggplot(mod, aes(y = slope, x=governance)) +
  geom_point() +
  geom_smooth()


# get the model values
intercept <- coef(mod)["(Intercept)"]
governance_coeff <- coef(mod)["governance"]

quantile(pressure_data$avg_percent_change, c(0.05)) # change will not fall below this rate
quantile(pressure_data$avg_percent_change, c(0.95)) # change will go above this rate

tmp <- filter(pressure_data, avg_percent_change<0)
quantile(tmp$avg_percent_change, c(0.95)) # nudge for SSP2
quantile(tmp$avg_percent_change, c(0.90)) # nudge for SSP5


gov_projections <- read_csv(here("prep/fisheries/data/governance_2000_2100.csv")) %>%
  mutate(governance = ifelse(governance < 0.2, 0.2, governance))


# Calculate the expected slopes for all rows in one operation
gov_projections <- gov_projections %>%
  rowwise() %>%
  mutate(acceleration_projected = intercept + governance_coeff * governance) %>% 
  arrange(iso3c, scenario, year) %>%
  filter(year >= 2005)


```

## Data for model
```{r}

## summarized data for each region (slope, avg_percent_change, etc.)
estimated_biomass <- fish_pressure %>%
  filter(iso3c %in% gov_projections$iso3c) %>%
  mutate(starting_prop = biomass_2005/historic_1900) %>%
  mutate(avg_percent_change = ifelse(avg_percent_change > 0 & starting_prop > 0.45, 0, avg_percent_change)) %>%  #assume no place will be increasing if meeting sustainability thresholds
  select(rgn_id, rgn_nam, iso3c, avg_biomass, avg_percent_change, historic_1900, biomass_2005, prop = prop_biomass, starting_prop)

model_data <- left_join(gov_projections, estimated_biomass, by=c("rgn_id", "rgn_nam", "iso3c"))
```

## Function for updating biomass based on governance
This is used in the next code chunk
```{r}

update_biomass <- function(df) {
  # df <- new_data_projected
  #print(df)
  # df <- filter(new_data_projected_v2, iso3c=="CCK", scenario == "SSP2")
  
  df$prop[1] <- df$biomass_2005[1] / df$historic_1900[1]
  df$avg_percent_change[1] <- df$avg_percent_change[1]
  df$avg_biomass[1] <- df$biomass_2005[1]  # starting value is 2005
  df$acceleration_projected[1] <- 0 ## starting value is zero for change
  
  scenario <- df$scenario2[1]
  starting_prop <- df$starting_prop[1]
  
  # print(scenario)

    if (scenario == "SSP1") {
      nudge <- 0
      upper_threshold <- 0.65
      lower_threshold <- 0.2
    } else if (scenario == "SSP2") {
      nudge <- -0.002  
      upper_threshold <- 0.45
      lower_threshold <- 0.15
    } else if (scenario == "SSP3" || scenario == "SSP4") {
      nudge <- -0.005  
      upper_threshold <- 0.45
      lower_threshold <- 0.15
    } else if (scenario == "SSP5") {
      nudge <- -0.005 
      upper_threshold <- 0.45
      lower_threshold <- 0.15
    } else {
    stop("Error: Unrecognized scenario value.")
    }
   
  for (i in 2:nrow(df)) { # i=2
    
    new_percent_change <- df$avg_percent_change[i - 1] + df$acceleration_projected[i] 
    
    if(new_percent_change < -0.03){ ## keep rate of change going beyond what is typically observed, 5th quantile
      new_percent_change <- -0.03
    }

    if(new_percent_change > 0.012){ ## keep rate of change going beyond what is typically observed, 95th quantile
      new_percent_change <- 0.012
    }
    
    potential_biomass <- df$avg_biomass[i - 1] * (1 + new_percent_change)
    potential_prop <- potential_biomass / df$historic_1900[i]


    ## Move high biomass lower towards a sustainable yield, if percent_change is positive
     if (potential_prop > upper_threshold & new_percent_change >= nudge) {
      new_percent_change <- nudge
      potential_biomass <- df$avg_biomass[i - 1] * (1 + new_percent_change)
      potential_prop <- potential_biomass / df$historic_1900[i]
    }
    
      ## keep from falling below lower threshold
     if (potential_prop < lower_threshold & new_percent_change < 0) {
      new_percent_change <- 0
      potential_biomass <- df$avg_biomass[i - 1] * (1 + new_percent_change)
      potential_prop <- potential_biomass / df$historic_1900[i]
     }
    
    ## keep high biomass countries from getting higher, with no nudge (SSP1)
      if (potential_prop >= starting_prop & potential_prop > upper_threshold & scenario=="SSP1") {
      new_percent_change <- 0
      potential_biomass <- df$avg_biomass[i - 1] * (1 + new_percent_change)
      potential_prop <- potential_biomass / df$historic_1900[i]
    }


    # Apply the adjusted percent change and update yearly values
    df$avg_percent_change[i] <- new_percent_change
    df$avg_biomass[i] <- df$avg_biomass[i - 1] * (1 + df$avg_percent_change[i])
    df$prop[i] <- df$avg_biomass[i] / df$historic_1900[i]
  }

  return(df)
}

```


## Run the function to get data

```{r}

## model 1: average biomass loss can accelerate and decelerate based on governance
new_data_projected <- model_data %>%
  mutate(scenario2 = scenario) %>%
  group_by(iso3c, rgn_id, rgn_nam, scenario) %>%
  arrange(iso3c, rgn_id, rgn_nam, scenario, year) %>%
  group_modify(~ update_biomass(.x)) %>%
  ungroup() %>%
  select(-scenario2)

## model 2: average biomass loss can only decelerate based on governance
new_data_projected_v2 <- model_data %>%
  mutate(acceleration_projected = ifelse(acceleration_projected < 0, 0, acceleration_projected)) %>%
  mutate(scenario2 = scenario) %>%
  group_by(iso3c, rgn_id, rgn_nam, scenario) %>%
  arrange(iso3c, rgn_id, rgn_nam, scenario, year) %>%
  group_modify(~ update_biomass(.x)) %>%
  ungroup() %>%
  select(iso3c, scenario, year, prop_no_accel = prop, avg_biomass_no_accel = avg_biomass)

proj_data <- left_join(new_data_projected, new_data_projected_v2) %>%
  select(scenario, iso3c, rgn_nam, rgn_id, year, governance, prop_accel = prop, prop_no_accel, avg_biomass_accel = avg_biomass, avg_biomass_no_accel, historic_1900)

write_csv(proj_data, here("prep/fisheries/data/habited_eez_biomass.csv"))

```

## Some explorations of the data
## this series explores the populated EEZs
```{r}

plot_data <- left_join(new_data_projected, new_data_projected_v2) %>% 
  filter(iso3c %in% c("USA", "CCK", "CHN", "BEL", "CIV", "COD", "DEU", "ECU") & scenario =="SSP2")

## plot 1: explore some countries with different tendencies
# observed data: this file created in step1
country_biomass_change <- read_csv("prep/fisheries/data/country_biomass_2000_2010.csv") %>%
  filter(iso3c %in% c("USA", "CCK", "CHN", "BEL", "CIV", "COD", "DEU", "ECU") & year %in% 2000:2010) %>%
  select(iso3c, year, biomass_obs = biomass) 

country_biomass_change <- left_join(country_biomass_change, fish_pressure, by="iso3c") %>%
  mutate(prop_obs = biomass_obs/historic_1900)
  

ggplot(plot_data, aes(x=year, y=prop, group=iso3c, color=iso3c)) +
  geom_line(linetype="dotted") +
  geom_line(aes(x=year, y=prop_no_accel, group=iso3c, color=iso3c)) +
  geom_hline(yintercept=0.45, color="orange") +
  geom_hline(yintercept=0.15, color="red") +
  geom_point(data = country_biomass_change, aes(x=year, y=prop_obs, color=iso3c)) +
  geom_text(
    data = plot_data %>% group_by(iso3c) %>% filter(year == max(year)), 
    aes(x = 2000, y = starting_prop, label = iso3c),
    hjust = 0,
    nudge_x = 0.5,
    size = 3
  )



## plot 1: explore some countries with different tendencies
# observed data: this file created in step1

plot_data <- left_join(new_data_projected, new_data_projected_v2) %>% 
  filter(iso3c %in% c("CCK", "NCL", "VUT", "SLB", "PNG", "FJI") & scenario =="SSP2")

country_biomass_change <- read_csv("prep/fisheries/data/country_biomass_2000_2010.csv") %>%
  filter(iso3c %in% c("CCK", "NCL", "VUT", "SLB", "PNG", "FJI") & year %in% 2000:2010) %>%
  select(iso3c, year, biomass_obs = biomass) 

country_biomass_change <- left_join(country_biomass_change, fish_pressure, by="iso3c") %>%
  mutate(prop_obs = biomass_obs/historic_1900)
  

ggplot(plot_data, aes(x=year, y=prop, group=iso3c, color=iso3c)) +
  geom_line(linetype="dotted") +
  geom_line(aes(x=year, y=prop_no_accel, group=iso3c, color=iso3c)) +
  geom_hline(yintercept=0.45, color="orange") +
  geom_hline(yintercept=0.15, color="red") +
  geom_point(data = country_biomass_change, aes(x=year, y=prop_obs, color=iso3c)) +
  geom_text(
    data = plot_data %>% group_by(iso3c) %>% filter(year == max(year)), 
    aes(x = 2000, y = starting_prop, label = iso3c),
    hjust = 0,
    nudge_x = 0.5,
    size = 3
  )
## plot 2: Compare 2 models (acceleration and deceleration vs. deceleration only)

## first a plot to see how the scenarios compare:
ssp_plot <- left_join(new_data_projected, new_data_projected_v2) %>%
  select(scenario, iso3c, rgn_nam, year, avg_biomass, avg_biomass_no_accel, historic_1900) %>%
  group_by(scenario, year) %>%
  summarize(biomass_accel = sum(avg_biomass, na.rm=TRUE),
            biomass_noaccel = sum(avg_biomass_no_accel, na.rm=TRUE),
            biomass_hist = sum(historic_1900)) %>%
  ungroup() %>%
  rowwise() %>% 
  mutate(prop_accel = biomass_accel/biomass_hist,
         prop_noaccel = biomass_noaccel/biomass_hist)


## add in observed
obs_biomass <- read_csv("prep/fisheries/data/country_1900_2010_biomass.csv") %>%
  filter(iso3c %in% proj_data$iso3c) %>%
  group_by(year) %>%
  summarize(biomass_obs = sum(biomass, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(obs_prop = biomass_obs/1148446559) %>%
  filter(year >=1950)

ggplot(ssp_plot, aes(x = year, color = scenario, group = scenario)) +
  geom_point(aes(y = prop_accel)) +
  geom_line(aes(y = prop_accel)) +
  geom_point(aes(y = prop_noaccel), alpha = 0.15) +
  geom_line(aes(y = prop_noaccel), alpha = 0.15) +
  geom_point(data=obs_biomass, aes(x=year, y=obs_prop), inherit.aes=FALSE)  
  

## plot 3: years 2005-2010, 

country_biomass_change <- read_csv("prep/fisheries/data/country_biomass_2000_2010.csv")  %>%
  select(iso3c, year, biomass_obs = biomass) %>%
  filter(!is.na(iso3c))

country_biomass_change <- left_join(country_biomass_change, select(fish_pressure, c(iso3c, historic_1900)), by="iso3c") %>%
  mutate(prop_obs = biomass_obs/historic_1900)

mod_data <- left_join(country_biomass_change, proj_data, by=c("iso3c", "year")) %>%
  filter(year >= 2005)

plot_data <- mod_data %>%
  group_by(scenario, year) %>%
  summarize(historic_biomass = sum(historic_1900.x, na.rm=TRUE),
            biomass_obs = sum(biomass_obs),
            biomass_accel = sum(avg_biomass_accel),
            biomass_no_accel = sum(avg_biomass_no_accel)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(obs_prop = biomass_obs/historic_biomass,
         accel_prop = biomass_accel/historic_biomass,
         no_accel_prop = biomass_no_accel/historic_biomass) %>%
   pivot_longer(
    cols = contains("prop"),  # Specify the columns to pivot
    names_to = "prop",      # Name for the new variable column
    values_to = "value"     # Name for the new value column
  )

ggplot(filter(plot_data, scenario=="SSP2"), aes(x=year, y=value, group = prop, color = prop)) +
  geom_point()

```






# unpopulated EEZs

## Get the data
```{r}
pressure_avg <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv")) %>%
  mutate(ant_typ = ifelse(!is.na(uninhabited), "eez_uninhabited", ant_typ))

biomass_rgn <- read_csv(here("prep/fisheries/data/country_biomass_2000_2010.csv")) %>%
  mutate(ant_typ = ifelse(!is.na(uninhabited), "eez_uninhabited", ant_typ))

```


Parameters for future predictions.
The mean  of the annual rate of change in biomass across uninhabited island regions is used to estimate future change.
```{r}


# using the 10 year average change
unin_model <- biomass_rgn %>%
  filter(year >= 2000) %>%
  filter(ant_typ == "eez_uninhabited") %>%
  group_by(rgn_id, rgn_nam, iso3c) %>%
  summarize(start_biomass = first(biomass[year == min(year)]),
            end_biomass = first(biomass[year == max(year)]),
            start_year = min(year),
            end_year = max(year),
            avg_biomass = mean(biomass),
            biomass_2005 = biomass[year==2005],
            biomass_2010 = biomass[year==2010]) %>%
  mutate(duration = end_year - start_year,
         avg_percent_change = ((end_biomass / start_biomass)^(1 / duration) - 1)) %>%
  select(rgn_id, rgn_nam, iso3c, avg_biomass, biomass_2005, biomass_2010, avg_percent_change) %>%
  ungroup()


mean_model <- mean(unin_model$avg_percent_change)

```

Creating the region specific data:
```{r}


scenario_year_template <- expand.grid(scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"),
                                      year = 2000:2100,
                                      rgn_id = unique(unin_model$rgn_id))

unin_average <- pressure_avg %>%
  filter(ant_typ %in% "eez_uninhabited") %>%
  select(rgn_id, iso3c, rgn_nam, biomass_2005, biomass_2010, avg_percent_change, historic_1900)

scenario_df <- left_join(scenario_year_template, unin_average, by=c("rgn_id")) %>%
  mutate(biomass = NA, prop=NA)


```

Function to update yearly biomass based on model:
```{r}

update_biomass <- function(df) {
  # df <- new_data_sim %>% filter(iso3c == "SLB" & scenario == "SSP2")
  #print(df)
  
  df$prop[1] <- df$biomass_2005[1] / df$historic_1900[1]
  df$biomass[1] <- df$biomass_2005[1]  # starting value is 2005

  scenario <- df$scenario2[1]
  
  # print(scenario)

    if (scenario == "SSP1") {
        lower_threshold <- 0.75
      } else if (scenario == "SSP2") {
      lower_threshold <- 0.65
    } else if (scenario == "SSP3" || scenario == "SSP4" || scenario == "SSP5") {
      lower_threshold <- 0.45
    } else {
    stop("Error: Unrecognized scenario value.")
    }

  
  for (i in 2:nrow(df)) { # i=2
    
    potential_biomass <- df$biomass[i - 1] * (1 + df$rate_change[i-1])
    potential_prop <- potential_biomass / df$historic_1900[i]


     adjusted_rate_change <- df$rate_change[i-1]

     
    # Constraints
      if (potential_prop < lower_threshold) {
      adjusted_rate_change <- 0         
        }
   
    # Apply the adjusted percent change and update yearly values
    df$biomass[i] <- df$biomass[i - 1] * (1 + adjusted_rate_change)
    df$prop[i] <- df$biomass[i] / df$historic_1900[i]
  }

  return(df)
}

```

Run function
```{r}

# Apply the function to each group
uninhab_model1_df <- scenario_df %>%
  filter(year>=2005) %>%
  ## add in annual change
  mutate(rate_change = mean_model) %>%
  mutate(scenario2 = scenario) %>%
  arrange(rgn_id, rgn_nam, scenario, year) %>%
  group_by(iso3c, rgn_id, rgn_nam, scenario) %>%
  group_modify(~ update_biomass(.x)) %>%
  ungroup()  %>%
  select(scenario, iso3c, rgn_nam, rgn_id, year, prop_accel=prop, prop_no_accel = prop, avg_biomass_accel = biomass, avg_biomass_no_accel = biomass, historic_1900)

write_csv(uninhab_model1_df, here("prep/fisheries/data/uninhabited_eez_biomass.csv"))


ggplot(filter(uninhab_model1_df, scenario=="SSP2"), aes(x = year, y = prop_accel, color =iso3c, group=iso3c)) +
  geom_line() +
  geom_point() 


## plot 2: Compare 2 models (acceleration and deceleration vs. deceleration only)

## first a plot to see how the scenarios compare:
uninhab_plot <- uninhab_model1_df %>%
  select(scenario, iso3c, rgn_nam, year, avg_biomass_accel, historic_1900) %>%
  group_by(scenario, year) %>%
  summarize(biomass = sum(avg_biomass_accel, na.rm=TRUE),
            biomass_hist = sum(historic_1900)) %>%
  ungroup() %>%
  rowwise() %>% 
  mutate(prop = biomass/biomass_hist)


## add in observed
obs_biomass <- read_csv("prep/fisheries/data/country_1900_2010_biomass.csv") %>%
  filter(iso3c %in% uninhab_model1_df$iso3c) %>%
  group_by(year) %>%
  summarize(biomass_obs = sum(biomass, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(obs_prop = biomass_obs/55749887) %>%
  filter(year >=1950)

ggplot(uninhab_plot, aes(x = year, color = scenario, group = scenario)) +
  geom_point(aes(y = prop)) +
  geom_line(aes(y = prop)) +
  geom_point(data=obs_biomass, aes(x=year, y=obs_prop), inherit.aes=FALSE)  
  

```




# high seas regions

## Get the data
```{r}

pressure_avg <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv")) %>%
  mutate(ant_typ = ifelse(!is.na(uninhabited), "eez_uninhabited", ant_typ))

biomass_rgn <- read_csv(here("prep/fisheries/data/country_biomass_2000_2010.csv")) %>%
  mutate(ant_typ = ifelse(!is.na(uninhabited), "eez_uninhabited", ant_typ))

ggplot(pressure_avg, aes(x=avg_percent_change, group=ant_typ, fill=ant_typ)) +
  geom_histogram()

hs_avg <- biomass_rgn %>%
  filter(ant_typ %in% c("fao", "eez-ccamlr")) %>%
  select(year, biomass) %>%
  group_by(year) %>%
  summarize(biomass=sum(biomass, na.rm=TRUE))

ggplot(hs_avg, aes(x=year, y=biomass)) + 
  geom_line() 


```


Parameters for future predictions.
The mean  of the annual rate of change in biomass across uninhabited island regions is used to estimate future change.
```{r}

hs_model1 <- pressure_avg %>%
  filter(ant_typ %in% c("fao", "eez-ccamlr"))


hs_mean_model1 <- mean(hs_model1$avg_percent_change)
```

Creating the high seas, region specific data:
```{r}


scenario_year_template <- expand.grid(scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"),
                                      year = 2000:2100,
                                      rgn_id = unique(hs_model1$rgn_id))

hs_average <- pressure_avg %>%
  filter(ant_typ %in% c("fao", "eez-ccamlr"))%>%
  select(rgn_id, rgn_nam, biomass_2005, biomass_2010, avg_percent_change, historic_1900)

scenario_df <- left_join(scenario_year_template, hs_average, by=c("rgn_id")) %>%
  mutate(biomass = NA, prop=NA)


```

Function to update yearly biomass based on model:
```{r}

update_biomass <- function(df) {
  # df <- hs_model1_df %>% filter(rgn_id == 213 & scenario == "SSP2")
  #print(df)
  
  df$prop[1] <- df$biomass_2005[1] / df$historic_1900[1]
  df$biomass[1] <- df$biomass_2005[1]  # starting value is 2005

  scenario <- df$scenario2[1]
  
  # print(scenario)

    if (scenario == "SSP1") {
        lower_threshold <- 0.75
      } else if (scenario == "SSP2") {
      lower_threshold <- 0.65
    } else if (scenario == "SSP3" || scenario == "SSP4" || scenario == "SSP5") {
      lower_threshold <- 0.55
    } else {
    stop("Error: Unrecognized scenario value.")
    }

  
  for (i in 2:nrow(df)) { # i=2
    
    potential_biomass <- df$biomass[i - 1] * (1 + df$rate_change[i-1])
    potential_prop <- potential_biomass / df$historic_1900[i]


     adjusted_rate_change <- df$rate_change[i-1]

     
    # Constraints
      if (scenario != "SSP1" && potential_prop < lower_threshold) {
      adjusted_rate_change <- 0         
        }
   
    # Apply the adjusted percent change and update yearly values
    df$biomass[i] <- df$biomass[i - 1] * (1 + adjusted_rate_change)
    df$prop[i] <- df$biomass[i] / df$historic_1900[i]
  }

  return(df)
}

```

Run function
```{r}

# Apply the function to each group
hs_model1_df <- scenario_df %>%
  filter(year>=2005) %>%
  ## add in annual change
  mutate(rate_change = hs_mean_model1) %>%
  mutate(scenario2 = scenario) %>%
  arrange(rgn_id, rgn_nam, scenario, year) %>%
  group_by(rgn_id, rgn_nam, scenario) %>%
  group_modify(~ update_biomass(.x)) %>%
  ungroup()  %>%
  select(scenario, rgn_nam, rgn_id, year, prop_accel=prop, prop_no_accel = prop, avg_biomass_accel = biomass, avg_biomass_no_accel = biomass, historic_1900)

write_csv(hs_model1_df, here("prep/fisheries/data/hs_biomass.csv"))




ggplot(filter(hs_model1_df, scenario=="SSP2"), aes(x = year, y = prop_no_accel, color =rgn_nam, group=rgn_nam)) +
  geom_line() +
  geom_point() 


## plot 2: Compare 2 models (acceleration and deceleration vs. deceleration only)

## first a plot to see how the scenarios compare:
hs_plot <- hs_model1_df %>%
  select(scenario, rgn_id, rgn_nam, year, avg_biomass_no_accel, historic_1900) %>%
  group_by(scenario, year) %>%
  summarize(biomass = sum(avg_biomass_no_accel, na.rm=TRUE),
            biomass_hist = sum(historic_1900)) %>%
  ungroup() %>%
  rowwise() %>% 
  mutate(prop = biomass/biomass_hist)


## add in observed
obs_biomass <- read_csv("prep/fisheries/data/country_1900_2010_biomass.csv") %>%
  filter(rgn_id %in% hs_model1_df$rgn_id) %>%
  group_by(year) %>%
  summarize(biomass_obs = sum(biomass, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(obs_prop = biomass_obs/868195557) %>%
  filter(year >=1950)

ggplot(hs_plot, aes(x = year, color = scenario, group = scenario)) +
  geom_point(aes(y = prop)) +
  geom_line(aes(y = prop)) +
  geom_point(data=obs_biomass, aes(x=year, y=obs_prop), inherit.aes=FALSE)  
  

```











