---
title: "combining_regions"
output: html_document
date: "2024-06-21"
---

This script combines the fishing pressure data for the three categories of regions:
habited EEZ, uninhabited EEZ, high seas.

The fishing pressures are then mapped by region and the data is formatted to align with the other pressures.

```{r setup, include=FALSE}

library(sf)
library(terra)
library(countrycode)
library(broom)
library(tidyverse)
library(here)
library(rnaturalearth)

```

## Get the data
```{r}


# eez
eez_mod <- read_csv(here("prep/fisheries/data/habited_eez_biomass.csv"))

## uninhabited
uninhabited_mod <- read_csv(here("prep/fisheries/data/uninhabited_eez_biomass.csv"))

## high seas
hs_mod <- read_csv(here("prep/fisheries/data/hs_biomass.csv"))

fish_data <- bind_rows(eez_mod, uninhabited_mod, hs_mod) %>%
  mutate(pressure = 1- avg_biomass_accel/historic_1900) %>%
  mutate(pressure = ifelse(pressure < 0, 0, pressure)) %>%
  mutate(pressure_no_accel = 1- avg_biomass_no_accel/historic_1900) %>%
  mutate(pressure_no_accel = ifelse(pressure_no_accel < 0, 0, pressure_no_accel))

```

Get the maps
```{r}
## EEZ regions
eez_shp <- sf::st_read("/home/shares/ohi/git-annex/globalprep/spatial/v2017/regions_2017_update.shp") 
#eez <- sf::st_transform(eez_shp, 4326)
eez_data <- read_csv(here("prep/fisheries/data/region_id.csv"))

rast_base_10km_file <- terra::rast(here('_spatial/rast_base_mol_10km.tif'))
ocean_template <- terra::rast(here('_spatial/ocean_area_mol.tif'))
ocean_template[ocean_template>0] <- 1
terra::plot(ocean_template)


```

Associate the data and make the maps
```{r}

# all the iterations I need
pressure_maps <- data.frame(scenarios = rep(c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"), each=4),
                            time_frame = rep(c("current", "near-term", "medium-term", "long-term"), 5),
                            years = rep(c("2010", "2021:2040", "2041:2060", "2080:2100"),5))

# function to make maps
mapping_fish_pressure <- function(scenarios, time_frame, years){
#scenarios = pressure_maps$scenarios[2]
#time_frame = pressure_maps$time_frame[2]
#years = eval(parse(text = pressure_maps$years[2]))

# create and save the model with acceleration and deceleration of biomass loss
  tmp_data <- fish_data %>%
  filter(scenario %in% scenarios) %>%
  filter(year %in% years) %>%
  group_by(rgn_id) %>%
  summarize(pressure = mean(pressure))
  
eez_tmp <- left_join(eez_shp, tmp_data, by="rgn_id")

# Rasterize the biomass data using the raster template
pressure_rast <- terra::rasterize(eez_tmp, ocean_template, field = "pressure", fun = mean, background = NA)
#terra::plot(pressure_rast)

pressure_rast_mask <- pressure_rast*ocean_template
#plot(pressure_rast_mask)
terra::writeRaster(pressure_rast_mask, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure/fisheries_%s_%s.tif", tolower(scenarios), time_frame), overwrite=TRUE)

## end of model 1

# create and save the model with only deceleration of biomass loss
  tmp_data_mod2 <- fish_data %>%
  filter(scenario %in% scenarios) %>%
  filter(year %in% years) %>%
  group_by(rgn_id) %>%
  summarize(pressure = mean(pressure_no_accel))
  
eez_tmp_mod2 <- left_join(eez_shp, tmp_data_mod2, by="rgn_id")

# Rasterize the biomass data using the raster template
pressure_rast_mod2 <- terra::rasterize(eez_tmp_mod2, ocean_template, field = "pressure", fun = mean, background = NA)
#terra::plot(pressure_rast)

pressure_rast_mask_mod2 <- pressure_rast_mod2*ocean_template
#plot(pressure_rast_mask)
terra::writeRaster(pressure_rast_mask_mod2, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2/fisheries_%s_%s.tif", tolower(scenarios), time_frame), overwrite=TRUE)


cat(i, "\n")
}

## loop through iterations
for(i in 1:dim(pressure_maps)[1]){ #i=1

tmp <- pressure_maps[i,]
  
scenarios_tmp = tmp$scenarios
time_frame_tmp = tmp$time_frame
years_tmp = eval(parse(text = tmp$years))  
 
mapping_fish_pressure(scenarios=scenarios_tmp, time_frame=time_frame_tmp, years=years_tmp) 
}

```

Look at some of the results
```{r}

list.files("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure/", full=TRUE)
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure//fisheries_ssp1_medium-term.tif"), main="SSP1, 2040-2060")
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure//fisheries_ssp2_medium-term.tif"), main="SSP2, 2040-2060")
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure//fisheries_ssp3_medium-term.tif"), main="SSP3, 2040-2060")
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure//fisheries_ssp4_medium-term.tif"), main="SSP4, 2040-2060")
v1 <- rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure//fisheries_ssp5_medium-term.tif")

list.files("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2/", full=TRUE)
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2//fisheries_ssp1_medium-term.tif"), main="SSP1, 2040-2060")
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2//fisheries_ssp2_medium-term.tif"), main="SSP2, 2040-2060")
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2//fisheries_ssp3_medium-term.tif"), main="SSP3, 2040-2060")
plot(rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2//fisheries_ssp4_medium-term.tif"), main="SSP4, 2040-2060")
v2 <- rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2//fisheries_ssp5_medium-term.tif")

plot(v1-v2)
