---
title: "Untitled"
output: html_document
date: "2024-03-20"
---

step1: extract biomass data for eezs.
step2: data and models to see if we can predict changes to fisheries pressures
This is mostly an information gathering step that can be skipped. 
step3: prepare governance data. I need two datasets: 
 - average governance from 2000 to 2010, with gapfilling, for estimating slope of change in fishing pressure from 2000 to 2010
 - 2010-2100 governance, with gapfilling

 
```{r setup, include=FALSE}

library(countrycode)
library(tidyverse)
library(here)
library(WDI)
```


## fisheries data
(generated in script: step 1)

```{r}

pressure_eez <- read_csv(here("prep/fisheries/data/region_biomass_trends.csv"))

pressure_eez <- pressure_eez %>%
  filter(ant_typ == "eez"|iso3c=="DIS") %>%
  filter(is.na(uninhabited)) %>%
  filter(!is.na(slope)) %>%
  select(-uninhabited)

```


Take a look at the data.
```{r}
# what governance looks like:
gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  group_by(scenario, year) %>%
  summarize(governance_mean = mean(governance, na.rm=TRUE),
            sd = sd(governance, na.rm=TRUE))
ggplot(gov_raw, aes(x=year, y=governance_mean, group=scenario, color=scenario)) +
  geom_point() +
  geom_line()

## note the weird jump
gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(scenario %in% c("Observed", "SSP5")) 
ggplot(gov_raw, aes(x=year, y=governance, group=country)) +
  geom_point(size=0.2) +
  geom_line(size=0.2)

```

## Get the observed governance data from 2000-2010

Take the average from 2000 to 2010
This will be used to develop the predictive model.

```{r}

gov_gf <- pressure_eez %>%
  select(iso3c, rgn_nam, rgn_id)


gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(scenario == "Observed") %>%
  filter(year >= 2000 & year <= 2010) %>%
  select(iso3c = countrycode, year, governance)

gov_gf <- left_join(gov_gf, gov_raw, by="iso3c")

## first couple years of Tonga missing, fill with closest year
gov_gf <- gov_gf %>%
  mutate(governance = ifelse(iso3c=="TON" & is.na(governance), gov_gf$governance[gov_gf$iso3c=="TON" & gov_gf$year==2003], governance))
summary(gov_gf)
# N = 61 (of 184) countries with no governance, included general disputed regions

## will be working with the avg. governance from 2000 to 2010

gov_gf_avg <- gov_gf %>%
  group_by(iso3c, rgn_id, rgn_nam) %>%
  summarize(governance = mean(governance)) %>%
  ungroup()

```


Gapfill missing governance data: average from 2000 - 2010
This is used to model the relationship between governance and change in fisheries loss from 2000 to 2010. 

I will use pcGDPppp and georegion to predict average governance from 2000-2010.

I attempted to use the SSP gdp and pop projection data (see step2: here("prep/fisheries/data/current_future_gdp_pop.csv")), but there were quite a lot of gaps
in these data, so I am just going to use World Bank data (even though there are no future 
projections).
```{r}

### GDP

#indicator <- "NY.GDP.PCAP.PP.CD"

# Fetch data
#gdp_pc <- WDI(
#  indicator = indicator,
#  start = 2000,  # Adjust the starting year as needed
#  end = 2010,    # Adjust the ending year as needed
#  extra = TRUE   # Include additional country metadata
#)

#write_csv(gdp_pc, here("prep/fisheries/data/inputs/world_bank_pcgdp.csv"))

gdp_pc <- read_csv(here("prep/fisheries/data/inputs/world_bank_pcgdp.csv"))

gdp_pc <- gdp_pc %>%
  select(iso3c, year, pcGDP=NY.GDP.PCAP.PP.CD) %>%
  filter(!is.na(pcGDP)) %>% #most have 11 years of data, but a few don't
  mutate(round(pcGDP,0)) %>%
  group_by(iso3c) %>%
  summarize(pcGDP = mean(pcGDP))

## rough estimate from google searches (Wikipedia and such)
gf_regions <- read_csv(here("prep/fisheries/data/country_pop_gf.csv")) %>%
  filter(population > 0) %>%
  mutate(pop_gf = population/1000000,
         gdp_gf = gdp_current_USD/1000000) %>%
  mutate(pcGDP = round(gdp_gf/pop_gf, 0)) %>%
  select(iso3c, pcGDP)

new <- setdiff(gf_regions$iso3c, gdp_pc$iso3c)


gdp_pc <- rbind(gdp_pc, gf_regions[gf_regions$iso3c %in% new,])

gov_gf_avg2 <- left_join(gov_gf_avg, gdp_pc, by="iso3c")  # N=41 of 61 have gdp data
filter(gov_gf_avg2, is.na(governance)) %>% print(n=100)

## Fill in disputed regions with low governance score
gov_gf_avg2$rgn_nam[gov_gf_avg2$iso3c == "DIS"] <- "Disputed"
gov_gf_avg2$governance[gov_gf_avg2$iso3c == "DIS"] <- quantile(gov_gf_avg2$governance, c(0.1), na.rm=TRUE)


## Georegions used to gapfill missing governance
georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv") %>%
  select(rgn_id, r1_label, r2_label)


gov_gf_avg3 <- left_join(gov_gf_avg2, georgns, by="rgn_id") %>%
  mutate(log_gdp = log(pcGDP))
plot(gov_gf_avg3$governance, log(gov_gf_avg3$pcGDP))
plot(gov_gf_avg3$governance, gov_gf_avg3$pcGDP)
summary(gov_gf_avg3)

### gapfill all regions (except Micronesia) that have gdp data
mod_gov1 <- lm(governance ~ log_gdp + r2_label, na.action=na.exclude, data=gov_gf_avg3)
summary(mod_gov1) # R2= 0.72


gov_gf_avg3$gov_gf1 <- predict(mod_gov1, newdata = gov_gf_avg3 %>%
                          mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

gov_gf_avg4 <- gov_gf_avg3 %>%
  mutate(gapfilled = ifelse(is.na(governance), "gapfilled", NA)) %>%
  mutate(governance = ifelse(is.na(governance), gov_gf1, governance)) # n missing now N= 24


### gapfill the Micronesia region
mod_gov2 <- lm(governance ~ log_gdp + r1_label, data=gov_gf_avg4, na.action=na.exclude)
summary(mod_gov2) # R2 = 0.70

gov_gf_avg4$gov_gf2 <- predict(mod_gov2, newdata = gov_gf_avg4)
gov_gf_avg5 <- gov_gf_avg4 %>%
  mutate(governance = ifelse(is.na(governance), gov_gf2, governance)) # n missing n= 19


### gapfill remaining regions based on georegion, r2 label
mod_gov3 <- lm(governance ~ r2_label, data=gov_gf_avg5, na.action=na.exclude)
summary(mod_gov3) # R2 = 0.63

gov_gf_avg5$gov_gf3 <- predict(mod_gov3, newdata = gov_gf_avg5)
gov_gf_avg6 <- gov_gf_avg5 %>%
  mutate(governance = ifelse(is.na(governance), gov_gf3, governance))

summary(gov_gf_avg6) # all have governance

avg_governance_gf <- gov_gf_avg6 %>%
  select(iso3c, rgn_id, rgn_nam, governance, gapfilled)

write_csv(avg_governance_gf, here("prep/fisheries/data/avg_2000_2010_governance_gf.csv"))

```

## Gapfilling governance 2000-2015 data

In this case, we will use the same general predictive model as before.

Gather the observed governance data

```{r}
regions <- pressure_eez %>%
  select(iso3c, rgn_nam, rgn_id)


gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(scenario == "Observed") %>%
  filter(year >= 2000) %>%
  select(iso3c = countrycode, year, governance)

gov_gf <- left_join(regions, gov_raw, by="iso3c") %>%
  group_by(iso3c, rgn_id, rgn_nam) %>%
  complete(year = 2000:2015) %>%
  arrange(iso3c, year) %>%
  ungroup() %>%
  filter(!is.na(year))



```

Now estimate the 2000 to 2015 years based on GDP and georegion:

```{r}


# Fetch data
# indicator <- "NY.GDP.PCAP.PP.CD"
#gdp_pc <- WDI(
#  indicator = indicator,
#  start = 2000,  # Adjust the starting year as needed
#  end = 2015,    # Adjust the ending year as needed
#  extra = TRUE   # Include additional country metadata
#)

#write_csv(gdp_pc, here("prep/fisheries/data/inputs/world_bank_pcgdp_2000_2015.csv"))

gdp_pc <- read_csv(here("prep/fisheries/data/inputs/world_bank_pcgdp_2000_2015.csv"))

# a few countries don't have all years of gdp data (NA N = 405 to N = 304), gapfill here:
gdp_pc <- gdp_pc %>%
  select(iso3c, year, pcGDP=NY.GDP.PCAP.PP.CD) %>%
  group_by(iso3c) %>%
  arrange(year) %>%
  tidyr::complete(year = 2000:2015) %>%
  tidyr::fill(pcGDP, .direction = "downup") %>%
  ungroup()

gov_gf2 <- left_join(gov_gf, gdp_pc, by=c("iso3c", "year")) # n=640 pcGDp



# Add in the values I looked up from other sources # n = 352 pcGDP
## rough estimate from google searches (Wikipedia and such)
gf_regions <- read_csv(here("prep/fisheries/data/country_pop_gf.csv")) %>%
  filter(population > 0) %>%
  mutate(pop_gf = population/1000000,
         gdp_gf = gdp_current_USD/1000000) %>%
  mutate(pcGDP_est = round(gdp_gf/pop_gf, 0)) %>%
  select(iso3c, pcGDP_est)

gov_gf3 <- left_join(gov_gf2, gf_regions, by="iso3c") %>%
  mutate(pcGDP = ifelse(is.na(pcGDP), pcGDP_est, pcGDP)) %>%
  select(-pcGDP_est)

## Georegions
georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv") %>%
  select(rgn_id, r1_label, r2_label)

## NA = N 1101
gov_gf4 <- left_join(gov_gf3, georgns, by="rgn_id") %>%
  mutate(log_gdp = log(pcGDP))
plot(gov_gf4$governance, log(gov_gf4$pcGDP))
plot(gov_gf4$governance, gov_gf4$pcGDP)

### gapfill all regions (except Micronesia) that have gdp data
# tried including year, but was not significant
mod_gov1 <- lm(governance ~ log_gdp + r2_label, na.action=na.exclude, data=gov_gf4)
summary(mod_gov1) # R2= 0.73

gov_gf4$gov_gf1 <- predict(mod_gov1, 
                          newdata = gov_gf4 %>%
                          mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

# NA gov = 402
gov_gf5 <- gov_gf4 %>%  
  mutate(gapfilled = ifelse(is.na(governance), "gapfilled", NA)) %>%
  mutate(governance = ifelse(is.na(governance), gov_gf1, governance))


### gapfill the Micronesia region
mod_gov2 <- lm(governance ~ log_gdp + r1_label, data=gov_gf5, na.action=na.exclude)
summary(mod_gov2) # R2 = 0.69

gov_gf5$gov_gf2 <- predict(mod_gov2, newdata = gov_gf5)

#NA gov = 322
gov_gf6 <- gov_gf5 %>%  
  mutate(governance = ifelse(is.na(governance), gov_gf2, governance))


### gapfill remaining regions based on georegion
mod_gov3 <- lm(governance ~ r2_label, data=gov_gf6, na.action=na.exclude)
summary(mod_gov3) # R2 = 0.68

gov_gf6$gov_gf3 <- predict(mod_gov3, newdata = gov_gf6)

## NA gov N = 16 (disputed areas)
gov_gf7 <- gov_gf6 %>%  
  mutate(governance = ifelse(is.na(governance), gov_gf3, governance))

## fill in disputed areas with bottom quantile
gov_gf7 <- gov_gf7 %>%
  mutate(governance = ifelse(is.na(governance), quantile(gov_gf7$governance, c(0.1), na.rm=TRUE), governance))

governance_gf_2000_2015 <- gov_gf7 %>%
  mutate(scenario = "Observed") %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled)



```

# Gapfill missing >2015 based on modeled average yearly change

First bias correct the governance data so the observed and scenario 2015 values align. 
```{r}

gov_raw <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv")

observed <- gov_raw %>%
  filter(year==2015 & scenario == "Observed") %>%
  select(iso3c=countrycode, observed_governance=governance)

scenario_correct <- gov_raw %>%
  filter(year==2015 & scenario != "Observed") %>%
  select(iso3c = countrycode, scenario, governance) %>%
  left_join(observed, by = "iso3c") %>%
  mutate(bias_correct = governance - observed_governance) %>%
  select(iso3c, scenario, bias_correct)

bias_corrected_future <- gov_raw %>%
  select(iso3c=countrycode, scenario, year, governance) %>%
  filter(year >= 2015 & scenario != "Observed") %>%
  left_join(scenario_correct, by=c("iso3c", "scenario")) %>%
  mutate(bias_correct = ifelse(is.na(bias_correct), 0, bias_correct)) %>% # COD, ESH, ROU have future data, but not past data
  mutate(governance = governance - bias_correct) %>%
  left_join(regions, by = "iso3c") %>%
  filter(!is.na(rgn_id)) %>%
  mutate(gapfilled = NA) %>%
  mutate(governance = ifelse(governance > 1, 1, governance)) %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled) 

```

Now use the average yearly change across regions for each SSP to estimate the regions that have current estimates (gapfilled) but do not have future data.
```{r}
# first figure out average yearly change for each scenario

yearly_change <- bias_corrected_future %>%
  arrange(iso3c, scenario, year) %>%
  group_by(iso3c, scenario, rgn_id, rgn_nam) %>%
  mutate(governance_change = governance - lag(governance, default = governance[1])) %>%
  ungroup()  

avg_yearly_change <- yearly_change %>%
  group_by(scenario, year) %>%
  filter(governance < 1) %>%
  summarize(avg_change = mean(governance_change)) %>%
  ungroup() %>%
  group_by(scenario) %>%
  mutate(cumulative_avg_change = cumsum(avg_change))

## figure out what is missing in the >2015 data
no_data_ids <- setdiff(governance_gf_2000_2015$iso3c, bias_corrected_future$iso3c)

scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5")
governance_2015 <- filter(governance_gf_2000_2015, iso3c %in% no_data_ids) %>%
  filter(year == 2015) %>%
  select(iso3c, rgn_id, rgn_nam, governance_2015 = governance) %>%
  crossing(scenario = scenario)

gov_future_gf <- full_join(avg_yearly_change, governance_2015, by=c("scenario")) %>%
  rowwise() %>%
  mutate(governance = governance_2015 + cumulative_avg_change) %>%
  mutate(governance = ifelse(governance > 1, 1, governance)) %>%
  mutate(gapfilled = "gapfilled") %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled)

## Now getting the observed data for 2000-2015 (duplicating for each scenario)
scenario <- c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5")
scenario_governance_gf_2000_2015 <- governance_gf_2000_2015 %>%
  select(-scenario) %>%
  expand_grid(scenario) %>%
  select(iso3c, scenario, year, rgn_id, rgn_nam, governance, gapfilled)



future_governance_gf <- rbind(bias_corrected_future, gov_future_gf) %>%
  filter(year != 2015) %>%
  rbind(scenario_governance_gf_2000_2015)

setdiff(pressure_eez$iso3c, future_governance_gf$iso3c)
## should be none! all accounted for!

write_csv(future_governance_gf, here("prep/fisheries/data/governance_2000_2100.csv"))

```


#### Stop here for now

## sampling


```{r}

sim_data <- avg_governance_gf %>%
  filter(ant_typ == "eez") %>%
  select(ant_typ, rgn_id, rgn_nam, iso3c, avg_biomass, avg_percent_change, slope, governance) %>%
  mutate(year = 2005) 

sim_data <- na.omit(sim_data)

sample_low_gov <- sim_data %>%
  filter(governance < 0.4) %>%
  pull(slope)
low_gov_avg <- mean(sample_low_gov)
low_gov_sd <- sd(sample_low_gov)

sample_mid_gov <- sim_data %>%
  filter(governance >= 0.4 & governance < 0.75) %>%
  pull(slope)
mid_gov_avg <- mean(sample_mid_gov)
mid_gov_sd <- sd(sample_mid_gov)

sample_high_gov <- sim_data %>%
  filter(governance >= 0.75) %>%
  pull(slope)
high_gov_avg <- mean(sample_high_gov)
high_gov_sd <- sd(sample_high_gov)

low_gov_avg
mid_gov_avg
high_gov_avg

sim_data_final <- sim_data %>%
  ungroup() %>%
  select(iso3c, biomass_t0 = avg_biomass, pct_change_t0 = avg_percent_change, slope, governance, year)


## prepping and gapfilling governance scenario data

#missing_recent <- data.frame(iso3c = c("ESH", "COD", "ROU"),
#                            governance = c(mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="ESH"]),
#                            mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="COD"]),
#                            mean(gov_scores$governance[gov_scores$year==2016 & gov_scores$iso3c=="ROU"])))
#missing_recent <- crossing(missing_recent, year = 2011:2015, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))
#gov_scores_gf <- rbind(gov_scores, missing_recent)

missing_gov_iso <- setdiff(sim_data_final$iso3c, gov_scores$iso3c)
missing_gov <- expand.grid(iso3c = missing_gov_iso, year = c(2011:2099), scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5"))

missing_gov_estimated <- filter(sim_data_final, iso3c %in% missing_gov_iso) %>%
  select(iso3c, governance)
missing_gov_estimated <- crossing(missing_gov_estimated, year= 2011:2099, scenario = c("SSP1", "SSP2", "SSP3", "SSP4", "SSP5")) %>%
  left_join(gov_ssp_change, by=c("scenario", "year")) %>%
  mutate(governance_yearly_change = ifelse(is.na(governance_yearly_change), 0, governance_yearly_change)) %>%
  mutate(governance2 = governance + governance_yearly_change) %>%
  select(iso3c, year, governance=governance2, scenario)
ggplot(filter(missing_gov_estimated, iso3c=="ABW"), aes(x=year, y=governance, group=scenario)) +
  geom_line()

gov_scores_gf <- rbind(gov_scores, missing_gov_estimated)
setdiff(gov_scores_gf$iso3c, sim_data_final$iso3c) # mostly landlocked
setdiff(sim_data_final$iso3c, gov_scores_gf$iso3c)
## now prepare sample data

########################################
## function to run randomization:
########################################
randomized_fishing <- function(scenario_selects="SSP1", sim_data_finals = sim_data_final, gov_scores_gfs=gov_scores_gf){

  for (year in 2005:2099){ #year = 2010 scenario_selects="SSP1"

scenario_select <- scenario_selects

if(year==2005){
t0 <- sim_data_finals %>%
  select(-slope) %>%
  mutate(scenario = scenario_select) %>%
  mutate(!!paste0("biomass_", year) := biomass_t0) %>%
  mutate(!!paste0("pct_change_", year) := pct_change_t0) 

} else{
  t0 <- simulation
}

  if(year %in% c(2005:2010)){
simulation <- t0 %>%
  rowwise() %>%
  mutate(
    change_pct_change_random = case_when(
      governance < 0.4 ~ rnorm(1, mean=low_gov_avg, sd=low_gov_sd),
      governance >= 0.4 & governance < 0.75 ~ rnorm(1, mean=mid_gov_avg, sd=mid_gov_sd),
      governance >= 0.75 ~ rnorm(1, mean=high_gov_avg, sd = high_gov_sd))) %>%
  mutate(pct_change_t1 = pct_change_t0 + change_pct_change_random) %>%
  mutate(!!paste0("pct_change_", year+1) := pct_change_t1) %>%
  mutate(!!paste0("change_pct_change", year+1) := change_pct_change_random) %>%
  mutate(biomass_t1 = biomass_t0 * (1 + pct_change_t1)) %>%
  mutate(!!paste0("biomass_", year+1) := biomass_t1) %>%
  mutate(year=year+1) %>%
  select(-biomass_t0, -pct_change_t0) %>%
  rename(biomass_t0 = biomass_t1, pct_change_t0=pct_change_t1) 

#cat(year, "\n", "dim = ", dim(simulation)[1], "\n") # for troubleshooting
  } else {
    
simulation <- t0 %>%
  select(-governance) %>%
  left_join(gov_scores_gfs, by=c("iso3c", "year", "scenario")) %>%
  rowwise() %>%
  mutate(
    change_pct_change_random = case_when(
      governance < 0.4 ~ rnorm(1, mean=low_gov_avg, sd=low_gov_sd),
      governance >= 0.4 & governance < 0.75 ~ rnorm(1, mean=mid_gov_avg, sd=mid_gov_sd),
      governance >= 0.75 ~ rnorm(1, mean=high_gov_avg, sd = high_gov_sd))) %>%
  mutate(pct_change_t1 = pct_change_t0 + change_pct_change_random) %>%
  mutate(!!paste0("pct_change_", year+1) := pct_change_t1) %>%
  mutate(!!paste0("change_pct_change", year+1) := change_pct_change_random) %>%
  mutate(biomass_t1 = biomass_t0 * (1 + pct_change_t1)) %>%
  mutate(!!paste0("biomass_", year+1) := biomass_t1) %>%
  mutate(year=year+1) %>%
  select(-biomass_t0, -pct_change_t0) %>%
  rename(biomass_t0 = biomass_t1, pct_change_t0=pct_change_t1) 
#cat(year, "\n", "dim = ", dim(simulation)[1], "\n") # for troubleshooting

  }
  }
  return(simulation)
}

########################################
## end of function to run randomization
########################################

summary(simulation_long)
ggplot(filter(simulation_long, year <=2060), aes(x=year, y=biomass, group=iso3c, col=governance)) +
  geom_line()

simulation_long %>%
  group_by(year) %>%
  summarize(sum(biomass, na.rm=TRUE)) %>% print(n=100)




## Run simulations

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP1", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)

scenario_select = "SSP1"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}


for(i in 1:100){
  
tmp <- randomized_fishing(scenario_selects="SSP2", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)

scenario_select = "SSP2"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP3", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select="SSP3"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP4", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select <- "SSP4"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

for(i in 1:100){
tmp <- randomized_fishing(scenario_selects="SSP5", sim_data_final = sim_data_final, gov_scores_gf=gov_scores_gf)
scenario_select <- "SSP5"
simulation_long <- tmp %>%
  select(iso3c, governance, scenario, starts_with("biomass")) %>%
  pivot_longer(
    cols = starts_with("biomass"),
    names_to = "year",
    names_prefix = "biomass_",
    values_to = "biomass"
  ) %>%
  filter(year != "t0")%>%
  mutate(simulation = i)
write_csv(simulation_long, sprintf("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_%s_s%s.csv", scenario_select, i))
}

```


```{r}
tmp <- read_csv("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/randomized_biomass_SSP1_s1.csv")

files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/", full=TRUE, pattern=".csv")

rand_biomass <- lapply(files, read.csv) %>% bind_rows()

## compare to global observed: 2005:2010
## ensure estimates are not going off the rails.
compare <- rand_biomass %>%
  group_by(simulation, scenario, year) %>%
  summarize(global_biomass = sum(biomass)) %>%
  filter(year %in% 2005:2010)
compare %>%
  group_by(year) %>%
  summarize(simulated_global_biomass = sum(global_biomass))

check <- country_biomass_change
check$iso3c <- countrycode(check$rgn_nam, "country.name", "iso3c")
check <- filter(check, iso3c %in% unique(tmp$iso3c))
observed <- check %>%
  group_by(year) %>%
  summarize(observed_biomass = sum(biomass, na.rm=TRUE) - 3865822)

ggplot() + 
  geom_line(data=compare, aes(x=year, y=global_biomass, group=simulation), alpha=0.2) +
  geom_point(data=filter(observed, year>=2005), aes(x=year, y=observed_biomass), col="red")

## note that the 2005 biomass estimate is lower in simulated data data due to averaging 2000-2010 
## to arrive at this value. (consider changing)
692322460-688456638

US_means <- rand_biomass %>%
  filter(iso3c=="USA" & scenario=="SSP1" & year %in% c(2010:2050)) %>%
  group_by(year) %>%
    summarize(mean_biomass = mean(biomass))

ggplot(data=filter(rand_biomass, iso3c=="USA" & scenario=="SSP1" & year%in%c(2010:2050)), aes(x=year, y=biomass)) +
  geom_point() +
  geom_point(data=US_means, aes(x=year, y=mean_biomass), col="red") +
    geom_line(data=US_means, aes(x=year, y=mean_biomass), col="red")
  


## constrain data, to not go above or below a particular percentage.
## get reference. biomass:
ref_times <- which(as.Date(time(biomass)) >= as.Date("1900-01-01") & as.Date(time(biomass)) < as.Date("1910-01-01")) 
ref_density_rasters <- subset(biomass, ref_times)
ref_biomass_rasters <- cellSize(ref_density_rasters) * ref_density_rasters * 1e-6
ref_biomass <- mean(ref_biomass_rasters)

ref_country_biomass <- terra::zonal(ref_biomass, terra::rast(zones_raster), fun = sum, na.rm=TRUE)

ref_country_biomass <- ref_country_biomass %>% 
 rename(rgn_id = layer, ref_biomass=mean) %>%
  left_join(rgn_data, by="rgn_id") %>%
  filter(!is.na(ref_biomass)) %>%
  filter(ant_typ == "eez")

ref_country_biomass$iso3c <- countrycode(ref_country_biomass$rgn_nam, "country.name", "iso3c") %>%
  filter(!is.na(iso3c))

## Get time 0 biomass
t0_times <- which(as.Date(time(biomass)) >= as.Date("1905-01-01") & as.Date(time(biomass)) < as.Date("1906-01-01")) 
t0_density_rasters <- subset(biomass, t0_times)
t0_biomass_rasters <- cellSize(biomass_c) * biomass_c * 1e-6
t0_biomass <- mean(ref_biomass_rasters)

ref_country_biomass <- terra::zonal(ref_biomass, terra::rast(zones_raster), fun = sum, na.rm=TRUE)

ref_country_biomass <- ref_country_biomass %>% 
 rename(rgn_id = layer, ref_biomass=mean) %>%
  left_join(rgn_data, by="rgn_id") %>%
  filter(!is.na(ref_biomass)) %>%
  filter(ant_typ == "eez")

ref_country_biomass$iso3c <- countrycode(ref_country_biomass$rgn_nam, "country.name", "iso3c") %>%
  filter(!is.na(iso3c))


```