---
title: "Calculating impacts"
output: html_document
date: "2024-09-16"
---

```{r setup, include=FALSE}
library(tidyverse)
library(terra)
library(here)
library(countrycode)

```

```{r}
vulnerability <- read_csv(here("impact/vulnerability_weighting_matrix.csv")) %>%
  mutate_if(is.numeric, ~ . / 4)

habitat_files <- list.files(here("habitats/data"), full=TRUE)

setdiff(names(vulnerability), gsub(".tif", "", basename(habitat_files)))
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

pressure_files <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures", full=TRUE)
setdiff(vulnerability$pressure, gsub(".tif", "", sub("_.*", "", basename(pressure_files))))
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

```

# combos
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
file.remove(clean)

for(habitat in names(vulnerability)[-1]){
  #habitat = names(vulnerability)[2]
  hab_file <- grep(habitat, habitat_files, value=TRUE)
  hab_raster <- rast(hab_file)
  cat(habitat, "\n")
  
  for(pressure in vulnerability$pressure){
  # pressure = vulnerability$pressure[7]
  # pressure = "sst-extremes"
  pressure_filter <- paste0(pressure, "_")
  p_files <- grep(pressure_filter, pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(pressure, " n = ", length(p_files), "\n")
  
  vulnerability_value <- vulnerability %>%
    filter(pressure == !!pressure) %>%
    select(habitat) %>%
    pull()
  
  pressure_vulnerability <- pressure_rasts * vulnerability_value
  
  pressure_vulnerability_habitat <- pressure_vulnerability * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/hab_vuln_pressure/"
  p_names <- basename(gsub(paste0(pressure, "_"), "", p_files))
  output_filenames <- sprintf("%s%s_%s_%s", save_loc, habitat, pressure, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}
}

```

Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
plot(rast(check[652]))

length(check)
 #should be 4,540
habitats <- sub("_.*", "", basename(check))
table(habitats)
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", basename(check))
table(pressure)

```

## adjust air temperature pressure
The pressure is a little different for salt-marsh, seagrass, and ice
```{r}

combo_files <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)

heatindex_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "salt-marsh"
#lower_threshold <- 0.2
#upper_threshold <- 0.25
combo_files <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)  
  
hi_files <- grep(sprintf("%s_air-heat-index", habitat), combo_files, value=TRUE)
file.remove(hi_files)
hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)
  
  p_files <- grep("air-heat-index", pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(" n = ", length(p_files), "\n")
  
  p <- pressure_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/hab_vuln_pressure/"
  p_names <- basename(p_files)
  output_filenames <- sprintf("%s%s_%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

heatindex_modify(habitat="salt-marsh", lower_threshold=0.2, upper_threshold=0.25) 
heatindex_modify(habitat="seagrass", lower_threshold=0.2, upper_threshold=0.25) 
#heatindex_modify(habitat="ice", lower_threshold=0, upper_threshold=0) 
```


## adjust SST pressure
The pressure is a little different for kelp and coral because there is more information about them.
First prepare the raw sst data. These impacts are based on absolute temperatures.
```{r}

# get the temperature data

sst_raw_prep <- function(scenario){
#scenario <- "ssp245"
sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average_final/", full=TRUE)
scenario_sst_raw_files <- grep(scenario, sst_raw_files, value=TRUE)

current_files <- grep("2015_2019", scenario_sst_raw_files, value=TRUE)
current_stack <- rast(current_files)
current_mean <- mean(current_stack) - 273.15

near_files <- grep("2020_2024|2025_2029|2030_2034|2035_2039", scenario_sst_raw_files, value=TRUE)
near_stack <- rast(near_files)
near_mean <- mean(near_stack) - 273.15

medium_files <- grep("2040_2044|2045_2049|2050_2054|2055_2059", scenario_sst_raw_files, value=TRUE)
medium_stack <- rast(medium_files)
medium_mean <- mean(medium_stack) - 273.15

long_files <- grep("2080_2084|2085_2089|2090_2094|2095_2099", scenario_sst_raw_files, value=TRUE)
long_stack <- rast(long_files)
long_mean <- mean(long_stack) - 273.15

writeRaster(current_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_current.tif", scenario), overwrite=TRUE)
writeRaster(near_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_near-term.tif", scenario), overwrite=TRUE)
writeRaster(medium_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_medium-term.tif", scenario), overwrite=TRUE)
writeRaster(long_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_long-term.tif", scenario), overwrite=TRUE)
}

sst_raw_prep(scenario="ssp245")
sst_raw_prep(scenario="ssp585")

```


Now apply to coral, kelp, and mangroves.
```{r}

sst_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "kelp"
#lower_threshold <- 18.3
#upper_threshold <- 24
combo_files <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)  
sst_files <- grep(sprintf("%s_sst-extremes", habitat), combo_files, value=TRUE)
file.remove(sst_files)

hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)

sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst", full=TRUE)    
sst_raw_rasts <- rast(sst_raw_files)
cat(" n = ", length(sst_raw_files), "\n")
  
  p <- sst_raw_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/hab_vuln_pressure/"
  p_names <- gsub("sst-raw", "", basename(sst_raw_files))
  output_filenames <- sprintf("%s%s_sst-extremes%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

sst_modify(habitat="kelp", lower_threshold=18.3, upper_threshold=24) 
sst_modify(habitat="coral-reef", lower_threshold=29, upper_threshold=30)  
sst_modify(habitat="mangroves", lower_threshold=32, upper_threshold=40)  
```

## Make ssp names consistent
Some pressures do not use RCP climate models. The difference in naming (ssp2 vs. ssp245) is throwing everything off!

```{r}

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.copy(from = files_ssp1, to = gsub("ssp1_", "ssp119_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.rename(from = files_ssp1, to = gsub("ssp1_", "ssp126_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp2 <- grep("ssp2_", files_fix, value=TRUE)
file.rename(from = files_ssp2, to = gsub("ssp2_", "ssp245_", files_ssp2))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp3 <- grep("ssp3_", files_fix, value=TRUE)
file.rename(from = files_ssp3, to = gsub("ssp3_", "ssp370_", files_ssp3))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp4 <- grep("ssp4_", files_fix, value=TRUE)
file.rename(from = files_ssp4, to = gsub("ssp4_", "ssp460_", files_ssp4))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp5 <- grep("ssp5_", files_fix, value=TRUE)
file.rename(from = files_ssp5, to = gsub("ssp5_", "ssp585_", files_ssp5))
  
```



Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/")
length(check)
 #should be 4,760
habitats <- sub("_.*", "", check)
table(habitats)
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", check)
table(pressure)

```


Create raster with total number of habitats in each cell.
```{r}
habitat_files <- list.files(here("habitats/data"), full=TRUE)
habitat_rast <- rast(habitat_files)
habitat_n <- sum(habitat_rast, na.rm=TRUE)
check <- ifel(habitat_n == 0, 1, NA)
plot(check, col="red")
global(check, "sum", na.rm=TRUE)
check <- ifel(habitat_n < 1, 1, NA)
plot(check, col="red")
habitat_n = ifel(habitat_n <1, 1, habitat_n)
writeRaster(habitat_n, here("habitats/habitat_n.tif"), overwrite=TRUE)

```

# Calculate cumulative impacts
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
file.remove(clean)

pressures <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)

cum_impact <- function(scen_time){
#scen_time <- "ssp585_medium-term"
habitat_n <- rast(here("habitats/habitat_n.tif"))

pressures_filter <- grep(scen_time, pressures, value=TRUE)
cat(paste0(scen_time, ", n = ", length(pressures_filter), "\n"))
pressure_stack <- rast(pressures_filter)
sum_pressure <- sum(pressure_stack, na.rm=TRUE)
impact <- sum_pressure/habitat_n
writeRaster(impact, sprintf("/home/shares/ohi/stressors_2021/impacts/cumulative/%s.tif", scen_time), overwrite=TRUE)
}
cum_impact(scen_time = "ssp585_medium-term")
cum_impact(scen_time = "ssp585_near-term")
cum_impact(scen_time = "ssp585_current")
cum_impact(scen_time = "ssp245_medium-term")
cum_impact(scen_time = "ssp245_near-term")
cum_impact(scen_time = "ssp245_current")

```

# For individual pressures, need to create rasters that correct for number of habitats
For regional summaries of pressure impacdts, the impact of the pressure will depend on the number of habitats. 


First step is to correct for the number of habitats (being sure to divide by the pressure). This is used for geographic summaries of pressures (but not habitat based summaries.).

```{r}

check <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure_habitat_corrected/", full=TRUE)
file.remove(check)

pressures <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

## have to correct the pressures by the number of habitats
habitat_n <- rast(here("habitats/habitat_n.tif"))

for(i in 1: nlyr(hab_impacts_stack)){ #i=1
current_layer <- hab_impacts_stack[[i]]

  # Perform the division
  corrected_layer <- current_layer / habitat_n

  # Define the output filename
  layer_name <- names(hab_impacts_stack)[i]

  # Write the output to disk
  writeRaster(corrected_layer, 
              filename = sprintf("/home/shares/ohi/stressors_2021/hab_vuln_pressure_habitat_corrected/%s.tif", layer_name), 
              overwrite = TRUE)

  # Optional: Print progress
  cat("Processed and saved:", layer_name, "\n")
}

```


# Statistic: Describe projected increase in impact over time.
```{r}
## Change in global pressure
## global
tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

global(stack, "sum", na.rm=TRUE)
2029599.1/890958.3
2385282.7/886167.4

```


### Figure 1
Maps of cumulative impact and change in cumulative impact
```{r}

library(paletteer)
library(png)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

```

Get land boundaries

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
# Define the Mollweide projection string
moll_crs <- "+proj=moll"

# Reproject the world data to Mollweide
world_moll <- st_transform(world, crs = moll_crs)
```

Get current and future cumulative impacts
```{r}
upperscale <- 1.5
cols_palette <- paletteer_c("grDevices::Lajolla", 200)
#cols_palette <- rev(paletteer_c("ggthemes::Classic Orange-Blue", 250)[1:150])
#cols_palette <- paletteer_c("ggthemes::Temperature Diverging", 100)

tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

ssp2_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")
quantile(data.frame(ssp2_med)$sum, c(0.9999))


stack_capped <- ifel(stack>upperscale, upperscale, stack)

for(i in 1:dim(stack_capped)[3]){
#i=5
zlim_values <- c(0, upperscale)
raster_data <- stack_capped[[i]]
df <- as.data.frame(raster_data, xy = TRUE, na.rm = TRUE)
colnames(df) <- c("x", "y", "value")

plot_cum <- ggplot(df, aes(x = x, y = y, fill = value)) +
  geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
  geom_sf(
    data = world_moll,
    fill = "gray80",       # Land fill color
    color = "gray85",      # Slightly lighter color for boundaries
    size = 0.2,            # Adjust line thickness
    inherit.aes = FALSE
  ) +
#  ggtitle(gsub(".tif", "", basename(tmp[i]))) +
geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
    theme_minimal() +
theme(
  panel.grid = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank(),
  axis.title = element_blank(),
  legend.position = "right",
  plot.title = element_text(hjust = 0.5)
)


ggsave(plot_cum, filename=here(sprintf("impact/figs/cumulative/%s", gsub(".tif", ".png", basename(tmp[i])))), width=5.5, height = 2.5, units = c("in"))
}

```

Make global plot of change in impact figures
```{r}

cols_palette <- c(paletteer_c("grDevices::BluGrn", 30)[21:30],
                  rev(paletteer_c("grDevices::Inferno", 110)))


diff_impact <- function(raster, lowerscale = -0.2, upperscale = 1.2, saveName){
#raster=ssp2_med
# fileName = "test"
  raster_data = raster
raster_data <- ifel(raster_data>upperscale, upperscale, raster_data)
raster_data <- ifel(raster_data<lowerscale, lowerscale, raster_data)
zlim_values <- c(lowerscale, upperscale)
df <- as.data.frame(raster_data, xy = TRUE, na.rm = TRUE)
colnames(df) <- c("x", "y", "value")

plot_cum <- ggplot(df, aes(x = x, y = y, fill = value)) +
  geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
  geom_sf(
    data = world_moll,
    fill = "gray80",       # Land fill color
    color = "gray85",      # Slightly lighter color for boundaries
    size = 0.2,            # Adjust line thickness
    inherit.aes = FALSE
  ) + 
   geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
#  ggtitle(gsub(".tif", "", basename(tmp[i]))) +
  theme_minimal() +
theme(
  panel.grid = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank(),
  axis.title = element_blank(),
  legend.position = "right",
  plot.title = element_text(hjust = 0.5)
)

ggsave(plot_cum, filename=here(sprintf("impact/figs/cumulative/%s", saveName)), width=4.5, height = 2.5, units = c("in"))
}


list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

ssp2_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif") 

global(ssp2_med, "min", na.rm=TRUE)
global(ssp2_med, "min", na.rm=TRUE)
plot(ssp2_med)
quantile(data.frame(ssp2_med)$sum, c(0.9999))
quantile(data.frame(ssp2_med)$sum, c(1-0.9999))
ssp5_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_current.tif") 

diff_impact(raster=ssp2_med, lowerscale = -0.1, upperscale = 1.25, saveName="sspp245_change.png")
diff_impact(raster=ssp5_med, lowerscale = -0.1, upperscale = 1.25, saveName="sspp585_change.png")
```

Global plot of hotspots
```{r}

ssp2_change <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif") -
               rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif")
ssp2_change_high_quant <-global(ssp2_change, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp2_change_high <- ifel(ssp2_change >= ssp2_change_high_quant, 1, 0)
ssp2_change_low_quant <-global(ssp2_change, fun = quantile, probs = 0.20, na.rm = TRUE) %>% pull()
ssp2_change_low <- ifel(ssp2_change <= ssp2_change_low_quant, 1, 0)

ssp2_current <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif")
ssp2_current_high_quant <-global(ssp2_current, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp2_current_high <- ifel(ssp2_current >= ssp2_current_high_quant, 1, 0)
ssp2_current_low_quant <-global(ssp2_current, fun = quantile, probs = 0.20, na.rm = TRUE) %>% pull()
ssp2_current_low <- ifel(ssp2_current <= ssp2_current_low_quant, 1, 0)

# category 1: current low and stable
cat1 <- ssp2_current_low + ssp2_change_low
cat1 <- ifel(cat1 == 2, 3, NA)
plot(cat1, col="#44ACC8")

# category 2: current high and stable
cat2 <- ssp2_current_high + ssp2_change_low
cat2 <- ifel(cat2 == 2, 4, NA)
plot(cat2, col="#E2D100", add=TRUE)

# category 3: current high and fast increase
cat3 <- ssp2_current_high + ssp2_change_high
cat3 <- ifel(cat3 == 2, 3, NA)
plot(cat3, col="#E58501", add=TRUE)

# category 4: current low and fast increase
cat4 <- ssp2_current_low + ssp2_change_high
cat4 <- ifel(cat4 == 2, 4, NA)
plot(cat4, col="#B41821", add=TRUE)

png(here("impact/figs/hotspots.png"), width = 2000, height = 1500, res = 300)

plot(st_geometry(world_moll), col="gray80", border="gray85")
plot(cat1, col="#44ACC8", add=TRUE, legend=FALSE)
plot(cat2, col="#E2D100", add=TRUE, legend=FALSE)
plot(cat3, col="#E58501", add=TRUE, legend=FALSE)
plot(cat4, col="#B41821", add=TRUE, legend=FALSE)

dev.off()

```

## Figure 2: coastal vs ocean

Some stats for the paper
```{r}


# relative pressures coastal vs global

## global coast
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(coast>0, 1, 0)

## offshore area only
ocean <- rast(here("_spatial/ocean_area_mol.tif"))
ocean <- ifel(ocean>0, 2, 0)
offshore <- ocean - ifel(is.na(coast), 0, coast)
offshore <- ifel(offshore ==2, 1, NA)


cumulative <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info

offshore_pressures <- cum_stack*offshore
total_offshore_pressures <- global(offshore_pressures, "sum", na.rm=TRUE)

offshore_cells_num <- global(offshore, "sum", na.rm=TRUE) %>% pull()

offshore_impact <- as.data.frame(total_offshore_pressures) %>%
  mutate(cells = offshore_cells_num) %>%
  mutate(average = sum/cells)

# figure out for coastal
coastal_pressures <- cum_stack*coast
total_coastal_pressures <- global(coastal_pressures, "sum", na.rm=TRUE)

coastal_cells_num <- global(coast, "sum", na.rm=TRUE) %>% pull()

coastal_impact <- as.data.frame(total_coastal_pressures) %>%
  mutate(cells = coastal_cells_num) %>%
  mutate(average = sum/cells)

```


Cumulative impact plot

```{r}

cumulative <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info

offshore_pressures <- cum_stack*offshore

offshore_points <- data.frame(offshore_pressures)
offshore_points$region <- "offshore"

coastal_pressures <- cum_stack*coast

coast_points <- data.frame(coastal_pressures)
coast_points$region <- "coastal"

point_data <- rbind(coast_points, offshore_points[sample(nrow(offshore_points), dim(coast_points)[1]), ])

colnames(point_data) <- gsub("\\.", "-", colnames(point_data))


point_data_long <- point_data %>%
  pivot_longer(
    cols = starts_with("ssp"),       # Columns to reshape
    names_to = c("scenario", "time_period"),  # New column names
    names_sep = "_",                 # Separator used in column names
    values_to = "value"              # Name of the values column
  )

point_data_plot <- filter(point_data_long, scenario == "ssp245", time_period %in% c("current", "medium-term"))

# Calculate the overall range of 'value' to set consistent axes
range(point_data_long$value, na.rm = TRUE)



cum <- ggplot(point_data_plot, aes(y = region, x = value, fill = region)) +
  geom_violin(trim = FALSE) +
  xlim(0, 2) + 
  facet_wrap(~ time_period, ncol = 1) +
  scale_fill_manual(values = c("offshore" = "#73A088", "coastal" = "#F98400")) +
  theme_minimal() +
  labs(
    title = "",
    x = "Cumulative impact",
    y = "",
    fill = ""
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.position = "none"
  )

ggsave(cum, filename=here("impact/figs/coastal_vs_offshore_cumulative_impact.png"), width=5.5, height =2, units = c("in"))

```

Now get pressure impacts data together by summarizing pressure impacts across habitats.

```{r}
# 1. summarize pressure rasters across habitats
summarize_pressures <- function(scenario ="ssp245", time = "medium-term"){
  # scenario = "ssp245"
  # time = "medium-term"
  
files <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure_habitat_corrected/", full=TRUE)

pressure_list <- data.frame(pressures=c("sst-extremes|temperature", "slr", "_oxygen-", "_oa_", "nutrient", "npp", "light-pollution", "fisheries", "coastal-disturbance", "air-heat-index"),
                            pressure_name = c("temperature", "slr", "oxygen", "oa", "nutrient", "npp", "light-pollution",
                                              "fisheries", "coastal-disturbance", "air-heat-index"))

  for(pressure in pressure_list$pressures){
    #pressure = pressure_list$pressures[2]
    pressure_files <- grep(scenario, files, value=TRUE)
    pressure_files <- grep(time, pressure_files, value=TRUE)
    pressure_files <- grep(pressure, pressure_files, value=TRUE)
    cat(pressure, "n = ", length(pressure_files), "\n")
    
    pressure_impact_rasts <- rast(pressure_files)
    pressure_impact_sum <- sum(pressure_impact_rasts, na.rm=TRUE)
   
    name_raster <- paste0(pressure_list$pressure_name[pressure_list$pressures==pressure],"_",
                          scenario, "_", time, ".tif")
    writeRaster(pressure_impact_sum, 
                paste0("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", name_raster),
                overwrite=TRUE) 
  }
}

summarize_pressures(scenario ="ssp245", time = "medium-term")
summarize_pressures(scenario ="ssp245", time = "current")

```

Now get the data for each pressure and plot by region
```{r}

list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/")

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", full=TRUE)
pressure_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(pressure_stack)))
names(pressure_stack) <- source_info

offshore2 <- ifel(offshore>=1, 2, 0)
offshore2 <- ifel(is.na(offshore2), 0, offshore2)
coast2 <- ifel(is.na(coast), 0, coast)
habitats <- offshore2 + coast2
habitats <- ifel(habitats == 0, NA, habitats)

coast_area <- global(coast, "sum", na.rm=TRUE) %>% pull()
offshore_area <- global(offshore, "sum", na.rm=TRUE) %>% pull()

area_df <- data.frame(region = c("coastal", "offshore"),
                      area = c(coast_area, offshore_area))

  # Compute the average pressure impact for each region
  zonal_sum <- zonal(pressure_stack, habitats, fun = "sum", na.rm=TRUE)


avg_pressure_impacts <- zonal_sum %>%
 pivot_longer(cols=-ocean_area_mol, names_to = "pressure_impact", values_to="value") %>%
separate(
    col = pressure_impact,
    into = c("pressure", "scenario", "time_period"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(region = ifelse(ocean_area_mol == 1, "coastal", "offshore")) %>%
  left_join(area_df, by="region") %>%
  mutate(pressure_impact = value/area) %>%
  select(pressure, scenario, time_period, region, pressure_impact)

# Calculate average impact for each pressure
average_impacts <- avg_pressure_impacts %>%
  group_by(pressure) %>%
  summarize(avg_impact = mean(pressure_impact, na.rm = TRUE)) %>%
  arrange(avg_impact)


# Prepare data
df <- avg_pressure_impacts %>%
  mutate(
    pressure = factor(pressure, levels = average_impacts$pressure),
    region = factor(region), 
    time_period = factor(time_period, levels = c("current", "medium-term"))
  )

# Reshape data to wide format
df_wide <- df %>%
  pivot_wider(
    id_cols = c(pressure, region),
    names_from = time_period,
    values_from = pressure_impact
  ) %>%
  arrange(pressure)

# Create combined factor and adjust positions
df_wide <- df_wide %>%
  mutate(
    pressure_position = as.numeric(factor(pressure, levels = levels(df$pressure))),
    y_position = pressure_position + ifelse(region == "coastal", -0.1, 0.1),
    delta = `medium-term` - current,
    offset = delta * 0.05,  # Adjust this fraction as needed
    x_start = current + sign(delta) * offset,
    x_end = `medium-term` - sign(delta) * offset
  )


# Create the plot
coast_compare <- ggplot() +
  geom_segment(
  data = df_wide,
  aes(
    x = x_start,
    xend = x_end,
    y = y_position,
    yend = y_position,
    color = region
  ),
  arrow = arrow(length = unit(0.2, "cm")),
  size = 0.5,
  alpha=0.5
)  +
  geom_point(
    data = df_wide,
    aes(x = current, y = y_position, shape = "Current", color = region),
    size = 2.5
  ) +
  geom_point(
    data = df_wide,
    aes(x = `medium-term`, y = y_position, shape = "Medium-Term", color = region),
    size = 2.5
  ) +
  scale_y_continuous(
    breaks = unique(df_wide$pressure_position),
    labels = levels(df$pressure)
  ) +
  scale_shape_manual(
    name = "Time period",
    values = c("Current" = 16, "Medium-Term" = 17)
  ) +
  scale_color_manual(
    name = "Region",
    values = c("offshore" = "#73A088", "coastal" = "#F98400")
  ) +
  theme_minimal() +
  labs(
    title = "",
    x = "Pressure impact",
    y = ""
  ) +
theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black")
  )
ggsave(coast_compare, filename ="impact/figs/coastal_vs_offshore_pressure_impact.png", width=6, height =3, units = c("in"))
```


## cumulative impact: country EEZ summaries
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))
## cut land
ocean_cells <- rast(here("_spatial/ocean_area_mol.tif"))
ocean_cells <- ifel(ocean_cells>0, 1, 0)
## add in the coastal area
eez_coast <- rast(here("_spatial/coastal_country_mol.tif"))
eez_coast <- ifel(eez_coast>0, 1, 0)
## combine coast and ocean
ocean_area <- app(c(ocean_cells, eez_coast), "sum", na.rm=TRUE)
ocean_area <- ifel(ocean_area>0, 1, NA)
plot(ocean_area)

#get the area
eez <- eez*ocean_area
plot(eez)

eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region

impact_files <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

impact_stack <- rast(impact_files)
source_info <- gsub(".tif", "", basename(sources(impact_stack)))
names(impact_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(impact_stack)) { # i = 1
  # Extract the current layer
  current_layer <- impact_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, eez, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
eez_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

eez_impact_result <- eez_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")

eez_impact_result <- eez_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
 eez_cell_num <- zonal(ifel(eez>0, 1, 0), eez, fun = "sum", na.rm=TRUE)
 names(eez_cell_num) <- c("rgn_id", "cells")

dim(eez_impact_result)
eez_impact_result <- eez_impact_result %>%
  left_join(eez_cell_num, by="rgn_id") %>%
  mutate(region = "eez") 


write_csv(eez_impact_result, here("impact/data/country_eez_impact.csv"))

```


## cumulative impact: country coastal region and  summaries
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(eez_coast>0, 1, 0)
#get the area
country_coast <- eez*coast
plot(country_coast)


eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region

impact_files <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

impact_stack <- rast(impact_files)
source_info <- gsub(".tif", "", basename(sources(impact_stack)))
names(impact_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(impact_stack)) { # i = 1
  # Extract the current layer
  current_layer <- impact_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, country_coast, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
coast_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

coast_impact_result <- coast_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")


coast_impact_result <- coast_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

dim(coast_impact_result)
coast_impact_result <- coast_impact_result %>%
  left_join(coast_cell_num, by="rgn_id") %>%
  mutate(region = "coast") 


write_csv(coast_impact_result, here("impact/data/country_coast_impact.csv"))

```

plot dependence vs. impacts
```{r}
#from paper_statistics_v2
depend <- read_csv(here("combining_pressures/data/selig_marine_dependence.csv"))

coast <- read_csv(here("impact/data/country_coast_impact.csv"))

impact_country <- coast %>%
  mutate(avg_value = value/cells) 
#%>%
#  separate(impact, into = c("scenario", "time_period"), sep = "_")

impact_country <- impact_country %>%
  select(region, rgn_label, iso3c, impact, avg_value) %>%
    pivot_wider(
    names_from = impact,
    values_from = avg_value
  )

impact_country$iso3c <- countrycode(impact_country$rgn_label, "country.name", "iso3c") 
impact_country <- filter(impact_country, rgn_label != "Bassas da India")


double_rgns <- data.frame(rgn_label = c("Guadeloupe and Martinique", "Guadeloupe and Martinique",
              "Northern Mariana Islands and Guam", "Northern Mariana Islands and Guam",
              "Puerto Rico and Virgin Islands of the United States", "Puerto Rico and Virgin Islands of the United States"),
                 iso3c = c("GLP", "MTQ", "GUM", "MNP", "PRI", "VIR"))

coastal_double_eez <- filter(impact_country, rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")) %>%
  select(-iso3c)

coastal_double_eez2 <- left_join(double_rgns, coastal_double_eez, by="rgn_label")

impact_country <- filter(impact_country, !(rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")))

coastal_eez_impact <- rbind(impact_country, coastal_double_eez2)
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label=="Northern Saint-Martin"] <- "NSM"
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label=="Micronesia"] <- "FSM"
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label %in% c("Crozet Islands",
                                       "Kerguelen Islands",
                                       "Amsterdam Island and Saint Paul Island",
                                       "Ile Europa",
                                       "Ile Tromelin",
                                       "Juan de Nova Island",
                                       "Glorioso Islands")] <- "ATF"
coastal_eez_impact <- filter(coastal_eez_impact, iso3c != "ATF")

coastal_eez_impact2 <- coastal_eez_impact %>%
  filter(!is.na(iso3c)) %>%
  select(-rgn_label)


missing_iso3c <- setdiff(depend$iso3c, coastal_eez_impact2$iso3c)
filter(depend, iso3c %in% missing_iso3c) %>% print(n=25)
setdiff(coastal_eez_impact2$iso3c, depend$iso3c)

coastal_compare <- coastal_eez_impact2 %>%
  filter(region=="coast") %>%
  left_join(depend, by="iso3c") %>%
  filter(!is.na(ssp585_current))

coastal_compare <- coastal_compare %>%
  mutate(change_ssp245_medium = `ssp245_medium-term` - ssp245_current) %>%
  mutate(pct_change_ssp245_medium = (`ssp245_medium-term` - ssp245_current)/ssp245_current) %>%
    mutate(change_ssp585_medium = `ssp585_medium-term` - ssp585_current) %>%
    mutate(pct_change_ssp585_medium = (`ssp585_medium-term` - ssp585_current)/ssp585_current)


write_csv(coastal_compare, here("impact/data/country_coast_cumulative_impact.csv"))


coastal_compare <- read_csv(here("impact/data/country_coast_cumulative_impact.csv")) 

quantile(coastal_compare$integrated_depend, c(0.75), na.rm=TRUE)

coastal_compare <- coastal_compare %>%
  mutate(depend_cat = ifelse(integrated_depend >= 0.3, 1, 0))


p <- ggplot(data=coastal_compare , aes(y=`ssp245_medium-term`, x=change_ssp245_medium, color=as.factor(depend_cat))) +
  geom_point(aes(text=country), alpha=0.7, size=1.1) +
  scale_color_manual(values = c("1" = "red", "0" = "gray")) + 
  geom_vline(xintercept = median(coastal_compare$change_ssp245_medium, na.rm=TRUE), linetype="dashed", color="orange") +
   geom_hline(yintercept = median(coastal_compare$`ssp245_medium-term`, na.rm=TRUE), linetype="dashed", color="orange") +
  theme_classic()  +
  theme(
    # Add major gridlines
    panel.grid.major = element_line(color = "lightgray", size = 0.1),
    # Remove minor gridlines
    panel.grid.minor = element_blank(),
    legend.position = "none"
  ) +
  labs(y="Future impact, 2040-2060", x = "\u0394 impact")

p
ggsave(here("impact/figs/current_future_depend_scatterplot.png"), plot = p, width = 5, height = 3, dpi = 300)


pp <- plotly::ggplotly(p, tooltip = "text")
pp
library(htmlwidgets)
saveWidget(pp, "interactive_plot.html", selfcontained = TRUE)

ggplot(data=filter(coastal_compare, !is.na(depend_cat)) , aes(x=change_ssp245_medium, fill=as.factor(depend_cat), color=as.factor(depend_cat))) +
  geom_density(alpha=0.7, size=2, position = "identity") +
  scale_fill_manual(values = c("1" = "red", "0" = "gray")) + 
  scale_color_manual(values = c("1" = "red", "0" = "gray")) 
x
mod <- lm(change_ssp245_medium ~ integrated_depend, data=coastal_compare)
summary(mod)
mod <- lm(`ssp245_medium-term` ~ integrated_depend, data=coastal_compare)
summary(mod)

```


### Zonal country pressure impacts
```{r}
pressures_correct <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure_habitat_corrected/", full=TRUE)

hab_impacts_stack_correct <- rast(pressures_correct)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack_correct)))
names(hab_impacts_stack_correct) <- source_info


## country coastal raster
eez <- rast(here("_spatial/eez_mol_w_land.tif"))
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(eez_coast>0, 1, 0)
#get the area
country_coast <- eez*coast
plot(country_coast)

## EEZ data
eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region


result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(hab_impacts_stack_correct)) { # i = 1
  # Extract the current layer
  current_layer <- hab_impacts_stack_correct[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, country_coast, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
  cat(i, "\n")
}

# Merge all data frames by region ID
eez_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

eez_impact_result <- eez_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")

eez_impact_result <- eez_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

write_csv(eez_impact_result, here("impact/data/country_eez_habxpressurexscenario_impact.csv"))


pressure_country <- read_csv(here("impact/data/country_eez_habxpressurexscenario_impact.csv"))


pressure_rgn <- pressure_country %>%
  separate(
    col = impact,
    into = c("habitat", "pressure", "scenario", "time_frame"),
    sep = "_",
    remove = FALSE,
    extra = "merge",
    fill = "right"
  )

## summarize each pressure (across all habitats) for each region
pressure_rgn <- pressure_rgn %>%
  group_by(pressure, scenario, time_frame, rgn_id, geosubrgn) %>%
  summarize(value = sum(value, na.rm=TRUE))

# get number of coastal cells in each country:
 coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

pressure_rgn <- pressure_rgn %>%
  left_join(coast_cell_num, by="rgn_id")

 pressure_georgn <- pressure_rgn %>%
   group_by(pressure, scenario, time_frame, geosubrgn) %>%
   summarize(value = sum(value, na.rm=TRUE),
             area = sum(cells, na.rm=TRUE))

pressure_georgn <- pressure_georgn %>%
  mutate(avg_value = value/area) 

plot_df <- pressure_georgn %>%
  filter(scenario %in% c("ssp245")) %>%
  select(pressure, time_frame, geosubrgn, avg_value)

# Spread the data by time_period
plot_df_wide <- plot_df %>%
  pivot_wider(names_from = time_frame, values_from = avg_value)

plot_df2 <- plot_df_wide %>%
  mutate(medium_change = `medium-term` - current) %>%
  select(scenario, pressure, geosubrgn, `medium-term`, medium_change) %>%
  pivot_longer(cols = c(`medium-term`, medium_change), 
               names_to = "variable", 
               values_to = "value")

colors <- read_csv(here("combining_pressures/data/pressure_colors.csv")) %>%
  mutate(region = ifelse(region=="coastal", "coast", "eez")) %>%
  filter(region =="eez") %>%
  select(color, pressure)

plot_df2 <- plot_df2 %>%
  left_join(colors, by = c("pressure"))

# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- plot_df2 %>%
  filter(variable == "medium_change") %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`value`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))

subregion_order <- plot_df2 %>%
  filter(variable == "medium_change") %>%
  group_by(geosubrgn) %>%
  summarize(total_value = sum(value), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(geosubrgn = factor(geosubrgn, levels = unique(geosubrgn)))

 # Step 2: Reorder the 'pressure' factor levels based on the total_value
plot_df2 <- plot_df2 %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(geosubrgn = factor(geosubrgn, levels = unique(subregion_order$geosubrgn))) %>%
  arrange(geosubrgn, pressure)

# Step 3: Plot the data

plots_bars <- ggplot(plot_df2, aes(y = geosubrgn, x = value, fill = pressure)) +
  geom_bar(stat = "identity") +
  facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_df2$color, plot_df2$pressure)) +
  theme(
    strip.text = element_blank(),  # Removes the facet strip titles
    legend.title = element_blank() # Removes the legend title
  )

plots_bars

ggsave(here("combining_pressures/paper_figs/georgn_pressure_change_barplot.png"), plot = plots_bars, width = 10, height = 4, dpi = 300)


```



## Get habitat impact data
```{r}

pressures <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

hab_impacts_sum <- global(hab_impacts_stack, "sum", na.rm=TRUE) 

hab_impacts <- hab_impacts_sum %>%
  rename(impact = sum) %>%
  mutate(hab_vul_scen = rownames(hab_impacts_sum)) %>%
    separate(
    col = hab_vul_scen,           
    into = c("habitat", "pressure", "scenario", "time_frame"),  
    sep = "_",                  
    remove = TRUE,                
    extra = "merge",              
    fill = "right"                
  )
rownames(hab_impacts) <- NULL


# get habitat areas
habitats <- list.files(here("habitats/data"), full=TRUE)
hab_stack <- rast(habitats)
source_info <- gsub(".tif", "", basename(sources(hab_stack)))
names(hab_stack) <- source_info

hab_area <- global(hab_stack, "sum", na.rm=TRUE)  
hab_area$habitat <- rownames(hab_area)
rownames(hab_area) <- NULL
hab_area <- hab_area %>%
  rename(area=sum)


## join with impact
avg_hab_impact <- left_join(hab_impacts, hab_area, by="habitat") %>%
  mutate(avg_impact = impact/area) %>%
  mutate(scen_time = paste(scenario, time_frame, sep="_")) %>%
  select(habitat, pressure, scen_time, avg_impact)

## make data wide
avg_hab_impact <- avg_hab_impact %>%
  pivot_wider(
    names_from = scen_time,
    values_from = avg_impact,
    values_fill = list(avg_impact = 0)) %>%
  mutate(ssp245_change = `ssp245_medium-term`-ssp245_current,
         ssp585_change = `ssp585_medium-term`-ssp585_current)

## add in color information
colors <- read_csv(here("combining_pressures/data/pressure_colors.csv")) %>%
  mutate(region = ifelse(region=="coastal", "coast", "eez")) %>%
  filter(region =="eez") %>%
  select(color, pressure)

avg_hab_impact <- avg_hab_impact %>%
  left_join(colors, by = c("pressure"))


# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- avg_hab_impact %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`ssp245_medium-term`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))


habitat_order <- avg_hab_impact %>%
  group_by(habitat) %>%
  summarize(total_value = sum(`ssp245_medium-term`), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat)))

 # Step 2: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp245_medium-term`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    strip.text = element_blank(),  # Removes the facet strip titles
    legend.title = element_blank() # Removes the legend title
  )
plots_bars
ggsave(here("combining_pressures/paper_figs/habitat_impact_future_barplot.png"), plot = plots_bars, width = 10, height = 4, dpi = 300)


# get the change
 # Step 2b: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp245_change`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    strip.text = element_blank(),  # Removes the facet strip titles
    legend.title = element_blank() # Removes the legend title
  )
plots_bars
ggsave(here("combining_pressures/paper_figs/habitat_impact_change_barplot.png"), plot = plots_bars, width = 10, height = 4, dpi = 300)


```

## Make a scatterplot of habitat impacts
```{r}

plot_data <- avg_hab_impact %>%
    select(habitat, pressure, ssp245_current, `ssp245_medium-term`, ssp245_change) %>%
  pivot_longer(
    cols = c(ssp245_current, `ssp245_medium-term`, ssp245_change),
    names_to = "scenario",
    values_to = "value"
  ) %>%
  group_by(habitat, scenario) %>%
  summarize(impact = sum(value)) %>%
  pivot_wider(values_from=impact, names_from=scenario)

ggplot(plot_data, aes(x=`ssp245_medium-term`, y=ssp245_change)) +
  geom_point()+
  geom_text(aes(label = habitat), vjust = -1, hjust = 0.5) + 
  theme_minimal()

ggplot(plot_data, aes(x=`ssp245_current`, y=ssp245_change)) +
  geom_point()+
  geom_text(aes(label = habitat), vjust = -1, hjust = 0.5) + 
  theme_minimal()

```


https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4452591/#:~:text=We%20found%20that%20the%20global,m%2D2%20s%2D1.