---
title: "Calculating impacts"
output: html_document
date: "2024-09-16"
---

```{r setup, include=FALSE}
library(tidyverse)
library(terra)
library(here)
library(countrycode)

```

```{r}
vulnerability <- read_csv(here("impact/vulnerability_weighting_matrix.csv")) %>%
  mutate_if(is.numeric, ~ . / 4)

habitat_files <- list.files(here("habitats/data"), full=TRUE)

setdiff(names(vulnerability), gsub(".tif", "", basename(habitat_files)))
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

pressure_files <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures", full=TRUE)
setdiff(vulnerability$pressure, gsub(".tif", "", sub("_.*", "", basename(pressure_files))))
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

```

# combos
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
file.remove(clean)

for(habitat in names(vulnerability)[-1]){
  #habitat = names(vulnerability)[2]
  hab_file <- grep(habitat, habitat_files, value=TRUE)
  hab_raster <- rast(hab_file)
  cat(habitat, "\n")
  
  for(pressure in vulnerability$pressure){
  # pressure = vulnerability$pressure[7]
  # pressure = "sst-extremes"
  pressure_filter <- paste0(pressure, "_")
  p_files <- grep(pressure_filter, pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(pressure, " n = ", length(p_files), "\n")
  
  vulnerability_value <- vulnerability %>%
    filter(pressure == !!pressure) %>%
    select(habitat) %>%
    pull()
  
  pressure_vulnerability <- pressure_rasts * vulnerability_value
  
  pressure_vulnerability_habitat <- pressure_vulnerability * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/hab_vuln_pressure/"
  p_names <- basename(gsub(paste0(pressure, "_"), "", p_files))
  output_filenames <- sprintf("%s%s_%s_%s", save_loc, habitat, pressure, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}
}

```

Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
plot(rast(check[652]))

length(check)
 #should be 4,767
habitats <- sub("_.*", "", basename(check))
table(habitats)
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", basename(check))
table(pressure)

```

## adjust air temperature pressure
The pressure is a little different for salt-marsh, seagrass, and ice
```{r}

combo_files <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)

heatindex_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "salt-marsh"
#lower_threshold <- 0.2
#upper_threshold <- 0.25
combo_files <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)  
  
hi_files <- grep(sprintf("%s_air-heat-index", habitat), combo_files, value=TRUE)
file.remove(hi_files)
hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)
  
  p_files <- grep("air-heat-index", pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(" n = ", length(p_files), "\n")
  
  p <- pressure_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/hab_vuln_pressure/"
  p_names <- basename(p_files)
  output_filenames <- sprintf("%s%s_%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

heatindex_modify(habitat="salt-marsh", lower_threshold=0.2, upper_threshold=0.25) 
heatindex_modify(habitat="seagrass", lower_threshold=0.2, upper_threshold=0.25) 
heatindex_modify(habitat="ice", lower_threshold=0, upper_threshold=0) 
```


## adjust SST pressure
The pressure is a little different for kelp and coral because there is more information about them.
First prepare the raw sst data. These impacts are based on absolute temperatures.
```{r}

# get the temperature data

sst_raw_prep <- function(scenario){
#scenario <- "ssp245"
sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average_final/", full=TRUE)
scenario_sst_raw_files <- grep(scenario, sst_raw_files, value=TRUE)

current_files <- grep("2015_2019", scenario_sst_raw_files, value=TRUE)
current_stack <- rast(current_files)
current_mean <- mean(current_stack) - 273.15

near_files <- grep("2020_2024|2025_2029|2030_2034|2035_2039", scenario_sst_raw_files, value=TRUE)
near_stack <- rast(near_files)
near_mean <- mean(near_stack) - 273.15

medium_files <- grep("2040_2044|2045_2049|2050_2054|2055_2059", scenario_sst_raw_files, value=TRUE)
medium_stack <- rast(medium_files)
medium_mean <- mean(medium_stack) - 273.15

long_files <- grep("2080_2084|2085_2089|2090_2094|2095_2099", scenario_sst_raw_files, value=TRUE)
long_stack <- rast(long_files)
long_mean <- mean(long_stack) - 273.15

writeRaster(current_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_current.tif", scenario), overwrite=TRUE)
writeRaster(near_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_near-term.tif", scenario), overwrite=TRUE)
writeRaster(medium_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_medium-term.tif", scenario), overwrite=TRUE)
writeRaster(long_mean, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst/sst-raw_%s_long-term.tif", scenario), overwrite=TRUE)
}

sst_raw_prep(scenario="ssp245")
sst_raw_prep(scenario="ssp585")

```


Now apply to coral, kelp, and mangroves.
```{r}

sst_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "kelp"
#lower_threshold <- 18.3
#upper_threshold <- 24
combo_files <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)  
sst_files <- grep(sprintf("%s_sst-extremes", habitat), combo_files, value=TRUE)
file.remove(sst_files)

hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)

sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/raw_sst", full=TRUE)    
sst_raw_rasts <- rast(sst_raw_files)
cat(" n = ", length(sst_raw_files), "\n")
  
  p <- sst_raw_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/hab_vuln_pressure/"
  p_names <- gsub("sst-raw", "", basename(sst_raw_files))
  output_filenames <- sprintf("%s%s_sst-extremes%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

sst_modify(habitat="kelp", lower_threshold=18.3, upper_threshold=24) 
sst_modify(habitat="coral-reef", lower_threshold=29, upper_threshold=30)  
sst_modify(habitat="mangroves", lower_threshold=32, upper_threshold=40)  
```

## Make ssp names consistent
Some pressures do not use RCP climate models. The difference in naming (ssp2 vs. ssp245) is throwing everything off!

```{r}

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.copy(from = files_ssp1, to = gsub("ssp1_", "ssp119_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.rename(from = files_ssp1, to = gsub("ssp1_", "ssp126_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp2 <- grep("ssp2_", files_fix, value=TRUE)
file.rename(from = files_ssp2, to = gsub("ssp2_", "ssp245_", files_ssp2))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp3 <- grep("ssp3_", files_fix, value=TRUE)
file.rename(from = files_ssp3, to = gsub("ssp3_", "ssp370_", files_ssp3))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp4 <- grep("ssp4_", files_fix, value=TRUE)
file.rename(from = files_ssp4, to = gsub("ssp4_", "ssp460_", files_ssp4))

files_fix <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp5 <- grep("ssp5_", files_fix, value=TRUE)
file.rename(from = files_ssp5, to = gsub("ssp5_", "ssp585_", files_ssp5))
  
```



Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/")
length(check)
 #should be 4,998
habitats <- sub("_.*", "", check)
table(habitats)
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", check)
table(pressure)

```


Create raster with total number of habitats in each cell.
```{r}
habitat_files <- list.files(here("habitats/data"), full=TRUE)
habitat_rast <- rast(habitat_files)
habitat_n <- sum(habitat_rast, na.rm=TRUE)
check <- ifel(habitat_n < 1, 1, NA)
plot(check, col="red")
habitat_n = ifel(habitat_n <1, 1, habitat_n)
writeRaster(habitat_n, here("habitats/habitat_n.tif"), overwrite=TRUE)

```

# Calculate cumulative impacts
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
file.remove(clean)

pressures <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)

cum_impact <- function(scen_time){
#scen_time <- "ssp585_medium-term"
habitat_n <- rast(here("habitats/habitat_n.tif"))

pressures_filter <- grep(scen_time, pressures, value=TRUE)
cat(paste0(scen_time, ", n = ", length(pressures_filter), "\n"))
pressure_stack <- rast(pressures_filter)
sum_pressure <- sum(pressure_stack, na.rm=TRUE)
impact <- sum_pressure/habitat_n
writeRaster(impact, sprintf("/home/shares/ohi/stressors_2021/impacts/cumulative/%s.tif", scen_time), overwrite=TRUE)
}
cum_impact(scen_time = "ssp585_medium-term")
cum_impact(scen_time = "ssp585_near-term")
cum_impact(scen_time = "ssp585_current")
cum_impact(scen_time = "ssp245_medium-term")
cum_impact(scen_time = "ssp245_near-term")
cum_impact(scen_time = "ssp245_current")

```

## zonal extracts of data
Cumulative pressures
```{r}
## Change in global pressure
## global
tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

global(stack, "sum", na.rm=TRUE)
2000034.7/852431.8
2353158.6/847087.1

## Change in coastal pressure
## global coast
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(coast>0, 1, 0)

tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

stack <- stack * coast
global(stack, "sum", na.rm=TRUE)
100920.63/ 55436.48
109594.93/55201.98


# relative pressures coastal vs global
cumulative <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info
tmp <- global(cum_stack, "sum", na.rm=TRUE)

ocean_cells <- rast(here("_spatial/ocean_area_mol.tif"))
ocean_cells <- ifel(ocean_cells>0, 1, 0)
ocean_cells_num <- global(ocean_cells, "sum", na.rm=TRUE)


cum_stack <- as.data.frame(tmp) %>%
mutate(cells = ocean_cells_num) %>%
  mutate(average = sum/cells)

# figure out for coastal
eez_coast <- rast(here("_spatial/coastal_country_mol.tif"))
eez_coast <- ifel(eez_coast>0, 1, NA)

p_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(p_stack)))
names(p_stack) <- source_info

p_stack_mask <- p_stack*eez_coast

coastal_stats <- global(p_stack_mask, "sum", na.rm=TRUE)
coastal_df <- data.frame(coastal_stats) 
coastal_df$filename <- rownames(coastal_df)
row.names(coastal_df) <- NULL

names(coastal_df) <- c("value", "filename")
coastal_df <- coastal_df %>%
  separate(filename, into=c("pressure", "scenario", "timeframe"), sep="_")

coast_cells <- ifel(eez_coast>0, 1, 0)
coast_cells_num <- global(coast_cells, "sum", na.rm=TRUE)

coastal_df <- coastal_df %>%
  mutate(cells = coast_cells_num$sum) %>%
  mutate(average = value/cells) %>%
  select(-value, -cells)




```

cumulative impact
global map future total impact.
```{r}

library(paletteer)
library(png)

upperscale <- 1.5
cols_palette <- paletteer_c("grDevices::Lajolla", 200)
#cols_palette <- rev(paletteer_c("ggthemes::Classic Orange-Blue", 250)[1:150])
#cols_palette <- paletteer_c("ggthemes::Temperature Diverging", 100)

tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

stack_capped <- ifel(stack>upperscale, upperscale, stack)

for(i in 1:dim(stack_capped)[3]){
#i=5
zlim_values <- c(0, upperscale)
raster_data <- stack_capped[[i]]
df <- as.data.frame(raster_data, xy = TRUE, na.rm = TRUE)
colnames(df) <- c("x", "y", "value")

plot_cum <- ggplot(df, aes(x = x, y = y, fill = value)) +
  geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
#  ggtitle(gsub(".tif", "", basename(tmp[i]))) +
  theme_minimal() +
theme(
  panel.grid = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank(),
  axis.title = element_blank(),
  legend.position = "right",
  plot.title = element_text(hjust = 0.5)
)

ggsave(plot_cum, filename=here(sprintf("impact/figs/cumulative/%s", gsub(".tif", ".png", basename(tmp[i])))), width=5, height = 2.5, units = c("in"))
}

```

Make global plot of change in impact figures
```{r}

cols_palette <- c(paletteer_c("grDevices::BluGrn", 30)[21:30],
                  rev(paletteer_c("grDevices::Inferno", 110)))


list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

ssp2_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif") 


plot(ssp2_med)
quantile(data.frame(ssp2_med)$sum, c(0.99999))

diff_impact <- function(raster, lowerscale = -0.2, upperscale = 1.2, saveName){
#raster=ssp2_med
# fileName = "test"
  raster_data = raster
raster_data <- ifel(raster_data>upperscale, upperscale, raster_data)
raster_data <- ifel(raster_data<lowerscale, lowerscale, raster_data)
zlim_values <- c(lowerscale, upperscale)
df <- as.data.frame(raster_data, xy = TRUE, na.rm = TRUE)
colnames(df) <- c("x", "y", "value")

plot_cum <- ggplot(df, aes(x = x, y = y, fill = value)) +
  geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
#  ggtitle(gsub(".tif", "", basename(tmp[i]))) +
  theme_minimal() +
theme(
  panel.grid = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank(),
  axis.title = element_blank(),
  legend.position = "right",
  plot.title = element_text(hjust = 0.5)
)

ggsave(plot_cum, filename=here(sprintf("impact/figs/cumulative/%s", saveName)), width=5, height = 2.5, units = c("in"))
}

diff_impact(raster=ssp2_med, lowerscale = -0.1, upperscale = 1.25, saveName="sspp245_change.png")
```

## Figure showing highest risk areas
```{r}

ssp2_future <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")
ssp2_future_quant <-global(ssp2_future, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp2_future_high <- ifel(ssp2_future >= ssp2_future_quant, 1, 0)

ssp2_change <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif")
ssp2_change_quant <-global(ssp2_change, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp2_change_high <- ifel(ssp2_change >= ssp2_change_quant, 1, 0)

high_risk <- ssp2_change_high + ssp2_future_high
high_risk <- ifel(high_risk == 0, NA, high_risk)
plot(high_risk, col=c("orange", "red"))
```

## cumulative impact: country EEZ summaries
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))
## cut land
ocean_cells <- rast(here("_spatial/ocean_area_mol.tif"))
ocean_cells <- ifel(ocean_cells>0, 1, 0)
## add in the coastal area
eez_coast <- rast(here("_spatial/coastal_country_mol.tif"))
eez_coast <- ifel(eez_coast>0, 1, 0)
## combine coast and ocean
ocean_area <- app(c(ocean_cells, eez_coast), "sum", na.rm=TRUE)
ocean_area <- ifel(ocean_area>0, 1, NA)
plot(ocean_area)

#get the area
eez <- eez*ocean_area
plot(eez)

eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region

impact_files <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

impact_stack <- rast(impact_files)
source_info <- gsub(".tif", "", basename(sources(impact_stack)))
names(impact_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(impact_stack)) { # i = 1
  # Extract the current layer
  current_layer <- impact_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, eez, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
eez_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

eez_impact_result <- eez_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")

eez_impact_result <- eez_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
 eez_cell_num <- zonal(ifel(eez>0, 1, 0), eez, fun = "sum", na.rm=TRUE)
 names(eez_cell_num) <- c("rgn_id", "cells")

dim(eez_impact_result)
eez_impact_result <- eez_impact_result %>%
  left_join(eez_cell_num, by="rgn_id") %>%
  mutate(region = "eez") 


write_csv(eez_impact_result, here("impact/data/country_eez_impact.csv"))

```


## cumulative impact: country coastal region and  summaries
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(eez_coast>0, 1, 0)
#get the area
country_coast <- eez*coast
plot(country_coast)


eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region

impact_files <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

impact_stack <- rast(impact_files)
source_info <- gsub(".tif", "", basename(sources(impact_stack)))
names(impact_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(impact_stack)) { # i = 1
  # Extract the current layer
  current_layer <- impact_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, country_coast, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
coast_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

coast_impact_result <- coast_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")


coast_impact_result <- coast_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

dim(coast_impact_result)
coast_impact_result <- coast_impact_result %>%
  left_join(coast_cell_num, by="rgn_id") %>%
  mutate(region = "coast") 


write_csv(coast_impact_result, here("impact/data/country_coast_impact.csv"))

```

plot dependence vs. impacts
```{r}
#from paper_statistics_v2
depend <- read_csv(here("combining_pressures/data/selig_marine_dependence.csv"))

coast <- read_csv(here("impact/data/country_coast_impact.csv"))

impact_country <- coast %>%
  mutate(avg_value = value/cells) 
#%>%
#  separate(impact, into = c("scenario", "time_period"), sep = "_")

impact_country <- impact_country %>%
  select(region, rgn_label, iso3c, impact, avg_value) %>%
    pivot_wider(
    names_from = impact,
    values_from = avg_value
  )

impact_country$iso3c <- countrycode(impact_country$rgn_label, "country.name", "iso3c") 
impact_country <- filter(impact_country, rgn_label != "Bassas da India")


double_rgns <- data.frame(rgn_label = c("Guadeloupe and Martinique", "Guadeloupe and Martinique",
              "Northern Mariana Islands and Guam", "Northern Mariana Islands and Guam",
              "Puerto Rico and Virgin Islands of the United States", "Puerto Rico and Virgin Islands of the United States"),
                 iso3c = c("GLP", "MTQ", "GUM", "MNP", "PRI", "VIR"))

coastal_double_eez <- filter(impact_country, rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")) %>%
  select(-iso3c)

coastal_double_eez2 <- left_join(double_rgns, coastal_double_eez, by="rgn_label")

impact_country <- filter(impact_country, !(rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")))

coastal_eez_impact <- rbind(impact_country, coastal_double_eez2)
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label=="Northern Saint-Martin"] <- "NSM"
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label=="Micronesia"] <- "FSM"
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label %in% c("Crozet Islands",
                                       "Kerguelen Islands",
                                       "Amsterdam Island and Saint Paul Island",
                                       "Ile Europa",
                                       "Ile Tromelin",
                                       "Juan de Nova Island",
                                       "Glorioso Islands")] <- "ATF"
coastal_eez_impact <- filter(coastal_eez_impact, iso3c != "ATF")

coastal_eez_impact2 <- coastal_eez_impact %>%
  filter(!is.na(iso3c)) %>%
  select(-rgn_label)


missing_iso3c <- setdiff(depend$iso3c, coastal_eez_impact2$iso3c)
filter(depend, iso3c %in% missing_iso3c) %>% print(n=25)
setdiff(coastal_eez_impact2$iso3c, depend$iso3c)

coastal_compare <- coastal_eez_impact2 %>%
  filter(region=="coast") %>%
  left_join(depend, by="iso3c") %>%
  filter(!is.na(ssp585_current))

coastal_compare <- coastal_compare %>%
  mutate(change_ssp245_medium = `ssp245_medium-term` - ssp245_current) %>%
  mutate(pct_change_ssp245_medium = (`ssp245_medium-term` - ssp245_current)/ssp245_current) %>%
    mutate(change_ssp585_medium = `ssp585_medium-term` - ssp585_current) %>%
    mutate(pct_change_ssp585_medium = (`ssp585_medium-term` - ssp585_current)/ssp585_current)


write_csv(coastal_compare, here("impact/data/country_coast_cumulative_impact.csv"))


coastal_compare <- read_csv(here("impact/data/country_coast_cumulative_impact.csv")) 

quantile(coastal_compare$integrated_depend, c(0.75), na.rm=TRUE)

coastal_compare <- coastal_compare %>%
  mutate(depend_cat = ifelse(integrated_depend >= 0.3, 1, 0))


p <- ggplot(data=coastal_compare , aes(y=`ssp245_medium-term`, x=change_ssp245_medium, color=as.factor(depend_cat))) +
  geom_point(aes(text=country), alpha=0.7, size=2) +
  scale_color_manual(values = c("1" = "red", "0" = "gray")) + 
#  scale_color_gradientn(colors=paletteer_c("ggthemes::Temperature Diverging", 30)) +
  geom_vline(xintercept = median(coastal_compare$change_ssp245_medium, na.rm=TRUE), linetype="dashed", color="orange") +
   geom_hline(yintercept = median(coastal_compare$`ssp245_medium-term`, na.rm=TRUE), linetype="dashed", color="orange") +
  theme_minimal()

p
ggsave(here("combining_pressures/paper_figs/current_future_scatterplot.png"), plot = p, width = 7, height = 4, dpi = 300)


plotly::ggplotly(ptooltip = "text")

ggplot(data=filter(coastal_compare, !is.na(depend_cat)) , aes(x=change_ssp245_medium, fill=as.factor(depend_cat), color=as.factor(depend_cat))) +
  geom_density(alpha=0.7, size=2, position = "identity") +
  scale_fill_manual(values = c("1" = "red", "0" = "gray")) + 
  scale_color_manual(values = c("1" = "red", "0" = "gray")) 
x
mod <- lm(change_ssp245_medium ~ integrated_depend, data=coastal_compare)
summary(mod)
mod <- lm(`ssp245_medium-term` ~ integrated_depend, data=coastal_compare)
summary(mod)

```


### Zonal country pressure impacts
First step is to summarize the pressures for each habitat (being sure to divide by the pressure).

```{r}

check <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure_habitat_corrected/", full=TRUE)
file.remove(check)

pressures <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

## have to correct the pressures by the number of habitats
habitat_n <- rast(here("habitats/habitat_n.tif"))

for(i in 1: nlyr(hab_impacts_stack)){ #i=1
current_layer <- hab_impacts_stack[[i]]

  # Perform the division
  corrected_layer <- current_layer / habitat_n

  # Define the output filename
  layer_name <- names(hab_impacts_stack)[i]

  # Write the output to disk
  writeRaster(corrected_layer, 
              filename = sprintf("/home/shares/ohi/stressors_2021/hab_vuln_pressure_habitat_corrected/%s.tif", layer_name), 
              overwrite = TRUE)

  # Optional: Print progress
  cat("Processed and saved:", layer_name, "\n")
}


pressures_correct <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure_habitat_corrected/", full=TRUE)

hab_impacts_stack_correct <- rast(pressures_correct)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack_correct)))
names(hab_impacts_stack_correct) <- source_info


## country coastal raster
eez <- rast(here("_spatial/eez_mol_w_land.tif"))
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(eez_coast>0, 1, 0)
#get the area
country_coast <- eez*coast
plot(country_coast)

## EEZ data
eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region


result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(hab_impacts_stack_correct)) { # i = 1
  # Extract the current layer
  current_layer <- hab_impacts_stack_correct[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, country_coast, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
  cat(i, "\n")
}

# Merge all data frames by region ID
eez_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

eez_impact_result <- eez_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")

eez_impact_result <- eez_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

write_csv(eez_impact_result, here("impact/data/country_eez_habxpressurexscenario_impact.csv"))


pressure_country <- read_csv(here("impact/data/country_eez_habxpressurexscenario_impact.csv"))


pressure_rgn <- pressure_country %>%
  separate(
    col = impact,
    into = c("habitat", "pressure", "scenario", "time_frame"),
    sep = "_",
    remove = FALSE,
    extra = "merge",
    fill = "right"
  )

## summarize each pressure (across all habitats) for each region
pressure_rgn <- pressure_rgn %>%
  group_by(pressure, scenario, time_frame, rgn_id, geosubrgn) %>%
  summarize(value = sum(value, na.rm=TRUE))

# get number of coastal cells in each country:
 coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

pressure_rgn <- pressure_rgn %>%
  left_join(coast_cell_num, by="rgn_id")

 pressure_georgn <- pressure_rgn %>%
   group_by(pressure, scenario, time_frame, geosubrgn) %>%
   summarize(value = sum(value, na.rm=TRUE),
             area = sum(cells, na.rm=TRUE))

pressure_georgn <- pressure_georgn %>%
  mutate(avg_value = value/area) 

plot_df <- pressure_georgn %>%
  filter(scenario %in% c("ssp245")) %>%
  select(pressure, time_frame, geosubrgn, avg_value)

# Spread the data by time_period
plot_df_wide <- plot_df %>%
  pivot_wider(names_from = time_frame, values_from = avg_value)

plot_df2 <- plot_df_wide %>%
  mutate(medium_change = `medium-term` - current) %>%
  select(scenario, pressure, geosubrgn, `medium-term`, medium_change) %>%
  pivot_longer(cols = c(`medium-term`, medium_change), 
               names_to = "variable", 
               values_to = "value")

colors <- read_csv(here("combining_pressures/data/pressure_colors.csv")) %>%
  mutate(region = ifelse(region=="coastal", "coast", "eez")) %>%
  filter(region =="eez") %>%
  select(color, pressure)

plot_df2 <- plot_df2 %>%
  left_join(colors, by = c("pressure"))

# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- plot_df2 %>%
  filter(variable == "medium_change") %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`value`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))

subregion_order <- plot_df2 %>%
  filter(variable == "medium_change") %>%
  group_by(geosubrgn) %>%
  summarize(total_value = sum(value), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(geosubrgn = factor(geosubrgn, levels = unique(geosubrgn)))

 # Step 2: Reorder the 'pressure' factor levels based on the total_value
plot_df2 <- plot_df2 %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(geosubrgn = factor(geosubrgn, levels = unique(subregion_order$geosubrgn))) %>%
  arrange(geosubrgn, pressure)

# Step 3: Plot the data

plots_bars <- ggplot(plot_df2, aes(y = geosubrgn, x = value, fill = pressure)) +
  geom_bar(stat = "identity") +
  facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_df2$color, plot_df2$pressure)) +
  theme(
    strip.text = element_blank(),  # Removes the facet strip titles
    legend.title = element_blank() # Removes the legend title
  )

plots_bars

ggsave(here("combining_pressures/paper_figs/georgn_pressure_change_barplot.png"), plot = plots_bars, width = 10, height = 4, dpi = 300)


```



## Get habitat impact data
```{r}

pressures <- list.files("/home/shares/ohi/stressors_2021/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

hab_impacts_sum <- global(hab_impacts_stack, "sum", na.rm=TRUE) 

hab_impacts <- hab_impacts_sum %>%
  rename(impact = sum) %>%
  mutate(hab_vul_scen = rownames(hab_impacts_sum)) %>%
    separate(
    col = hab_vul_scen,           
    into = c("habitat", "pressure", "scenario", "time_frame"),  
    sep = "_",                  
    remove = TRUE,                
    extra = "merge",              
    fill = "right"                
  )
rownames(hab_impacts) <- NULL


# get habitat areas
habitats <- list.files(here("habitats/data"), full=TRUE)
hab_stack <- rast(habitats)
source_info <- gsub(".tif", "", basename(sources(hab_stack)))
names(hab_stack) <- source_info

hab_area <- global(hab_stack, "sum", na.rm=TRUE)  
hab_area$habitat <- rownames(hab_area)
rownames(hab_area) <- NULL
hab_area <- hab_area %>%
  rename(area=sum)


## join with impact
avg_hab_impact <- left_join(hab_impacts, hab_area, by="habitat") %>%
  mutate(avg_impact = impact/area) %>%
  mutate(scen_time = paste(scenario, time_frame, sep="_")) %>%
  select(habitat, pressure, scen_time, avg_impact)

## make data wide
avg_hab_impact <- avg_hab_impact %>%
  pivot_wider(
    names_from = scen_time,
    values_from = avg_impact,
    values_fill = list(avg_impact = 0)) %>%
  mutate(ssp245_change = `ssp245_medium-term`-ssp245_current,
         ssp585_change = `ssp585_medium-term`-ssp585_current)

## add in color information
colors <- read_csv(here("combining_pressures/data/pressure_colors.csv")) %>%
  mutate(region = ifelse(region=="coastal", "coast", "eez")) %>%
  filter(region =="eez") %>%
  select(color, pressure)

avg_hab_impact <- avg_hab_impact %>%
  left_join(colors, by = c("pressure"))


# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- avg_hab_impact %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`ssp245_medium-term`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))


habitat_order <- avg_hab_impact %>%
  group_by(habitat) %>%
  summarize(total_value = sum(`ssp245_medium-term`), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat)))

 # Step 2: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp245_medium-term`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    strip.text = element_blank(),  # Removes the facet strip titles
    legend.title = element_blank() # Removes the legend title
  )
plots_bars
ggsave(here("combining_pressures/paper_figs/habitat_impact_future_barplot.png"), plot = plots_bars, width = 10, height = 4, dpi = 300)


# get the change
 # Step 2b: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp245_change`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    strip.text = element_blank(),  # Removes the facet strip titles
    legend.title = element_blank() # Removes the legend title
  )
plots_bars
ggsave(here("combining_pressures/paper_figs/habitat_impact_change_barplot.png"), plot = plots_bars, width = 10, height = 4, dpi = 300)


```

## Make a scatterplot of habitat impacts
```{r}

plot_data <- avg_hab_impact %>%
    select(habitat, pressure, ssp245_current, `ssp245_medium-term`, ssp245_change) %>%
  pivot_longer(
    cols = c(ssp245_current, `ssp245_medium-term`, ssp245_change),
    names_to = "scenario",
    values_to = "value"
  ) %>%
  group_by(habitat, scenario) %>%
  summarize(impact = sum(value)) %>%
  pivot_wider(values_from=impact, names_from=scenario)

ggplot(plot_data, aes(x=`ssp245_medium-term`, y=ssp245_change)) +
  geom_point()+
  geom_text(aes(label = habitat), vjust = -1, hjust = 0.5) + 
  theme_minimal()

ggplot(plot_data, aes(x=`ssp245_current`, y=ssp245_change)) +
  geom_point()+
  geom_text(aes(label = habitat), vjust = -1, hjust = 0.5) + 
  theme_minimal()

```


https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4452591/#:~:text=We%20found%20that%20the%20global,m%2D2%20s%2D1.