---
title: "Calculating impacts"
output: html_document
date: "2024-09-16"
---

```{r setup, include=FALSE}
library(tidyverse)
library(terra)
library(here)
library(countrycode)

```

```{r}
vulnerability <- read_csv(here("impact/vulnerability_weighting_matrix.csv")) %>%
  mutate_if(is.numeric, ~ . / 4)

habitat_files <- list.files(here("habitats/data"), full=TRUE)

setdiff(names(vulnerability), gsub(".tif", "", basename(habitat_files))) # "pressure" ok!
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

pressure_files <- list.files("/home/shares/ohi/stressors_2021/_final_data/rescaled_pressures", full=TRUE)
setdiff(vulnerability$pressure, gsub(".tif", "", sub("_.*", "", basename(pressure_files))))
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

```

# combos
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
file.remove(clean)

for(habitat in names(vulnerability)[-1]){
  #habitat = names(vulnerability)[2]
  hab_file <- grep(habitat, habitat_files, value=TRUE)
  hab_raster <- rast(hab_file)
  cat(habitat, "\n")
  
  for(pressure in vulnerability$pressure){
  # pressure = vulnerability$pressure[7]
  # pressure = "sst-extremes"
  pressure_filter <- paste0(pressure, "_")
  p_files <- grep(pressure_filter, pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(pressure, " n = ", length(p_files), "\n")
  
  vulnerability_value <- vulnerability %>%
    filter(pressure == !!pressure) %>%
    select(habitat) %>%
    pull()
  
  pressure_vulnerability <- pressure_rasts * vulnerability_value
  
  pressure_vulnerability_habitat <- pressure_vulnerability * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/"
  p_names <- basename(gsub(paste0(pressure, "_"), "", p_files))
  output_filenames <- sprintf("%s%s_%s_%s", save_loc, habitat, pressure, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}
}

```

Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
plot(rast(check[652]))

length(check)
 #should be 4,620
habitats <- sub("_.*", "", basename(check))
table(habitats) # should be the same
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", basename(check))
table(pressure)

```

## adjust air temperature impact
The impact is a little different for salt-marsh, seagrass, and ice
```{r}

combo_files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)

heatindex_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "salt-marsh"
#lower_threshold <- 0.2
#upper_threshold <- 0.25
combo_files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)  
  
hi_files <- grep(sprintf("%s_air-heat-index", habitat), combo_files, value=TRUE)
file.remove(hi_files)
hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)
  
  p_files <- grep("air-heat-index", pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(" n = ", length(p_files), "\n")
  
  p <- pressure_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/"
  p_names <- basename(p_files)
  output_filenames <- sprintf("%s%s_%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

heatindex_modify(habitat="salt-marsh", lower_threshold=0.2, upper_threshold=0.25) 
heatindex_modify(habitat="seagrass", lower_threshold=0.2, upper_threshold=0.25) 

```


## adjust SST pressure
The pressure is a little different for kelp and coral because there is more information about them.
First prepare the raw sst data. These impacts are based on absolute temperatures.
```{r}

# get the temperature data

sst_raw_prep <- function(scenario){
#scenario <- "ssp245"
sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average_final/", full=TRUE)
scenario_sst_raw_files <- grep(scenario, sst_raw_files, value=TRUE)

current_files <- grep("2015_2019", scenario_sst_raw_files, value=TRUE)
current_stack <- rast(current_files)
current_mean <- mean(current_stack) - 273.15

near_files <- grep("2020_2024|2025_2029|2030_2034|2035_2039", scenario_sst_raw_files, value=TRUE)
near_stack <- rast(near_files)
near_mean <- mean(near_stack) - 273.15

medium_files <- grep("2040_2044|2045_2049|2050_2054|2055_2059", scenario_sst_raw_files, value=TRUE)
medium_stack <- rast(medium_files)
medium_mean <- mean(medium_stack) - 273.15

long_files <- grep("2080_2084|2085_2089|2090_2094|2095_2099", scenario_sst_raw_files, value=TRUE)
long_stack <- rast(long_files)
long_mean <- mean(long_stack) - 273.15

writeRaster(current_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_current.tif", scenario), overwrite=TRUE)
writeRaster(near_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_near-term.tif", scenario), overwrite=TRUE)
writeRaster(medium_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_medium-term.tif", scenario), overwrite=TRUE)
writeRaster(long_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_long-term.tif", scenario), overwrite=TRUE)
}

sst_raw_prep(scenario="ssp245")
sst_raw_prep(scenario="ssp585")

```


Now apply to coral, kelp, and mangroves.
```{r}

sst_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "kelp"
#lower_threshold <- 18.3
#upper_threshold <- 24
combo_files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)  
sst_files <- grep(sprintf("%s_sst-extremes", habitat), combo_files, value=TRUE)
file.remove(sst_files)

hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)

sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/impacts/raw_sst", full=TRUE)    
sst_raw_rasts <- rast(sst_raw_files)
cat(" n = ", length(sst_raw_files), "\n")
  
  p <- sst_raw_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/"
  p_names <- gsub("sst-raw", "", basename(sst_raw_files))
  output_filenames <- sprintf("%s%s_sst-extremes%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

sst_modify(habitat="kelp", lower_threshold=18.3, upper_threshold=24) 
sst_modify(habitat="coral-reef", lower_threshold=29, upper_threshold=30)  
sst_modify(habitat="mangroves", lower_threshold=32, upper_threshold=40)  
```

## Make ssp names consistent
Some pressures do not use RCP climate models. The difference in naming (ssp2 vs. ssp245) is throwing everything off!

```{r}

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.copy(from = files_ssp1, to = gsub("ssp1_", "ssp119_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.rename(from = files_ssp1, to = gsub("ssp1_", "ssp126_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp2 <- grep("ssp2_", files_fix, value=TRUE)
file.rename(from = files_ssp2, to = gsub("ssp2_", "ssp245_", files_ssp2))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp3 <- grep("ssp3_", files_fix, value=TRUE)
file.rename(from = files_ssp3, to = gsub("ssp3_", "ssp370_", files_ssp3))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp4 <- grep("ssp4_", files_fix, value=TRUE)
file.rename(from = files_ssp4, to = gsub("ssp4_", "ssp460_", files_ssp4))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp5 <- grep("ssp5_", files_fix, value=TRUE)
file.rename(from = files_ssp5, to = gsub("ssp5_", "ssp585_", files_ssp5))


```



Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/")
length(check)
 #should be 4,840
habitats <- sub("_.*", "", check)
table(habitats)
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", check)
table(pressure)

```


Create raster with total number of habitats in each cell.
```{r}
habitat_files <- list.files(here("habitats/data"), full=TRUE)
habitat_rast <- rast(habitat_files)
habitat_n <- sum(habitat_rast, na.rm=TRUE)
check <- ifel(habitat_n == 0, 1, NA)
plot(check, col="red")
global(check, "sum", na.rm=TRUE)

writeRaster(habitat_n, here("habitats/habitat_n.tif"), overwrite=TRUE)

```

# Calculate cumulative impacts
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
file.remove(clean)

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)

cum_impact <- function(scen_time){
#scen_time <- "ssp585_medium-term"
habitat_n <- rast(here("habitats/habitat_n.tif"))

pressures_filter <- grep(scen_time, pressures, value=TRUE)
cat(paste0(scen_time, ", n = ", length(pressures_filter), "\n"))
pressure_stack <- rast(pressures_filter)
sum_pressure <- sum(pressure_stack, na.rm=TRUE)
impact <- sum_pressure/habitat_n
writeRaster(impact, sprintf("/home/shares/ohi/stressors_2021/impacts/cumulative/%s.tif", scen_time), overwrite=TRUE)
}
cum_impact(scen_time = "ssp585_medium-term")
cum_impact(scen_time = "ssp585_near-term")
cum_impact(scen_time = "ssp585_current")
cum_impact(scen_time = "ssp245_medium-term")
cum_impact(scen_time = "ssp245_near-term")
cum_impact(scen_time = "ssp245_current")



```

# For individual pressures, need to create rasters that correct for number of habitats
For regional summaries of pressure impacts, the impact of the pressure will depend on the number of habitats. 
First step is to correct for the number of habitats (being sure to divide by the pressure). This is used for geographic summaries of pressures (but not habitat based summaries.).

```{r}

check <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure_habitat_corrected/", full=TRUE)
file.remove(check)

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

## have to correct the pressures by the number of habitats
habitat_n <- rast(here("habitats/habitat_n.tif"))

for(i in 1: nlyr(hab_impacts_stack)){ #i=1
current_layer <- hab_impacts_stack[[i]]

  # Perform the division
  corrected_layer <- current_layer / habitat_n

  # Define the output filename
  layer_name <- names(hab_impacts_stack)[i]

  # Write the output to disk
  writeRaster(corrected_layer, 
              filename = sprintf("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure_habitat_corrected/%s.tif", layer_name), 
              overwrite = TRUE)

  # Optional: Print progress
  cat("Processed and saved:", layer_name, "\n")
}

```


# Statistic: Describe projected increase in impact over time.
```{r}
## Change in global pressure
## global
tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

# paper statistics
global(stack, "sum", na.rm=TRUE)
1957006.7/890768.3
2330441.5/888146.8

#1958809.2/892389.5 ##ssp245
#2331685.3/889717.7 ##ssp585


```


## Vulnerability table (supplement)
```{r}

habitats <- read_csv(here("impact/habitat_list.csv")) %>%
  select(file_name, plot_name_habitat)
pressures <- read_csv(here("impact/pressure_list.csv")) %>%
  select(pressure, plot_name_pressure) %>%
  mutate(plot_name_pressure = ifelse(pressure == "sst-extremes", "ocean warming, surface", plot_name_pressure))  %>%
  mutate(plot_name_pressure = ifelse(pressure=="temperature-benthic", "ocean warming, benthic", plot_name_pressure)) %>%
 mutate(plot_name_pressure = ifelse(pressure == "oxygen-benthic", "reduced oxygen, benthic", plot_name_pressure)) %>%
        mutate(plot_name_pressure = ifelse(pressure == "oxygen-surface", "reduced oxygen, surface", plot_name_pressure))

vuln <- read_csv(here("impact/vulnerability_weighting_matrix.csv")) %>%
pivot_longer(
  cols = -pressure, 
  names_to = "file_name",
  values_to = "value"
) %>%
  mutate(value = round(value/4, 2)) %>%
  left_join(habitats) %>%
  left_join(pressures) %>%
  select(pressure = plot_name_pressure, habitat=plot_name_habitat, value) %>%
  pivot_wider(names_from = pressure,
              values_from = value)
  
write_csv(vuln, here("impact/tables/vulnerability.csv"))



```

### Figure 1
Maps of cumulative impact and change in cumulative impact
```{r}

library(paletteer)
library(png)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

```

Get land boundaries

```{r}

world <- ne_countries(scale = "medium", returnclass = "sf")
# Define the Mollweide projection string
moll_crs <- "+proj=moll"

# Reproject the world data to Mollweide
world_moll <- st_transform(world, crs = moll_crs)

```

Get current and future cumulative impacts
```{r}

upperscale <- 1.5
cols_palette <- paletteer_c("grDevices::Lajolla", 200)
#cols_palette <- rev(paletteer_c("ggthemes::Classic Orange-Blue", 250)[1:150])
#cols_palette <- paletteer_c("ggthemes::Temperature Diverging", 100)

tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

ssp2_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")
quantile(data.frame(ssp2_med)$sum, c(0.999))


stack_capped <- ifel(stack>upperscale, upperscale, stack)

for(i in 1:dim(stack_capped)[3]){
#i=5
zlim_values <- c(0, upperscale)
raster_data <- stack_capped[[i]]
df <- as.data.frame(raster_data, xy = TRUE, na.rm = TRUE)
colnames(df) <- c("x", "y", "value")

plot_cum <- ggplot(df, aes(x = x, y = y, fill = value)) +
  geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
  geom_sf(
    data = world_moll,
    fill = "gray80",       # Land fill color
    color = "gray85",      # Slightly lighter color for boundaries
    size = 0.2,            # Adjust line thickness
    inherit.aes = FALSE
  ) +
#  ggtitle(gsub(".tif", "", basename(tmp[i]))) +
geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
    theme_minimal() +
theme(
  panel.grid = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank(),
  axis.title = element_blank(),
  legend.position = "right",
  plot.title = element_text(hjust = 0.5)
)


ggsave(plot_cum, filename=here(sprintf("impact/figs/cumulative/%s", gsub(".tif", ".png", basename(tmp[i])))), width=5.5, height = 2.5, units = c("in"))
}

```

Function to make global plot of change in impact figures. 
```{r}

cols_palette <- c(paletteer_c("grDevices::BluGrn", 30)[21:30],
                  colorRampPalette(c("#F5F5F5", "#FFFE9EFF"))(10),
                  rev(paletteer_c("grDevices::Inferno", 100)))
#library(colorspace)
#cols_all <- diverging_hcl("Red-Green", n=220)
#cols_palette <- c(cols_all[151:160], rev(cols_all[1:90]))

diff_impact <- function(raster, lowerscale = -0.1, upperscale = 1.1, saveName){
#raster=ssp2_med
# fileName = "test"
  raster_data = raster
raster_data <- ifel(raster_data>upperscale, upperscale, raster_data)
raster_data <- ifel(raster_data<lowerscale, lowerscale, raster_data)
zlim_values <- c(lowerscale, upperscale)
df <- as.data.frame(raster_data, xy = TRUE, na.rm = TRUE)
colnames(df) <- c("x", "y", "value")

plot_cum <- ggplot(df, aes(x = x, y = y, fill = value)) +
  geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
  geom_sf(
    data = world_moll,
    fill = "gray80",       # Land fill color
    color = "gray85",      # Slightly lighter color for boundaries
    size = 0.2,            # Adjust line thickness
    inherit.aes = FALSE
  ) + 
   geom_raster() +
  scale_fill_gradientn(
    colours = cols_palette,
    limits = zlim_values,
    name = ""
  ) +
#  ggtitle(gsub(".tif", "", basename(tmp[i]))) +
  theme_minimal() +
theme(
  panel.grid = element_blank(),
  panel.border = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank(),
  axis.title = element_blank(),
  legend.position = "right",
  plot.title = element_text(hjust = 0.5)
)

ggsave(plot_cum, filename=here(sprintf("impact/figs/cumulative/%s", saveName)), width=4.5, height = 2.5, units = c("in"))
}


list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

ssp2_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif") 

global(ssp2_med, "min", na.rm=TRUE)
global(ssp2_med, "max", na.rm=TRUE)
plot(ssp2_med)

quantile(data.frame(ssp2_med)$sum, c(0.9999))
quantile(data.frame(ssp2_med)$sum, c(1-0.9999))
ssp5_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_current.tif") 

diff_impact(raster=ssp2_med, lowerscale = -0.1, upperscale = 1, saveName="sspp245_change.png")
diff_impact(raster=ssp5_med, lowerscale = -0.1, upperscale = 1, saveName="sspp585_change.png")
```


histogram and statistics for latitudinal patterns
```{r}

tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE, pattern = "current|medium-term")
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

global_df <- as.data.frame(stack, xy=TRUE) %>%
  pivot_longer(
    cols = starts_with("ssp"),  
    names_to = "scenario",      
    values_to = "value"         
  ) %>%
  group_by(y, scenario) %>%
  summarize(n = n(),
            sum_impact = sum(value)) %>%
  ungroup() %>%
  mutate(avg_impact = sum_impact/n) %>%
  select(y, scenario, avg_impact) 


plot_data <- global_df %>%
  filter(scenario %in% c("ssp245_current", "ssp245_medium-term"))

hist <- ggplot(plot_data, aes(x = y, y = avg_impact, fill = scenario)) +
  geom_area(
    data = filter(plot_data, scenario == "ssp245_current"),
    aes(x = y, y = avg_impact, fill = "ssp245_current"),
    alpha = 0.5, color = NA
  ) +
  geom_area(
    data = filter(plot_data, scenario == "ssp245_medium-term"),
    aes(x = y, y = avg_impact, fill = "ssp245_medium-term"),
    alpha = 0.5, color = NA
  ) +
  scale_fill_manual(
    values = c("ssp245_current" = "blue", "ssp245_medium-term" = "red"),
    guide = "none"  # Remove legend
  ) +
  theme_void() +  # Removes all background, grids, and axis lines
  theme(
    plot.title = element_blank(),    # Remove title
    axis.title = element_blank(),    # Remove axis titles
    axis.text = element_blank(),     # Remove axis text
    axis.ticks = element_blank(),    # Remove axis ticks
    panel.grid = element_blank(),    # Remove grid lines
    panel.border = element_blank()   # Remove border
  )

ggsave(hist, filename=here("impact/figs/latitudinal_impacts.png"), width=5.5, height =2, units = c("in"))


## upper
130/180 #+ 65 degrees
50/180 #-65 degrees

lats <- as.data.frame(stack, xy=TRUE) %>%
  pivot_longer(
    cols = starts_with("ssp"),  
    names_to = "scenario",      
    values_to = "value"         
  )

lat_values <- unique(lats$y)
low_lat <- quantile(lat_values, c(50/180))
high_lat <- quantile(lat_values, c(130/180))
eq_lower <- quantile(lat_values, c(70/180))
eq_upper <- quantile(lat_values, c(110/180))

ssp2_lowlat <- filter(lats, scenario == "ssp245_current" & y <= low_lat)
mean(ssp2_lowlat$value)
ssp2_highlat <- filter(lats, scenario == "ssp245_current" & y >= high_lat)
mean(ssp2_highlat$value)
ssp2_eq <- filter(lats, scenario == "ssp245_current" & y >= eq_lower & y <= eq_upper)
mean(ssp2_eq$value)


ssp2_lowlat <- filter(lats, scenario == "ssp245_medium-term" & y <= low_lat)
mean(ssp2_lowlat$value)
ssp2_highlat <- filter(lats, scenario == "ssp245_medium-term" & y >= high_lat)
mean(ssp2_highlat$value)
ssp2_eq <- filter(lats, scenario == "ssp245_medium-term" & y >= eq_lower & y <= eq_upper)
mean(ssp2_eq$value)

0.5298306/0.2543985
0.6631517/0.3858724
0.5305982/0.1964062

#0.5305011/0.2547932 # antarctic
#0.6625507/0.3853962 #arctic
#0.53126/ 0.1969383 # equator


ssp5_lowlat <- filter(lats, scenario == "ssp585_current" & y <= low_lat)
mean(ssp5_lowlat$value)
ssp5_highlat <- filter(lats, scenario == "ssp585_current" & y >= high_lat)
mean(ssp5_highlat$value)
ssp5_eq <- filter(lats, scenario == "ssp585_current" & y >= eq_lower & y <= eq_upper)
mean(ssp5_eq$value)


ssp5_lowlat <- filter(lats, scenario == "ssp585_medium-term" & y <= low_lat)
mean(ssp5_lowlat$value)
ssp5_highlat <- filter(lats, scenario == "ssp585_medium-term" & y >= high_lat)
mean(ssp5_highlat$value)
ssp5_eq <- filter(lats, scenario == "ssp585_medium-term" & y >= eq_lower & y <= eq_upper)
mean(ssp5_eq$value)

0.6248193/0.2492332
 0.7541939/0.3852165
0.6350968/ 0.1968663


#0.6254916/0.2498347  #antarctic
#0.7532798/0.3847086 #arctic
#0.6355367/0.1973814 #equator

```
Hotspot statistics

```{r}

## find the highest spots
list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

current <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif")
max(current$sum)
quantile(as.data.frame(current)$sum, c(0.99999))
quantile(as.data.frame(current)$sum, c(0.00001))
current_high <- ifel(current>1.8 , 1, NA)
current_high_df <- as.data.frame(current_high, xy=TRUE)
plot(current)
points(current_high_df$x, current_high_df$y, col="red")

mollweide_coords <- data.frame(
  x = c(11878718, 11878718),
  y = c(1805048, 1795048)
)

# Define the Mollweide projection (EPSG:54009)
mollweide_crs <- "+proj=moll +lon_0=0 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs"

# Create a simple feature object
mollweide_sf <- st_as_sf(mollweide_coords, coords = c("x", "y"), crs = mollweide_crs)

# Convert to WGS84 (latitude and longitude)
latlong_sf <- st_transform(mollweide_sf, crs = 4326)

# Extract the converted coordinates
latlong_coords <- st_coordinates(latlong_sf)
latlong_coords

future <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")

quantile(as.data.frame(future)$sum, c(0.99999))
quantile(as.data.frame(future)$sum, c(0.00001))
future_high <- ifel(future>2.01, 1, NA)
future_high_df <- as.data.frame(future_high, xy=TRUE)
plot(future)
points(future_high_df$x, future_high_df$y, col="red", pch=16)

```


Global plot of hotspots
```{r}

ssp2_change <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif") -
               rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif")
ssp2_change_high_quant <-global(ssp2_change, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp2_change_high <- ifel(ssp2_change >= ssp2_change_high_quant, 1, 0)
ssp2_change_low_quant <-global(ssp2_change, fun = quantile, probs = 0.20, na.rm = TRUE) %>% pull()
ssp2_change_low <- ifel(ssp2_change <= ssp2_change_low_quant, 1, 0)

ssp2_current <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif")
ssp2_current_high_quant <-global(ssp2_current, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp2_current_high <- ifel(ssp2_current >= ssp2_current_high_quant, 1, 0)
ssp2_current_low_quant <-global(ssp2_current, fun = quantile, probs = 0.20, na.rm = TRUE) %>% pull()
ssp2_current_low <- ifel(ssp2_current <= ssp2_current_low_quant, 1, 0)

# category 1: current low and stable
cat1 <- ssp2_current_low + ssp2_change_low
cat1 <- ifel(cat1 == 2, 3, NA)
plot(cat1, col="#44ACC8")

# category 2: current high and stable
cat2 <- ssp2_current_high + ssp2_change_low
cat2 <- ifel(cat2 == 2, 4, NA)
plot(cat2, col="#E2D100", add=TRUE)

# category 3: current high and fast increase
cat3 <- ssp2_current_high + ssp2_change_high
cat3 <- ifel(cat3 == 2, 3, NA)
plot(cat3, col="#E58501", add=TRUE)

# category 4: current low and fast increase
cat4 <- ssp2_current_low + ssp2_change_high
cat4 <- ifel(cat4 == 2, 4, NA)
plot(cat4, col="#B41821", add=TRUE)

png(here("impact/figs/hotspots.png"), width = 2000, height = 1500, res = 300)

plot(st_geometry(world_moll), col="gray80", border="gray85")
plot(cat1, col="#44ACC8", add=TRUE, legend=FALSE)
plot(cat2, col="#E2D100", add=TRUE, legend=FALSE)
plot(cat3, col="#B41821", add=TRUE, legend=FALSE)
plot(cat4, col="#E58501", add=TRUE, legend=FALSE)

dev.off()

```

## Figure 2: coastal vs ocean

Some stats for the paper
```{r}


# relative pressures coastal vs global

## global coast
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(coast>0, 1, 0)

## offshore area only
ocean <- rast(here("_spatial/ocean_area_mol.tif"))
ocean <- ifel(ocean>0, 2, 0)
offshore <- ocean - ifel(is.na(coast), 0, coast)
offshore <- ifel(offshore ==2, 1, NA)


cumulative <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info

offshore_pressures <- cum_stack*offshore
total_offshore_pressures <- global(offshore_pressures, "sum", na.rm=TRUE)

offshore_cells_num <- global(offshore, "sum", na.rm=TRUE) %>% pull()

offshore_impact <- as.data.frame(total_offshore_pressures) %>%
  mutate(cells = offshore_cells_num) %>%
  mutate(average = sum/cells)

# figure out for coastal
coastal_pressures <- cum_stack*coast
total_coastal_pressures <- global(coastal_pressures, "sum", na.rm=TRUE)

coastal_cells_num <- global(coast, "sum", na.rm=TRUE) %>% pull()

coastal_impact <- as.data.frame(total_coastal_pressures) %>%
  mutate(cells = coastal_cells_num) %>%
  mutate(average = sum/cells)
0.3804300/0.2355148
#0.3689473/ 0.2364934

```


Cumulative impact plot

```{r}

cumulative <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info

offshore_pressures <- cum_stack*offshore

offshore_points <- data.frame(offshore_pressures)
offshore_points$region <- "offshore"

coastal_pressures <- cum_stack*coast

coast_points <- data.frame(coastal_pressures)
coast_points$region <- "coastal"

point_data <- rbind(coast_points, offshore_points[sample(nrow(offshore_points), dim(coast_points)[1]), ])

colnames(point_data) <- gsub("\\.", "-", colnames(point_data))


point_data_long <- point_data %>%
  pivot_longer(
    cols = starts_with("ssp"),       # Columns to reshape
    names_to = c("scenario", "time_period"),  # New column names
    names_sep = "_",                 # Separator used in column names
    values_to = "value"              # Name of the values column
  )

point_data_plot <- filter(point_data_long, scenario == "ssp245", time_period %in% c("current", "medium-term"))

# Calculate the overall range of 'value' to set consistent axes
range(point_data_long$value, na.rm = TRUE)


cum <- ggplot(point_data_plot, aes(y = region, x = value, fill = region)) +
  geom_violin(trim = FALSE) +
  xlim(0, 2) + 
  facet_wrap(~ time_period, ncol = 1) +
  scale_fill_manual(values = c("offshore" = "#3CA8BCFF", "coastal" = "#EF8A0CFF")) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "",
    fill = ""
  ) +
  scale_x_continuous(breaks = c(0, 1, 2)) +  # Customize x-axis values
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.position = "none",
 #   axis.text.y = element_blank(),  # Remove y-axis labels
    axis.ticks.y = element_blank(), # Remove y-axis ticks
    strip.text = element_blank()    # Remove facet plot titles
  )
cum
ggsave(cum, filename=here("impact/figs/coastal_vs_offshore_cumulative_impact.png"), width=3, height =1.5, units = c("in"))

```

Now get pressure impacts data together by summarizing pressure impacts across habitats.

```{r}
check <- list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", full=TRUE)
file.remove(check)


# 1. summarize pressure rasters across habitats
summarize_pressures <- function(scenario ="ssp245", time = "medium-term"){
  # scenario = "ssp245"
  # time = "medium-term"
  
files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure_habitat_corrected/", full=TRUE)

pressure_list <- data.frame(pressures=c("sst-extremes|temperature", "slr", "_oxygen-", "_oa_", "nutrient", "npp", "light-pollution", "fisheries", "coastal-disturbance", "air-heat-index"),
                            pressure_name = c("ocean-warming", "sea-level-rise", "reduced-oxygen", "acidification", "nutrient-pollution", "primary-productivity", "light-pollution",
                                              "fisheries-biomass-loss", "coastal-disturbance", "air-heat-index"))

  for(pressure in pressure_list$pressures){
    #pressure = pressure_list$pressures[9]
    pressure_files <- grep(scenario, files, value=TRUE)
    pressure_files <- grep(time, pressure_files, value=TRUE)
    pressure_files <- grep(pressure, pressure_files, value=TRUE)
    cat(pressure, "n = ", length(pressure_files), "\n")
    
    pressure_impact_rasts <- rast(pressure_files)
    pressure_impact_sum <- sum(pressure_impact_rasts, na.rm=TRUE)
   
    name_raster <- paste0(pressure_list$pressure_name[pressure_list$pressures==pressure],"_",
                          scenario, "_", time, ".tif")
    writeRaster(pressure_impact_sum, 
                paste0("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", name_raster),
                overwrite=TRUE) 
  }
}

summarize_pressures(scenario ="ssp245", time = "medium-term")
summarize_pressures(scenario ="ssp245", time = "current")
summarize_pressures(scenario ="ssp585", time = "medium-term")
summarize_pressures(scenario ="ssp585", time = "current")

```

Now get the data for each pressure and plot by region
```{r}

list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/")

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", full=TRUE, pattern="ssp245")
pressure_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(pressure_stack)))
names(pressure_stack) <- source_info


offshore2 <- ifel(offshore>=1, 2, 0)
offshore2 <- ifel(is.na(offshore2), 0, offshore2)
coast2 <- ifel(is.na(coast), 0, coast)
habitats <- offshore2 + coast2
habitats <- ifel(habitats == 0, NA, habitats)

coast_area <- global(coast, "sum", na.rm=TRUE) %>% pull()
offshore_area <- global(offshore, "sum", na.rm=TRUE) %>% pull()

area_df <- data.frame(region = c("coastal", "offshore"),
                      area = c(coast_area, offshore_area))

  # Compute the average pressure impact for each region
  zonal_sum <- zonal(pressure_stack, habitats, fun = "sum", na.rm=TRUE)


avg_pressure_impacts <- zonal_sum %>%
 pivot_longer(cols=-ocean_area_mol, names_to = "pressure_impact", values_to="value") %>%
separate(
    col = pressure_impact,
    into = c("pressure", "scenario", "time_period"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(region = ifelse(ocean_area_mol == 1, "coastal", "offshore")) %>%
  mutate(pressure = gsub("-", " ", pressure)) %>%
  left_join(area_df, by="region") %>%
  mutate(pressure_impact = value/area) %>%
  select(pressure, scenario, time_period, region, pressure_impact)

# Calculate average impact for each pressure
average_impacts <- avg_pressure_impacts %>%
  group_by(pressure) %>%
  summarize(avg_impact = mean(pressure_impact, na.rm = TRUE)) %>%
  arrange(avg_impact)


# Prepare data
df <- avg_pressure_impacts %>%
  mutate(
    pressure = factor(pressure, levels = average_impacts$pressure),
    region = factor(region), 
    time_period = factor(time_period, levels = c("current", "medium-term"))
  )

# Reshape data to wide format
df_wide <- df %>%
  pivot_wider(
    id_cols = c(pressure, region),
    names_from = time_period,
    values_from = pressure_impact
  ) %>%
  arrange(pressure)


# Stats for Table S3
df_wide %>% filter(region == "offshore") %>%
  mutate(dif = `medium-term` - current)
df_wide %>% filter(region == "coastal") %>%
  mutate(dif = `medium-term` - current)

# Create combined factor and adjust positions
df_wide <- df_wide %>%
  mutate(
    pressure_position = as.numeric(factor(pressure, levels = levels(df$pressure))),
    y_position = pressure_position + ifelse(region == "coastal", -0.1, 0.1),
    delta = `medium-term` - current,
    offset = delta * 0.05,  # Adjust this fraction as needed
    x_start = current + sign(delta) * offset,
    x_end = `medium-term` - sign(delta) * offset
  )


# Create the plot
coast_compare <- ggplot() +
  geom_segment(
    data = df_wide,
    aes(
      x = x_start,
      xend = x_end,
      y = y_position,
      yend = y_position,
      color = region
    ),
    arrow = arrow(length = unit(0.2, "cm")),
    linewidth = 0.5,
    alpha = 0.5
  ) +
  geom_point(
    data = df_wide,
    aes(x = current, y = y_position, shape = "Current", color = region),
    size = 2.5
  ) +
  geom_point(
    data = df_wide,
    aes(x = `medium-term`, y = y_position, shape = "Medium-Term", color = region),
    size = 2.5
  ) +
  scale_y_continuous(
    breaks = unique(df_wide$pressure_position),
    labels = levels(df$pressure)
  ) +
  scale_shape_manual(
    name = "",  # Legend title
    values = c("Current" = 16, "Medium-Term" = 17),
    labels = c("Current" = "current", "Medium-Term" = "future")  # Customized labels
  ) +
  scale_color_manual(
    name = "",
    values = c("offshore" = "#3CA8BCFF", "coastal" = "#EF8A0CFF")
  ) +
  theme_minimal() +
  labs(
    title = "",
    x = "Impact",
    y = ""
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.position = "none"  )
coast_compare
ggsave(coast_compare, filename ="impact/figs/coastal_vs_offshore_pressure_impact.png", width=6, height =3, units = c("in"))

combined_plot <- cowplot::plot_grid(cum, coast_compare, ncol = 1,  rel_heights = c(1, 2))  # Use ncol = 1 for stacking
combined_plot
ggsave(combined_plot, filename ="impact/figs/coastal_vs_offshore.png", width=5, height =5, units = c("in"))


```

2nd part of stats for table S3
```{r}

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", full=TRUE, pattern="ssp585")
pressure_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(pressure_stack)))
names(pressure_stack) <- source_info


offshore2 <- ifel(offshore>=1, 2, 0)
offshore2 <- ifel(is.na(offshore2), 0, offshore2)
coast2 <- ifel(is.na(coast), 0, coast)
habitats <- offshore2 + coast2
habitats <- ifel(habitats == 0, NA, habitats)

coast_area <- global(coast, "sum", na.rm=TRUE) %>% pull()
offshore_area <- global(offshore, "sum", na.rm=TRUE) %>% pull()

area_df <- data.frame(region = c("coastal", "offshore"),
                      area = c(coast_area, offshore_area))

  # Compute the average pressure impact for each region
  zonal_sum <- zonal(pressure_stack, habitats, fun = "sum", na.rm=TRUE)


avg_pressure_impacts <- zonal_sum %>%
 pivot_longer(cols=-ocean_area_mol, names_to = "pressure_impact", values_to="value") %>%
separate(
    col = pressure_impact,
    into = c("pressure", "scenario", "time_period"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(region = ifelse(ocean_area_mol == 1, "coastal", "offshore")) %>%
  mutate(pressure = gsub("-", " ", pressure)) %>%
  left_join(area_df, by="region") %>%
  mutate(pressure_impact = value/area) %>%
  select(pressure, scenario, time_period, region, pressure_impact)

# Calculate average impact for each pressure
average_impacts <- avg_pressure_impacts %>%
  group_by(pressure) %>%
  summarize(avg_impact = mean(pressure_impact, na.rm = TRUE)) %>%
  arrange(avg_impact)


# Prepare data
df <- avg_pressure_impacts %>%
  mutate(
    pressure = factor(pressure, levels = average_impacts$pressure),
    region = factor(region), 
    time_period = factor(time_period, levels = c("current", "medium-term"))
  )

# Reshape data to wide format
df_wide <- df %>%
  pivot_wider(
    id_cols = c(pressure, region),
    names_from = time_period,
    values_from = pressure_impact
  ) %>%
  arrange(pressure)

df_wide %>% filter(region == "offshore") %>%
  mutate(dif = `medium-term` - current)
df_wide %>% filter(region == "coastal") %>%
  mutate(dif = `medium-term` - current)



```


## Figure 3 Habitat comparison

Get habitat impact data
```{r}

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

hab_impacts_sum <- global(hab_impacts_stack, "sum", na.rm=TRUE) 

hab_impacts <- hab_impacts_sum %>%
  rename(impact = sum) %>%
  mutate(hab_vul_scen = rownames(hab_impacts_sum)) %>%
    separate(
    col = hab_vul_scen,           
    into = c("habitat", "pressure", "scenario", "time_frame"),  
    sep = "_",                  
    remove = TRUE,                
    extra = "merge",              
    fill = "right"                
  )
rownames(hab_impacts) <- NULL

# join with names and condense benthic and surface oxygen and benthic and surface temperatures
habitats <- read_csv(here("impact/habitat_list.csv"))
pressures <- read_csv(here("impact/pressure_list.csv"))
setdiff(hab_impacts$habitat, habitats$file_name)
setdiff(habitats$file_name, hab_impacts$habitat)
setdiff(hab_impacts$pressure, pressures$pressure)
setdiff(pressures$pressure, hab_impacts$pressure)

hab_impacts <- hab_impacts %>%
  full_join(pressures, by=c("pressure")) %>%
  group_by(habitat, plot_name_pressure, scenario, time_frame, color) %>%
  summarize(impact = sum(impact, na.rm=TRUE)) %>%
  ungroup() %>%
  full_join(habitats, by=c("habitat" = "file_name"))

# get habitat areas
habitats <- list.files(here("habitats/data"), full=TRUE)
hab_stack <- rast(habitats)
source_info <- gsub(".tif", "", basename(sources(hab_stack)))
names(hab_stack) <- source_info

hab_area <- global(hab_stack, "sum", na.rm=TRUE)  
hab_area$habitat <- rownames(hab_area)
rownames(hab_area) <- NULL
hab_area <- hab_area %>%
  rename(area=sum)


## join with impact
avg_hab_impact_tmp <- left_join(hab_impacts, hab_area, by="habitat") %>%
  mutate(avg_impact = impact/area) %>%
  mutate(scen_time = paste(scenario, time_frame, sep="_")) %>%
  select(habitat = plot_name_habitat, pressure = plot_name_pressure, scen_time, color, avg_impact)

## make data wide
avg_hab_impact <- avg_hab_impact_tmp %>%
  pivot_wider(
    names_from = scen_time,
    values_from = avg_impact,
    values_fill = list(avg_impact = 0)) %>%
  mutate(ssp245_change = `ssp245_medium-term`-ssp245_current,
         ssp585_change = `ssp585_medium-term`-ssp585_current)

filter(avg_hab_impact, ssp245_change <0)

### prepare supplemental data
dataS1 <- avg_hab_impact %>%
  select(-color, -`ssp245_near-term`, -`ssp585_near-term`)
write_csv(dataS1, "impact/data/dataS1.csv")
### 

# Step 1: Summarize the data to get total values for each pressure within each habitat
pressure_order <- avg_hab_impact %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`ssp245_change`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))


habitat_order <- avg_hab_impact %>%
  group_by(habitat) %>%
  summarize(total_value = sum(`ssp245_change`), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat)))

 # Step 2: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp245_medium-term`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/habitat_impact_future_barplot.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


# get the change
 # Step 2b: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp245_change`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/habitat_impact_change_barplot.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


```

Make the second part of the plot that shows overall change in cumulative pressure.
```{r}
cum_impact_habitat <- avg_hab_impact_tmp %>%
  group_by(habitat, scen_time) %>%
  summarize(avg_impact = sum(avg_impact, na.rm=TRUE))

# Reshape data to wide format
cum_wide <- cum_impact_habitat %>%
  pivot_wider(
    id_cols = c(habitat),
    names_from = scen_time,
    values_from = avg_impact
  ) 

min(cum_wide$"ssp245_medium-term")
max(cum_wide$"ssp245_medium-term")
1.32/0.33
#1.29/0.34


# Create combined factor and adjust positions
cum_wide <- cum_wide %>%
  mutate(
    habitat = factor(habitat, levels = unique(habitat_order$habitat)),
    habitat_position = as.numeric(factor(habitat, levels = unique(habitat_order$habitat))),
    y_position = habitat_position,
    delta = `ssp245_medium-term` - ssp245_current,
    offset = delta * 0.05,  # Adjust this fraction as needed
    x_start = ssp245_current + sign(delta) * offset,
    x_end = `ssp245_medium-term` - sign(delta) * offset
  )


# Create the plot
habitat_compare <- ggplot() +
  # Add segments with arrows
  geom_segment(
    data = cum_wide,
    aes(
      x = x_start,
      xend = x_end,
      y = habitat,
      yend = habitat
    ),
    arrow = arrow(length = unit(0.2, "cm")),
    size = 0.5,
    alpha = 0.5
  ) +
  # Add points for current impacts
  geom_point(
    data = cum_wide,
    aes(x = ssp245_current, y = habitat, shape = "current"),
    size = 2.5
  ) +
  # Add points for future impacts
  geom_point(
    data = cum_wide,
    aes(x = `ssp245_medium-term`, y = habitat, shape = "future"),
    size = 2.5
  ) +
  theme_minimal() +
  labs(
    title = "",
    x = "Cumulative impact",
    y = ""
  ) +
  # Customize the legend title and labels
  scale_shape_manual(
    name = "",  # Set the legend title here
    values = c("current" = 16, "future" = 17)  # Use different shapes for "Current" and "Future"
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
habitat_compare
ggsave(habitat_compare, filename ="impact/figs/habitat_pressure_cumulative_impact.png", width=8, height =4, units = c("in"))

```

## Figure 4:  Zonal region pressure impacts
```{r}

pressures_correct <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure_habitat_corrected/", full=TRUE)

hab_impacts_stack_correct <- rast(pressures_correct)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack_correct)))
names(hab_impacts_stack_correct) <- source_info


## country coastal raster
eez <- rast(here("_spatial/eez_mol_w_land.tif"))
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(coast>0, 1, 0)
#get the area
country_coast <- eez*coast
plot(country_coast)

## EEZ data
eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region


result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(hab_impacts_stack_correct)) { # i = 1
  # Extract the current layer
  current_layer <- hab_impacts_stack_correct[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, country_coast, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
  cat(i, "\n")
}

# Merge all data frames by region ID
eez_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

eez_impact_result <- eez_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")

eez_impact_result <- eez_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

write_csv(eez_impact_result, here("impact/data/country_eez_habxpressurexscenario_impact.csv"))


pressure_country <- read_csv(here("impact/data/country_eez_habxpressurexscenario_impact.csv"))


pressure_rgn <- pressure_country %>%
  separate(
    col = impact,
    into = c("habitat", "pressure", "scenario", "time_frame"),
    sep = "_",
    remove = FALSE,
    extra = "merge",
    fill = "right"
  )

## summarize each pressure (across all habitats) for each region
pressure_rgn <- pressure_rgn %>%
  group_by(pressure, scenario, time_frame, rgn_id, geosubrgn) %>%
  summarize(value = sum(value, na.rm=TRUE))

# join with names and condense benthic and surface oxygen and benthic and surface temperatures
pressures <- read_csv(here("impact/pressure_list.csv"))
setdiff(pressure_rgn$pressure, pressures$pressure)
setdiff(pressures$pressure, pressure_rgn$pressure)

pressure_rgn <- pressure_rgn %>%
  full_join(pressures, by=c("pressure")) %>%
  group_by(geosubrgn, rgn_id, plot_name_pressure, scenario, time_frame, color) %>%
  summarize(impact = sum(value, na.rm=TRUE)) %>%
  ungroup() 



# get number of coastal cells in each country:
 coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

pressure_rgn <- pressure_rgn %>%
  left_join(coast_cell_num, by="rgn_id")

### Save supplement of coastal data for supplmenet
rgn_names <- read_csv("_spatial/eez_rgn_names/rgn_names_ohi.csv") %>%
  select(rgn_id, rgn_label) %>%
  unique() # guam and northern mariana islands have same ID


datas4 <- left_join(pressure_rgn, rgn_names, by= "rgn_id") %>%
  filter(time_frame %in% c("current", "medium-term")) %>%
  mutate(average_impact = impact/cells) %>%
  select(region = rgn_label, scenario, time_frame, pressure = plot_name_pressure, average_impact)

filter(datas4, pressure == "fisheries biomass loss", region %in% c("New Caledonia", "Vanuatu",
                                                                   "Soloman", "Papua New Guinae", "Fiji"))

write_csv(datas4, here("impact/data/dataS4.csv"))
#####

 pressure_georgn <- pressure_rgn %>%
   group_by(plot_name_pressure, scenario, time_frame, geosubrgn, color) %>%
   summarize(value = sum(impact, na.rm=TRUE),
             area = sum(cells, na.rm=TRUE)) %>%
   ungroup() %>%
  mutate(avg_impact = value/area) %>%
  mutate(scen_time = paste(scenario, time_frame, sep="_")) %>%
  select(georgn = geosubrgn, pressure = plot_name_pressure, scen_time, color, avg_impact)

## make data wide
avg_georgn_impact <- pressure_georgn %>%
  pivot_wider(
    names_from = scen_time,
    values_from = avg_impact,
    values_fill = list(avg_impact = 0)) %>%
  mutate(ssp245_change = `ssp245_medium-term`-ssp245_current,
         ssp585_change = `ssp585_medium-term`-ssp585_current)


### prepare supplemental data
dataS2 <- avg_georgn_impact %>%
  select(-color, -`ssp245_near-term`, -`ssp585_near-term`)
write_csv(dataS2, "impact/data/dataS2.csv")
### 

# paper statistics
filter(avg_georgn_impact, ssp245_change<0)


# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- avg_georgn_impact %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`ssp245_change`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))


georgn_order <- avg_georgn_impact %>%
  group_by(georgn) %>%
  summarize(total_value = sum(`ssp245_change`), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(georgn = factor(georgn, levels = unique(georgn)))

 # Step 2: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_georgn_impact %>%
  select(georgn, pressure, impact =`ssp245_medium-term`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(georgn = factor(georgn, levels = unique(georgn_order$georgn))) %>%
  arrange(georgn, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = georgn, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/subregion_impact_future_barplot.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


# get the change
 # Step 2b: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_georgn_impact %>%
  select(georgn, pressure, impact =`ssp245_change`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(georgn = factor(georgn, levels = unique(georgn_order$georgn))) %>%
  arrange(georgn, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = georgn, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/subregion_impact_change_barplot.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


```

Make the second part of the plot that shows overall change in cumulative pressure.
```{r}
cum_impact_habitat <- pressure_georgn %>%
  group_by(georgn, scen_time) %>%
  summarize(avg_impact = sum(avg_impact, na.rm=TRUE))

# Reshape data to wide format
cum_wide <- cum_impact_habitat %>%
  pivot_wider(
    id_cols = c(georgn),
    names_from = scen_time,
    values_from = avg_impact
  ) 

# Create combined factor and adjust positions
cum_wide <- cum_wide %>%
  mutate(
    georgn = factor(georgn, levels = unique(georgn_order$georgn)),
    georgn_position = as.numeric(factor(georgn, levels = unique(georgn_order$georgn))),
    y_position = georgn_position,
    delta = `ssp245_medium-term` - ssp245_current,
    offset = delta * 0.05,  # Adjust this fraction as needed
    x_start = ssp245_current + sign(delta) * offset,
    x_end = `ssp245_medium-term` - sign(delta) * offset
  )


# Create the plot
georgn_compare <- ggplot() +
  # Add segments with arrows
  geom_segment(
    data = cum_wide,
    aes(
      x = x_start,
      xend = x_end,
      y = georgn,
      yend = georgn
    ),
    arrow = arrow(length = unit(0.2, "cm")),
    size = 0.5,
    alpha = 0.5
  ) +
  # Add points for current impacts
  geom_point(
    data = cum_wide,
    aes(x = ssp245_current, y = georgn, shape = "current"),
    size = 2.5
  ) +
  # Add points for future impacts
  geom_point(
    data = cum_wide,
    aes(x = `ssp245_medium-term`, y = georgn, shape = "future"),
    size = 2.5
  ) +
  theme_minimal() +
  labs(
    title = "",
    x = "Cumulative impact",
    y = ""
  ) +
  # Customize the legend title and labels
  scale_shape_manual(
    name = "",  # Set the legend title here
    values = c("current" = 16, "future" = 17)  # Use different shapes for "Current" and "Future"
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
georgn_compare
ggsave(georgn_compare, filename ="impact/figs/georegion_pressure_cumulative_impact.png", width=8, height =4, units = c("in"))

```

 

## Figure 5: 

cumulative impact: country EEZ summaries (don't end up using this, go with coastal area instead)
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))

ocean_cells<- rast(here("_spatial/ocean_area_mol.tif"))
ocean_area <- ifel(ocean_cells>0, 1, 0)


#get the area
eez <- eez*ocean_area
plot(eez)

eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region

impact_files <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

impact_stack <- rast(impact_files)
source_info <- gsub(".tif", "", basename(sources(impact_stack)))
names(impact_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(impact_stack)) { # i = 1
  # Extract the current layer
  current_layer <- impact_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, eez, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
eez_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

eez_impact_result <- eez_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")

eez_impact_result <- eez_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
 eez_cell_num <- zonal(ifel(eez>0, 1, 0), eez, fun = "sum", na.rm=TRUE)
 names(eez_cell_num) <- c("rgn_id", "cells")

dim(eez_impact_result)
eez_impact_result <- eez_impact_result %>%
  left_join(eez_cell_num, by="rgn_id") %>%
  mutate(region = "eez") 


write_csv(eez_impact_result, here("impact/data/country_eez_impact.csv"))

```


cumulative impact: country coastal region 
```{r}

country_coast <- rast(here("_spatial/coastal_country_mol.tif"))


eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(!(iso3c %in% "GUM")) # cut out duplicated dual region

impact_files <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)

impact_stack <- rast(impact_files)
source_info <- gsub(".tif", "", basename(sources(impact_stack)))
names(impact_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(impact_stack)) { # i = 1
  # Extract the current layer
  current_layer <- impact_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, country_coast, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
coast_impact_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

coast_impact_result <- coast_impact_result %>%
 pivot_longer(cols=-rgn_id, names_to = "impact", values_to="value")


coast_impact_result <- coast_impact_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

dim(coast_impact_result)
coast_impact_result <- coast_impact_result %>%
  left_join(coast_cell_num, by="rgn_id") %>%
  mutate(region = "coast") 


```

plot dependence vs. impacts
```{r}
#from paper_statistics_v2
depend <- read_csv(here("impact/data/selig_marine_dependence.csv"))

impact_country <- coast_impact_result %>%
  mutate(avg_value = value/cells) 
#%>%
#  separate(impact, into = c("scenario", "time_period"), sep = "_")

impact_country <- impact_country %>%
  select(region, rgn_label, iso3c, impact, avg_value) %>%
    pivot_wider(
    names_from = impact,
    values_from = avg_value
  )

impact_country$iso3c <- countrycode(impact_country$rgn_label, "country.name", "iso3c") 
impact_country <- filter(impact_country, rgn_label != "Bassas da India")


double_rgns <- data.frame(rgn_label = c("Guadeloupe and Martinique", "Guadeloupe and Martinique",
              "Northern Mariana Islands and Guam", "Northern Mariana Islands and Guam",
              "Puerto Rico and Virgin Islands of the United States", "Puerto Rico and Virgin Islands of the United States"),
                 iso3c = c("GLP", "MTQ", "GUM", "MNP", "PRI", "VIR"))

coastal_double_eez <- filter(impact_country, rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")) %>%
  select(-iso3c)

coastal_double_eez2 <- left_join(double_rgns, coastal_double_eez, by="rgn_label")

impact_country <- filter(impact_country, !(rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")))

coastal_eez_impact <- rbind(impact_country, coastal_double_eez2)
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label=="Northern Saint-Martin"] <- "NSM"
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label=="Micronesia"] <- "FSM"
coastal_eez_impact$iso3c[coastal_eez_impact$rgn_label %in% c("Crozet Islands",
                                       "Kerguelen Islands",
                                       "Amsterdam Island and Saint Paul Island",
                                       "Ile Europa",
                                       "Ile Tromelin",
                                       "Juan de Nova Island",
                                       "Glorioso Islands")] <- "ATF"
coastal_eez_impact <- filter(coastal_eez_impact, iso3c != "ATF")

coastal_eez_impact2 <- coastal_eez_impact %>%
  filter(!is.na(iso3c)) %>%
  select(-rgn_label)


missing_iso3c <- setdiff(depend$iso3c, coastal_eez_impact2$iso3c)
filter(depend, iso3c %in% missing_iso3c) %>% print(n=25)
setdiff(coastal_eez_impact2$iso3c, depend$iso3c)

coastal_compare <- coastal_eez_impact2 %>%
  filter(region=="coast") %>%
  left_join(depend, by="iso3c") %>%
  filter(!is.na(ssp585_current))

coastal_compare <- coastal_compare %>%
  mutate(change_ssp245_medium = `ssp245_medium-term` - ssp245_current) %>%
    mutate(change_ssp585_medium = `ssp585_medium-term` - ssp585_current) %>%
  select(region, iso3c, ssp245_current, `ssp245_medium-term`, change_ssp245_medium, ssp585_current, `ssp585_medium-term`, change_ssp585_medium, marine_dependence = integrated_depend)

rgn_names <- read_csv("_spatial/eez_rgn_names/rgn_names_ohi.csv") %>%
  select(rgn_label, iso3c)

coastal_compare <- left_join(coastal_compare, rgn_names, by="iso3c") %>%
  select(region, iso3c, region_name = rgn_label,  ssp245_current, `ssp245_medium-term`, change_ssp245_medium, ssp585_current, `ssp585_medium-term`, change_ssp585_medium, marine_dependence)


write_csv(coastal_compare, here("impact/data/country_coast_cumulative_impact.csv"))
write_csv(coastal_compare, here("impact/data/dataS3.csv"))


coastal_compare <- read_csv(here("impact/data/country_coast_cumulative_impact.csv")) 

quants <- c(-Inf, quantile(coastal_compare$marine_dependence, c(0.2, 0.4, 0.6, 0.8), na.rm=TRUE), Inf)

coastal_compare <- coastal_compare %>%
  mutate(
    integrated_depend_category = cut(
     marine_dependence,
      breaks = quants,
      labels = c("1", "2", "3", "4", "5"),
      include.lowest = TRUE
    )
  )

p <- ggplot(data=coastal_compare , aes(y=`ssp245_medium-term`, x=change_ssp245_medium)) +
  geom_point(aes(text=region_name), alpha=0.7, size = 1.5, color="gray") +
  geom_vline(xintercept = median(coastal_compare$change_ssp245_medium, na.rm=TRUE), linetype="dashed", color="orange") +
   geom_hline(yintercept = median(coastal_compare$`ssp245_medium-term`, na.rm=TRUE), linetype="dashed", color="orange") +
  theme_classic()  +
  theme(
    # Add major gridlines
    panel.grid.major = element_line(color = "lightgray", size = 0.1),
    # Remove minor gridlines
    panel.grid.minor = element_blank(),
    legend.position = "none"
  ) +
  labs(y="Future impact", x = "\u0394 impact")
p
ggsave(here("impact/figs/current_future_scatterplot.png"), plot = p, width = 5, height = 3, dpi = 300)


## color version, not currently used
p <- ggplot(data=coastal_compare , aes(y=`ssp245_medium-term`, x=change_ssp245_medium, color=integrated_depend_category)) +
  geom_point(aes(text=iso3c), alpha=0.7, size = 1.5) +
  scale_color_manual(values = c("1" = "darkgray", "2" = "darkgray", "3" = "darkgray", "4"="orange", "5"= "red")) + 
  geom_vline(xintercept = median(coastal_compare$change_ssp245_medium, na.rm=TRUE), linetype="dashed", color="orange") +
   geom_hline(yintercept = median(coastal_compare$`ssp245_medium-term`, na.rm=TRUE), linetype="dashed", color="orange") +
  theme_classic()  +
  theme(
    # Add major gridlines
    panel.grid.major = element_line(color = "lightgray", size = 0.1),
    # Remove minor gridlines
    panel.grid.minor = element_blank(),
    legend.position = "none"
  ) +
  labs(y="Future impact", x = "\u0394 impact")
ggsave(here("impact/figs/current_future_scatterplot.png"), plot = p, width = 5, height = 3, dpi = 300)

p
pp <- plotly::ggplotly(p, tooltip = "text")
pp


p <- ggplot(data=coastal_compare , aes(y=`marine_dependence`, x=change_ssp245_medium, color=`integrated_depend_category`)) +
  geom_point(aes(text=region_name), alpha=0.7, size = 1.5) +
    geom_smooth(method = "lm", se = TRUE, color = "darkgray") +  # Add linear regression line
  scale_color_manual(values = c("1" = "gray", "2" = "gray", "3"="gray", "4"="gray", "5"= "gray")) +
  theme_classic()  +
  theme(
    # Add major gridlines
    panel.grid.major = element_line(color = "lightgray", size = 0.1),
    # Remove minor gridlines
    panel.grid.minor = element_blank(),
    legend.position = "none"
  ) +
  labs(y="Marine dependence", x = "\u0394 impact")

p
ggsave(here("impact/figs/current_depend_scatterplot.png"), plot = p, width = 5, height = 3, dpi = 300)




pp <- plotly::ggplotly(p, tooltip = "text")
pp
library(htmlwidgets)
saveWidget(pp, "interactive_plot.html", selfcontained = TRUE)

mod <- lm(change_ssp245_medium ~ marine_dependence, data=coastal_compare)
summary(mod)
mod <- lm(`ssp245_medium-term` ~ marine_dependence, data=coastal_compare)
summary(mod)

```


# Fisheries model comparisons
```{r}
v1 <- rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure//fisheries_ssp2_medium-term.tif")
v2 <- rast("/home/shares/ohi/stressors_2021/_dataprep/fisheries_biomass/pressure_mod2//fisheries_ssp2_medium-term.tif")

result = v1-v2
color_gradient <- colorRampPalette(c("lightgray", "orange", "red"))

# Plot without boxes and with the desired gradient
plot(result, col = color_gradient(100), box = FALSE, axes = FALSE, main = "")



```
