---
title: "Calculating impacts"
output: html_document
date: "2024-09-16"
---

```{r setup, include=FALSE}
library(tidyverse)
library(terra)
library(here)
library(countrycode)

```

```{r}
vulnerability <- read_csv(here("impact/vulnerability_weighting_matrix.csv")) %>%
  mutate_if(is.numeric, ~ . / 4)

habitat_files <- list.files(here("habitats/data"), full=TRUE)

setdiff(names(vulnerability), gsub(".tif", "", basename(habitat_files))) # "pressure" ok!
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

pressure_files <- list.files("/home/shares/ohi/stressors_2021/_final_data/rescaled_pressures", full=TRUE)
setdiff(vulnerability$pressure, gsub(".tif", "", sub("_.*", "", basename(pressure_files))))
setdiff(gsub(".tif", "", basename(habitat_files)), names(vulnerability))

```

# combos
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
file.remove(clean)

for(habitat in names(vulnerability)[-1]){
  #habitat = names(vulnerability)[2]
  hab_file <- grep(habitat, habitat_files, value=TRUE)
  hab_raster <- rast(hab_file)
  cat(habitat, "\n")
  
  for(pressure in vulnerability$pressure){
  # pressure = vulnerability$pressure[7]
  # pressure = "sst-extremes"
  pressure_filter <- paste0(pressure, "_")
  p_files <- grep(pressure_filter, pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(pressure, " n = ", length(p_files), "\n")
  
  vulnerability_value <- vulnerability %>%
    filter(pressure == !!pressure) %>%
    select(habitat) %>%
    pull()
  
  pressure_vulnerability <- pressure_rasts * vulnerability_value
  
  pressure_vulnerability_habitat <- pressure_vulnerability * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/"
  p_names <- basename(gsub(paste0(pressure, "_"), "", p_files))
  output_filenames <- sprintf("%s%s_%s_%s", save_loc, habitat, pressure, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}
}

```

Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
plot(rast(check[652]))

length(check)
 #should be 4,620
habitats <- sub("_.*", "", basename(check))
table(habitats) # should be the same
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", basename(check))
table(pressure)

```

## adjust air temperature impact
The impact is a little different for salt-marsh, and seagrass
```{r}

combo_files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)

heatindex_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "salt-marsh"
#lower_threshold <- 0.2
#upper_threshold <- 0.25
combo_files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)  
  
hi_files <- grep(sprintf("%s_air-heat-index", habitat), combo_files, value=TRUE)
file.remove(hi_files)
hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)
  
  p_files <- grep("air-heat-index", pressure_files, value=TRUE)
  pressure_rasts <- rast(p_files)
  cat(" n = ", length(p_files), "\n")
  
  p <- pressure_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/"
  p_names <- basename(p_files)
  output_filenames <- sprintf("%s%s_%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

heatindex_modify(habitat="salt-marsh", lower_threshold=0.2, upper_threshold=0.25) 
heatindex_modify(habitat="seagrass", lower_threshold=0.2, upper_threshold=0.25) 

```


## adjust SST pressure
The pressure is a little different for kelp and coral because there is more information about them.
First prepare the raw sst data. These impacts are based on absolute temperatures.
```{r}

# get the temperature data

sst_raw_prep <- function(scenario){
#scenario <- "ssp245"
sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average_final/", full=TRUE)
scenario_sst_raw_files <- grep(scenario, sst_raw_files, value=TRUE)

current_files <- grep("2015_2019", scenario_sst_raw_files, value=TRUE)
current_stack <- rast(current_files)
current_mean <- mean(current_stack) - 273.15

near_files <- grep("2020_2024|2025_2029|2030_2034|2035_2039", scenario_sst_raw_files, value=TRUE)
near_stack <- rast(near_files)
near_mean <- mean(near_stack) - 273.15

medium_files <- grep("2040_2044|2045_2049|2050_2054|2055_2059", scenario_sst_raw_files, value=TRUE)
medium_stack <- rast(medium_files)
medium_mean <- mean(medium_stack) - 273.15

long_files <- grep("2080_2084|2085_2089|2090_2094|2095_2099", scenario_sst_raw_files, value=TRUE)
long_stack <- rast(long_files)
long_mean <- mean(long_stack) - 273.15

writeRaster(current_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_current.tif", scenario), overwrite=TRUE)
writeRaster(near_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_near-term.tif", scenario), overwrite=TRUE)
writeRaster(medium_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_medium-term.tif", scenario), overwrite=TRUE)
writeRaster(long_mean, sprintf("/home/shares/ohi/stressors_2021/impacts/raw_sst/sst-raw_%s_long-term.tif", scenario), overwrite=TRUE)
}

sst_raw_prep(scenario="ssp245")
sst_raw_prep(scenario="ssp585")

```


Now apply to coral, kelp, and mangroves.
```{r}

sst_modify <- function(habitat, lower_threshold, upper_threshold){ 
#habitat <- "coral"
#lower_threshold <- 18.3
#upper_threshold <- 24
combo_files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)  
sst_files <- grep(sprintf("%s_sst-extremes", habitat), combo_files, value=TRUE)
file.remove(sst_files)

hab_file <- grep(habitat, habitat_files, value=TRUE)
hab_raster <- rast(hab_file)

sst_raw_files <- list.files("/home/shares/ohi/stressors_2021/impacts/raw_sst", full=TRUE)    
sst_raw_rasts <- rast(sst_raw_files)
cat(" n = ", length(sst_raw_files), "\n")

## habitat <- ifel(hab_raster>0, 1, 0)  
## sst_curr <- rast(sst_raw_files[1])
## current_kelp <- habitat*sst_curr
## sst_future <- rast(sst_raw_files[7])
## future_kelp <- habitat*sst_future
##tmp <- c(current_kelp, future_kelp)
##tmp_df <- data.frame(tmp)
##plot(tmp_df$kelp, tmp_df$kelp.1, pch=16, col = rgb(0.2, 0.2, 1, 0.05), cex=0.6, ##xlab="current_temp", ylab="ssp585-2050")
##abline(0,1, col="black")
##abline(v=18.3, col="orange")
##abline(h=18.3, col="orange")
##abline(v=24, col="red")
##abline(h=24, col="red")

  p <- sst_raw_rasts
  p <- ifel(p <= lower_threshold, 0, p)
  p <- ifel(p >= upper_threshold, 1, p)
  p <- ifel(p > lower_threshold & p < upper_threshold, (p - lower_threshold) / (upper_threshold - lower_threshold), p)
  
  pressure_vulnerability_habitat <- p * hab_raster
  
  save_loc <- "/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/"
  p_names <- gsub("sst-raw", "", basename(sst_raw_files))
  output_filenames <- sprintf("%s%s_sst-extremes%s", save_loc, habitat, p_names)
writeRaster(pressure_vulnerability_habitat, filename = output_filenames, overwrite = TRUE)

}

sst_modify(habitat="kelp", lower_threshold=18.3, upper_threshold=24) 
sst_modify(habitat="coral-reef", lower_threshold=29, upper_threshold=30)  
sst_modify(habitat="mangroves", lower_threshold=32, upper_threshold=40)  
```

## Make ssp names consistent
Some pressures do not use RCP climate models. The difference in naming (ssp2 vs. ssp245) is throwing everything off!

```{r}

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.copy(from = files_ssp1, to = gsub("ssp1_", "ssp119_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
files_ssp1 <- grep("ssp1_", files_fix, value=TRUE)
file.rename(from = files_ssp1, to = gsub("ssp1_", "ssp126_", files_ssp1))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp2 <- grep("ssp2_", files_fix, value=TRUE)
file.rename(from = files_ssp2, to = gsub("ssp2_", "ssp245_", files_ssp2))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp3 <- grep("ssp3_", files_fix, value=TRUE)
file.rename(from = files_ssp3, to = gsub("ssp3_", "ssp370_", files_ssp3))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp4 <- grep("ssp4_", files_fix, value=TRUE)
file.rename(from = files_ssp4, to = gsub("ssp4_", "ssp460_", files_ssp4))

files_fix <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
length(files_fix)
files_ssp5 <- grep("ssp5_", files_fix, value=TRUE)
file.rename(from = files_ssp5, to = gsub("ssp5_", "ssp585_", files_ssp5))


```



Do a little check.
```{r}
check <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/")
length(check)
 #should be 4,840
habitats <- sub("_.*", "", check)
table(habitats)
pressure <- sub("^[^_]*_([^_]*)_.*", "\\1", check)
table(pressure)

```


Create raster with total number of habitats in each cell.
```{r}
habitat_files <- list.files(here("habitats/data"), full=TRUE)
habitat_rast <- rast(habitat_files)
habitat_n <- sum(habitat_rast, na.rm=TRUE)
plot(habitat_n)
writeRaster(habitat_n, here("habitats/habitat_n.tif"), overwrite=TRUE)


```

# Calculate cumulative impacts
```{r}

clean <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
file.remove(clean)

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)

cum_impact <- function(scen_time){
#scen_time <- "ssp585_medium-term"
habitat_n <- rast(here("habitats/habitat_n.tif"))

pressures_filter <- grep(scen_time, pressures, value=TRUE)
cat(paste0(scen_time, ", n = ", length(pressures_filter), "\n"))
pressure_stack <- rast(pressures_filter)
sum_pressure <- sum(pressure_stack, na.rm=TRUE)
impact <- sum_pressure/habitat_n
writeRaster(impact, sprintf("/home/shares/ohi/stressors_2021/impacts/cumulative/%s.tif", scen_time), overwrite=TRUE)
}
cum_impact(scen_time = "ssp585_medium-term")
cum_impact(scen_time = "ssp585_near-term")
cum_impact(scen_time = "ssp585_current")
cum_impact(scen_time = "ssp245_medium-term")
cum_impact(scen_time = "ssp245_near-term")
cum_impact(scen_time = "ssp245_current")



```

# For individual pressures, need to create rasters that correct for number of habitats
For regional summaries of pressure impacts, the impact of the pressure will depend on the number of habitats. 
First step is to correct for the number of habitats (being sure to divide by the pressure). This is used for geographic summaries of pressures (but not habitat based summaries.).

```{r}

check <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure_habitat_corrected/", full=TRUE)
file.remove(check)

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

## have to correct the pressures by the number of habitats
habitat_n <- rast(here("habitats/habitat_n.tif"))

for(i in 1: nlyr(hab_impacts_stack)){ #i=1
current_layer <- hab_impacts_stack[[i]]

  # Perform the division
  corrected_layer <- current_layer / habitat_n

  # Define the output filename
  layer_name <- names(hab_impacts_stack)[i]

  # Write the output to disk
  writeRaster(corrected_layer, 
              filename = sprintf("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure_habitat_corrected/%s.tif", layer_name), 
              overwrite = TRUE)

  # Optional: Print progress
  cat("Processed and saved:", layer_name, "\n")
}

```


