---
title: "supplement_figs"
output: html_document
date: "2024-11-14"
---

### Figure 1
Maps of cumulative impact and change in cumulative impact
```{r}

library(paletteer)
library(png)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(tidyverse)
library(terra)
library(here)
library(countrycode)

```

Get land boundaries

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
# Define the Mollweide projection string
moll_crs <- "+proj=moll"

# Reproject the world data to Mollweide
world_moll <- st_transform(world, crs = moll_crs)

```



Map of decreased areas
```{r}

ssp2_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp245_current.tif") 

decrease <- ifel(ssp2_med <= 0, 1, NA)
ocean <- rast(here("_spatial/ocean_area_mol.tif"))
area <- decrease*ocean
global(area, "sum", na.rm=TRUE)/global(ocean, "sum", na.rm=TRUE)

decrease_df <- as.data.frame(decrease, xy=TRUE)
ocean2 <- ifel(ocean>0, 1, NA)
# Remove borders and labels
par(mar = c(0, 0, 0, 0), xaxt = 'n', yaxt = 'n', bty = 'n')
plot(ocean2, col="lightblue", axes=FALSE, box=FALSE, legend=FALSE )
points(decrease_df$x, decrease_df$y, col="red", pch=16, cex=0.2)


ssp5_med <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_medium-term.tif")- 
            rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_current.tif") 

decrease <- ifel(ssp5_med <= 0, 1, NA)

decrease_df <- as.data.frame(decrease, xy=TRUE)
# Remove borders and labels
par(mar = c(0, 0, 0, 0), xaxt = 'n', yaxt = 'n', bty = 'n')
plot(ocean2, col="lightblue", axes=FALSE, box=FALSE, legend=FALSE )
points(decrease_df$x, decrease_df$y, col="red", pch=16, cex=0.2)


# determine what is happening in these regions
pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", full=TRUE, pattern="ssp245") # created below
pressure_stack <- rast(pressures)
names(pressure_stack) <- basename(sources(pressure_stack))
tmp <- zonal(pressure_stack, decrease, fun="mean", na.rm=TRUE) %>%
  pivot_longer(cols=-sum,
               names_to="variable",
               values_to = "value") %>%
  separate(col=variable, into=c("pressure", "scenario", "timeframe"),
           sep="_")



```

Some figures made in impacts.Rmd for SSP5-8.5 scenarios. They are in impact/figs/cumulative

histogram and statistics for latitudinal patterns
```{r}

tmp <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE, pattern = "current|medium-term")
stack <- rast(tmp)
new_names <- gsub(".tif", "", basename(tmp))
names(stack) <- new_names

global_df <- as.data.frame(stack, xy=TRUE) %>%
  pivot_longer(
    cols = starts_with("ssp"),  
    names_to = "scenario",      
    values_to = "value"         
  ) %>%
  group_by(y, scenario) %>%
  summarize(n = n(),
            sum_impact = sum(value)) %>%
  ungroup() %>%
  mutate(avg_impact = sum_impact/n) %>%
  select(y, scenario, avg_impact) 


plot_data <- global_df %>%
  filter(scenario %in% c("ssp585_current", "ssp585_medium-term"))

hist <- ggplot(plot_data, aes(x = y, y = avg_impact, fill = scenario)) +
  geom_area(
    data = filter(plot_data, scenario == "ssp585_current"),
    aes(x = y, y = avg_impact, fill = "ssp585_current"),
    alpha = 0.5, color = NA
  ) +
  geom_area(
    data = filter(plot_data, scenario == "ssp585_medium-term"),
    aes(x = y, y = avg_impact, fill = "ssp585_medium-term"),
    alpha = 0.5, color = NA
  ) +
  scale_fill_manual(
    values = c("ssp585_current" = "blue", "ssp585_medium-term" = "red"),
    guide = "none"  # Remove legend
  ) +
  theme_void() +  # Removes all background, grids, and axis lines
  theme(
    plot.title = element_blank(),    # Remove title
    axis.title = element_blank(),    # Remove axis titles
    axis.text = element_blank(),     # Remove axis text
    axis.ticks = element_blank(),    # Remove axis ticks
    panel.grid = element_blank(),    # Remove grid lines
    panel.border = element_blank()   # Remove border
  )
hist
ggsave(hist, filename=here("impact/figs/latitudinal_impacts_ssp585.png"), width=5.5, height =2, units = c("in"))

```


Global plot of hotspots
```{r}

ssp5_change <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_medium-term.tif") -
               rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_current.tif")
ssp5_change_high_quant <-global(ssp5_change, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp5_change_high <- ifel(ssp5_change >= ssp5_change_high_quant, 1, 0)
ssp5_change_low_quant <-global(ssp5_change, fun = quantile, probs = 0.20, na.rm = TRUE) %>% pull()
ssp5_change_low <- ifel(ssp5_change <= ssp5_change_low_quant, 1, 0)

ssp5_current <- rast("/home/shares/ohi/stressors_2021/impacts/cumulative/ssp585_current.tif")
ssp5_current_high_quant <-global(ssp5_current, fun = quantile, probs = 0.80, na.rm = TRUE) %>% pull()
ssp5_current_high <- ifel(ssp5_current >= ssp5_current_high_quant, 1, 0)
ssp5_current_low_quant <-global(ssp5_current, fun = quantile, probs = 0.20, na.rm = TRUE) %>% pull()
ssp5_current_low <- ifel(ssp5_current <= ssp5_current_low_quant, 1, 0)

# category 1: current low and stable
cat1 <- ssp5_current_low + ssp5_change_low
cat1 <- ifel(cat1 == 2, 3, NA)
plot(cat1, col="#44ACC8")

# category 2: current high and stable
cat2 <- ssp5_current_high + ssp5_change_low
cat2 <- ifel(cat2 == 2, 4, NA)
plot(cat2, col="#E2D100", add=TRUE)

# category 3: current high and fast increase
cat3 <- ssp5_current_high + ssp5_change_high
cat3 <- ifel(cat3 == 2, 3, NA)
plot(cat3, col="#E58501", add=TRUE)

# category 4: current low and fast increase
cat4 <- ssp5_current_low + ssp5_change_high
cat4 <- ifel(cat4 == 2, 4, NA)
plot(cat4, col="#B41821", add=TRUE)

png(here("impact/figs/hotspots_ssp585.png"), width = 2000, height = 1500, res = 300)

plot(st_geometry(world_moll), col="gray80", border="gray85")
plot(cat1, col="#44ACC8", add=TRUE, legend=FALSE)
plot(cat2, col="#E2D100", add=TRUE, legend=FALSE)
plot(cat3, col="#B41821", add=TRUE, legend=FALSE)
plot(cat4, col="#E58501", add=TRUE, legend=FALSE)

dev.off()

```

## Figure 2: coastal vs ocean

Cumulative impact plots
```{r}
# relative pressures coastal vs global

## global coast
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(coast>0, 1, 0)

## offshore area only
ocean <- rast(here("_spatial/ocean_area_mol.tif"))
ocean <- ifel(ocean>0, 2, 0)
offshore <- ocean - ifel(is.na(coast), 0, coast)
offshore <- ifel(offshore ==2, 1, NA)


cumulative <- list.files("/home/shares/ohi/stressors_2021/impacts/cumulative", full=TRUE)
cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info

offshore_pressures <- cum_stack*offshore

offshore_points <- data.frame(offshore_pressures)
offshore_points$region <- "offshore"

coastal_pressures <- cum_stack*coast

coast_points <- data.frame(coastal_pressures)
coast_points$region <- "coastal"

point_data <- rbind(coast_points, offshore_points[sample(nrow(offshore_points), dim(coast_points)[1]), ])

colnames(point_data) <- gsub("\\.", "-", colnames(point_data))


point_data_long <- point_data %>%
  pivot_longer(
    cols = starts_with("ssp"),       # Columns to reshape
    names_to = c("scenario", "time_period"),  # New column names
    names_sep = "_",                 # Separator used in column names
    values_to = "value"              # Name of the values column
  )

point_data_plot <- filter(point_data_long, scenario == "ssp585", time_period %in% c("current", "medium-term"))

# Calculate the overall range of 'value' to set consistent axes
range(point_data_long$value, na.rm = TRUE)


cum <- ggplot(point_data_plot, aes(y = region, x = value, fill = region)) +
  geom_violin(trim = FALSE) +
  xlim(0, 2) + 
  facet_wrap(~ time_period, ncol = 1) +
  scale_fill_manual(values = c("offshore" = "#3CA8BCFF", "coastal" = "#EF8A0CFF")) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "",
    fill = ""
  ) +
  scale_x_continuous(breaks = c(0, 1, 2)) +  # Customize x-axis values
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.position = "none",
 #   axis.text.y = element_blank(),  # Remove y-axis labels
    axis.ticks.y = element_blank(), # Remove y-axis ticks
    strip.text = element_blank()    # Remove facet plot titles
  )
cum
ggsave(cum, filename=here("impact/figs/coastal_vs_offshore_cumulative_impact_ssp585.png"), width=3, height =1.5, units = c("in"))

```

Now get pressure impacts data together by summarizing pressure impacts across habitats.

```{r}
# 1. summarize pressure rasters across habitats
summarize_pressures <- function(scenario ="ssp585", time = "medium-term"){
  # scenario = "ssp245"
  # time = "medium-term"
  
files <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure_habitat_corrected/", full=TRUE)

pressure_list <- data.frame(pressures=c("sst-extremes|temperature", "slr", "_oxygen-", "_oa_", "nutrient", "npp", "light-pollution", "fisheries", "coastal-disturbance", "air-heat-index"),
                            pressure_name = c("ocean-warming", "sea-level-rise", "reduced-oxygen", "acidification", "nutrient-pollution", "primary-productivity", "light-pollution",
                                              "fisheries-biomass-loss", "coastal-disturbance", "air-heat-index"))

  for(pressure in pressure_list$pressures){
    #pressure = pressure_list$pressures[2]
    pressure_files <- grep(scenario, files, value=TRUE)
    pressure_files <- grep(time, pressure_files, value=TRUE)
    pressure_files <- grep(pressure, pressure_files, value=TRUE)
    cat(pressure, "n = ", length(pressure_files), "\n")
    
    pressure_impact_rasts <- rast(pressure_files)
    pressure_impact_sum <- sum(pressure_impact_rasts, na.rm=TRUE)
   
    name_raster <- paste0(pressure_list$pressure_name[pressure_list$pressures==pressure],"_",
                          scenario, "_", time, ".tif")
    writeRaster(pressure_impact_sum, 
                paste0("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", name_raster),
                overwrite=TRUE) 
  }
}

summarize_pressures(scenario ="ssp585", time = "medium-term")
summarize_pressures(scenario ="ssp585", time = "current")

```

Now get the data for each pressure and plot by region
```{r}

list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/")

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/pressure_impacts/", full=TRUE, pattern="ssp585")
pressure_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(pressure_stack)))
names(pressure_stack) <- source_info

offshore2 <- ifel(offshore>=1, 2, 0)
offshore2 <- ifel(is.na(offshore2), 0, offshore2)
coast2 <- ifel(is.na(coast), 0, coast)
habitats <- offshore2 + coast2
habitats <- ifel(habitats == 0, NA, habitats)

coast_area <- global(coast, "sum", na.rm=TRUE) %>% pull()
offshore_area <- global(offshore, "sum", na.rm=TRUE) %>% pull()

area_df <- data.frame(region = c("coastal", "offshore"),
                      area = c(coast_area, offshore_area))

  # Compute the average pressure impact for each region
  zonal_sum <- zonal(pressure_stack, habitats, fun = "sum", na.rm=TRUE)


avg_pressure_impacts <- zonal_sum %>%
 pivot_longer(cols=-ocean_area_mol, names_to = "pressure_impact", values_to="value") %>%
separate(
    col = pressure_impact,
    into = c("pressure", "scenario", "time_period"),
    sep = "_",
    remove = FALSE
  ) %>%
  mutate(region = ifelse(ocean_area_mol == 1, "coastal", "offshore")) %>%
  mutate(pressure = gsub("-", " ", pressure)) %>%
  left_join(area_df, by="region") %>%
  mutate(pressure_impact = value/area) %>%
  select(pressure, scenario, time_period, region, pressure_impact)

# Calculate average impact for each pressure
average_impacts <- avg_pressure_impacts %>%
  group_by(pressure) %>%
  summarize(avg_impact = mean(pressure_impact, na.rm = TRUE)) %>%
  arrange(avg_impact)


# Prepare data
df <- avg_pressure_impacts %>%
  mutate(
    pressure = factor(pressure, levels = average_impacts$pressure),
    region = factor(region), 
    time_period = factor(time_period, levels = c("current", "medium-term"))
  )

# Reshape data to wide format
df_wide <- df %>%
  pivot_wider(
    id_cols = c(pressure, region),
    names_from = time_period,
    values_from = pressure_impact
  ) %>%
  arrange(pressure)

# Create combined factor and adjust positions
df_wide <- df_wide %>%
  mutate(
    pressure_position = as.numeric(factor(pressure, levels = levels(df$pressure))),
    y_position = pressure_position + ifelse(region == "coastal", -0.1, 0.1),
    delta = `medium-term` - current,
    offset = delta * 0.05,  # Adjust this fraction as needed
    x_start = current + sign(delta) * offset,
    x_end = `medium-term` - sign(delta) * offset
  )


# Create the plot
coast_compare <- ggplot() +
  geom_segment(
    data = df_wide,
    aes(
      x = x_start,
      xend = x_end,
      y = y_position,
      yend = y_position,
      color = region
    ),
    arrow = arrow(length = unit(0.2, "cm")),
    linewidth = 0.5,
    alpha = 0.5
  ) +
  geom_point(
    data = df_wide,
    aes(x = current, y = y_position, shape = "Current", color = region),
    size = 2.5
  ) +
  geom_point(
    data = df_wide,
    aes(x = `medium-term`, y = y_position, shape = "Medium-Term", color = region),
    size = 2.5
  ) +
  scale_y_continuous(
    breaks = unique(df_wide$pressure_position),
    labels = levels(df$pressure)
  ) +
  scale_shape_manual(
    name = "",  # Legend title
    values = c("Current" = 16, "Medium-Term" = 17),
    labels = c("Current" = "current", "Medium-Term" = "future, 2041-2060")  # Customized labels
  ) +
  scale_color_manual(
    name = "",
    values = c("offshore" = "#3CA8BCFF", "coastal" = "#EF8A0CFF")
  ) +
  theme_minimal() +
  labs(
    title = "",
    x = "Impact",
    y = ""
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    legend.position = "none"  )
coast_compare
ggsave(coast_compare, filename ="impact/figs/coastal_vs_offshore_pressure_impact_ssp585.png", width=6, height =3, units = c("in"))

combined_plot <- cowplot::plot_grid(cum, coast_compare, ncol = 1,  rel_heights = c(1, 2))  # Use ncol = 1 for stacking
combined_plot
ggsave(combined_plot, filename ="impact/figs/coastal_vs_offshoret_ssp585.png", width=5, height =5, units = c("in"))


```

## Figure 3 Habitat comparison

Get habitat impact data
```{r}

pressures <- list.files("/home/shares/ohi/stressors_2021/impacts/hab_vuln_pressure/", full=TRUE)
pressures <- grep("ssp245|ssp585", pressures, value=TRUE)
pressures <- grep("_current|_near-term|_medium-term", pressures, value=TRUE)

hab_impacts_stack <- rast(pressures)
source_info <- gsub(".tif", "", basename(sources(hab_impacts_stack)))
names(hab_impacts_stack) <- source_info

hab_impacts_sum <- global(hab_impacts_stack, "sum", na.rm=TRUE) 

hab_impacts <- hab_impacts_sum %>%
  rename(impact = sum) %>%
  mutate(hab_vul_scen = rownames(hab_impacts_sum)) %>%
    separate(
    col = hab_vul_scen,           
    into = c("habitat", "pressure", "scenario", "time_frame"),  
    sep = "_",                  
    remove = TRUE,                
    extra = "merge",              
    fill = "right"                
  )
rownames(hab_impacts) <- NULL

# join with names and condense benthic and surface oxygen and benthic and surface temperatures
habitats <- read_csv(here("impact/habitat_list.csv"))
pressures <- read_csv(here("impact/pressure_list.csv"))
setdiff(hab_impacts$habitat, habitats$file_name)
setdiff(habitats$file_name, hab_impacts$habitat)
setdiff(hab_impacts$pressure, pressures$pressure)
setdiff(pressures$pressure, hab_impacts$pressure)

hab_impacts <- hab_impacts %>%
  full_join(pressures, by=c("pressure")) %>%
  group_by(habitat, plot_name_pressure, scenario, time_frame, color) %>%
  summarize(impact = sum(impact, na.rm=TRUE)) %>%
  ungroup() %>%
  full_join(habitats, by=c("habitat" = "file_name"))

# get habitat areas
habitats <- list.files(here("habitats/data"), full=TRUE)
hab_stack <- rast(habitats)
source_info <- gsub(".tif", "", basename(sources(hab_stack)))
names(hab_stack) <- source_info

hab_area <- global(hab_stack, "sum", na.rm=TRUE)  
hab_area$habitat <- rownames(hab_area)
rownames(hab_area) <- NULL
hab_area <- hab_area %>%
  rename(area=sum)


## join with impact
avg_hab_impact_tmp <- left_join(hab_impacts, hab_area, by="habitat") %>%
  mutate(avg_impact = impact/area) %>%
  mutate(scen_time = paste(scenario, time_frame, sep="_")) %>%
  select(habitat = plot_name_habitat, pressure = plot_name_pressure, scen_time, color, avg_impact)

## make data wide
avg_hab_impact <- avg_hab_impact_tmp %>%
  pivot_wider(
    names_from = scen_time,
    values_from = avg_impact,
    values_fill = list(avg_impact = 0)) %>%
  mutate(ssp245_change = `ssp245_medium-term`-ssp245_current,
         ssp585_change = `ssp585_medium-term`-ssp585_current)

filter(avg_hab_impact, ssp585_change <0)

# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- avg_hab_impact %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`ssp585_change`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))


habitat_order <- avg_hab_impact %>%
  group_by(habitat) %>%
  summarize(total_value = sum(`ssp585_change`), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat)))

 # Step 2: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp585_medium-term`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/habitat_impact_future_barplot_ssp585.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


# get the change
 # Step 2b: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp585_change`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/habitat_impact_change_barplot_ssp585.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


```

Make the second part of the plot that shows overall change in cumulative pressure.
```{r}
cum_impact_habitat <- avg_hab_impact_tmp %>%
  group_by(habitat, scen_time) %>%
  summarize(avg_impact = sum(avg_impact, na.rm=TRUE))

# Reshape data to wide format
cum_wide <- cum_impact_habitat %>%
  pivot_wider(
    id_cols = c(habitat),
    names_from = scen_time,
    values_from = avg_impact
  ) 


# Create combined factor and adjust positions
cum_wide <- cum_wide %>%
  mutate(
    habitat = factor(habitat, levels = unique(habitat_order$habitat)),
    habitat_position = as.numeric(factor(habitat, levels = unique(habitat_order$habitat))),
    y_position = habitat_position,
    delta = `ssp585_medium-term` - ssp585_current,
    offset = delta * 0.05,  # Adjust this fraction as needed
    x_start = ssp585_current + sign(delta) * offset,
    x_end = `ssp585_medium-term` - sign(delta) * offset
  )


# Create the plot
habitat_compare <- ggplot() +
  # Add segments with arrows
  geom_segment(
    data = cum_wide,
    aes(
      x = x_start,
      xend = x_end,
      y = habitat,
      yend = habitat
    ),
    arrow = arrow(length = unit(0.2, "cm")),
    size = 0.5,
    alpha = 0.5
  ) +
  # Add points for current impacts
  geom_point(
    data = cum_wide,
    aes(x = ssp585_current, y = habitat, shape = "current"),
    size = 2.5
  ) +
  # Add points for future impacts
  geom_point(
    data = cum_wide,
    aes(x = `ssp585_medium-term`, y = habitat, shape = "future"),
    size = 2.5
  ) +
  theme_minimal() +
  labs(
    title = "",
    x = "Cumulative impact",
    y = ""
  ) +
  # Customize the legend title and labels
  scale_shape_manual(
    name = "",  # Set the legend title here
    values = c("current" = 16, "future" = 17)  # Use different shapes for "Current" and "Future"
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
habitat_compare
ggsave(habitat_compare, filename ="impact/figs/habitat_pressure_cumulative_impact_ssp585.png", width=8, height =4, units = c("in"))

```


## Figure 4:  Zonal region pressure impacts
```{r}

pressure_country <- read_csv(here("impact/data/country_eez_habxpressurexscenario_impact.csv"))

pressure_rgn <- pressure_country %>%
  separate(
    col = impact,
    into = c("habitat", "pressure", "scenario", "time_frame"),
    sep = "_",
    remove = FALSE,
    extra = "merge",
    fill = "right"
  )

## summarize each pressure (across all habitats) for each region
pressure_rgn <- pressure_rgn %>%
  group_by(pressure, scenario, time_frame, rgn_id, geosubrgn) %>%
  summarize(value = sum(value, na.rm=TRUE))

# join with names and condense benthic and surface oxygen and benthic and surface temperatures
pressures <- read_csv(here("impact/pressure_list.csv"))
setdiff(pressure_rgn$pressure, pressures$pressure)
setdiff(pressures$pressure, pressure_rgn$pressure)

pressure_rgn <- pressure_rgn %>%
  full_join(pressures, by=c("pressure")) %>%
  group_by(geosubrgn, rgn_id, plot_name_pressure, scenario, time_frame, color) %>%
  summarize(impact = sum(value, na.rm=TRUE)) %>%
  ungroup() 


## country coastal raster
eez <- rast(here("_spatial/eez_mol_w_land.tif"))
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(coast>0, 1, 0)
#get the area
country_coast <- eez*coast
plot(country_coast)

# get number of coastal cells in each country:
 coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

pressure_rgn <- pressure_rgn %>%
  left_join(coast_cell_num, by="rgn_id")

 pressure_georgn <- pressure_rgn %>%
   group_by(plot_name_pressure, scenario, time_frame, geosubrgn, color) %>%
   summarize(value = sum(impact, na.rm=TRUE),
             area = sum(cells, na.rm=TRUE)) %>%
   ungroup() %>%
  mutate(avg_impact = value/area) %>%
  mutate(scen_time = paste(scenario, time_frame, sep="_")) %>%
  select(habitat = geosubrgn, pressure = plot_name_pressure, scen_time, color, avg_impact)

## make data wide
avg_hab_impact <- pressure_georgn %>%
  pivot_wider(
    names_from = scen_time,
    values_from = avg_impact,
    values_fill = list(avg_impact = 0)) %>%
  mutate(ssp245_change = `ssp245_medium-term`-ssp245_current,
         ssp585_change = `ssp585_medium-term`-ssp585_current)


# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- avg_hab_impact %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`ssp585_change`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))


habitat_order <- avg_hab_impact %>%
  group_by(habitat) %>%
  summarize(total_value = sum(`ssp585_change`), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat)))

 # Step 2: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp585_medium-term`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/subregion_impact_future_barplot_ssp585.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


# get the change
 # Step 2b: Get plot data and reorder the 'pressure' factor levels based on the total_value
plot_data <- avg_hab_impact %>%
  select(habitat, pressure, impact =`ssp585_change`, color)

plot_data <- plot_data %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(habitat = factor(habitat, levels = unique(habitat_order$habitat))) %>%
  arrange(habitat, pressure)


# Step 3: Plot the data

plots_bars <- ggplot(plot_data, aes(y = habitat, x = impact, fill = pressure)) +
  geom_bar(stat = "identity") +
  #facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_data$color, plot_data$pressure)) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
plots_bars
ggsave(here("impact/figs/subregion_impact_change_barplot_ssp585.png"), plot = plots_bars, width = 8, height = 4, dpi = 300)


```

Make the second part of the plot that shows overall change in cumulative pressure.
```{r}
cum_impact_habitat <- pressure_georgn %>%
  group_by(habitat, scen_time) %>%
  summarize(avg_impact = sum(avg_impact, na.rm=TRUE))

# Reshape data to wide format
cum_wide <- cum_impact_habitat %>%
  pivot_wider(
    id_cols = c(habitat),
    names_from = scen_time,
    values_from = avg_impact
  ) 

# Create combined factor and adjust positions
cum_wide <- cum_wide %>%
  mutate(
    habitat = factor(habitat, levels = unique(habitat_order$habitat)),
    habitat_position = as.numeric(factor(habitat, levels = unique(habitat_order$habitat))),
    y_position = habitat_position,
    delta = `ssp585_medium-term` - ssp585_current,
    offset = delta * 0.05,  # Adjust this fraction as needed
    x_start = ssp585_current + sign(delta) * offset,
    x_end = `ssp585_medium-term` - sign(delta) * offset
  )


# Create the plot
habitat_compare <- ggplot() +
  # Add segments with arrows
  geom_segment(
    data = cum_wide,
    aes(
      x = x_start,
      xend = x_end,
      y = habitat,
      yend = habitat
    ),
    arrow = arrow(length = unit(0.2, "cm")),
    size = 0.5,
    alpha = 0.5
  ) +
  # Add points for current impacts
  geom_point(
    data = cum_wide,
    aes(x = ssp585_current, y = habitat, shape = "current"),
    size = 2.5
  ) +
  # Add points for future impacts
  geom_point(
    data = cum_wide,
    aes(x = `ssp585_medium-term`, y = habitat, shape = "future"),
    size = 2.5
  ) +
  theme_minimal() +
  labs(
    title = "",
    x = "Cumulative impact",
    y = ""
  ) +
  # Customize the legend title and labels
  scale_shape_manual(
    name = "",  # Set the legend title here
    values = c("current" = 16, "future" = 17)  # Use different shapes for "Current" and "Future"
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "white"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(angle = 20, hjust = 1)
  )
habitat_compare
ggsave(habitat_compare, filename ="impact/figs/georegion_pressure_cumulative_impact_ssp585.png", width=8, height =4, units = c("in"))

```

