---
title: "step4"
format: html
editor: visual
---

This script rescales the plumed nutrient data.

```{r}
library(tidyverse)
library(terra)
library(sf)
library(mapview)
library(here)

ocean_template <- terra::rast(here('_spatial/ocean_area_mol.tif'))
ocean_template[ocean_template>0] <- 1
terra::plot(ocean_template)

```

Project final combined plume models:

```{r}

plumed_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/nutrients/output/N_plume", full=TRUE)

for(plume in plumed_list){ # plume = plumed_list[1]
filename <- tolower(gsub("_joined", "", basename(plume)))
plumed_proj <- project(rast(plume), ocean_template, method="sum")
plumed_proj_mask <- plumed_proj*ocean_template

#plumed_latlon <- terra::project(plumed_proj_mask, "EPSG:4236")
#NA_extent <- ext(-170, -50, 5, 85) 
#plumed_latlon <- crop(plumed_latlon, NA_extent)
#mapview::mapview(log(plumed_latlon+1), maxpixels=970830)
#pal <- leaflet::colorNumeric(palette = "viridis", domain = values(plumed_latlon), alpha = TRUE)
#mapview::mapview(plumed_latlon, maxpixels=970830, col.regions=pal)

writeRaster(plumed_proj_mask, sprintf("/home/shares/ohi/stressors_2021/_dataprep/nutrients/projected/%s", filename), overwrite=TRUE)

}
```

## Get the rescaling value

Rescale: I am only going to use the nearshore values to determine the rescaling point. I will use the mean of the scenarios for the current timeframe.

Using area of deadzones to base rescaling on: https://www.wri.org/news/new-web-based-map-tracks-marine-dead-zones-worldwide 530 dead zones 95,000 miles 2 or \~ 250,000 km2

```{r}

## global coast
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(coast>0, 1, 0)
plot(coast)

# Will use the value associated with the 98.6th quantile as reference point
global(coast, "sum", na.rm=TRUE)
250000/15923300
# area of dead zones and area of marine coastal area (shoreline and 10km offshore)
# 1.6%

file_paths <- list.files("/home/shares/ohi/stressors_2021/_dataprep/nutrients/projected/", full=TRUE, pattern = "current")

current_plumes <- rast(file_paths)
avg_current_plumes <- mean(current_plumes)

ref_raster <- avg_current_plumes*coast
values <- values(ref_raster)
values <- na.omit(values)
hist(log(values+1))
hist((values))
1-0.0157
percentile_98_6 <- quantile(values, probs = 0.984, na.rm = TRUE)
#10782
```

## Rescale the rasters using the reference value

```{r}

file_paths <- list.files("/home/shares/ohi/stressors_2021/_dataprep/nutrients/projected/", full=TRUE)

for(file in file_paths){ #file= file_paths[1]
clamped_raster <- clamp(rast(file), lower = 0, upper = percentile_98_6)
rescaled_raster <- (clamped_raster - 0) / (percentile_98_6 - 0)
writeRaster(rescaled_raster, sprintf("/home/shares/ohi/stressors_2021/_dataprep/nutrients/rescaled/nutrient_%s", basename(file)), overwrite = TRUE)
            }

list.files("/home/shares/ohi/stressors_2021/_dataprep/nutrients/rescaled", full=TRUE)

```
