---
title: "Prepping population data"
output: html_document
date: '2022-09-07'
---

#Data Source
**Reference**: Olen and Lehsten 2022, https://www.sciencedirect.com/science/article/pii/S2352340922000166
**Downloaded**: 9/2022
**Description**:  Population count per cell, modelled to be consistent with both the CMIP6 RCP-specific urban fraction dataset (LUH2-v2f, luh.umd.edu) and the country level SSP population and urban fraction scenarios from the SSP database 
**Native data resolution**: 30 arcseconds (~1km equator) 
**Time range**: projection 2010-2100 (SSP1_2.6, SSP2_3.4, SSP3_7.0, SSP4_3.4, SSP5_8.5), yearly
**Format**:  tif


##Setup  

```{r setup,message=FALSE,warning=FALSE,verbose=FALSE}
#set options for all chunks in code
knitr::opts_chunk$set(warning=FALSE, message=FALSE,fig.width=6, fig.height=6)
#libraries
library(maps)
library(tidyverse)
library(RColorBrewer)
library(terra)
library(here)

# color
cols = rev(colorRampPalette(brewer.pal(9, 'Spectral'))(255)) # rainbow color scheme

# spatial map data
ocean_raw <- terra::rast(here("_spatial/ocean_area_mol.tif"))
ocean <- ocean_raw
ocean[ocean >0] <- 1

## need land area (no longer used:
## Originally, I thought it would be better to divide the
## population by the amount of actual land in the cell
## (vs. the total cell area),
## but then I realized we are calculating pressure across the entire cell...not just the land, so if we calculated density based on land area, but land occupies 25% of the cell, the pressure in the cell would then be multiplied by 0.25.)
prop_land <- 1-ocean_raw 
prop_land[is.na(prop_land)] <- 1
land_area_km2 <- prop_land*100 # each cell is 100 km2

## boundary mask
boundary_mask <- rast(nrow=2160, ncol=4320, xmin=-180, xmax=180, ymin=-90, ymax=90)
values(boundary_mask) <- 1
boundary_mask_mol <- terra::project(boundary_mask, "+proj=moll", res=10000)

plot(boundary_mask_mol)

```

## Defining years to use
```{r}
years_select <- data.frame(
  time_frame = c("current", "near-term", "medium-term", "long-term"),
  year_range = c("2011:2020",
                 "2021:2040",
                 "2041:2060",
                 "2081:2100"))

scenarios <- c("SSP1_RCP2_6", 
               "SSP2_RCP4_5",
               "SSP3_RCP7",
               "SSP4_RCP3_4",
               "SSP4_RCP6",
               "SSP5_RCP8_5")

```


Explore: can be skipped
```{r}

# see what is tin directory
files <- dir("/home/shares/ohi/stressors_2021/_raw_data/olen_population", full=TRUE)

pop <- raster::raster(files[1]) 
raster::cellStats(pop, "sum", na.rm=TRUE)
## world population in 2010 was 6.922 billion (worldbank)
## estimate here: 6.872, looks good!
plot(log(pop+1))

```


## Function to prep raster
```{r}
#terraOptions(progress=0)

#dir.create("/home/shares/ohi/stressors_2021/_dataprep/population")
prep_rast <- function(scen = "SSP5_RCP8_5"){ # scen = scenarios[3]
    for(yrs in years_select$year_range){ # yrs = years_select$year_range[1]
  file_years <- eval(parse(text=paste0("c(", yrs, ")")))
  file_years <- paste0("count", file_years, collapse = "|")
  
  all_files <- list.files("/home/shares/ohi/stressors_2021/_raw_data/olen_population", full=TRUE)
  subset_files <- grep(scen, all_files, value=TRUE)
  subset_files <- grep(file_years, subset_files, value=TRUE)
  
  stack_rasters <- terra::rast(subset_files)
  cat(scen, ", ", yrs, ",  N=", length(subset_files), "\n")
  
  # get mean number of people per cell across relevant window of years
  avg_pop <- terra::app(stack_rasters, "mean", na.rm=FALSE)
  
  # aggregate the data to be similar to our target resolution
  avg_pop_agg <- terra::aggregate(avg_pop, fact=10, fun="sum", na.rm=TRUE)
  
  # convert to density (people/km2) in order to safely project
  avg_pop_density <- avg_pop_agg/terra::cellSize(avg_pop_agg, unit="km")

      shoreline_density_km2 <- terra::project(avg_pop_density, ocean)
  # terra::global(shoreline_density_km2, "sum", na.rm=TRUE)*100
#cellSize(shoreline_density_km2, unit="km")
      
       time_frame <- years_select$time_frame[years_select$year_range==yrs]
  terra::writeRaster(shoreline_density_km2, filename=sprintf("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2/human-pop-density-km2_%s_%s.tif",
              time_frame, scen), overwrite=TRUE)
      
  ## get the nearshore disturbance estimate, average population density of nearby land and ocean cells (these are zero population) 
  offshore_density_km2_noNA <- ifel(is.na(shoreline_density_km2), 0, shoreline_density_km2) 

offshore_density_km2_noNA_smoothed <-   terra::focal(offshore_density_km2_noNA, w=3, fun="mean", na.rm=TRUE)
      

# check that things look ok!    
#  extent <- terra::ext(1e7, 1.25e7, 3e6, 5e6)  
# extent <- terra::ext(1e7, 1.15e7, 4e6, 5e6) 
  # extent <- terra::ext(1.09e7, 1.12e7, 4.4e6, 4.7e6) 
#plot(terra::crop(log(shoreline_density_km2_cell+1), extent))
#plot(terra::crop(log(offshore_density_km2_cell+1), extent))
# plot(terra::crop(terra::ifel(offshore_density_km2_cell>0, 1, NA), extent))     
# plot(terra::ifel(offshore_density_km2_cell>0, 1, NA))     

    ## combine shoreline and offshore regions
  total <- terra::ifel(is.na(shoreline_density_km2), offshore_density_km2_noNA_smoothed, shoreline_density_km2)  
    
  # global(total, "sum", na.rm=TRUE)*100
  #plot(terra::crop(log(total+1), extent))
  #plot(ocean_moll_sf, add=TRUE, color=NA) # from below

  terra::writeRaster(total, filename=sprintf("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand/human-pop-density-km2_%s_%s.tif",
              time_frame, scen), overwrite=TRUE)

  }
  }


```

Run function on all scenarios.
```{r}
for(scenario in scenarios){ #scenario= scenarios[1]
  prep_rast(scen=scenario)
  }

```

## We interupt this program....
Here I am going to save an additional iteration on the population data that will be useful for predicting other pressures (such as, light pollution).

In this case we extend the population data a bit further into the ocean.

```{r}

pop_density_files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2/", full=TRUE)
 
for(pop_density_file in pop_density_files){ # pop_density_file = pop_density_files[1]
  
  saveName <- basename(pop_density_file)
  saveName <- paste0("extend_", saveName)
  
  pop_raster <- terra::rast(pop_density_file)
  pop_no_NA <- ifel(is.na(pop_raster), 0, pop_raster) 

  
for(i in 1:3){
  i=1+i  
offshore_pop <-   terra::focal(pop_no_NA, w=3, fun=mean, na.rm=TRUE)
}      
    ## combine shoreline and offshore regions
  total <- terra::ifel(is.na(pop_raster), offshore_pop, pop_raster)  

  
terra::writeRaster(total, sprintf("/home/shares/ohi/stressors_2021/_dataprep/population/other_pop_files/%s", saveName), overwrite=TRUE)
}




```

## Back to our regularly scheduled program....

## Log transform data
This paper shows how log tranforming population data is linearly correlated with several stressors.
https://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1000606
```{r}
files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand/", full=TRUE)

for(file in files){
  # file <- files[1]
  file_name <- basename(file)
  file_name <- gsub("-km2", "-km2-log", file_name)
  
  raster <- terra::rast(file)
  ln_raster <- log(raster + 1)
  
    terra::writeRaster(ln_raster, filename=sprintf("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand_log/%s",
              file_name), overwrite=TRUE)
  
}
 
```

## Rescaling data

Ideally, we would want to rescale this based on something biologically meaningful,
for example, the relationship between human density and biological stress.

But I haven't been able to find anything. Given this, I will rescale these data based on a confidence interval.

Step 1: 
Create a mask so rescaling is based only on coastal rasters. 

```{r}

moll_crs <- '+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m'

ocean_sf <- sf::read_sf(here('_spatial', 'ne_10m_ocean', 'ne_10m_ocean.shp'))

  ocean_moll_sf <- ocean_sf %>%
    sf::st_transform(crs = moll_crs)

ocean_moll_line <- sf::st_cast(ocean_moll_sf,"MULTILINESTRING")  
raster_coast <- terra::rasterize(ocean_moll_line, ocean, touches=TRUE) 
plot(raster_coast, col="red") 
#  extent <- terra::ext(1e7, 1.15e7, 4e6, 5e6)  
# plot(terra::crop(raster_coast, extent))

coastal_mask <- focal(raster_coast, w=3, "max", na.policy="only", na.rm=TRUE)

plot(coastal_mask, col="red")
# plot(terra::crop(coastal_mask, extent))
# plot(ocean_moll_sf, add=TRUE, color=NA)

```

Get scaling value from current year. 
(use only coastal cells to determine scaling value, 
but rescale all population data)

```{r}
# using this value as information to determine best quantile to use.
# how many coastal cells are there?
global(coastal_mask, sum, na.rm=TRUE)
#255352

rescale_raster <- terra::rast("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand_log/human-pop-density-km2-log_current_SSP1_RCP2_6.tif")

rescale_raster_coast <- rescale_raster*coastal_mask

extent <- terra::ext(1e7, 1.15e7, 4e6, 5e6) 
plot(terra::crop(rescale_raster_coast, extent))

## determine quantile value used for rescaling
global(rescale_raster_coast, fun= function(x) {quantile(x, probs=c(0.99, 0.999, 0.9999), na.rm=TRUE)})

test1 <- terra::ifel(rescale_raster_coast >= 10.52521, 1, NA)
plot(as.points(test1), col="red")
plot(ocean_moll_sf, add=TRUE, color=NA)

test2 <- terra::ifel(rescale_raster_coast >= 9.169312, 1, NA)
plot(as.points(test2), col="red")
plot(ocean_moll_sf, add=TRUE, color=NA)


# will go with the 99.9th quantile value
rescale_value <- global(rescale_raster_coast, fun= function(x) {quantile(x, probs=c(0.999), na.rm=TRUE)})
# rescale_value <- 9.169315
```

## rescaling the data!
```{r}

files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand_log", full=TRUE)


for(file in files){
  #file <- files[1]
  save_name <- basename(file)
  save_name <- gsub("human-pop-density-km2-log", "disturbance", basename(file))
  save_name <- tolower(save_name)
  save_name <- gsub("_rcp", "", save_name)
  save_name <- gsub("([0-9])_([0-9])", "\\1\\2", save_name)
  save_name <- gsub("_ssp37", "_ssp370", save_name)

  # Function to switch the 2nd and 3rd sections
swap_sections <- function(filename) {
  sub("^([^_]+)_([^_]+)_([^_]+)(\\..*)$", "\\1_\\3_\\2\\4", filename)
}
save_name <- swap_sections(save_name)

  tmp <- terra::rast(file)
#  rescaled <- tmp/pull(rescale_value)
    rescaled <- tmp/rescale_value
  rescaled <- ifel(rescaled > 1, 1, rescaled)
# plot(rescaled, col=cols)
  
 terra::writeRaster(rescaled, filename=sprintf("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand_log_rescaled/%s",
              save_name), overwrite=TRUE)  
    
}

```

## clip to area 1 cell inland, coastline and 1 cell offshore
```{r}

## prepare raster mask
ocean <- terra::rast(here('_spatial/ocean_area_mol.tif'))

## get coastline
coastline <- terra::ifel(ocean<1 & ocean > 0,  1, 0)
coastline <- terra::ifel(coastline ==1, 1, NA)
global(coastline, "sum", na.rm=TRUE)

## get one cell inland
land <- terra::ifel(is.na(ocean) , 1, NA)
one_inland <- boundaries(land, falseval=NA, inner=TRUE)
global(one_inland, "sum", na.rm=TRUE)

## get one cell offshore
full_ocean <- terra::ifel(ocean==1, 1, NA)
one_offshore<- boundaries(full_ocean, falseval=NA, inner=TRUE)
global(one_offshore, "sum", na.rm=TRUE)

mask <- app(c(coastline, one_inland, one_offshore), "sum", na.rm=TRUE)
plot(mask)

# get rid of outer boundary
boundary_mask <- rast(nrow=1814, ncol=3617, xmin=-180, xmax=180, ymin=-90, ymax=90)
values(boundary_mask) <- 1
boundary_mask_mol <- terra::project(boundary_mask, "+proj=moll", res=10000)
boundary_mask_mol <- extend(boundary_mask_mol, ext(ocean), fill=NA)
boundary_mask_mol <- resample(boundary_mask_mol, ocean, method="near")
inner_boundary <-  boundaries(boundary_mask_mol, inner=TRUE, falseval=0)
inner_boundary <- ifel(inner_boundary==0, 1, NA) # cuts off out layer
inner_inner_boundary <- boundaries(inner_boundary, inner=TRUE, falseval=0)
inner_inner_boundary <- ifel(inner_inner_boundary==0, 1, NA) # cuts off next outer layer
i_i_i_boundary <- boundaries(inner_inner_boundary, inner=TRUE, falseval=0)
boundary_mask <- ifel(i_i_i_boundary==0, 1, NA) # one more layer removed!

mask <- mask*boundary_mask
plot(mask)

raster_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand_log_rescaled/", full=TRUE)

for(rstr in raster_list){ # rstr <- raster_list[20]
  save_name <- basename(rstr)
  rstr_tmp <- terra::rast(rstr)
  rstr_tmp <- rstr_tmp * mask
  
  writeRaster(rstr_tmp, sprintf("/home/shares/ohi/stressors_2021/_dataprep/population/pop_density_km2_ocean_expand_log_rescaled_crop/%s", save_name), overwrite=TRUE)
  
}



***

