---
title: "OA pressure"
output: html_document
date: '2022-08-12'
---
#Data Source
**Reference**: From: Liqing JIANG - NOAA Affiliate <liqing.jiang@noaa.gov>
https://www.ncei.noaa.gov/data/oceans/ncei/ocads/data/0259391/
https://www.ncei.noaa.gov/data/oceans/ncei/ocads/metadata/0259391.html
**Downloaded**: May 2 2023
**Description**:  Aragonite Saturation State  $\Omega_{arg}$
**Native data resolution**: 1 [20.5:1:379.5] degree longitude x 1 [-89.5:1:89.75] degree latitude grid in the global surface ocean
**Time range**: historical 1750 and average decadal from 1840-2010 (n = 18 layers);
average decadal future scenarios 2010-2100 (n=9 layers)
**Format**:  NetCDF

This script takes the raw netCDF data and does the following:

   a. Calculates the historic mean for a 50 year period from 1840-1890 
   b. Rescales each averaged decadal raster layer from 0 to 1 based on biological thresholds (&#937; <= 1 and <=3) and the proportional change compared to a historical mean
   c. Resamples/reprojects to our CRS
   c. Interpolates the data to gap-fill for cells where there is no data
   e. Mask the resampled data to select only those cells within the ocean

## Rescaling
This pressure layer is rescaled so that all values lie between 0 and 1 using both a historical reference period and biological reference points. All cells with values less than one, indicating an undersaturated state, are set equal to the highest stressor level, 1. All cells with values > 3 have the lowest stressor level of 0. For all other cells, rescaling the aragonite staturation state value to between 0 and 1 relies upon the change in saturation relative to the reference period.

Deviation from aragonite saturation state is determined for each year in the study period using this equation:

$$\Delta \Omega_{year} = \frac{(\Omega_{base} - \Omega_{year})}{(\Omega_{base} - 1)}$$

Note that the current value is subtracted from the baseline; this way, a reduction in $\Omega$ becomes a positive pressure value.  It is then normalized by the current mean state; so a decrease in $\Omega$ while the current state is high indicates less pressure than the same decrease when the current state is near 1. 

$\Delta \Omega_{year}$ is then modified to account for increases in aragonite saturation state (pressure = 0) and arag sat state less than 1 (pressure = 1).

If the current value is less than or equal to 1, it is set to 1. If the value is >3 the pressure is set to 0. Otherwise, the value is calculated from the above equation.

These thresholds are based on this research: https://www.sciencedirect.com/science/article/pii/S0921818198000356

##Setup  

```{r setup,message=FALSE,warning=FALSE,verbose=FALSE}
#set options for all chunks in code
knitr::opts_chunk$set(warning=FALSE, message=FALSE,fig.width=6, fig.height=6)
#libraries
library(maps)
library(tidyverse)
library(RColorBrewer)
library(terra)
library(here)
library(ncdf4)

# color
cols = rev(colorRampPalette(brewer.pal(9, 'Spectral'))(255)) # rainbow color scheme

# spatial map data
ocean <- terra::rast(here("_spatial/ocean_area_mol.tif"))
ocean[ocean >0] <- 1

```

## Defining years to use
```{r}
years_select <- data.frame(
  time_frame = c(rep("historic", 5), "current", rep(c("near-term", "medium-term", "long-term"), each=2)),
  year_range = c("1841-1850",
                 "1851-1860",
                 "1861-1870",
                 "1871-1880",
                 "1881-1890",
                 "2011-2020",
                 "2021-2030", "2031-2040",
                 "2041-2050", "2051-2060",
                 "2081-2090", "2091-2100"))


```


Explore: can be skipped
```{r}
# see what is in directory
dir("/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median")

tmp <- readtext::readtext("/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/old/README.txt")
tmp$text

# get description of data
oa <- nc_open("/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp245.nc")
oa
names(oa$var)

# Get the data
ssp245_stack <- raster::stack("/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp245.nc")


```


## Function to align raster data
```{r}
align_rasts <- function(hist_stack){
  #hist_stack <- terra::rast("/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp245.nc")

terra::ext(hist_stack)<- c(0, 360, -90, 90)
hist_stack <- terra::rotate(hist_stack)
hist_stack <- terra::shift(hist_stack, dx=20)
x1 <- terra::crop(hist_stack, terra::ext(180, 200, -90, 90))
x2 <- terra::crop(hist_stack, terra::ext(-160, 180, -90, 90))
terra::ext(x1) <- c(-180, -160, -90, 90)
terra::ext(x2) <- c(-160, 180, -90, 90)
m <- terra::merge(x1, x2)
terra::crs(m) <- "epsg:4326"
m
#plot(m[[1]])
#maps::map('world', col='gray95', fill=T, border='gray80', add=T)
}

#dir.create("/home/shares/ohi/stressors_2021/_dataprep/oa")
```

***

# Methods

## Get historical values
This will be average of decadal data from 1840-1890.
```{r load_data}
# list of rasters in historical nc files
historic_stack_list <- data.frame(rasters_name = paste0("X",1:18),
                         years=c("years_1750",
                                 "years_1841-1850",
                                 "years_1851-1860",
                                 "years_1861-1870",
                                 "years_1871-1880",
                                 "years_1881-1890",
                                 "years_1891-1900",
                                 "years_1901-1910",
                                 "years_1911-1920",
                                 "years_1921-1930",
                                 "years_1931-1940",
                                 "years_1941-1950",
                                 "years_1951-1960",
                                 "years_1961-1970",
                                 "years_1971-1980",
                                 "years_1981-1990",
                                 "years_1991-2000",
                                 "years_2001-2010"))


# read in historical data
hist_stack <- terra::rast("/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_historical.nc")
hist_stack <- hist_stack[[grep("aragonite", names(hist_stack))]]
names(hist_stack) <- historic_stack_list$years
plot(hist_stack[[1]]) #we'll have to rotate the rasters

hist_stack_align <- align_rasts(hist_stack)
plot(hist_stack_align[[1]])
maps::map('world', col='gray95', fill=T, border='gray80', add=T)

years_select_hist <- years_select %>%
  filter(time_frame == "historic") %>%
  pull(year_range)
years_select_hist <- paste(years_select_hist, collapse="|")

hist_stack_historical <- hist_stack_align[[grep(years_select_hist, names(hist_stack_align))]]

# Take avg of 5 time periods
historical_avg <- terra::app(hist_stack_historical, mean)
plot(historical_avg)

# check to see how much is <1
tmp <- terra::ifel(historical_avg <= 1, 1, NA)
plot(tmp, col="red")

tmp <- terra::ifel(historical_avg >3, 1, NA)
plot(tmp, col="red")
```


## Select relevant years and rescale

### Function for future scenarios
The `oa_rescale` function averages the appropriate rasters for each time series and then rescales and saves. 

```{r rescale-explore,eval=F}
# this just is a space to explore how the pressures change with different inputs.

old <- seq(1,3, by=0.1)
new <- 1.2
new_method <- (old - new)/(old-1)
old_method <- (new-1)/(3-1)
predict <- data.frame(old, new, new_method, old_method)

predict
```

```{r rescale,eval=F}

#for each layer, all values <=1 are assigned a 1, otherwise old-new/(old-1)
oa_rescale <- function(raster_path, stack_list=stack_list_spp, historical_avg){ #raster_path <- "/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp245.nc"

  saveName <- basename(raster_path)
  saveName <- gsub(".nc", "", saveName)
  saveName <- gsub("Aragonite_median_", "", saveName)
  
  raster_stack <- terra::rast(raster_path)
  raster_stack <- raster_stack[[grep("aragonite", names(raster_stack))]]
  names(raster_stack) <- stack_list$years
  raster_stack_align <- align_rasts(raster_stack)

  scenario <- c("current", "near-term", "medium-term", "long-term")

for(select_time_frame in scenario){
          #select_time_frame = "current"

    ## Select relevant years for time period
    select_year_range_tmp <- years_select %>%
      filter(time_frame == select_time_frame) %>%
      pull(year_range)
    select_year_range <- paste(select_year_range_tmp, collapse="|")

    year_range_stack <- raster_stack_align[[grep(select_year_range, names(raster_stack_align))]]
    cat(sprintf("third dimension should be 2: observed is %s", dim(year_range_stack)[3]))
    oa_predict <- terra::app(year_range_stack, mean)

    ## rescale
  diff = (historical_avg - oa_predict)/(historical_avg - 1)
  
  oa_pressure <- terra::ifel(oa_predict<=1, 1, oa_predict)                                 #all values at or less than 1 are given a value of 1
  
  oa_pressure <- terra::ifel(oa_predict>3, 0, oa_pressure)                                 #all values greater than 3 are given a value of 0

    
  oa_pressure <- terra::ifel(oa_pressure>1 & oa_pressure<=3, 1-(oa_pressure-1)/(3-1), oa_pressure)                     #all cells with values between 1 and 3 are scaled linearly between 0 and 1.
  
    oa_pressure <- terra::ifel(historical_avg>1 & historical_avg<=3, diff, oa_pressure)  
    # all cells where the historical values between 0 and 3 are scaled relative to the historical value.
    
    #all cells with values greater than 1 are capped at 1
    oa_pressure <- terra::ifel(oa_pressure>1 , 1, oa_pressure)

  
## format raster
oa_pressure_crs <- terra::project(oa_pressure, ocean, method="bilinear")
oa_pressure_crs[is.nan(oa_pressure_crs)] <- NA

# gapfill coastline
oa_pressure_gf <- oa_pressure_crs
for(i in 1:200){
  i=1+i
oa_pressure_gf <- terra::focal(oa_pressure_gf, w=3, fun=mean, na.rm=TRUE, na.policy="only")
}

oa_mask <- terra::mask(oa_pressure_gf, ocean)

terra::writeRaster(oa_mask, filename=sprintf("/home/shares/ohi/stressors_2021/_dataprep/oa/oa_%s_%s.tif",
              saveName, select_time_frame), overwrite=TRUE)
}
}
```

### Run function for future years

```{r}
## Clean

#rm_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/oa", full=TRUE)
#file.remove(rm_list)


#Associate raster layers with years
stack_list_spp <- data.frame(rasters_name = paste0("X",1:9),
                         years=c("years_2011-2020",
                                 "years_2021-2030",
                                 "years_2031-2040",
                                 "years_2041-2050",
                                 "years_2051-2060",
                                 "years_2061-2070",
                                 "years_2071-2080",
                                 "years_2081-2090",
                                 "years_2091-2100"))

ssp245_stack <- "/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp245.nc"
oa_rescale(raster_path = ssp245_stack, stack_list=stack_list_spp, historical_avg=historical_avg)

ssp119_stack <- "/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp119.nc"
oa_rescale(raster_path = ssp119_stack, stack_list=stack_list_spp, historical_avg=historical_avg)

ssp126_stack <- "/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp126.nc"
oa_rescale(raster_path = ssp126_stack, stack_list=stack_list_spp, historical_avg=historical_avg)

ssp370_stack <- "/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp370.nc"
oa_rescale(raster_path = ssp370_stack, stack_list=stack_list_spp, historical_avg=historical_avg)

ssp585_stack <- "/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_ssp585.nc"
oa_rescale(raster_path = ssp585_stack, stack_list=stack_list_spp, historical_avg=historical_avg)

```



## Some checks
```{r plot rescale}
rast_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/oa", full=TRUE)
library(ggplot2)
library(tidyterra)
library(ggmap)
library(gridExtra)
 library(rnaturalearth)
# For stamen map, you have to give the extremity of the window you are looking at. here is an example with the watercolor background (Around brisbane)
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")

ggplot() +
  geom_spatraster(data = terra::rast(rast_list[12])) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "")
  ) +
  geom_sf(data=world, fill="lightgray", color="white", size=0.1) +
  labs(fill = "Rescaled aragonite",
       title = "Current") +
    geom_spatraster(data = terra::rast(rast_list[12])) +
  theme_minimal() 
```


# Gap-filled cells

We want to create a raster layer that shows all cells that were gap-filled. Since they were the same cells interpolated across all years, we only need to create one raster.

```{r,eval=F}
#Data before interpolation 
historical_stack <- "/home/shares/ohi/stressors_2021/_raw_data/oa_arag_noaa_jiang/nc/median/Aragonite_median_historical.nc"
## get the years in historical file
raster_stack_his <- terra::rast(historical_stack)
  raster_stack_his <- raster_stack_his[[grep("aragonite", names(raster_stack_his))]]
  names(raster_stack_his) <- historic_stack_list$years
  raster_stack_align_his <- align_rasts(raster_stack_his)
  hist_rast <- raster_stack_align_his[[grep("2001-2010", names(raster_stack_his))]]
plot(hist_rast)

hist_rast_mol <- terra::project(hist_rast, ocean, method="bilinear")
oa_org <- hist_rast_mol
oa_org[oa_org >=0] <- 1
oa_org[is.na(oa_org)] <- 0

#interpolated (or gap-filled) cells    
gf <- terra::rast(list.files("/home/shares/ohi/stressors_2021/_dataprep/oa", full=TRUE)[1])
gf[gf >= 0] <- 1
gf[is.na(gf)] <- 0

gf_locations <- gf-oa_org
plot(gf_locations)


terra::writeRaster(gf_locations, "/home/shares/ohi/stressors_2021/_dataprep/oa/oa_interpolated_cells.tif", overwrite=TRUE)
```

```{r plot_interp_cells}
plot(raster(file.path(oagit_dir, 'v2017/output/oa_interpolated_cells.tif')), col=cols, box=F, axes=F, main='Interpolated cells')
```



