---
title: "SST average temp"
output: html_document
date: '2022-08-09'
---

This is another form of the SST pressure. In STEP2_extreme_events.Rmd, we calculate the extreme events.  Here we are going to calculate annual mean temperatures, this allows us to convert the raw temperatures into taxa specific pressures based on their thermal performance curves.  
 
issues: 51

```{r setup, include=FALSE}
library(raster)
library(ncdf4)
library(here)
library(tidyverse)
library(foreach)
library(doParallel)

#year ranges and scenarios:
yr_range <- data.frame(period = c(rep("historical", each=15), rep("current", each=6), rep(c("near-term", "medium-term", "long-term"), each=20)),
                       year = c(2000:2014, 2015:2020, 2021:2040, 2041:2060, 2081:2100))


yr_bundles <- data.frame(scenario = rep(c("ssp585", "ssp245"), each=17),
                       year = c("2015:2019",  
                                "2020:2024", 
                                "2025:2029",
                                "2030:2034",
                                "2035:2039",
                                "2040:2044",
                                "2045:2049",
                                "2050:2054",
                                "2055:2059",
                                "2060:2064",
                                "2065:2069",
                                "2070:2074",
                                "2075:2079",
                                "2080:2084",
                                "2085:2089",
                                "2090:2094",
                                "2095:2099"))

```

In previous script, we converted the 6 hour time slots to daily values. Here we average the daily values for each year to get the yearly average.

```{r}
file_names <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/", full=TRUE)

# take care of historical data
for(year in 1979:2014){
  #year = 1979
yr_stack <- stack(sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/hist_%s.nc", year), varname="tos")


yearly_average <- mean(yr_stack, na.rm=TRUE)

writeRaster(yearly_average, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_average/hist_%s.tif", year), bylayer=TRUE, format='GTiff', overwrite=TRUE)
cat(year)
}

# ssp245
for(year in 2015:2100){
  #year = 2015
yr_stack <- stack(sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/ssp245_%s.nc", year), varname="tos")
yearly_average <- mean(yr_stack, na.rm=TRUE)

writeRaster(yearly_average, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_average/ssp245_%s.tif", year), bylayer=TRUE, format='GTiff', overwrite=TRUE)
cat(year)
}

# ssp585
for(year in 2015:2100){
  #year = 1979
yr_stack <- stack(sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/ssp585_%s.nc", year), varname="tos")
yearly_average <- mean(yr_stack, na.rm=TRUE)

writeRaster(yearly_average, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_average/ssp585_%s.tif", year), bylayer=TRUE, format='GTiff', overwrite=TRUE)
cat(year)

}

tmp <- terra::rast(list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/year_average", full=TRUE)[1])
terra::plot(tmp)
terra::click(tmp)
```

Average for five year periods of time.
```{r}
list_files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/year_average", full=TRUE)

for(i in 1:dim(yr_bundles)[1]){ #i = 2
scen = yr_bundles$scenario[i]
yr_range = yr_bundles$year[i]
yr_range_start= as.numeric(str_split(yr_range, pattern=":")[[1]])[1] 
yr_range_end = as.numeric(str_split(yr_range, pattern=":")[[1]])[2]
all_years = seq(yr_range_start, yr_range_end)
all_years = paste(paste0(all_years, ".tif"), collapse="|")

rast_yrs <- grep(all_years, list_files, value=TRUE)
rast_scen_yrs <- grep(scen, rast_yrs, value=TRUE)
cat("n= ", length(rast_scen_yrs), "for scenario = ", scen, "\n")

sst_avg <- mean(stack(rast_scen_yrs), na.rm=TRUE)

writeRaster(sst_avg, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average/sst_avg_%s_%s_%s.tif", scen, yr_range_start, yr_range_end), overwrite=TRUE)

}

check <- raster("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average/sst_avg_ssp245_2095_2099.tif")
plot(check)
```


# final formatting
1. gapfill to get coastline
2. project raster to appropriate resolution

```{r}

rast_base_10km_file <- terra::rast(here('_spatial/rast_base_mol_10km.tif'))
ocean_template <- terra::rast(here('_spatial/ocean_area_mol.tif'))
ocean_template[ocean_template>0] <- 1
plot(ocean_template)

avg_paths <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average", full=TRUE)

for(avg_path in avg_paths){ # avg_path = avg_paths[1]
avg_rast <- terra::rast(avg_path)

avg_rast <- terra::rotate(avg_rast)

terra::ext(avg_rast) <- c(-180, 180, -90, 90)
#plot(avg_rast)
#maps::map('world', col='black', fill=FALSE, border='gray80', add=TRUE, size=0.5)

mol_rast <- terra::project(avg_rast, rast_base_10km_file, method="near", over=TRUE)
#plot(mol_rast)


## gapfill
gf_raster <- function(x){raster::focal(x, w = matrix(1,3,3), fun = mean, na.rm=TRUE, na.policy="only")}

r = mol_rast
  
## Repeat 100 times (most is immediately gapfilled but the antarctic area was persistent)
i <- 0
while (i <= 50){
r <- gf_raster(r)
i <- i + 1
print(i)
}

# plot(r)
r <- r*ocean_template
# diff_rast <- r - mol_rast
#plot(diff_rast)

writeRaster(r, filename = sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/five_year_average_final/%s", basename(avg_path)), overwrite = TRUE)
}

```



