---
title: "Downloading and prepping SST future data"
output: html_document
date: '2022-08-09'
---
 
Downloaded from here (Aug 9 2022): 
https://www.scidb.cn/en/detail?dataSetId=791587189614968832&dataSetType=personal
 
Bias-corrected CMIP6 global dataset for dynamical downscaling of the Earth’s historical and future climate (1979–2100) 

1. Xu, Z., Han, Y., Tam, C.-Y., Yang, Z.-L. & Fu, C. Bias-corrected CMIP6 global dataset for dynamical downscaling of the historical and future climate (1979–2100). Sci Data 8, 293 (2021).

Dynamical downscaling is an important approach to obtaining fine-scale weather and climate information. However, dynamical downscaling simulations are often degraded by biases in the large-scale forcing itself. We constructed a bias-corrected global dataset based on 18 models from the Coupled Model Intercomparison Project Phase 6 (CMIP6) and the European Centre for Medium-Range Weather Forecasts Reanalysis 5 (ERA5) dataset. The bias-corrected data have an ERA5-based mean climate and interannual variance, but with a non-linear trend from the ensemble mean of the 18 CMIP6 models. The dataset spans the historical time period 1979–2014 and future scenarios (SSP245 and SSP585) for 2015–2100 with a horizontal grid spacing of (1.25° × 1.25°) at six-hourly intervals. Our evaluation suggests that the bias-corrected data are of better quality than the individual CMIP6 models in terms of the climatological mean, interannual variance and extreme events. This dataset will be useful for dynamical downscaling projections of the Earth’s future climate, atmospheric environment, hydrology, agriculture, wind power, etc.

 
Issue: https://github.com/OHI-Science/stressors_issues/issues/24
 

```{r setup, include=FALSE}
library(raster)
library(ncdf4)
library(here)
library(tidyverse)
library(foreach)
library(doParallel)

#year ranges and scenarios:
yr_range <- data.frame(period = c(rep("historical", each=15), rep("current", each=6), rep(c("near-term", "medium-term", "long-term"), each=20)),
                       year = c(2000:2014, 2015:2020, 2021:2040, 2041:2060, 2081:2100))

```

Explore the rasters a bit:
```{r}
tmp <- nc_open("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_1979_01.nc4")
tmp$dim$time$vals
read_sst <- ncvar_get(tmp, "tos")
names(tmp$var)
tmp$dim$time$vals # origin 1850-1-1
tmp <- raster::raster("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_1979_01.nc4", varname="tos")

```

Data is reported every 6 hours. Average this to get the daily value. Using cdo system commands to do this.

```{r}
file_names <- list.files("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/")

registerDoParallel(6)

foreach(file_name=file_names) %do% {
#for(file_name %in% files_names)
#file_name = file_names[1]
file_rename <- gsub("atm", "daymean", file_name)

cdo_command <- sprintf("cdo daymean /home/shares/ohi/stressors_2021/_raw_data/sst_Xu/%s /home/shares/ohi/stressors_2021/_dataprep/SST/day_mean/%s", file_name, file_rename)
system(cdo_command)
}

system("cdo daymean /home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_1979_01.nc4 /home/shares/ohi/stressors_2021/_dataprep/SST/day_mean/${name}")
check <- raster("/home/shares/ohi/stressors_2021/_dataprep/SST/day_mean/daymean_1979_1.nc", varname="tos")

```

Next step is to merge the daily values into one yearly netcdf file.
```{r}

registerDoParallel(6)

foreach(year=1979:2014) %do% {
#year = 1979
cdo_command <- sprintf("cdo -b F64 mergetime /home/shares/ohi/stressors_2021/_dataprep/SST/day_mean/daymean_hist_%s_*.nc4  /home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/hist_%s.nc", year, year)

system(cdo_command)
}

foreach(year=2015:2100) %do% {
#year = 1979
cdo_command <- sprintf("cdo -b F64 mergetime /home/shares/ohi/stressors_2021/_dataprep/SST/day_mean/daymean_ssp245_%s_*.nc4  /home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/ssp245_%s.nc", year, year)

system(cdo_command)
}

foreach(year=2015:2100) %do% {
#year = 1979
cdo_command <- sprintf("cdo -b F64 mergetime /home/shares/ohi/stressors_2021/_dataprep/SST/day_mean/daymean_ssp585_%s_*.nc4  /home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/ssp585_%s.nc", year, year)

system(cdo_command)
}

```


Next step is to convert to weekly averages.
```{r}
# take care of historical data
for(year in 1979:2014){
  #year = 1979
yr_stack <- stack(sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/hist_%s.nc", year), varname="tos")
weeks <- lubridate::week(lubridate::ymd(gsub("X", "", names(yr_stack))))
weekly_mean <- stackApply(yr_stack, weeks, fun=mean, na.rm=TRUE)
names(weekly_mean) <- gsub("index", sprintf("hist_y%s", year), names(weekly_mean))
cat("weeks = ", length(names(weekly_mean)), "\n")
writeRaster(weekly_mean, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_mean/%s", names(weekly_mean)), bylayer=TRUE, format='GTiff')
}

# ssp245
for(year in 2015:2100){
  #year = 1979
yr_stack <- stack(sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/ssp245_%s.nc", year), varname="tos")
weeks <- lubridate::week(lubridate::ymd(gsub("X", "", names(yr_stack))))
weekly_mean <- stackApply(yr_stack, weeks, fun=mean, na.rm=TRUE)
names(weekly_mean) <- gsub("index", sprintf("ssp245_y%s", year), names(weekly_mean))
cat("weeks = ", length(names(weekly_mean)), "\n")
writeRaster(weekly_mean, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_mean/%s", names(weekly_mean)), bylayer=TRUE, format='GTiff')
}

# ssp585
for(year in 2015:2100){
  #year = 1979
yr_stack <- stack(sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_merge/ssp585_%s.nc", year), varname="tos")
weeks <- lubridate::week(lubridate::ymd(gsub("X", "", names(yr_stack))))
weekly_mean <- stackApply(yr_stack, weeks, fun=mean, na.rm=TRUE)
names(weekly_mean) <- gsub("index", sprintf("ssp585_y%s", year), names(weekly_mean))
cat("weeks = ", length(names(weekly_mean)), "\n")
writeRaster(weekly_mean, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_mean/%s", names(weekly_mean)), bylayer=TRUE, format='GTiff')
}

tmp <- terra::rast(list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_mean", full=TRUE)[1])
terra::plot(tmp)
terra::click(tmp)
```

Next use historical data to estimate sd (or, 90th quantile) for each week using years 1979-2010.
```{r}
hist_yrs <- paste(1979:2010, collapse="|")

weekly_mean_files <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_mean/", full=TRUE)
plot(raster(weekly_mean_files[1]))

hist_rasts <- grep(hist_yrs, weekly_mean_files, value=TRUE)

for(week_num in 1:52){
# week_num <- 1

  week <- paste0("_", week_num, ".tif")
  
hist_rasts_week <- grep(week, hist_rasts, value=TRUE)
stack_week <- stack(hist_rasts_week)
cat(week, " n= ", dim(stack_week)[3], "\n")
#fun_90 <- function(x){quantile(x, probs = c(0.90), na.rm=TRUE)}
fun_sd <- function(x){sd(x, na.rm=TRUE)}

beginCluster(n=8)
#quant_raster <- raster::clusterR(stack_week, fun = calc, args=list(fun=fun_90))
sd_raster <- raster::clusterR(stack_week, fun = calc, args=list(fun=fun_sd))
endCluster()

#writeRaster(quant_raster, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_quant90/ref_point_week%s", week), overwrite=TRUE)
writeRaster(sd_raster, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_sd/sd_week%s", week), overwrite=TRUE)

}


## now get the corresponding weekly mean for the historical data
for(week_num in 1:52){
# week_num <- 1

  week <- paste0("_", week_num, ".tif")
  
hist_rasts_week <- grep(week, hist_rasts, value=TRUE)
stack_week <- stack(hist_rasts_week)
cat(week, " n= ", dim(stack_week)[3], "\n")
fun_mean <- function(x){mean(x, na.rm=TRUE)}

beginCluster(n=8)
mean_raster <- raster::clusterR(stack_week, fun = calc, args=list(fun=fun_mean))
endCluster()

#writeRaster(quant_raster, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_quant90/ref_point_week%s", week), overwrite=TRUE)
writeRaster(mean_raster, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_mean/mean_week%s", week), overwrite=TRUE)

}

check <-raster("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_mean/mean_week_52.tif")
plot(check)

check <-terra::rast("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_mean/mean_week_26.tif")
plot(check)
click(check)
```

Calculate the weekly thresholds that, when exceeded, indicate extreme events.
```{r}

for(week_num in 1:52){
# week_num <- 1

  week <- paste0("_", week_num, ".tif")
  
  mean_sst <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_mean", pattern = week, full=TRUE)
  sd_sst <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/historical_sd", pattern = week, full=TRUE)
  

threshold_sst <- raster(mean_sst) + 2*raster(sd_sst)

writeRaster(threshold_sst, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_sd_threshold/weekly_threshold_sd%s", week), overwrite=TRUE)

}


tmp <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_sd_threshold/", full=TRUE)

plot(raster(tmp[26]))
```

Calculate whether each week exceeds the relevant week's theshold value.
(takes about 2 days to run)
```{r}

for(week_num in 1:52){ #week_num = 1
ref_val <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_sd_threshold/", full=TRUE, pattern = sprintf("_%s.tif", week_num))
ref_raster <- raster(ref_val)

sst_week <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_mean", pattern = sprintf("_%s.tif", week_num), full=TRUE)
for(file in sst_week){ # file <- sst_week[200]
  sst_raster <- raster(file)
  save_name <- basename(file)
  save_name <- gsub(".tif", "_exceed_boolean", save_name)
  
  fun_exceed <- function(x, y){ifelse(is.na(x) | is.na(y), NA, ifelse(x > y, 1, 0))}
    
    beginCluster(n=8)
    weekly_exceed <- raster::clusterR(stack(sst_raster, ref_raster),
                     overlay, 
                    arg = list(fun = fun_exceed))
    endCluster()
#plot(weekly_exceed)
cat("save_name = ", save_name, "   exceeds=", cellStats(weekly_exceed, "sum", na.rm=TRUE), "\n")
  writeRaster(weekly_exceed, filename =  sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_exceed_two_sd/sd_exceed_%s.tif", save_name), overwrite=TRUE)
}
}

tmp <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_exceed_two_sd/", full=TRUE)

plot(raster(tmp[500]))

```

Stack the weeks for each year and determine total exceedances.
```{r}

# ssp 5-8.5
for(year in c(2015:2100)){
#year <- "2015"
year_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_exceed_two_sd/", pattern=as.character(year), full=TRUE)
year_list <- grep("ssp585", year_list, value=TRUE)

#plot(raster(year_list[1]))

week_stack <- raster::stack(year_list)

total_yr_exceed <- calc(week_stack, sum)
cat(year, cellStats(total_yr_exceed, max), "\n")

writeRaster(total_yr_exceed, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd/total_yr_exceed_ssp585_%s", year), format='GTiff', overwrite=TRUE)
}
#plot(total_yr_exceed)

# ssp 2-4.5
for(year in c(2015:2100)){
#year <- "2015"
year_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_exceed_two_sd/", pattern=as.character(year), full=TRUE)
year_list <- grep("ssp245", year_list, value=TRUE)
week_stack <- raster::stack(year_list)

total_yr_exceed <- calc(week_stack, sum)
cat(year, cellStats(total_yr_exceed, max), "\n")

writeRaster(total_yr_exceed, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd/total_yr_exceed_ssp245_%s", year), format='GTiff', overwrite=TRUE)
}

# historical
for(year in c(1979:2014)){
#year <- 1976
year_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/weekly_exceed_two_sd/", pattern=as.character(year), full=TRUE)

week_stack <- raster::stack(year_list)

total_yr_exceed <- calc(week_stack, sum)
cat(year, cellStats(total_yr_exceed, max), "\n")

writeRaster(total_yr_exceed, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd/total_yr_exceed_hist_%s", year), format='GTiff', overwrite=TRUE)
}


check <- terra::rast("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd/total_yr_exceed_hist_2010.tif")
plot(check)

check <- raster("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd/total_yr_exceed_ssp585_2050.tif")
plot(check)

check <- raster("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd/total_yr_exceed_ssp585_2085.tif")
plot(check)

```

Sum the total exceedences across the scenarios and year ranges.

```{r}
yr_range
ssps <- c("ssp245", "ssp585")

## historical
year_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd", pattern = paste(yr_range$year[yr_range$period=="historical"], collapse="|"), full=TRUE)
year_list <- grep("hist", year_list, value=TRUE)

year_stack <- raster::stack(year_list)

prop_yr_exceed <- calc(year_stack, mean, na.rm=TRUE)
prop_yr_exceed <- prop_yr_exceed/52

writeRaster(prop_yr_exceed, "/home/shares/ohi/stressors_2021/_dataprep/SST/pressure_twoSD/sst-extreme_NA_historical.tif", format='GTiff', overwrite=TRUE)


## scenarios

for(period in c("current", "near","medium", "long")){
  # period = "current"
for(scenario in ssps){ #scenario = ssps[2]
year_list <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/year_exceed_two_sd", pattern = paste(yr_range$year[yr_range$period==period], collapse="|"), full=TRUE)
  
year_list <- grep(scenario, year_list, value=TRUE)

year_stack <- raster::stack(year_list)

prop_yr_exceed <- calc(year_stack, mean)
prop_yr_exceed <- prop_yr_exceed/52

writeRaster(prop_yr_exceed, sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/pressure_twoSD/sst-extreme_%s_%s.tif", scenario, period), format='GTiff', overwrite=TRUE)
}
}


```

# final formatting
1. gapfill to get coastline
2. project raster to appropriate resolution

```{r}

rast_base_10km_file <- terra::rast(here('_spatial/rast_base_mol_10km.tif'))
ocean_template <- terra::rast(here('_spatial/ocean_area_mol.tif'))
ocean_template[ocean_template>0] <- 1
plot(ocean_template)

exceed_paths <- list.files("/home/shares/ohi/stressors_2021/_dataprep/SST/pressure_twoSD", full=TRUE)

for(exceed_path in exceed_paths){ # exceed_path = exceed_paths[1]
exceed_path <- exceed_path
exceed_rast <- terra::rast(exceed_path)

exceed_rast <- terra::rotate(exceed_rast)

terra::ext(exceed_rast) <- c(-180, 180, -90, 90)
#plot(exceed_rast)
#maps::map('world', col='black', fill=FALSE, border='gray80', add=TRUE, size=0.5)

mol_rast <- terra::project(exceed_rast, rast_base_10km_file, method="near", over=TRUE)
#plot(mol_rast)


## gapfill
gf_raster <- function(x){raster::focal(x, w = matrix(1,3,3), fun = mean, na.rm=TRUE, na.policy="only")}

r = mol_rast
  
## Repeat 100 times (most is immediately gapfilled but the antarctic area was persistent)
i <- 0
while (i <= 50){
r <- gf_raster(r)
i <- i + 1
print(i)
}

# plot(r)
r <- r*ocean_template
# diff_rast <- r - mol_rast
#plot(diff_rast)

writeRaster(r, filename = sprintf("/home/shares/ohi/stressors_2021/_dataprep/SST/final_extreme/%s", basename(exceed_path)), overwrite = TRUE)
}

```



