---
title: "NPP pressure"
format: html
editor: visual
---

## NPP

The net primary productivity data are bio-oracle data from here:

https://bio-oracle.org/downloads-to-email.php

units: mmol . m-3

average of a given decade, and long-term average of the yearly maxima and minima of a given decade (e.g., the average temperature of the warmest month in the period 2000-2010).

Resolution: 0.05 degree resolution

For benthic: because focal cells at 0.05 degree resolution comprise a wide range of depth values, the benthic layers were developed for the minimum, average and maximum depth within focal cells.

Downloaded June 11 2024.

```{r}
library(tidyverse)
library(terra)
library(ncdf4)
library(biooracler)
library(here)
# devtools::install_github("bio-oracle/biooracler")
```

Downloading the data (https://github.com/bio-oracle/biooracler):

```{r}
data.frame(list_layers())

list_layers("Phytoplankton")$dataset_id
dataset_id <- "phyc_baseline_2000_2020_depthsurf"
info_layer(dataset_id)


# make a function to download the data from site
# (called by getdata function after this)
layer_download <- function(time, latitude, longitude, dataset_id){
constraints = list(time, latitude, longitude)
names(constraints) = c("time", "latitude", "longitude")
variables = paste0(variable, "_ltmin")
layers <- download_layers(dataset_id, variables, constraints)
return(layers)
}

# make a function to get oxygen and depth data for each time period.
getdata <- function(variable = "o2", depth = "depthmean"){

latitude = c(-89.975, 89.975)
longitude = c(-179.975, 179.975)

 scenarios = c("ssp119", "ssp126", "ssp245", "ssp370", "ssp460", "ssp585")

#historical
dataset_id <- paste0(variable, "_baseline_2000_2018_", depth)
time = c('2000-01-01T00:00:00Z', '2000-01-01T00:00:00Z')

layers <- layer_download(time=time, latitude=latitude, longitude = longitude, dataset_id = dataset_id)

writeRaster(layers, filename = file.path("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw", paste0(variable, "-", depth, "_hist_hist.tif")), overwrite=TRUE)

#current
dataset_id <- paste0(variable, "_baseline_2000_2018_", depth)
time = c('2010-01-01T00:00:00Z', '2010-01-01T00:00:00Z')

layers <- layer_download(time=time, latitude=latitude, longitude = longitude, dataset_id = dataset_id)

writeRaster(layers, filename = file.path("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw", paste0(variable, "-", depth, "_current_current.tif")), overwrite=TRUE)

# scenarios
for(scenario in scenarios){ # scenario=scenarios[3]
  
dataset_id <- paste0(variable, "_", scenario, "_2020_2100_", depth)

##near
time = c('2020-01-01T00:00:00Z', '2030-01-01T00:00:00Z')
layers <- layer_download(time=time, latitude=latitude, longitude = longitude, dataset_id = dataset_id)
rmean <- app(layers, mean)
writeRaster(rmean, filename = file.path("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw", paste0(variable, "-", depth, "_", scenario, "_near.tif")), overwrite=TRUE)
  
##medium
time = c('2040-01-01T00:00:00Z', '2050-01-01T00:00:00Z')
layers <- layer_download(time=time, latitude=latitude, longitude = longitude, dataset_id = dataset_id)
rmean <- app(layers, mean)
writeRaster(rmean, filename = file.path("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw", paste0(variable, "-", depth, "_", scenario, "_medium.tif")), overwrite=TRUE)

##long
time = c('2080-01-01T00:00:00Z', '2090-01-01T00:00:00Z')
layers <- layer_download(time=time, latitude=latitude, longitude = longitude, dataset_id = dataset_id)
rmean <- app(layers, mean)
writeRaster(rmean, filename = file.path("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw", paste0(variable, "-", depth, "_", scenario, "_long.tif")), overwrite=TRUE)
  
}
}


#### This will update the data in the folder!
# surface data
#getdata(variable = "o2", depth="depthsurf")
# benthic (average depth) data
#getdata(variable = "o2", depth="depthmean")

```

# Rescaling

```{r}
min_threshold <- 61 
max_threshold <- 183

# min_threshold <- 2 
# max_threshold <- 6

o2rescale <- function(fp_current = files2scale[20], 
                      historic = historic_rast){
  current <- rast(fp_current)
  
  #. current = 3
  # historic = 7
  stress_1_boolean <- ifel(current <= min_threshold, 1, 0)
  stress_2_boolean <- ifel(current >min_threshold & current<max_threshold, 1, 0)
  stress_3_boolean <- ifel(current >=max_threshold, 1, 0)
  
  stress_2_c_rescale <- (current - min_threshold)/(max_threshold - min_threshold)
  stress_2_h_rescale <- (historic - min_threshold)/(max_threshold - min_threshold)
  stress_2 <- (1 - stress_2_c_rescale/stress_2_h_rescale)*stress_2_boolean
  stress_2 <- ifel(stress_2<0, 0, stress_2)
  stress_2 <- ifel(stress_2>1, 1, stress_2)
  
  stress_3 <- (1 - current/historic) * stress_3_boolean
  stress_3 <- ifel(stress_3<0, 0, stress_3)
  stress_3 <- ifel(stress_3>1, 1, stress_3)

  stress <- stress_1_boolean + stress_2 + stress_3
  return(stress)
} 


# deep water
files2scale <- list.files("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw", pattern = "depthmean", full=TRUE)

for(file2scale in files2scale){ # file2scale <- files2scale[1]
historic_rast <- rast("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw/o2-depthmean_hist_hist.tif")
stress_rescale <- o2rescale(fp_current = file2scale, 
          historic = historic_rast)
writeRaster(stress_rescale, file.path("/home/shares/ohi/stressors_2021/_dataprep/oxygen/rescaled", basename(file2scale)), overwrite=TRUE)
}

# surface water
files2scale <- list.files("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw", pattern = "depthsurf", full=TRUE)

for(file2scale in files2scale){ # file2scale <- files2scale[1]
historic_rast <- rast("/home/shares/ohi/stressors_2021/_dataprep/oxygen/raw/o2-depthsurf_hist_hist.tif")
stress_rescale <- o2rescale(fp_current = file2scale, 
          historic = historic_rast)
writeRaster(stress_rescale, file.path("/home/shares/ohi/stressors_2021/_dataprep/oxygen/rescaled", basename(file2scale)), overwrite=TRUE)
}

```

Final formatting

```{r}
rast_base_10km_file <- terra::rast(here('_spatial/rast_base_mol_10km.tif'))
ocean_template <- terra::rast(here('_spatial/ocean_area_mol.tif'))
ocean_template[ocean_template>0] <- 1
plot(ocean_template)

oxy_paths <- list.files("/home/shares/ohi/stressors_2021/_dataprep/oxygen/rescaled", full=TRUE)

for(oxy_path in oxy_paths){ # oxy_path = oxy_paths[1]
oxy_rast <- terra::rast(oxy_path)

mol_rast <- terra::project(oxy_rast, rast_base_10km_file, method="bilinear", over=TRUE)
#plot(mol_rast)

## gapfill
gf_raster <- function(x){raster::focal(x, w = matrix(1,3,3), fun = mean, na.rm=TRUE, na.policy="only")}

r = mol_rast
  
## Repeat 50 times (most is immediately gapfilled but the antarctic area was persistent)
i <- 0
while (i <= 50){
r <- gf_raster(r)
i <- i + 1
print(i)
}

# plot(r)
r <- r*ocean_template
# diff_rast <- r - mol_rast
#plot(diff_rast) # should be 0 

writeRaster(r, filename = sprintf("/home/shares/ohi/stressors_2021/_dataprep/oxygen/rescaled_mol/%s", basename(oxy_path)), overwrite = TRUE)
}

```

Quick look at some!

```{r}
raws <- list.files("/home/shares/ohi/stressors_2021/_dataprep/oxygen/rescaled_mol", full=TRUE)

tmp <- rast(oxy_paths[10]) - rast(oxy_paths[1])
tmp <- ifel(tmp>0.5, 0.5, tmp)
plot(tmp[tmp<1])



```
