---
title: "Set up master rasters"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/src/templates/ohara_hdr.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(sf)
library(fasterize)
library(here)

```

# Summary

Set up rasters for analysis.

* Master rasters at 10 x 10 km and 1 x 1 km resolution Mollweide CRS.
    * 10 x 10 km raster will double as a cell ID raster.
* Ocean area proportion $\in$ (0, 1], 10 km x 10 km Mollweide
    * create global raster from NE ocean at 1 km x 1 km resolution; aggregate up to 10 km x 10 km
* Coastal area proportion $\in$ (0, 1], 10 km x 10 km Mollweide
    * create global raster from buffered land, at 1 x 1 km resolution; aggregate up to 10 km x 10 km.
    * can also create one at 1 km buffer.
    * do we need inland buffers?
* Using the 10 km x 10 km ocean area raster as a base, create:
    * EEZ raster

# Data

Land and ocean shapefiles from Natural Earth.  Free vector and raster map data @ naturalearthdata.com.

# Methods

## Master raster 1 km and 10 km

``` {r}
moll_crs <- '+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m'

rast_base_10km_file <- here('_spatial/rast_base_mol_10km.tif')

if(!file.exists(rast_base_10km_file)) {
  x <- raster() %>%
    setValues(1)

  rast_base_10km <- projectRaster(x, crs = moll_crs, res = 10000, progress="text")

  writeRaster(rast_base_10km, rast_base_10km_file)
}

### disaggregate 10 x 10 km cells to 1 x 1 km cells
rast_base_1km_file <- here('_spatial/rast_base_moll_1km.tif')

if(!file.exists(rast_base_1km_file)) {
  rast_base_10km <- raster(rast_base_10km_file)

  rast_base_1km <- disaggregate(rast_base_10km, fact = 10, progress = 'text')
  
  writeRaster(rast_base_1km, rast_base_1km_file)
}

```

## Ocean area raster

* Reproject NE ocean shapefile to Mollweide.
* Use `fasterize::fasterize` to rasterize the reprojected ocean shapefile to Mollweide CRS with a 1 km resolution.
    * This will be saved on the server, too large for Github.
    * Values will be 1 (ocean) and 0 (not ocean).
* Aggregate the 1 km raster to 10 km cells, taking the sum of ocean cells and divide by 100 to result in values $\in$ (0, 1].  Cells with no ocean presence will result in NA.

```{r ocean 1 km}

ocean_1km_file <- here('_spatial/ocean_moll_1km.tif')
# unlink(ocean_1km_file)

if(!file.exists(ocean_1km_file)) {
  message('Processing ocean area at 1 km resolution...')
  ocean_sf <- read_sf(here('_spatial', 'ne_10m_ocean', 'ne_10m_ocean.shp'))
  
  ocean_moll_sf <- ocean_sf %>%
    st_transform(crs = moll_crs)
  
  rast_base_moll_1km <- raster(rast_base_1km_file)
  
  ocean_moll_1km <- fasterize::fasterize(ocean_moll_sf, rast_base_1km, 
                                         field = NULL, ### NULL sets all polygons to 1
                                         background = NA)
    
  writeRaster(ocean_moll_1km, filename = ocean_1km_file)
}
```

```{r ocean 10 km}
ocean_10km_file <- here('_spatial', 'ocean_area_mol.tif')
# unlink(ocean_10km_file)

if(!file.exists(ocean_10km_file)) {
  message('Processing ocean area at 10 km resolution...')
  ocean_moll_1km <- raster(ocean_1km_file)
  
  ocean_moll_10km <- aggregate(ocean_moll_1km, fact = 10, fun = sum, progress = 'text')
  
  # rast_base_10km <- raster(rast_base_10km_file)
  # compareRaster(ocean_moll_10km, rast_base_10km)
  # [1] TRUE
  
  ### drop the Caspian Sea:
  casp_bbox <- c(xmin = 3700000, xmax = 4800000, ymin = 4300000, ymax = 5650000)
  ocean_moll_df <- as.data.frame(ocean_moll_10km, xy = TRUE) %>%
    setNames(c('x', 'y', 'z')) %>%
    mutate(caspian_x = between(x, casp_bbox['xmin'], casp_bbox['xmax']),
           caspian_y = between(y, casp_bbox['ymin'], casp_bbox['ymax']),
           caspian   = caspian_x & caspian_y) %>%
    mutate(z = ifelse(caspian, NA, z)) %>%
    select(-contains('caspian'))
  
  ocean_moll_10km_no_casp <- rasterFromXYZ(ocean_moll_df, res = 10000, crs = crs(ocean_moll_10km))
  
  # plot(ocean_moll_10km_no_casp)
  ocean_moll_10km_no_casp <- ocean_moll_10km_no_casp / 100
  
  writeRaster(ocean_moll_10km_no_casp, ocean_10km_file, overwrite = TRUE)
}

ocean_moll_10km <- rast(ocean_10km_file)
plot(ocean_moll_10km, main = 'ocean area')
```


## EEZ raster

``` {r set up EEZ raster}

eez_rast_file <- here('_spatial', 'eez_mol.tif')

if(!file.exists(eez_rast_file)) {
  
  eez_shp_file  <- file.path(here('_spatial/rgns_mol_1k.gpkg'))
  eez_sf <- st_read(eez_shp_file)

  eez_df <- eez_sf %>%
    as.data.frame() %>%
    dplyr::select(-geom)
  eez_ids <- eez_df %>%
    select(rgn_id, rgn_name, rgn_ant_id) %>%
    distinct()
  ### rgn_ant_id is same as rgn_id except for Antarctica.  We don't really
  ### need to chunk out Antarctica at this point so just use rgn_id.

  ocean_moll_10km <- raster(ocean_10km_file)
  
  eez_rast <- rasterize(eez_sf, ocean_moll_10km,
                                   field = 'rgn_id', 
                                   fun = 'min')
  
  writeRaster(eez_rast, here("_spatial/eez_mol_w_land.tif"))
  
  eez_rast_ocean <- eez_rast %>%
    mask(ocean_moll_10km)
  
  writeRaster(eez_rast_ocean, eez_rast_file, overwrite = TRUE)
  
}
eez_rast_ocean <- rast(eez_rast_file)
plot(eez_rast_ocean, main = 'EEZ raster 10 km^2 Mollweide projection')


```

