---
title: "Indicators of fisheries governance"
output: html_document
date: "2023-10-24"
---
Another index: https://www.nature.com/articles/s41598-023-37048-6
fishing effort
catch/productivity
change in trophic level

Governance
Where: https://github.com/marina-andrijevic/governance2019
https://www.nature.com/articles/s41893-019-0405-0
```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(countrycode)
library(sf)
library(fasterize)

```

Estimate current average pressure:
```{r}

template <- terra::rast(here("stressors/prep/fisheries/data/tonnes_catch.tif"))
zones_shp <- sf::st_read("/home/shares/ohi/git-annex/globalprep/spatial/v2017/regions_2017_update.shp") 
zones_lat_lon <- sf::st_transform(zones_shp, 4326)
zones_raster <- fasterize(zones_lat_lon, raster::raster(template), field="rgn_id")

rgn_data <- read.csv("/home/shares/ohi/git-annex/globalprep/spatial/d2014/data/rgn_mol_raster_1km/regionData.csv")

```

## country scale

catch (tonnes, Watson) and npp (1e-9 mol/m2/s)
1 x 10-9 mol m-2s-1 total organic carbon is about equal to 378.432 tonnes of organic carbon per yer per km2
```{r}

catch <- terra::rast(here("stressors/prep/fisheries/data/tonnes_catch.tif"))
#biomass <- terra::rast(here("prep/fisheries/data/tonnes_biomass.tif"))

npp_tmp <- terra::rast("/home/shares/ohi/stressors_2021/_dataprep/npp/ensembles/npp_ssp126_historical.tif")
npp_crs <- terra::resample(npp_tmp, catch, method = "bilinear")
npp_gf <- npp_crs
for(i in 1:200){
  i=1+i
npp_gf <- terra::focal(npp_gf, w=3, fun=mean, na.rm=TRUE, na.policy="only")
}

ocean <- terra::rast(here("stressors/_spatial/ocean_area_mol.tif"))
ocean[ocean >0] <- 1
ocean <- terra::project(ocean, npp_gf, method="max")

npp_mask <- terra::mask(npp_gf, ocean)
npp_mask <- npp_mask*378.432 # convert to tonnes per km2 per year

npp <- npp_mask*terra::cellSize(npp_mask, unit="km") # Units are area dependent, want total throughout EEZ.
biomass <- terra::rast(here("stressors/prep/fisheries/data/tonnes_biomass.tif"))

c_to_b <- catch/biomass
c_to_b <- terra::ifel(c_to_b > 2, 2, c_to_b)
plot(c_to_b)

plot(log(biomass+1)/log(npp +1))
plot(log(catch + 1))
plot(npp)
plot(log(npp +1))
plot(log(catch+1)/log(biomass+1))
plot(catch/npp)


ln_npp <- log(npp +1)
ln_catch <- log(catch + 1)
ln_biomass <- log(biomass +1)

stack_lns <- c(ln_npp, ln_catch, ln_biomass)
stack_df <- as.data.frame(stack_lns)
names(stack_df) <- c("ln_npp", "ln_catch", "ln_biomass")
mod <- lm(ln_biomass ~ ln_npp, data=stack_df)
summary(mod)

mod <- lm(ln_catch ~ ln_biomass, data=stack_df)
summary(mod)
plot(stack_df$ln_biomass ~ stack_df$ln_npp)

stack_o_rast <- c(catch, npp, biomass)



data1 <- terra::zonal(stack_o_rast, terra::rast(zones_raster), fun = sum, na.rm=TRUE)

names(data1) <- c("rgn_id", "catch", "npp", "biomass")
plot(log(data1$catch+1) ~ log(data1$npp +1))

data1 <- left_join(data1, rgn_data, by="rgn_id")
data1$iso3c <- countrycode(data1$rgn_nam, "country.name", "iso3c") 

# cut out uninhabited (or, mostly uninhabited) islands. These have
# fisheries and should be modeled, but will need a different model.
data1 <- data1 %>%
  filter(!is.na(iso3c),
         iso3c!= "ATA", 
         rgn_nam != "Bassas da India", 
         iso3c != "IOT",  # British Indian Ocean Territory
         iso3c != "SGS",  # South Georgia and the South Sandwich Islands 
         iso3c != "HMD",  # Heard and McDonald Islands
         iso3c != "BVT")  # Bouvet Island
sum(duplicated(data1$iso3c)) 

data1 <- data1 %>%
mutate(prop_catch_npp = catch/npp) %>%
  mutate(prop_catch_biomass = catch/biomass)

summary(data1)
# losing Slovenia and Jordan due to very small EEZs. Not much to do about that, unless we want to do a smaller resolution, but this seems unadviseable because none of the data are at this resolution.
```

Information on islands:
islands <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/rgn_uninhabited_islands.csv") %>%
  select(rgn_id) %>%
  mutate(island = "island")



Add in GDP and population.
https://tntcat.iiasa.ac.at/SspDb/dsd?Action=htmlpage&page=60

```{r}

gdp <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "GDP|PPP") %>%
  filter(MODEL == "OECD Env-Growth") %>%
  filter(SCENARIO == "SSP1_v9_130325") %>%
  select(iso3c = REGION, gdp_ppp = `2020`)

pop <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/SspDb_country_data_2013-06-12.csv") %>%
  filter(VARIABLE == "Population") %>%
  filter(SCENARIO == "SSP1_v9_130115") %>%
  select(MODEL, SCENARIO, REGION, `2020`) %>%
  pivot_wider(names_from = MODEL, values_from = `2020`) %>%
  mutate(NCAR = ifelse(is.na(NCAR), `IIASA-WiC POP`, NCAR)) %>% # estimates from two models are nearly the same, will use NCAR, but gapfill with IIASA 
  select(iso3c = REGION, population = NCAR)
  
data2 <- left_join(data1, gdp, by="iso3c") %>%
         left_join(pop, by="iso3c") 

```

Gini and poverty indices
```{r}
gini <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/NRao_et_al_GiniProjections_2018/Gini_projections_SSPs.csv") %>%
  filter(year == 2020) %>%
  filter(scenario == "SSP1") %>%
  pivot_longer( cols = -c(scenario, year),  
  names_to = "iso3c",
  values_to = "gini") %>%
  select(iso3c, gini)

# poverty ends up being more complicated because the level shifts depending on the income of the country based on GNI (GDP + foreign)
#poverty <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/country_pop_gini_IIASA/NRao_et_al_GiniProjections_2018/Poverty_headcounts_SSPs.csv") %>%
#
#  filter(year == 2020) %>%
#  filter(scenario == "SSP1") 

data3 <- left_join(data2, gini, by="iso3c") 

```

## Governance
https://www.nature.com/articles/s41893-019-0405-0
https://github.com/marina-andrijevic/governance2019/tree/master

```{r}
gov <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/governance_andrijevic/governance2019/data/master_proj_obs.csv") %>%
  filter(year==2015 & scenario=="Observed") %>%
  rename("iso3c" = "countrycode") %>%
  select(iso3c, governance)


data4 <- left_join(data3, gov, by="iso3c")

```

## eez area
```{r}

eez_area <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/rgn_area.csv")

data5 <- left_join(data4, eez_area, by="rgn_id")


```

## gapfilling!
```{r}

# current data from Wikipedia, will need to estimate future scenarios using simple models. Also need to adjust units to match above data. 

# for future, the other data is billion US$ 2005, so I will need to adjust this if I move forward with this method.

gf_regions <- read_csv(here("stressors/prep/fisheries/data/country_pop_gf.csv")) %>%
  select(iso3c, pop_gf = population, gdp_gf = gdp_current_USD) %>%
  mutate(pop_gf = pop_gf/1000000,
         gdp_gf = gdp_gf/1000000000)

data6 <- left_join(data5, gf_regions, by="iso3c") %>%
  mutate(population = ifelse(is.na(population), pop_gf, population),
         gdp_ppp = ifelse(is.na(gdp_ppp), gdp_gf, gdp_ppp)) %>%
         mutate(pc_gdp = gdp_ppp/population) %>%
  select(-rgn_typ, -ant_id, -pop_gf, -gdp_gf)

georgns <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/georegion_labels.csv")

data7 <- left_join(data6, georgns, by="rgn_id")

mod_gov1 <- lm(governance ~ pc_gdp + r2_label, data=data7, na.action=na.exclude)
summary(mod_gov1)

data7$gov_gf1 <- predict(mod_gov1, newdata = data7 %>%
                                               mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

mod_gov2 <- lm(governance ~ pc_gdp + r1_label, data=data7, na.action=na.exclude)
summary(mod_gov2)

data7$gov_gf2 <- predict(mod_gov2, newdata = data7)

data7 <- data7 %>%
  mutate(governance = ifelse(is.na(governance), gov_gf1, governance),
         governance = ifelse(is.na(governance), gov_gf2, governance))

mod_gini1 <- lm(gini ~ pc_gdp + governance + r2_label, data=data7, na.action=na.exclude)
summary(mod_gini1)

data7$gini_gf1 <- predict(mod_gini1, newdata = data7 %>%
                                               mutate(r2_label = ifelse(r2_label == "Micronesia", NA, r2_label)))

mod_gini2 <- lm(gini ~ pc_gdp + governance + r1_label, data=data7, na.action=na.exclude)
summary(mod_gini2)

data7$gini_gf2 <- predict(mod_gini2, newdata = data7)

data7 <- data7 %>%
  mutate(gini = ifelse(is.na(gini), gini_gf1, gini),
           gini = ifelse(is.na(gini), gini_gf2, gini)) %>%
  select(-gini_gf1, -gini_gf2, -gov_gf1, -gov_gf2, -ant_typ)

```

Predict catch at country level.
```{r}
library(ranger)

library(caret)
# Set up training control
train_control <- trainControl(method = "LOOCV", savePredictions=TRUE)

# Train the model with leave-one-out cross-validation
data8 <- na.omit(data7)
data8 <- data8 %>%
  mutate(ln_catch = log(catch + 1),
         ln_npp = log(npp + 1),
         ln_gdp_ppp = log(gdp_ppp+1),
         ln_biomass = log(biomass + 1),
         catch_to_biomass = catch/biomass, 
         ln_population = log(population + 1),
         ln_catch_ln_npp = log(catch +1)/log(npp + 1),
         ln_pop_ln_area = log(population)/log(area_km2),
         ln_catch_ln_biomass = log(catch + 1)/ log(biomass +1))

model <- train(ln_catch_ln_biomass ~  ln_pop_ln_area + gini + governance, 
               data = na.omit(data8), 
               method = "lm", 
               trControl = train_control)
model
plot(model$pred$pred ~ model$pred$obs)




model <- train(ln_catch_ln_npp ~  ln_gdp_ppp + r1_label + ln_pop_ln_area, 
               data = na.omit(data8), 
               method = "lm", 
               trControl = train_control)
model


plot(model$pred$pred ~ model$pred$obs)
sum(exp(model$pred$pred))
sum(exp(model$pred$obs))

# gini, area_km2, pc_gdp, npp, gdp_ppp, 
rf_model <- ranger(
  formula         = catch ~ npp + gdp_ppp + population + governance + area_km2, 
  data            = na.omit(data7),
  num.trees       = 500,  # Number of trees to grow - can be adjusted
  importance      = 'impurity', # Variable importance based on impurity
  write.forest    = TRUE # If you want to make predictions later
)

rf_model
```

# other indicators, but no future data
fi <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/global_fishing_index/Global Fishing Index 2021 Data for Download V1.1.csv") 

Get the Global Fishing Index. This includes several indicators, but I am excluding the indicators describing how well the stock is doing. I am more interested in how the government is performing. I include governance capacity (score on several indicators and an assessment of evenness) and data availability on the stocks.  
I rescaled and averaged these to get an average indicator.
```{r}
fi <- read_csv("/home/shares/ohi/stressors_2021/_raw_data/global_fishing_index/Global Fishing Index 2021 Data for Download V1.1.csv") %>%
  rename("iso3c" = `ISO Code`,
         "gfi_governance" =  "Governance capacity",
         "data_availability" = "Data availability (%)") %>%
  select(iso3c, gfi_governance, data_availability) %>%
  mutate(gfi_gov_rescaled = gfi_governance/12) %>%
  mutate(data_avail_rescaled = data_availability/100) %>%
  mutate(average_index = (gfi_gov_rescaled+data_avail_rescaled)/2)

```

Dependence on fisheries
https://conbio.onlinelibrary.wiley.com/doi/full/10.1111/conl.12617
```{r}

depend <- read_csv(here("prep/fisheries/marine_dependence_Selig.csv")) %>%
  select(country = Country, econ_depend = `Economic dependence`, 
         nutrition_depend = `Nutritional dependence`) %>%
  slice(-1)

depend$iso3c <- countrycode(depend$country, "country.name", "iso3c") 


```    


OHI fishery status score
```{r}

scores <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohi-global/draft/yearly_results/global2023/OHI_final_formatted_scores_2023-10-02.csv") %>%
  filter(goal == "FIS" & scenario==2017 & dimension == "status") %>%
  rename("country" = region_name)

scores$iso3c <- countrycode(scores$country, "country.name", "iso3c") 

```

Coastal population
```{r}

pop <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/mar_prs_population/v2021/output/mar_pop_25mi.csv") %>%
  filter(year==2017)
area <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2019/output/area_km2_25mi_inland.csv") 
regions <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/spatial/v2017/output/regionData.csv") %>%
  filter(rgn_type=="eez") %>%
  select(rgn_id, country=rgn_name) %>%
  unique()

pop_density <- left_join(area, pop) %>%
  left_join(regions) %>%
  mutate(density = popsum/area_km2)

pop_density$iso3c <- countrycode(pop_density$country, "country.name", "iso3c") 


```

world government indicator
```{r}

wgi <- read_csv("https://raw.githubusercontent.com/OHI-Science/ohiprep_v2023/gh-pages/globalprep/prs_res_wgi/v2023/intermediate/wgi_combined_scores_by_country.csv") %>%
  filter(year==2018)
wgi$iso3c <- countrycode(wgi$country, "country.name", "iso3c") 
```

