---
title: "RAM_spatial"
output: html_document
date: "2024-02-16"
---
We have the original spatial data from Chris Free, but many RAM stocks no longer match.

Here we are determining which stocks we have spatial data for, and working to get spatial data for the remaining.

RAM regions are from this source (lots of weirdness, so check carefully):
https://zenodo.org/records/834755

```{r}

library(leaflet)
library(sf)
library(tidyverse)
library(here)

## ram data (area and metadata)
load(file.path("/home/shares/ohi/stressors_2021/_raw_data/fisheries_RAMLDBv4_64/R Data/DBdata[asmt][v4.64].RData"))

## created in ram_prep_gf.Rmd
ram_biomass_gf2 <- read_csv(here("prep/fisheries/data/RAM_gapfilled.csv"))

## eez regions
eez <- st_read("/home/shares/ohi/stressors_2021/_raw_data/eez/eez_v12.shp")
plot(st_geometry(eez))
#eez <- st_read("/home/shares/ohi/git-annex/globalprep/spatial/v2017/regions_2017_update.shp")
#eez <- st_transform(eez, crs = 4326)

map <- leaflet(filter(eez, GEONAME=="United States Exclusive Economic Zone (Alaska)")) %>%
  addProviderTiles("OpenStreetMap") %>%
  addPolygons(    
    fillColor = "orange", # Example: color by 'AREA' column
    weight = 2,
    opacity = 1,
    color = 'blue',
    dashArray = '3',
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = '',
      fillOpacity = 0.7,
      bringToFront = TRUE),
    popup = ~paste("<strong>SP_ID:</strong>",  GEONAME) # And this one as per your data
  )
map

### meow regions
meow <- st_read(here("_spatial/meow_rgns/meow_rgns.shp"))

# ### ram regions:  (don't think this is proving to be useful)
# ram_rgns <- st_read("/home/shares/ohi/stressors_2021/_raw_data/RAM_spatial/RAM_Geography/results/ram.shp") %>%
#   mutate(SP_ID = as.numeric(SP_ID))
# #I created this data frame:
# ram_rgns_names <- read_csv(here("prep/fisheries/data/reglocs_updated.csv")) %>%
#   select(names, size = sizes, lat=lats, lon=lons, SP_ID)
# 
# ram_rgns <- left_join(ram_rgns, ram_rgns_names, by = "SP_ID")
# 
# map <- leaflet(ram_rgns) %>%
#   addProviderTiles("OpenStreetMap") %>%
#   addPolygons(    
#     fillColor = "orange", # Example: color by 'AREA' column
#     weight = 2,
#     opacity = 1,
#     color = 'blue',
#     dashArray = '3',
#     fillOpacity = 0.7,
#     highlightOptions = highlightOptions(
#       weight = 5,
#       color = "#666",
#       dashArray = '',
#       fillOpacity = 0.7,
#       bringToFront = TRUE),
#     popup = ~paste("<strong>SP_ID:</strong>",  SP_ID) # And this one as per your data
#   )
# map
# Print the map
iho <- st_read(here("_spatial/World_Seas_IHO_v3/World_Seas_IHO_v3.shp"))
##

ices <- st_read("/home/shares/ohi/stressors_2021/_raw_data/ICES_areas/ICES_Areas_20160601_cut_dense_3857.shp")
ices <- st_transform(ices, crs = 4326)
ices <- st_make_valid(ices)
mapview::mapview(ices)

nafo <- st_read("/home/shares/ohi/stressors_2021/_raw_data/NAFO_areas/NAFO_Divisions_2021.gdb")
nafo <- st_cast(nafo, "POLYGON")
nafo <- st_transform(nafo, crs = 4326)
mapview::mapview(nafo)
```

## Copy over RAM stocks

https://existencia.org/pro/?p=60

```{r}

## copy RAM spatial data to our folders
#RAM_spatial_dir <- file.path("/home/shares/ohi/git-annex/globalprep/_raw_data/RAM/d2017/ramldb_boundaries/ramldb_boundaries") 

#find missing spatial
#ram_sf <- list.files(RAM_spatial_dir, full=TRUE)

#dest_dir <- "/home/shares/ohi/stressors_2021/_raw_data/RAM_spatial/free_org/"
#file.copy(from=ram_sf, to=dest_dir)

tmp <- st_read("/home/shares/ohi/stressors_2021/_raw_data/RAM_spatial/free_org/CCSBT-SC-SBT-1952-2014-PONS.shp")
plot(tmp)

map <- leaflet(tmp) %>%
  addProviderTiles("OpenStreetMap") %>%
  addPolygons(    
    fillColor = "orange", # Example: color by 'AREA' column
    weight = 2,
    opacity = 1,
    color = 'blue',
    dashArray = '3',
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = '',
      fillOpacity = 0.7,
      bringToFront = TRUE),
    popup = ~paste("<strong>name:</strong>",  comm_name1) # And this one as per your data
  )
map

```

## Identify missing files
1. Determine which stocks we need spatial data for:
```{r}

# these are the current Chris Free spatial data files
free_loc <- file.path("/home/shares/ohi/stressors_2021/_raw_data/RAM_spatial/free_org") 
## put spatial data into one file
#find missing spatial
free_sf <- list.files(free_loc, pattern = "shp", full=TRUE)
free_sf <- grep(".xml", free_sf, value=TRUE, invert=TRUE)
tmp <- st_read(free_sf[1])

#Initialize an empty list to store sf objects
free_sf_objects <- list()

# Loop through the list of shapefile paths
for (shp_path in free_sf) {
  # Read the shapefile into an sf object
  sf_obj <- st_read(shp_path)
  sf_obj <- st_set_crs(sf_obj, 4326)
  # Add the sf object to the list
  free_sf_objects[[length(free_sf_objects) + 1]] <- sf_obj
}

# Combine all sf objects into one
free_ram <- bind_rows(free_sf_objects) %>%
  filter(!is.na(stockid))
dim(free_ram) # n = 664
length(unique(free_ram$stockid)) # n= 664


free_ram_df <- st_drop_geometry(free_ram)

length(unique(free_ram$stockid)) #n=664
length(unique(ram_biomass_gf2$stockid)) #n = 387
intersect(ram_biomass_gf2$stockid, free_ram$stockid) # N=261
# these are the ones I need to find (n = 126):
missing_area <- data.frame(stockid = setdiff(ram_biomass_gf2$stockid, free_ram$stockid)) %>%
  left_join(metadata, by="stockid") %>%  # n = 126
  left_join(area, by="areaname")  # n = 151
#write_csv(missing_area, here("prep/fisheries/data/ram_missing_boundary.csv"))

```

2. Determine the regions we have based on the Chris Free data
```{r}

intersect(ram_biomass_gf2$stockid, free_ram$stockid) # N=261
# these are the ones I need to find (n = 126):
with_area <- filter(free_ram, stockid %in% intersect(ram_biomass_gf2$stockid, free_ram$stockid)) %>%
  left_join(metadata, by="stockid") %>%  # n = 261
  left_join(area, by="areaname")  # n = 296

with_area$stockid[duplicated(with_area$stockid)]
filter(with_area, stockid=="AFLONCH")

plot(st_geometry(filter(with_area, stockid == "AFLONCH")))
plot(st_geometry(filter(ram_rgns, SP_ID == 126)), add=TRUE)
```


## Go through each missing stock
1. See if the region is available in Chris Free's ranges for another stock with the same location 
2. See if the region is available from the other data source.


```{r}

##SARDNWAZAB
## boundaries: https://firms.fao.org/firms/resource/10086/en#:~:text=In%20the%20north%20zone%2C%20the,by%20traditional%20coastal%20purse%20seiners.

rgn1 <- filter(eez, GEONAME %in% c("Portuguese Exclusive Economic Zone (Madeira)",
                                   "Spanish Exclusive Economic Zone (Canary Islands)",
                                   "Moroccan Exclusive Economic Zone", 
                                   "Overlapping claim Western Sahara: Western Sahara / Morocco", 
                                   "Mauritanian Exclusive Economic Zone"))

##RSROCKBCWS
can <- filter(eez, GEONAME == "Canadian Exclusive Economic Zone")
can_map <- mapview::mapview(can)
subset <- mapedit::editMap(can_map)
bc_south <- st_intersection(subset$drawn, st_make_valid(can_eez))
mapview::mapview(bc_south)

st_write(bc_south, here("prep/fisheries/data/maps/Canada-DFO-BCWS.shp"), append=FALSE)

##RSROCKBCWN
can <- filter(eez, GEONAME == "Canadian Exclusive Economic Zone")
can_map <- mapview::mapview(can) + mapview::mapview(bc_south)
subset <- mapedit::editMap(can_map)
bc_north <- st_intersection(subset$drawn, st_make_valid(can_eez))
mapview::mapview(bc_north)

st_write(bc_north, here("prep/fisheries/data/maps/Canada-DFO-BCWN.shp"), append=FALSE)


## REDFEAUS
# https://en.wikipedia.org/wiki/Eastern_states_of_Australia
tmp <- filter(eez, GEONAME == "Australian Exclusive Economic Zone")
tmp <- st_make_valid(tmp)
plot_tmp <- mapview::mapview(tmp) 
subset <- mapedit::editMap(plot_tmp)
aus_east <- st_intersection(subset$drawn, tmp)
mapview::mapview(aus_east)
st_write(aus_east, here("prep/fisheries/data/maps/Australia-AFMA-EAUS.shp"), append=FALSE)


## PLAIC2123
rgn1 <- filter(ices, SubDivisio %in% c("21","22", "23"))
region <- st_union(rgn1)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-2123.shp"), append=FALSE)


##PHFLOUNSOJ
# sea of japan
js <- filter(iho, NAME == "Japan Sea")
mapview::mapview(js)
st_write(js, here("prep/fisheries/data/maps/Japan-FAJ-SOJ.shp"), overwrite=TRUE)

## PCODAI
ak_eez <-filter(eez, GEONAME=="United States Exclusive Economic Zone (Alaska)")
meow_ai <- filter(meow, ECOREGION %in% c("Aleutian Islands"))

mapview::mapview(meow_ai) + 
  mapview::mapview(ak_eez)
ai <- st_intersection(meow_ai, ak_eez)
mapview::mapview(ai)

st_write(ai, here("prep/fisheries/data/maps/USA-NMFS-AI.shp"), overwrite=TRUE)
  
# PCODAI
ak_eez <-filter(eez, GEONAME=="United States Exclusive Economic Zone (Alaska)")
meow_bs <- filter(meow, ECOREGION %in% c("Eastern Bering Sea"))

mapview::mapview(meow_bs) + 
  mapview::mapview(ak_eez)
bs <- st_intersection(meow_bs, ak_eez)
mapview::mapview(bs)
st_write(bs, here("prep/fisheries/data/maps/USA-NMFS-BS.shp"), overwrite=TRUE)



##PCOD5ABCD
# https://publications.gc.ca/collections/collection_2018/mpo-dfo/fs70-6/Fs70-6-2018-020-eng.pdf
can_eez <- filter(eez, GEONAME == "Canadian Exclusive Economic Zone")
can_eez <- st_make_valid(can_eez)
can_map <- mapview::mapview(can_eez)
subset <- mapedit::editMap(can_map)
rgn5 <- st_intersection(subset$drawn, st_make_valid(can_eez))
mapview::mapview(rgn5)

st_write(rgn5, here("prep/fisheries/data/maps/Canada-DFO-5ABCD.shp"), append=FALSE)



## PCEELSCH
tmpCS <- st_read(here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-CS.shp"))
tmpC <- st_read(here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-C.shp"))
map_chile <- mapview::mapview(tmpCS) + mapview::mapview(tmpC)
subset <- mapedit::editMap(map_chile)
chile_s <- st_intersection(subset$drawn, tmpCS)
mapview::mapview(chile_s)

st_write(chile_s, here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-S.shp"), append=FALSE)

##PANDALIIIaW-IVaE
rgn1 <- filter(ices, SubArea == "3" & Division %in% c("a"))
rgn2 <- filter(ices, SubArea == "4" & Division %in% c("a"))
region <- rbind(rgn1, rgn2)
region <- st_union(region)
map_region <- mapview::mapview(region)

subset <- mapedit::editMap(map_region)
ices34 <- st_intersection(subset$drawn, region)
mapview::mapview(ices34)
st_write(ices34, here("prep/fisheries/data/maps/multinational-ICES-IIIaW-IVaE.shp"), append=FALSE)


## PANCHNCHSP
# Northern Chile Southern Peru
# https://en.wikipedia.org/wiki/Central_Chile
peru_eez <- filter(eez, GEONAME == "Peruvian Exclusive Economic Zone")
meow_south <- filter(meow, ECOREGION == "Humboldtian")
s_peru <- st_intersection(peru_eez, meow_south)

tmpN <- st_read(here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-N.shp"))
sp_nc <- st_union(st_geometry(s_peru), st_geometry(tmpN))
mapview::mapview(sp_nc)
st_write(sp_nc, here("prep/fisheries/data/maps/Chile-SUBPESCA-NCH-SP.shp"), append=FALSE)


## PANCHCCH
# https://en.wikipedia.org/wiki/Central_Chile
chile_eez <- filter(eez, GEONAME == "Chilean Exclusive Economic Zone")
tmpCS <- st_read(here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-CS.shp"))
tmpN <- st_read(here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-N.shp"))
CS_map <- mapview::mapview(tmpCS)
subset <- mapedit::editMap(CS_map)
c <- st_intersection(subset$drawn, tmpCS)
mapview::mapview(c)
st_write(c, here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-C.shp"), append=FALSE)



## NEPHMEDGSA17-18
# https://eur-lex.europa.eu/legal-content/EN/TXT/HTML/?uri=CELEX%3A52017SC0064
rgn <- filter(iho, NAME == "Adriatic Sea")
mapview::mapview(rgn)
st_write(rgn, here("prep/fisheries/data/maps/multinational-GFCM-GSA17-18.shp"), append=FALSE)


## NPOUTIIIa-IV
rgn1 <- filter(ices, SubArea == "4")
rgn2 <- filter(ices, SubArea == "3" & Division %in% c("a"))
region <- rbind(rgn1, rgn2)
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/NPOUTIIIa-IV.shp"), append=FALSE)



##MUTSNAPSATLCGM
satlc <- st_read(here("prep/fisheries/data/maps/USA-NMFS-SATLC.shp"))
gom <- st_read(here("prep/fisheries/data/maps/USA-NMFS-GM.shp"))
mapview::mapview(satlc) + mapview::mapview(gom)
satlcgom <- st_union(st_geometry(satlc), st_geometry(gom))
mapview::mapview(satlcgom)
st_write(satlcgom, here("prep/fisheries/data/maps/USA-NMFS-SATLCGM.shp"), append=FALSE)

## MORWONGWSE
# https://www.frdc.com.au/sites/default/files/products/2002-223-Marine%20Matters-%20Atlas%20of%20Australian%20Marine%20Fishing%20and%20Costal%20Communities.pdf
tmp <- filter(eez, GEONAME == "Australian Exclusive Economic Zone")
plot_tmp <- mapview::mapview(tmp) + 
  mapview::mapview(meow)
subset <- mapedit::editMap(plot_tmp)
wse <- st_intersection(subset$drawn, tmp)
mapview::mapview(wse)
st_write(wse, here("prep/fisheries/data/maps/Australia-AFMA-WSE.shp"), append=FALSE)
st_write(wse, here("prep/fisheries/data/maps/Australia-SARDI-WSE.shp"), append=FALSE)

## MORWONGESE
tmp <- filter(eez, GEONAME == "Australian Exclusive Economic Zone")
plot_tmp <- mapview::mapview(wse) + 
  mapview::mapview(meow)
subset <- mapedit::editMap(plot_tmp)
ese <- st_intersection(subset$drawn, tmp)
mapview::mapview(ese)
st_write(ese, here("prep/fisheries/data/maps/Australia-AFMA-ESE.shp"), append=FALSE)

## MEGVII-VIIIabd
rgn1 <- filter(ices, SubArea == "7")
rgn2 <- filter(ices, SubArea == "8" & Division %in% c("a", "b", "d"))
region <- rbind(rgn1, rgn2)
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-VII-VIIIabd.shp"), append=FALSE)


## LINGVa
rgn1 <- filter(ices, SubArea == "5" & Division %in% c("a"))
rgn1 <- st_union(rgn1)
mapview::mapview(rgn1)
st_write(rgn1, here("prep/fisheries/data/maps/multinational-ICES-Va.shp"), append=FALSE)


## KCROCKNPAC
iho_pac <- filter(iho, NAME == "North Pacific Ocean") 
iho_pac <- st_make_valid(iho_pac)
st_write(iho_pac, here("prep/fisheries/data/maps/Japan-FAJ-NPAC.shp"), append=FALSE)


## JPUFFIMKB
jpn_eez <- filter(eez, GEONAME == "Japanese Exclusive Economic Zone")
jpn_eez <- st_make_valid(jpn_eez)
jpn_map <- mapview::mapview(jpn_eez)
subset <- mapedit::editMap(jpn_map)
imb <- st_intersection(subset$drawn, jpn_eez)
mapview::mapview(imb)
st_write(imb, here("prep/fisheries/data/maps/Japan-FAJ-IMKB.shp"), append=FALSE)

## JMACKPJPN
jpn_eez <- filter(eez, GEONAME == "Japanese Exclusive Economic Zone")
jpn_eez <- st_make_valid(jpn_eez)
iho_pac <- filter(iho, NAME == "North Pacific Ocean") 
iho_pac <- st_make_valid(iho_pac)
jpn_pc <- st_intersection(iho_pac, jpn_eez)
mapview::mapview(jpn_pc)
st_write(jpn_pc, here("prep/fisheries/data/maps/Japan-FAJ-PJPN.shp"), append=FALSE)


## JAMBERJPN
grep("Japan", eez$GEONAME, value=TRUE)
tmp <- filter(eez, GEONAME == "Japanese Exclusive Economic Zone")
mapview::mapview(tmp)
st_write(tmp, here("prep/fisheries/data/maps/Japan-FAJ-JPN.shp"), append=FALSE)
st_write(tmp, here("prep/fisheries/data/maps/Japan-FAJ-JPNAR.shp"), append=FALSE)
st_write(tmp, here("prep/fisheries/data/maps/Japan-FAJ-JPNWR.shp"), append=FALSE)

## HMACKIIa-IVa-Vb-VIa-VII-VIII
rgn1 <- filter(ices, SubArea == "2" & Division %in% c("a"))
rgn2 <- filter(ices, SubArea == "4" & Division %in% c("a"))
rgn3 <- filter(ices, SubArea == "5" & Division %in% c("b"))
rgn4 <- filter(ices, SubArea == "6" & Division %in% c("a"))
rgn5 <- filter(ices, SubArea %in% c("7"))
rgn6 <- filter(ices, SubArea %in% c("8"))
region <- rbind(rgn1, rgn2, rgn3, rgn4, rgn5, rgn6)
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-IIa-IVa-Vb-VIa-VII-VIII.shp"), append=FALSE)


## HERRNS-IIIa-VIId
rgn1 <- filter(ices, SubArea == "3" & Division %in% c("a"))
rgn2 <- filter(ices, SubArea %in% c("4"))
rgn3 <- filter(ices, SubArea == "7" & Division %in% c("d"))
region <- rbind(rgn1, rgn2, rgn3)
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-IIIa-IV-VIId.shp"), append=FALSE)



## HERRHG
# https://www.pac.dfo-mpo.gc.ca/fm-gp/maps-cartes/areas-secteurs/index-eng.html#area-table
can_eez <- filter(eez, GEONAME == "Canadian Exclusive Economic Zone")
can_eez <- st_make_valid(can_eez)
can <- mapview::mapview(can_eez)
subset <- mapedit::editMap(can)
hg <- st_intersection(subset$drawn, can_eez)
mapview::mapview(hg)
st_write(hg, here("prep/fisheries/data/maps/Canada-DFO-HG.shp"), append=FALSE)


## HERR2529-32
rgn1 <- filter(ices, SubDivisio %in% c("30", "31"))
region <- st_union(rgn1)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-30-31.shp"), append=FALSE)


## HERR2529-32
rgn1 <- filter(ices, SubDivisio %in% c("25", "27", "28", "29", "32"))
region <- st_union(rgn1)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-2529-32.shp"), append=FALSE)

## DVANGSASC
grep("Namibia", eez$GEONAME, value=TRUE)
tmp <- filter(eez, GEONAME == "Namibian Exclusive Economic Zone")
mapview::mapview(tmp)
st_write(tmp, here("prep/fisheries/data/maps/Namibia-NMFMR-SAF.shp"), append=FALSE)



## HADNS-IIIa-VIa
rgn1 <- filter(ices, SubArea == "3" & Division %in% c("a"))
rgn2 <- filter(ices, SubArea %in% c("4"))
rgn3 <- filter(ices, SubArea == "6" & Division %in% c("a"))
region <- rbind(rgn1, rgn2, rgn3)
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-NS-IIIa-VIa.shp"), append=FALSE)


## GOLDREDV-VI-XII-XIV
rgn1 <- filter(ices, SubArea %in% c("5", "6", "12", "14"))
region <- st_union(rgn1)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-V-VI-XII-XIV.shp"), append=FALSE)


## EBASSIVbc-VII
rgn1 <- filter(ices, SubArea == "4" & Division %in% c("b", "c"))
rgn2 <- filter(ices, SubArea %in% c("7"))
region <- rbind(st_make_valid(rgn1), st_make_valid(rgn2))
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-IVbc-VII.shp"), append=FALSE)


## DVANGSASC
grep("South Africa", eez$GEONAME, value=TRUE)
tmp <- filter(eez, GEONAME == "South African Exclusive Economic Zone")
south <- filter(meow, ECOREGION == "Agulhas Bank")
south_af <- st_intersection(tmp, south)
mapview::mapview(south_af)
st_write(south_af, here("prep/fisheries/data/maps/South Africa-DEFF-SASC.shp"), append=FALSE)
st_write(south_af, here("prep/fisheries/data/maps/South Africa-DETMCM-SASC.shp"), append=FALSE)



## DVANGSAWC
grep("South Africa", eez$GEONAME, value=TRUE)
tmp <- filter(eez, GEONAME == "South African Exclusive Economic Zone")
west <- filter(meow, ECOREGION == "Namaqua")
west_af <- st_intersection(tmp, west)
mapview::mapview(west_af)
st_write(west_af, here("prep/fisheries/data/maps/South Africa-DEFF-SAWC.shp"), append=FALSE)
st_write(west_af, here("prep/fisheries/data/maps/South Africa-DETMCM-SAWC.shp"), append=FALSE)


## CUSKVa-XIV
tmp <- filter(ices, SubArea %in% c("5") & Division =="a")
tmp2 <- filter(ices, SubArea %in% c("14"))
region <- rbind(st_make_valid(tmp), st_make_valid(tmp2))
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-Va-XIV.shp"), append=FALSE)


## CODIIIaW-IV-VIId
tmp <- filter(ices, SubArea %in% c("4"))
tmp2 <- filter(ices, SubArea %in% c("3") & Division %in% c("a"))
tmp3 <- filter(ices, SubArea %in% c("7") & Division %in% c("d"))
region <- rbind(st_make_valid(tmp), st_make_valid(tmp2), st_make_valid(tmp3))
region <- st_union(region)
mapview::mapview(region)
st_write(region, here("prep/fisheries/data/maps/multinational-ICES-IIIaW-IV-VIId.shp"), append=FALSE)


## COD1IN
tmp <- filter(nafo, NAFO_SubArea == "1")
tmp <- st_union(tmp)
mapview::mapview(tmp)
st_write(tmp, here("prep/fisheries/data/maps/multinational-ICES-1IN.shp"), append=FALSE)


##COD1F-XIV
tmp <- filter(ices, SubArea %in% c("14"))
tmp2 <- filter(nafo, Label == "1F")
area <- c(st_geometry(tmp), st_geometry(tmp2))
area <- st_union(area)
mapview::mapview(area)
st_write(area, here("prep/fisheries/data/maps/multinational-ICES-1F-XIV.shp"), append=FALSE)

## COBGMT
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
tmp_iho <- filter(iho, NAME == "Gulf of Mexico")
mapview::mapview(tmp) + 
  mapview::mapview(tmp_iho)
gm <- st_intersection(st_geometry(tmp_iho), st_geometry(tmp))
mapview::mapview(gm)
st_write(gm, here("prep/fisheries/data/maps/USA-NMFS-GM.shp"), append=FALSE)


## CMACKPJPN
tmp <- filter(eez, GEONAME == "Japanese Exclusive Economic Zone")
np <- filter(iho, NAME == "North Pacific Ocean")
np <- st_make_valid(np)
pac_japan <- st_intersection(tmp, np)
mapview::mapview(pac_japan)
st_write(pac_japan, here("prep/fisheries/data/maps/Japan-FAJ-PJPN.shp"), append=FALSE)


## CHROCKNPCOAST
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
plot_tmp <- mapview::mapview(tmp) + 
  mapview::mapview(meow)
subset <- mapedit::editMap(plot_tmp)
npc <- st_intersection(subset$drawn, tmp)
mapview::mapview(npc)
st_write(npc, here("prep/fisheries/data/maps/USA-NMFS-NPCOAST.shp"), append=FALSE)

## CHROCKNPCOAST
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
plot_tmp <- mapview::mapview(tmp) + 
  mapview::mapview(meow)
subset <- mapedit::editMap(plot_tmp)
spc <- st_intersection(subset$drawn, tmp)
mapview::mapview(spc)
st_write(spc, here("prep/fisheries/data/maps/USA-NMFS-SPCOAST.shp"), append=FALSE)

## CHROCKCPCOAST
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
plot_tmp <- mapview::mapview(tmp) + 
  mapview::mapview(npc) + 
  mapview::mapview(spc)
subset <- mapedit::editMap(plot_tmp)
cpc <- st_intersection(subset$drawn, tmp)
mapview::mapview(cpc)
st_write(cpc, here("prep/fisheries/data/maps/USA-NMFS-CPCOAST.shp"), append=FALSE)

## BSQLOBSTERCSCH
grep("Chile", eez$GEONAME, value = TRUE)
tmp <- filter(eez, GEONAME == "Chilean Exclusive Economic Zone")
meow_plot <- mapview::mapview(tmp) +
  mapview::mapview(meow)
subset <- mapedit::editMap(meow_plot)
schile <- st_intersection(subset$drawn, tmp)
mapview::mapview(schile)
st_write(schile, here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-CS.shp"), append=FALSE)

## BSQLOBSTERNCH
tmp <- filter(eez, GEONAME == "Chilean Exclusive Economic Zone")
meow_plot <- mapview::mapview(tmp) +
  mapview::mapview(schile)
subset <- mapedit::editMap(meow_plot)
nchile <- st_intersection(subset$drawn, tmp)
mapview::mapview(nchile)
st_write(nchile, here("prep/fisheries/data/maps/Chile-SUBPESCA-CH-N.shp"), append=FALSE)


## BRNROCKPCOAST
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
plot_tmp <- mapview::mapview(tmp)
subset <- mapedit::editMap(plot_tmp)
uswc <- st_intersection(subset$drawn, tmp)
mapview::mapview(uswc)
st_write(uswc, here("prep/fisheries/data/maps/USA-NMFS-PCOAST.shp"), append=FALSE)


## NMPHS SATLC
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
plot_tmp <- mapview::mapview(tmp)
subset <- mapedit::editMap(plot_tmp)
segm <- st_intersection(subset$drawn, tmp)
mapview::mapview(segm)

st_write(segm, here("prep/fisheries/data/maps/USA-NMFS-SATLC.shp"), append=FALSE)
st_write(segm, here("prep/fisheries/data/maps/USA-NMFS-SATL.shp"), append=FALSE)



## BLSHARIO
tmp <- filter(iho, NAME == "Indian Ocean")
mapview::mapview(tmp)
st_write(tmp, here("prep/fisheries/data/maps/multinational-IOTC-IO.shp"))

## BLACKROCKWASH
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
plot_tmp <- mapview::mapview(tmp)
#subset <- mapedit::editMap(plot_tmp)
wa <- st_intersection(subset$drawn, tmp)
mapview::mapview(wa)
st_write(wa, here("prep/fisheries/data/maps/USA-NMFS-WASH.shp"), append=FALSE)


## BLACKROCKCAL
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
mapview::mapview(tmp)
#subset <- mapedit::editMap()
ca <- st_intersection(subset, tmp)
mapview::mapview(ca)
st_write(ca, here("prep/fisheries/data/maps/USA-NMFS-CAL.shp"))

## BLACKROCKORECOAST
eez <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
mapview::mapview(eez)
m <- mapview::mapview(eez) +
  mapview::mapview(wa) + 
  mapview::mapview(ca)
tmp <- mapedit::editMap(m)
or <- st_intersection(eez, tmp$drawn)

mapview::mapview(or)
st_write(or, here("prep/fisheries/data/maps/USA-NMFS-ORECOAST.shp"))


## BIGSKAGA

goa <- filter(meow, ECOREGION %in% c("Gulf of Alaska"))

mapview::mapview(goa)

st_write(goa, here("prep/fisheries/data/maps/USA-NMFS-GA.shp"), overwrite=TRUE)



## BGRDRNSWWA

boundary <- filter(eez, GEONAME == "Australian Exclusive Economic Zone")
mapview::mapview(boundary)
st_write(boundary, here("prep/fisheries/data/maps/Australia-AFMA-NSWWA.shp"))



# AUROCKPCOAST
tmp <- filter(eez, GEONAME == "United States Exclusive Economic Zone")
mapview::mapview(tmp)
#subset <- mapedit::editMap()
pacific_coast <- st_intersection(subset, tmp)
mapview::mapview(pacific_coast)
st_write(pacific_coast, here("prep/fisheries/data/maps/USA-NMFS-PCOAST.shp"))

## APOLLNSJ
boundary <- filter(with_area, areaid =="Japan-FAJ-NSJ")
mapview::mapview(boundary)
st_write(boundary, here("prep/fisheries/data/maps/Japan-FAJ-NSJ.shp"))

## ANCHONWA
boundary <- filter(meow, ECOREGION %in% c("Azores Canaries Madeira", "Saharan Upwelling", "Sahelian Upwelling", "Cape Verde"))
boundary <- st_union(boundary)
mapview::mapview(boundary)
st_write(boundary, here("prep/fisheries/data/maps/multinational-FAO-NWA.shp"))

## ACMACKSARG
boundary <- filter(with_area, areaid =="Argentina-CFP-ARG-S")
plot(st_geometry(boundary))
st_write(boundary[1, ], here("prep/fisheries/data/maps/Argentina-CFP-ARG-S.shp"))

# ALSKABSAI

ak_eez <-filter(eez, GEONAME=="United States Exclusive Economic Zone (Alaska)")
meow_bsai <- filter(meow, ECOREGION %in% c("Aleutian Islands", "Eastern Bering Sea"))

mapview::mapview(meow_bsai) + 
  mapview::mapview(ak_eez)
bsai <- st_intersection(meow_bsai, ak_eez)
bsai <- st_union(bsai)
mapview::mapview(bsai)

st_write(bsai, here("prep/fisheries/data/maps/USA-NMFS-BSAI.shp"), overwrite=TRUE)
  

