---
title: "Downloading and prepping extreme air temperature events"
output: html_document
date: '2022-08-09'
---

I will be using the IPCC ensemble models for number of days >35 and >40 C. 

I originally attempted to use the Xu data (used for SST data), but this will not work for this purpose.

I format these data in two ways: 
1) proportion of days within a year (based on days) in which temperature exceeds 35.
2) proportion of months within a year in which at least 2 days exceed 35.

There is no need to gapfill these data because it extends all the way into the ocean.  However, the IPCC website indicates low model agreement for global coastlines!

```{r setup, include=FALSE}
library(raster)
library(ncdf4)
library(here)
library(tidyverse)
library(foreach)
library(doParallel)

#year ranges and scenarios:
yr_35_range <- data.frame(filename = c("TX35  days - Long Term (2081-2100) SSP1-2.6",
                                    "TX35  days - Long Term (2081-2100) SSP3-7.0",
                                    "TX35  days - Long Term (2081-2100) SSP5-8.5",
                                    "TX35  days - Medium Term (2041-2060) SSP1-2.6",
                                    "TX35  days - Medium Term (2041-2060) SSP3-7.0",
                                    "TX35  days - Medium Term (2041-2060) SSP5-8.5",
                                    "TX35  days - Near Term (2021-2040) SSP1-2.6",
                                    "TX35  days - Near Term (2021-2040) SSP3-7.0",
                                    "TX35  days - Near Term (2021-2040) SSP5-8.5"),
                       scenario = c("ssp126", "ssp370", "ssp585", "ssp126", "ssp370", "ssp585", "ssp126", "ssp370", "ssp585"),
                       time_frame = c("long-term", "long-term", "long-term", "medium-term", "medium-term", "medium-term",
                                      "near-term", "near-term", "near-term"))


yr_40_range <- data.frame(filename = c("TX40  days - Long Term (2081-2100) SSP1-2.6",
                                    "TX40  days - Long Term (2081-2100) SSP5-8.5",
                                    "TX40  days - Medium Term (2041-2060) SSP1-2.6",
                                    "TX40  days - Medium Term (2041-2060) SSP5-8.5",
                                    "TX40  days - Near Term (2021-2040) SSP1-2.6",
                                    "TX40  days - Near Term (2021-2040) SSP5-8.5"),
                       scenario = c("ssp126", "ssp585", "ssp126", "ssp585", "ssp126",  "ssp585"),
                       time_frame = c("long-term", "long-term", "medium-term", "medium-term",
                                       "near-term", "near-term"))



rast_base_10km_file <- terra::rast(here('_spatial/rast_base_mol_10km.tif'))
ocean <- terra::rast(here('_spatial/ocean_area_mol.tif'))
land_coast <- terra::ifel(ocean<1,  1, 0)
land <- terra::ifel(is.na(land_coast), 1, land_coast)
plot(land)
```


I tried to use the Xu data, but it is not going to work for this purpose. The near ground air temperatures
are only monthly (need higher temporal resolution to get extreme events) and the atmospheric air temperatures do not 
align with near ground temps. So going back to the drawing board on this one.
```{r}
tmp <- nc_open("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_1979_01.nc4")
names(tmp$var)
tmp$dim$time$vals # origin 1850-1-1
tmp$dim$vals
tmp <- raster::raster("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_2000_01.nc4", varname="ps")
cellStats(tmp, "mean")

ocean_proj <- raster(ocean_mask)
ocean_proj <- projectRaster(ocean_proj, tmp)

tmp_mask <- ocean_proj*tmp
cellStats(tmp_mask, "mean", na.rm=TRUE)
plot(tmp_mask/100)
ta_1 <- terra::rast("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_2000_06.nc4")
ta_1 <- ta_1[[grep("ta_", names(ta_1))]]
ta_1 <- ta_1[[grep("lev=100000_", names(ta_1))]]
ta_1 <- mean(ta_1)



humidity_1 <- raster::brick("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_2000_06.nc4", varname="hur", level=1) 
humidity_14 <- raster::brick("/home/shares/ohi/stressors_2021/_raw_data/sst_Xu/atm_hist_2000_06.nc4", varname="hur", level=14)
plot(humidity_1[[1]])
plot(humidity_14[[1]])

land_temp <- terra::rast("/home/shares/ohi/stressors_2021/_raw_data/land_temp_Xu/lnd.historical.200006.nc")
land_temp <- land_temp[[grep("tas", names(land_temp))]]
plot(land_temp - 273)
plot(ta_1 -273)

plot(land_temp - ta_1)

system("ncdump -h /home/shares/ohi/stressors_2021/_raw_data/land_temp_Xu/lnd.historical.198905.nc")
system("ncdump -h /home/shares/ohi/stressors_2021/_raw_data/land_temp_Xu/lnd.historical.198905.nc")

```


## Proportion of days

35C
```{r}
# scenarios
for (file in yr_35_range$filename){ #file = yr_35_range$filename[1]

all_files <- list.files("/home/shares/ohi/stressors_2021/_raw_data/Tair_35d_ipcc", full=TRUE)

file_path <- grep(file, all_files, fixed=TRUE, value=TRUE)
cat("N= ", length(file_path), " raster file was discovered")

raster <- terra::rast(file_path)
scaled_raster <- raster/365

proj_scaled_raster <- terra::project(scaled_raster, rast_base_10km_file, method="near")

scen <- yr_35_range$scenario[yr_35_range$filename==file]
timeframe <-  yr_35_range$time_frame[yr_35_range$filename==file]

writeRaster(proj_scaled_raster, filename = sprintf("/home/shares/ohi/stressors_2021/_dataprep/T_air/proportion_days_35C/days-35_%s_%s.tif", scen, timeframe), overwrite = TRUE)

}

#historical
scen <- terra::rast("/home/shares/ohi/stressors_2021/_raw_data/Tair_35d_ipcc/CMIP6 - Bias Adjusted TX35  days - Near Term (2021-2040) SSP1-2.6  - Annual (24 models).tiff")
hist_diff <- terra::rast( "/home/shares/ohi/stressors_2021/_raw_data/Tair_35d_ipcc/CMIP6 - Bias Adjusted TX35 Change days - Near Term (2021-2040) SSP1-2.6 (rel. to 1995-2014) - Annual (24 models).tiff")
historical <- scen - hist_diff

scaled_historical <- historical/365

proj_scaled_historical <- terra::project(scaled_historical, rast_base_10km_file, method="near")
writeRaster(proj_scaled_historical, "/home/shares/ohi/stressors_2021/_dataprep/T_air/proportion_days_35C/days-35_NA_historical.tif", overwrite = TRUE)

```


40C
```{r}

# scenarios
for (file in yr_40_range$filename){ #file = yr_40_range$filename[1]

all_files <- list.files("/home/shares/ohi/stressors_2021/_raw_data/Tair_40d_ipcc", full=TRUE)

file_path <- grep(file, all_files, fixed=TRUE, value=TRUE)
cat("N= ", length(file_path), " raster file was discovered")

raster <- terra::rast(file_path)
scaled_raster <- raster/365

proj_scaled_raster <- terra::project(scaled_raster, rast_base_10km_file, method="near")

scen <- yr_40_range$scenario[yr_40_range$filename==file]
timeframe <-  yr_40_range$time_frame[yr_40_range$filename==file]

writeRaster(proj_scaled_raster, filename = sprintf("/home/shares/ohi/stressors_2021/_dataprep/T_air/proportion_days_40C/days-40_%s_%s.tif", scen, timeframe), overwrite = TRUE)

}
#historical
scen <- terra::rast("/home/shares/ohi/stressors_2021/_raw_data/Tair_40d_ipcc/CMIP6 - Bias Adjusted TX40  days - Near Term (2021-2040) SSP1-2.6  - Annual (24 models).tiff")
hist_diff <- terra::rast( "/home/shares/ohi/stressors_2021/_raw_data/Tair_40d_ipcc/CMIP6 - Bias Adjusted TX40 Change days - Near Term (2021-2040) SSP1-2.6 (rel. to 1995-2014) - Annual (24 models).tiff")
historical <- scen - hist_diff

scaled_historical <- historical/365

proj_scaled_historical <- terra::project(scaled_historical, rast_base_10km_file, method="near")
writeRaster(proj_scaled_historical, "/home/shares/ohi/stressors_2021/_dataprep/T_air/proportion_days_40C/days-40_NA_historical.tif", overwrite = TRUE)

```

## Proportion of months

35C


First calculate exceedences (>= 2 days in a month) of the historical data.  

```{r}

all_files <- list.files("/home/shares/ohi/stressors_2021/_raw_data/Tair_35d_ipcc/monthly", full=TRUE)

month_list <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

for(month in month_list){ # month = month_list[1]

# Define the directory where the files are located
directory <- "/home/shares/ohi/stressors_2021/_raw_data/Tair_35d_ipcc/monthly"
# Define the file name pattern
pattern <- paste0("CMIP6 - Bias Adjusted TX35  days - Near Term \\(2021-2040\\) SSP1-2\\.6  - ", month, " \\(.*models\\)\\.tiff$")
# List the files that match the pattern
scen_path <- list.files(path = directory, pattern = pattern, full=TRUE)
scen_raster <- terra::rast(scen_path)


pattern <- paste0("CMIP6 - Bias Adjusted TX35 Change days - Near Term \\(2021-2040\\) SSP1-2\\.6 \\(rel\\. to 1995-2014\\) - ", month, " \\(.*models\\)\\.tiff$")
hist_path <- list.files(path = directory, pattern = pattern, full=TRUE)
hist_diff_raster <- terra::rast(hist_diff)

cat("For month = ", month, ", hist_diff is = ", class(hist_diff_raster)[1], "and scen_raster is = ", class(scen_raster)[1], "\n")

historical <- scen_raster - hist_diff_raster
historical_exceed <- terra::ifel(historical>=2, 1, 0)

proj_exceed_historical <- terra::project(historical_exceed, rast_base_10km_file, method="near")


writeRaster(proj_exceed_historical, sprintf("/home/shares/ohi/stressors_2021/_dataprep/T_air/proportion_months_35C/exceed_2days_monthly/month-exceed-35_%s_NA_historical.tif", month), overwrite = TRUE)
}

```


First calculate exceedences (>= 2 days in a month) of the other data.  

```{r}

all_files <- list.files("/home/shares/ohi/stressors_2021/_raw_data/Tair_35d_ipcc/monthly", full=TRUE, pattern = "TX35  days")


for(file in all_files){ # file = all_files[2]

saveName <- basename(file)
scen_raster <- terra::rast(file)
max_scen_raster <- terra::global(scen_raster, "max")$max
cat(saveName, " has max = ", max_scen_raster, "\n")

scen_raster_exceed <- terra::ifel(scen_raster>=2, 1, 0)

proj_exceed_scen <- terra::project(scen_raster_exceed, rast_base_10km_file, method="near")


writeRaster(proj_exceed_scen, sprintf("/home/shares/ohi/stressors_2021/_dataprep/T_air/proportion_months_35C/exceed_2days_monthly/month-exceed_%s", saveName), overwrite = TRUE)
}

```




