---
title: "Untitled"
output: html_document
date: "2024-07-17"
---

```{r setup, include=FALSE}

library(tidyverse)
library(terra)
library(here)

```

Save final pressure files to consistent file location and adjust filenaming.
```{r}

pressure_save <- "/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures"
pressure_list <- read_csv(here("combining_pressures/pressure_list.csv"))

files_to_save <- paste0(pressure_list$f_path, "/", pressure_list$fn_pressure, "_",  pressure_list$fn_scenario, "_", pressure_list$fn_time, ".tif")
files_new_loc <- paste0(pressure_save, "/", pressure_list$new_fn_pressure, "_", pressure_list$fn_scenario, "_", pressure_list$fn_time, ".tif")
for (i in seq_along(files_to_save)) { # i = 178
  file.copy(from = files_to_save[i], to = files_new_loc[i], overwrite = TRUE)
}

length(list.files(pressure_save)) # currently should be 203

```


Calculate raster differences
```{r}

files_df <- data.frame(fp = list.files("/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures", full=TRUE))
files_df$fn <- basename(files_df$fp) 
files_df <- files_df %>%
  tidyr::separate(fn, into=c("df_pressure", "df_scenario", "df_time_frame"), extra="merge", sep="_") %>%
  mutate(df_time_frame = gsub(".tif", "", df_time_frame))

pressure_list = unique(files_df$df_pressure)

for(pressure in pressure_list){ # pressure = pressure_list[1]
  tmp_list <- filter(files_df, df_pressure  == pressure)
  
  tmp_scenarios = unique(tmp_list$df_scenario)
  for(scenario in tmp_scenarios){ #scenario = tmp_scenarios[1]
    tmp_list_scenario = filter(tmp_list, df_scenario == scenario)
    current_path <- grep("current", tmp_list_scenario$fp, value=TRUE)
    current_rast <- terra::rast(current_path)
    near_path <- grep("near-term", tmp_list_scenario$fp, value=TRUE)
    near_rast <- terra::rast(near_path)
    medium_path <- grep("medium-term", tmp_list_scenario$fp, value=TRUE)
    medium_rast <- terra::rast(medium_path)
    near_change <- near_rast-current_rast
    writeRaster(near_change, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/change/%s_%s_near-change.tif", pressure, scenario), overwrite=TRUE)
    medium_change <- medium_rast-current_rast
    writeRaster(medium_change, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/change/%s_%s_medium-change.tif", pressure, scenario), overwrite=TRUE)
    
    if(length(grep("long-term", tmp_list_scenario$fp, value=TRUE))>0){
    long_path <- grep("long-term", tmp_list_scenario$fp, value=TRUE)
    long_rast <- terra::rast(long_path)  
    long_change <- long_rast-current_rast
    writeRaster(long_change, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/change/%s_%s_long-change.tif", pressure, scenario), overwrite=TRUE)
    }
  }
}

```


Obtain cumulative pressures
```{r}

files_df <- data.frame(fp = list.files("/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures", full=TRUE))
files_df$fn <- basename(files_df$fp) 
files_df <- files_df %>%
  tidyr::separate(fn, into=c("df_pressure", "df_scenario", "df_time_frame"), extra="merge", sep="_") %>%
  mutate(df_time_frame = gsub(".tif", "", df_time_frame))

cumulative_list <- expand.grid(scenario = c("ssp245", "ssp585"),
                               time_frame = c("current", "near-term", "medium-term"))

for (i in 1:dim(cumulative_list)[1]){ # i = 1
  scenario_time <- cumulative_list[i,]
  scenario = scenario_time$scenario
  scenario2 = substring(scenario, 1, 4)
  timeframe = scenario_time$time_frame
  file_list <- filter(files_df, 
                      df_scenario == scenario | df_scenario == scenario2,
                      df_time_frame == timeframe)
  #file_list <- filter(df_pressure != "oxygen-deep") # decided to keep for now, question about this!
   cat(as.character(scenario), " ", as.character(timeframe), " ", length(file_list$df_pressure), "\n")
  stack <- rast(file_list$fp)
  cumulative <- app(stack, "sum", na.rm=TRUE)
  writeRaster(cumulative, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/%s_%s.tif", scenario, timeframe))
  
}

#check that all outputs have N=11 (or, whatever!)

list.files("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/", full=TRUE)

ssp245_t1 <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp245_near-term.tif") - 
  rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp245_current.tif")
plot(ssp245_t1)
writeRaster(ssp245_t1, "/home/shares/ohi/stressors_2021/combining_pressures/cumulative_change/cumulative_ssp245_near-change.tif")

ssp245_t2 <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp245_medium-term.tif") - 
  rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp245_current.tif")
plot(ssp245_t2)
writeRaster(ssp245_t2, "/home/shares/ohi/stressors_2021/combining_pressures/cumulative_change/cumulative_ssp245_medium-change.tif")


ssp585_t1 <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp585_near-term.tif") - 
  rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp585_current.tif")
plot(ssp585_t1)
writeRaster(ssp585_t1, "/home/shares/ohi/stressors_2021/combining_pressures/cumulative_change/cumulative_ssp585_near-change.tif")

ssp585_t2 <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp585_medium-term.tif") - 
  rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative//ssp585_current.tif")
plot(ssp585_t2)
writeRaster(ssp585_t2, "/home/shares/ohi/stressors_2021/combining_pressures/cumulative_change/cumulative_ssp585_medium-change.tif")


```

Make cumulative figures
```{r}

library(paletteer)
library(png)

cols_palette <- c(paletteer_c("ggthemes::Sunset-Sunrise Diverging", 100), "#D14644FF")

upperscale <- 6

tmp <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/cumulative", full=TRUE)
stack <- rast(tmp)
stack_capped <- ifel(stack>upperscale, upperscale, stack)

for(i in 1:dim(stack_capped)[3]){ #i = 1

png(filename = here(sprintf("combining_pressures/figs_cumulative/%s.png", 
             basename(tmp[i]))), width = 800, height = 600)

par(mar = c(0, 0, 0, 0))

plot(stack_capped[[i]], col=cols_palette, main=gsub(".tif", "", basename(tmp[i])),
     axes=FALSE, box=FALSE, zlim=c(0,upperscale), breaks = seq(0, upperscale, length.out = 101), legend=FALSE)
#legend_image <- as.raster(matrix(cols_palette, ncol=1))
 # par(new = TRUE, mar = c(5, 5, 5, 7), xpd = TRUE)
  #plot(c(0, 1), c(0, 6), type = "n", xaxs = "i", yaxs = "i", xaxt = "n", yaxt = "n", xlab = "", ylab = "")
  #rasterImage(legend_image, 1.05, 0, 1.15, 6)
  #axis(4, at = seq(0, 6, length.out = 6), labels = seq(0, 6, length.out = 6), las = 1)
dev.off()
}


```

Make cumulative difference raster plots.

```{r}

tmp <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/cumulative", full=TRUE)

ssp245_near_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_near-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_current.tif")
ssp245_medium_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_medium-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_current.tif")
ssp585_near_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_near-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_current.tif")
ssp585_medium_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_medium-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_current.tif")

diff_function <- function(diff_raster, filename){
  # diff_raster = ssp245_near_change
  # filename = "ssp585_medium-change"
  
  diff_raster <- ifel(diff_raster>3, 3, diff_raster)
  png(filename = here(sprintf("combining_pressures/figs_cumulative_diff/%s.png", 
             filename)), width = 800, height = 600)

par(mar = c(0, 0, 0, 0))

cols_palette <- c(paletteer_c("grDevices::BluGrn", 100),
                  rev(paletteer_c("grDevices::Inferno", 301)))

plot(diff_raster, col=cols_palette, main=filename, zlim=c(-1, 3),
     axes=FALSE, box=FALSE, breaks = seq(-1, 3, length.out = 401), legend=FALSE)

dev.off()
}

diff_function(ssp245_near_change, filename="ssp245_near-change")
diff_function(ssp245_medium_change, filename="ssp245_medium-change")
diff_function(ssp585_near_change, filename="ssp585_near-change")
diff_function(ssp585_medium_change, filename="ssp585_medium-change")