---
title: "paper_statistics_v2"
output: html_document
date: "2024-08-09"
---

```{r}
library(terra)
library(tidyverse)
library(here)
library(cowplot)
library(ggplot2)
library(paletteer)
library(countrycode)

```



## EEZ region and country summaries
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))
## cut land
ocean_cells <- rast(here("_spatial/ocean_area_mol.tif"))
ocean_cells <- ifel(ocean_cells>0, 1, 0)
## add in the coastal area
eez_coast <- rast(here("_spatial/coastal_country_mol.tif"))
eez_coast <- ifel(eez_coast>0, 1, 0)
## combine coast and ocean
ocean_area <- app(c(ocean_cells, eez_coast), "sum", na.rm=TRUE)
ocean_area <- ifel(ocean_area>0, 1, NA)
plot(ocean_area)

#get the area
eez <- eez*ocean_area
plot(eez)


eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(iso3c != "GUM") # cut out duplicated dual region

pressure_files <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures", full=TRUE)

pressure_stack <- rast(pressure_files)
source_info <- gsub(".tif", "", basename(sources(pressure_stack)))
names(pressure_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(pressure_stack)) { # i = 1
  # Extract the current layer
  current_layer <- pressure_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, eez, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
eez_pressure_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

eez_pressure_result <- eez_pressure_result %>%
 pivot_longer(cols=-rgn_id, names_to = "pressure", values_to="value")


eez_pressure_result <- eez_pressure_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
 eez_cell_num <- zonal(ifel(eez>0, 1, 0), eez, fun = "sum", na.rm=TRUE)
 names(eez_cell_num) <- c("rgn_id", "cells")

dim(eez_pressure_result)
eez_pressure_result <- eez_pressure_result %>%
  left_join(eez_cell_num, by="rgn_id") %>%
  mutate(region = "eez") 


write_csv(eez_pressure_result, here("combining_pressures/data/country_eez_pressure.csv"))

```


## coastal region and country summaries
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))
coast <- rast(here("_spatial/coastal_country_mol.tif"))
coast <- ifel(eez_coast>0, 1, 0)
#get the area
country_coast <- eez*coast
plot(country_coast)


eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  filter(iso3c != "GUM") # cut out duplicated dual region

pressure_files <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures", full=TRUE)

pressure_stack <- rast(pressure_files)
source_info <- gsub(".tif", "", basename(sources(pressure_stack)))
names(pressure_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(pressure_stack)) { # i = 1
  # Extract the current layer
  current_layer <- pressure_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, country_coast, fun = "sum", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
coast_pressure_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

coast_pressure_result <- coast_pressure_result %>%
 pivot_longer(cols=-rgn_id, names_to = "pressure", values_to="value")


coast_pressure_result <- coast_pressure_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label)) %>%
   filter(rgn_label != "disputed")

# get number of cells in each eez:
coast_cell_num <- zonal(ifel(country_coast>0, 1, 0), country_coast, fun = "sum", na.rm=TRUE)
 names(coast_cell_num) <- c("rgn_id", "cells")

dim(coast_pressure_result)
coast_pressure_result <- coast_pressure_result %>%
  left_join(coast_cell_num, by="rgn_id") %>%
  mutate(region = "coast") 


write_csv(coast_pressure_result, here("combining_pressures/data/country_coast_pressure.csv"))

```


Make a stacked bar plot of regions
```{r}

coast <- read_csv(here("combining_pressures/data/country_coast_pressure.csv"))
eez <- read_csv(here("combining_pressures/data/country_eez_pressure.csv"))

pressure_country <- rbind(coast, eez)

pressure_rgn <- pressure_country %>%
  group_by(pressure, geosubrgn, region) %>%
  summarize(value = sum(value, na.rm=TRUE),
            cells = sum(cells, na.rm=TRUE))

pressure_rgn <- pressure_rgn %>%
  mutate(avg_value = value/cells) %>%
  separate(pressure, into = c("pressure", "scenario", "time_period"), sep = "_")

plot_df <- pressure_rgn %>%
  filter(scenario %in% c("ssp245", "ssp2")) %>%
  select(pressure, time_period, geosubrgn, region, avg_value)

# Spread the data by time_period
plot_df_wide <- plot_df %>%
  pivot_wider(names_from = time_period, values_from = avg_value)

plot_df2 <- plot_df_wide %>%
  mutate(medium_change = `medium-term` - current) %>%
  select(pressure, geosubrgn, region, `medium-term`, medium_change) %>%
  pivot_longer(cols = c(`medium-term`, medium_change), 
               names_to = "variable", 
               values_to = "value")

colors <- read_csv(here("combining_pressures/data/pressure_colors.csv")) %>%
  mutate(region = ifelse(region=="coastal", "coast", "eez")) %>%
  filter(region =="eez") %>%
  select(color, pressure)

plot_df2 <- plot_df2 %>%
  left_join(colors, by = c("pressure"))

# Step 1: Summarize the data to get total values for each pressure within each geosubrgn
pressure_order <- plot_df2 %>%
  filter(variable == "medium_change") %>%
  group_by(pressure) %>%
  summarize(total_value = sum(`value`), .groups = 'drop') %>%
  arrange(total_value) %>%
  mutate(pressure = factor(pressure, levels = unique(pressure)))

subregion_order <- plot_df2 %>%
  filter(variable == "medium_change") %>%
  filter(region == "coast") %>%
  group_by(geosubrgn) %>%
  summarize(total_value = sum(value), .groups='drop') %>%
  arrange(desc(total_value)) %>%
  mutate(geosubrgn = factor(geosubrgn, levels = unique(geosubrgn)))

 # Step 2: Reorder the 'pressure' factor levels based on the total_value
plot_df2 <- plot_df2 %>%
  mutate(pressure = factor(pressure, levels = unique(pressure_order$pressure))) %>%
  mutate(geosubrgn = factor(geosubrgn, levels = unique(subregion_order$geosubrgn))) %>%
  arrange(region, geosubrgn, pressure)

# Step 3: Plot the data

plots_bars <- ggplot(filter(plot_df2, region == "coast"), aes(y = geosubrgn, x = value, fill = pressure)) +
  geom_bar(stat = "identity") +
  facet_wrap(~variable, nrow = 1) +
  labs(x = "", y = "", title = "", fill = NULL) +  # fill = NULL removes the legend title
  theme_minimal() +
  scale_fill_manual(values = setNames(plot_df2$color, plot_df2$pressure)) +
  theme(
    strip.text = element_blank(),  # Removes the facet strip titles
    legend.title = element_blank() # Removes the legend title
  )

ggsave(here("combining_pressures/paper_figs/georgn_pressure_change_barplot.png"), plot = plots_bars, width = 10, height = 4, dpi = 300)
```




## Pressures vs. dependence

Format marine dependence data.

https://conbio.onlinelibrary.wiley.com/doi/full/10.1111/conl.12617
```{r}

depend <- read_csv(here("prep/fisheries/data/marine_dependence_Selig.csv")) %>%
  select(country = Country, 
         integrated_depend = `Integrated dependence`,
         econ_depend = `Economic dependence`, 
         nutrition_depend = `Nutritional dependence`) %>%
  slice(-1)
depend$iso3c <- countrycode(depend$country, "country.name", "iso3c") 

depend$iso3c[depend$country=="Saint Martin"] <- "NSM"

write_csv(depend, here("combining_pressures/data/selig_marine_dependence.csv"))

```



coastline: Relationship between dependence on marine environments and cumulative pressures

```{r}
eez_coast <- rast(here("_spatial/coastal_country_mol.tif"))

cumulative <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/", full=TRUE)

cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(cum_stack)) {
  # Extract the current layer
  current_layer <- cum_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, eez_coast, fun = "mean", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
coastal_combined_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)


eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  select(rgn_id, rgn_label) %>%
  unique()

coastal_combined_result <- coastal_combined_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label))

coastal_combined_result$iso3c <- countrycode(coastal_combined_result$rgn_label, "country.name", "iso3c") 
coastal_combined_result <- filter(coastal_combined_result, rgn_label != "Bassas da India")


double_rgns <- data.frame(rgn_label = c("Guadeloupe and Martinique", "Guadeloupe and Martinique",
              "Northern Mariana Islands and Guam", "Northern Mariana Islands and Guam",
              "Puerto Rico and Virgin Islands of the United States", "Puerto Rico and Virgin Islands of the United States"),
                 iso3c = c("GLP", "MTQ", "GUM", "MNP", "PRI", "VIR"))

coastal_double_eez <- filter(coastal_combined_result, rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")) %>%
  select(-iso3c)

coastal_double_eez2 <- left_join(double_rgns, coastal_double_eez, by="rgn_label")

coastal_combined_result <- filter(coastal_combined_result, !(rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")))

coastal_eez_cum <- rbind(coastal_combined_result, coastal_double_eez2)
coastal_eez_cum$iso3c[coastal_eez_cum$rgn_label=="Northern Saint-Martin"] <- "NSM"
coastal_eez_cum$iso3c[coastal_eez_cum$rgn_label=="Micronesia"] <- "FSM"
coastal_eez_cum$iso3c[coastal_eez_cum$rgn_label %in% c("Crozet Islands",
                                       "Kerguelen Islands",
                                       "Amsterdam Island and Saint Paul Island",
                                       "Ile Europa",
                                       "Ile Tromelin",
                                       "Juan de Nova Island",
                                       "Glorioso Islands")] <- "ATF"
coastal_eez_cum <- filter(coastal_eez_cum, iso3c != "ATF")

coastal_eez_cum2 <- coastal_eez_cum %>%
  filter(!is.na(iso3c)) %>%
  select(-rgn_id, -rgn_label)

write_csv(coastal_eez_cum2, here("combining_pressures/data/coast_cumulative.csv"))

```

Combine with coastal dependence data
```{r}
coastal_eez_cum2 <- read_csv(here("combining_pressures/data/coast_cumulative.csv"))
depend <- read_csv(here("combining_pressures/data/selig_marine_dependence.csv"))

missing_iso3c <- setdiff(depend$iso3c, coastal_eez_cum2$iso3c)
filter(depend, iso3c %in% missing_iso3c) %>% print(n=25)
setdiff(coastal_eez_cum2$iso3c, depend$iso3c)

coastal_compare <- left_join(depend, coastal_eez_cum2, by="iso3c") %>%
  filter(!is.na(ssp585_current))

coastal_compare <- coastal_compare %>%
  mutate(change_ssp245_medium = `ssp245_medium-term` - ssp245_current) %>%
  mutate(pct_change_ssp245_medium = (`ssp245_medium-term` - ssp245_current)/ssp245_current) %>%
    mutate(change_ssp585_medium = `ssp585_medium-term` - ssp585_current) %>%
    mutate(pct_change_ssp585_medium = (`ssp585_medium-term` - ssp585_current)/ssp585_current)

write_csv(coastal_compare, here("combining_pressures/data/country_coastal_cumulative_pressures.csv"))

```


Plot and analyze the data
```{r}

coastal_compare <- read_csv(here("combining_pressures/data/country_coastal_cumulative_pressures.csv")) %>%
  mutate(integrated_depend = ifelse(integrated_depend > 0.8, 0.8, integrated_depend))

p <- ggplot(data=coastal_compare , aes(x=`ssp245_medium-term`, y=change_ssp245_medium, color=integrated_depend, size=integrated_depend)) +
  geom_point(aes(text=country), alpha=0.7) +
  scale_color_gradientn(colors=paletteer_c("ggthemes::Temperature Diverging", 30)) +
  geom_hline(yintercept = median(coastal_compare$change_ssp245_medium, na.rm=TRUE), linetype="dashed", color="orange") +
   geom_vline(xintercept = median(coastal_compare$`ssp245_medium-term`, na.rm=TRUE), linetype="dashed", color="orange") +
  theme_minimal()

p

plotly::ggplotly(ptooltip = "text")

mod <- lm(change_ssp245_medium ~ integrated_depend, data=coastal_compare)
summary(mod)
mod <- lm(`ssp245_medium-term` ~ integrated_depend, data=coastal_compare)
summary(mod)

ggsave(here("combining_pressures/paper_figs/current_future_scatterplot.png"), plot = p, width = 7, height = 4, dpi = 300)

```


Doing the same thing, but for EEZ.
EEZ: Relationship between dependence on marine environments and cumulative pressures
```{r}

eez <- rast(here("_spatial/eez_mol_w_land.tif"))
eez_id <- read_csv(here("_spatial/eez_rgn_names/rgn_names_ohi.csv")) %>%
  select(rgn_id, rgn_label) %>%
  unique()

cumulative <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/", full=TRUE)

cum_stack <- rast(cumulative)
source_info <- gsub(".tif", "", basename(sources(cum_stack)))
names(cum_stack) <- source_info

result_list <- list()

# Loop through each layer in the cumulative stack
for (i in 1:nlyr(cum_stack)) {
  # Extract the current layer
  current_layer <- cum_stack[[i]]
  
  # Compute the sum for each region using the ID raster
  zonal_sum <- zonal(current_layer, eez, fun = "mean", na.rm=TRUE)
  
  # Add the result to the list
  result_list[[i]] <- zonal_sum
}

# Merge all data frames by region ID
combined_result <- Reduce(function(x, y) merge(x, y, by = "rgn_id"), result_list)

combined_result <- combined_result %>%
  left_join(eez_id, by="rgn_id") %>%
  filter(!is.na(rgn_label))

eez_cum <- combined_result 

eez_cum$iso3c <- countrycode(eez_cum$rgn_label, "country.name", "iso3c") 
eez_cum <- filter(eez_cum, rgn_label != "Bassas da India")

double_rgns <- data.frame(rgn_label = c("Guadeloupe and Martinique", "Guadeloupe and Martinique",
              "Northern Mariana Islands and Guam", "Northern Mariana Islands and Guam",
              "Puerto Rico and Virgin Islands of the United States", "Puerto Rico and Virgin Islands of the United States"),
                 iso3c = c("GLP", "MTQ", "GUM", "MNP", "PRI", "VIR"))

double_eez <- filter(eez_cum, rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")) %>%
  select(-iso3c)
double_eez2 <- left_join(double_rgns, double_eez, by="rgn_label")

eez_cum <- filter(eez_cum, !(rgn_label %in% c("Guadeloupe and Martinique",
                                               "Northern Mariana Islands and Guam",
                                               "Puerto Rico and Virgin Islands of the United States")))

eez_cum <- rbind(eez_cum, double_eez2)
eez_cum$iso3c[eez_cum$rgn_label=="Northern Saint-Martin"] <- "NSM"
eez_cum$iso3c[eez_cum$rgn_label=="Micronesia"] <- "FSM"
eez_cum$iso3c[eez_cum$rgn_label %in% c("Crozet Islands",
                                       "Kerguelen Islands",
                                       "Amsterdam Island and Saint Paul Island",
                                       "Ile Europa",
                                       "Ile Tromelin",
                                       "Juan de Nova Island",
                                       "Glorioso Islands")] <- "ATF"
eez_cum <- filter(eez_cum, iso3c != "ATF")

eez_cum2 <- eez_cum %>%
  filter(!is.na(iso3c)) %>%
  select(-rgn_id, -rgn_label) 

sum(duplicated(eez_cum2$iso3c))

write_csv(eez_cum, here("combining_pressures/data/eez_cumulative.csv"))

```

Combine eez and marine dependence data.
```{r}
eez_cumulative <- read_csv(here("combining_pressures/data/eez_cumulative.csv"))
depend <- read_csv(here("combining_pressures/data/selig_marine_dependence.csv"))

missing_iso3c <- setdiff(depend$iso3c, eez_cum2$iso3c)
filter(depend, iso3c %in% missing_iso3c) %>% print(n=25)
setdiff(eez_cum$iso3c, depend$iso3c)

compare <- left_join(depend, eez_cum2, by="iso3c") %>%
  filter(!is.na(ssp585_current))

compare <- compare %>%
  mutate(change_ssp245_medium = `ssp245_medium-term` - ssp245_current) %>%
  mutate(pct_change_ssp245_medium = (`ssp245_medium-term` - ssp245_current)/ssp245_current) %>%
    mutate(change_ssp585_medium = `ssp585_medium-term` - ssp585_current) %>%
    mutate(pct_change_ssp585_medium = (`ssp585_medium-term` - ssp585_current)/ssp585_current)

ggplot(data=compare , aes(x=integrated_depend, y=change_ssp245_medium)) +
  geom_point() 

ggplot(data=compare , aes(x=integrated_depend, y=`ssp245_current`)) +
  geom_point() 

ggplot(data=compare , aes(x=integrated_depend, y=`ssp245_medium-term`)) +
  geom_point() 

ggplot(data=compare , aes(y=change_ssp245_medium, x=`ssp245_current`, col=integrated_depend)) +
  geom_point() +
  scale_color_gradientn(colors = rev(rainbow(7)[1:6]))

mod <- lm(change_ssp245_medium ~ integrated_depend, data=compare)
summary(mod)
mod <- lm(`ssp245_medium-term` ~ integrated_depend, data=compare)
summary(mod)
mod <- lm(`ssp245_current` ~ integrated_depend, data=compare)
summary(mod)


```



# making a raster that shows vulnerable areas
```{r}

list.files("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/", full=TRUE)

ssp245_medium <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_medium-term.tif")

ssp245_near_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_near-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_current.tif")
ssp245_medium_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_medium-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp245_current.tif")
ssp585_near_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_near-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_current.tif")
ssp585_medium_change <- rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_medium-term.tif") - rast("/home/shares/ohi/stressors_2021/combining_pressures/cumulative/ssp585_current.tif")

plot(ssp245_medium)
terra::quantile(values(ssp245_medium, na.rm=TRUE), probs = c(0.5, 0.75, 0.9, 0.95))
terra::quantile(values(ssp245_medium_change, na.rm=TRUE), probs = c(0.5, 0.75, 0.9, 0.95))

high_ssp245_medium <- ifel(ssp245_medium>1.38, 1, 0)
high_ssp245_medium_change <- ifel(ssp245_medium_change > 0.76, 1, 0)

high_stress <- high_ssp245_medium + high_ssp245_medium_change
plot(high_stress)


```

Make and save change rasters.
```{r}

pressure_files <- list.files("/home/shares/ohi/stressors_2021/combining_pressures/rescaled_pressures", full=TRUE)
pressure_data <- gsub(".tif$", "", basename(pressure_files))

# Split the filenames by "_"
pressure_data <- strsplit(pressure_data, "_")

# Convert the list to a dataframe
pressure_dataframe <- do.call(rbind, lapply(pressure_data, function(x) data.frame(t(x), stringsAsFactors = FALSE)))

# Rename the columns
colnames(pressure_dataframe) <- c("pressure", "scenario", "time")

combos <- expand.grid(pressure = unique(pressure_dataframe$pressure),
            scenario = c("ssp2", "ssp5"))

for(i in 1:dim(combos)[1]){ # i = 7
  comparison <- combos[i, ]
 scenario_tmp <- comparison$scenario
 scenario_tmp_name <- ifelse(scenario_tmp == "ssp2", "ssp245", "ssp585")
 pressure_tmp <- paste0(comparison$pressure, "_")
 pressure_tmp_name <- gsub("_", "", pressure_tmp)
 pressure_files_tmp <- grep(pressure_tmp, pressure_files, value=TRUE)
 pressure_files_tmp <- grep(scenario_tmp, pressure_files_tmp, value=TRUE)

 current_tmp <- grep("current", pressure_files_tmp, value=TRUE)    
 near_tmp <- grep("near-term", pressure_files_tmp, value=TRUE)
 medium_tmp <- grep("medium-term", pressure_files_tmp, value=TRUE)
 
 near_change <- rast(near_tmp) - rast(current_tmp)
 writeRaster(near_change, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/pressure_change/%s_%s_near-change.tif", pressure_tmp_name, scenario_tmp_name), overwrite=TRUE)
 
  medium_change <- rast(medium_tmp) - rast(current_tmp)
 writeRaster(medium_change, sprintf("/home/shares/ohi/stressors_2021/combining_pressures/pressure_change/%s_%s_medium-change.tif", pressure_tmp_name, scenario_tmp_name), overwrite=TRUE)
 
  
}

```

